### 2023-02-07 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
first commit

---
title: Assessing linkages between watershed nutrient loading and estuary water quality in Lavaca Bay, Texas
author:
  - name: Michael Schramm
    affil: 1
    orcid: 0000-0003-1876-6592
affiliation:
  - num: 1
    address: |
      Texas A&M AgriLife Research - 
      Texas Water Resources Institute
      1001 Holleman Dr. E. 
      College Station, TX 77840-2118
    email: michael.schramm@ag.tamu.edu
# author citation list in chicago format
authorcitation: |
  Schramm, M.
correspondence: |
  michael.schramm@ag.tamu.edu
# specify the mdpi cls and location of file
# as of 2021 the template files are located in
# a subdirectory and should be retained there 
# since file locations are hard coded in the cls file.
cls: Definitions/mdpi
# class options
classoption:
  - water #first option is the journal name
  - article #second option is the type of manuscript
  - submit #do not change
  - oneauthor #change to oneauthor for single author papers
  #- pdftex #remove if compiling with XeLaTeX
# front matter
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
# back matter
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
funding: |
  This paper was funded by financial assistance provided
  by the Coastal Zone Management Act of 1972, as amended, administered by the
  National Oceanic and Atmospheric Administration (NOAA), Office for Coastal
  Management, pursuant to NOAA Award No. NA21NOS4190136. The views
  expressed herein are those of the author(s) and do not necessarily 
  reflect the views of NOAA, the U.S. Department of Commerce, or any of their 
  subagencies.
dataavailability: |
  Data and code are openly available in Zenodo at 
  https://doi.org/10.5281/zenodo.7330754.
conflictsofinterest: |
  The author declares no conflict of interest. The founding sponsors had no 
  role in the design of the study; in the collection, analyses, or 
  interpretation of data; in the writing of the manuscript, an in the decision 
  to publish the results.
bibliography: mybibfile.bib
endnotes: false
output: 
  rticles::mdpi_article:
    latex_engine: xelatex #pdflatex or XeLaTeX allowed
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(targets)
#library(unitar)
library(knitr)
library(tidyverse)
library(kableExtra)
library(patchwork)
library(ggrepel)
library(rcartocolor)
library(twriTemplates)
library(ggspatial)
library(ragg)
library(flextable)
library(modelsummary)
library(scico)
library(ggridges)
library(ggtext)
library(lubridate)
library(units)

## this sets our default code chunk options
knitr::opts_chunk$set(dev = "cairo_pdf",
                      echo = FALSE,
                      ## figure chunk options
                      dpi = 300)

store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
```

# Introduction

Like many estuaries globally, estuaries along the Texas Gulf coast are facing 
pressures from increasing population, increases in point source 
and non-point source pollution and alterations to freshwater inflows leading to
increases in the occurrences and risks of algal blooms and eutrophication [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; 
@bugica_water_2020]. Recent studies indicate that estuary water quality
dynamics in both agriculturally dominated and urban watersheds within Texas are
displaying signals of conditions increasingly conducive to eutrophication 
[@wetzWaterQualityDynamics2016; @wetz_exceptionally_2017; 
@bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022].

# Materials and Methods

## Study Area and Data

Lavaca Bay is a secondary bay in the Matagorda Bay system located on the 
Texas Gulf coast, roughly halfway between the cities of Houston and Corpus 
Christi (Figure  \ref{fig:fig1}). Lavaca Bay is 190 km^2^ with the majority
of freshwater inflow provided by the Lavaca and Navidad River systems. The 
Garcitas-Arenosa, Placedo Creek, and Cox Bay watersheds provide additional 
freshwater inflows. The entire watershed land area for Lavaca Bay is 8,149 km^2^. 
The Lavaca and Navidad River watersheds are a combined 5,966 km^2^, or 
approximately 73% of the entire Lavaca Bay watershed area. Discharge from the 
Navidad River is regulated by Lake Texana which has been in operation since 1980.
Lake Texana provides 170,000 acre-feet of water storage and discharges into the 
tidal section of the Navidad River which ultimately joins the tidal section of 
the Lavaca River 15 km upstream of the confluence with the Bay.

\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
\end{adjustwidth}
\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
\end{figure}

Daily discharges for the Lavaca River (USGS-08164000) were obtained from the 
United States Geologic Survey (USGS) National Water Information System using the 
*dataRetrieval* R package [@deciccoDataRetrievalPackagesDiscovering2022]. Gaged
dailly discharges from Lake Texana (USGS-0816425) were provided by the Texas
Water Development Board (TWDB) (April 21, 2022 email from R. Neupane, TWDB).

Water quality sample data for both freshwater and estuary locations were o
btained from the Texas Commission on Environmental Quality (TCEQ) Surface Water 
Quality Monitoring Information System. Data submitted through the system 
are required to be collected under Quality Assurance Project Plans and lab 
method procedures outlined by the TCEQ's procedures manual. 
The QAPP and procedures manuals ensure the consistent collection and laboratory 
methods are applied between samples collected by different entities and under 
different projects. 
For freshwater locations, total phosphorus (TP) and nitrate-nitrogen (NO~3~) 
data were downloaded (Table \ref{tab:fwsummary}). Unfortunately, insufficient data was available for 
assessment of total nitrogen (TN) and total Kjeldahl nitrogen (TKN) loadings, 
so analysis was restricted to TP and NO~3~ loads. For estuary locations, we 
obtained data for TP, Nitrite+Nitrate (NO*~x~*), TKN, chlorophyll-*a*, and
dissolved oxygen (Table \ref{tab:estuarysummary}).

```{r fwsummary}
df <- tar_read(model_data, store = store)

df <- df |>  
  select(Date, site_no, Flow, original_NO3, original_TP) |>
  mutate(site_no = case_when(
    site_no == "lktexana_g" ~ "usgs08164525",
    site_no != "lktexana_g" ~ site_no
  )) |>
  filter(site_no == "usgs08164000" |
           site_no == "usgs08164525") |> 
  rename(`NO\\textsubscript{3} (mg/L)` = original_NO3,
         `TP (mg/L)` = original_TP,
         `Station ID` = site_no,
         `Mean Daily Streamflow (cfs)` = Flow)
df$`Station ID` <- str_replace(df$`Station ID`, "usgs", "USGS-")

datasummary(`Station ID` * (`TP (mg/L)` + `NO\\textsubscript{3} (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
               Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
            data = df, output = "kableExtra",
            booktabs = TRUE, table.envir = "table", 
            position = "H", escape = FALSE,
            caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.")
```

```{r estuarysummary}
df <- tar_read(estuary_model_data, store = store)

df <- df |> 
  filter(station_id == "13384" |
         station_id == "13383" |
         station_id == "13563") |>
  filter(parameter_code == "00630" |
           parameter_code == "00625" |
           parameter_code =="00665" |
           parameter_code == "70953" |
           parameter_code == "00300") |> 
  filter(end_date >= as.Date("2005-01-01")) |> 
  select(end_date, station_id, station_description, parameter_description, parameter_code, value) |> 
  mutate(parameter_description =
           case_when(
             parameter_code == "70953" ~ "Chlorophyll-\\emph{a} ($\\mu$g/L)",
             parameter_code == "00625" ~ "TKN (mg/L)",
             parameter_code == "00630" ~ "NO\\textsubscript{\\emph{x}} (mg/L)",
             parameter_code == "00665" ~ "TP (mg/L)",
             parameter_code == "00300" ~ "DO (mg/L)"
           )) |>
  mutate(station_id = paste0("TCEQ-", station_id)) |> 
  rename(`Station ID` = station_id) |> 
  pivot_wider(names_from = parameter_description,
              values_from = value)

 modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
                            `NO\\textsubscript{\\emph{x}} (mg/L)` +
                            `TKN (mg/L)` +
                            `Chlorophyll-\\emph{a} ($\\mu$g/L)` +
                            `DO (mg/L)`) ~
                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
                          data = df, output = "dataframe") |> 
   mutate(`Station ID` = case_when(
     `Station ID` == "" ~ NA_character_,
     `Station ID` != "" ~ `Station ID`
     )) |>
   tidyr::fill(`Station ID`, .direction = "down") |> 
   kbl(format = "latex", booktabs = TRUE, table.envir = "table", 
       position = "H", escape = FALSE,
       caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.") |> 
   collapse_rows(columns = 1, latex_hline = "major",
                 valign = "middle")

```

## Estimating Watershed Based Nutrient Loads

Estimates of nutrient loads were developed using Generalized Additive Models
(GAMs) relating nutrient concentration to river discharge, season, and time.
Separate models were fit at each station for each parameter and used to predict
nutrient concentrations for each day in the study period. GAMs can be specified
in a functionally similar manner to the commonly used LOADEST 
[@cohn_validity_1992] or WRTDS 
[@hirsch_weighted_2010] regression models and have been shown to
produce reliable estimates of nutrient and sediment loadings 
[@wangLoadEstimationUncertainties2011; @kroonRiverLoadsSuspended2012; 
@kuhnert_quantifying_2012; @robson_prediction_2015-1; 
@hagemannEstimatingNutrientOrganic2016; @mcdowell_implications_2021; 
@biagi_novel_2022]. GAMs are a semiparametric extension of generalized linear 
models where the linear predictor is represented as the sum of multiple unknown 
smooth functions and parametric linear predictors [@wood_fast_2011]. Although 
the underlying parameter estimation procedure of GAMs is substantially different 
than WRTDS, both the functional form and results are demonstrated to be similar [@beckNumericalQualitativeContrasts2017]. The use of GAMs over other 
regression-based approaches was (1) the ability to easily explore and 
incorporate different model terms, (2) the ability to incorporate non-linear
smooth function without explicit apriori knowledge of the expect shape, and (3)
the ability to specify a link function that relates the expected value of the
response to the linear predictors and allows use to avoid data transformations
as much as possible.

GAMs were fit using the *mgcv* package in R which makes available multiple 
types of smooth functions with automatic smoothness selection [@wood_fast_2011]. 
The general form of the model relating NO~3~ and TP concentration to streamflow,
season, and time was:

\begin{equation}\label{eq:1}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  \\
y &\sim \mathcal{N}(\mu,\,\sigma^{2}),
\end{aligned}
\end{equation}

where $\mu$ is the conditional expected NO~3~-N or TP concentration, *g()* is the 
log-link, $\alpha$ is the intercept, *f~n~()* are smoothing functions. *y* is the 
response variable (NO~3~ or TP concentration) modeled as normally distributed 
with mean $\mu$ and standard deviation $\sigma$. *ddate* is the date converted to 
decimal notation, *yday* is numeric day of year (1-366), and *log1p(Q)* is the 
natural log of mean daily streamflow plus 1.

Moving average (*ma*) is an exponentially smoothed moving average that attempts 
to incorporate the influence of prior streamflow events on concentration at the 
current time period. @wangLoadEstimationUncertainties2011, 
@kuhnert_quantifying_2012 and @zhang_improving_2017 refer to this as averaged or 
smoothed discounted flow and demonstrated improvements in nutrient loading 
models by including the term. @kuhnert_quantifying_2012 expresses MA as:

\begin{equation}\label{eq:2}
ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m,
\end{equation}

where $\delta$ is the discount factor (here, set equal to 0.95), $\kappa_i$ 
is the cumulative flow (*Q*) up to the *i*th day.

Flow anomaly (*fa*) is a unitless term that represents how wet or dry the 
current time period is from a previous time period [@vecchia_trends_2009; 
@zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
the previous year relative to the entire period and calculated as described 
by @zhang_improving_2017:

\begin{equation}\label{eq:3}
ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period} 
\end{equation}

and the short-term flow anomaly (*stfa*) calculated as the current day flow 
compared to the preceding 1-month streamflow:

\begin{equation}\label{eq:4}
stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t) 
\end{equation}

where *x* are the averages of log-transformed streamflow over the antecedent 
period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3 models 
and *stfa* in TP models based on results from @zhang_improving_2017 
demonstrating major improvements in NO~x~ regression models that incorporated 
*ltfa* and moderate improvements in TP regression models that incorporated 
*stfa*.

The calculation of model terms for the Lake Texana site were slightly modified
because daily loads are not a function of natural stream flow processes alone,
but of dam releases and nutrient concentrations at the discharge point of the 
lake. *Q*, *ma*, and *fa* terms were calculated based on total gaged inflow from
the 4 major tributaries to the lake. Thin-plate regression splines were used 
for *ddate*, *log1p(Q)*, *fa*, and *ma*. A cyclic cubic regression spline was 
used for *yday* to ensure the ends of the spline match 
(day 1 and day 366 are expected to match). First order penalties were applied 
to the smooths of flow-based variables which penalize departures from a 
flat function to help constrain extrapolations for high flow measurements.

Left-censored data were not uncommon in this dataset. Several methods are
available to account for censored data. We transformed left-censored nutrient 
concentrations to one-half the detection limit. Although this simple approach 
can introduce bias [@hornungEstimationAverageConcentration1990],  we considered 
it acceptable because high concentrations and loadings are associated with 
high-flow events and low-flow/low-concentration events will account for a small 
proportion of total loadings [@mcdowell_implications_2021].

Daily loads were estimated as the predicted concentration multiplied by the 
daily streamflow. For the Lake Texana site, model terms were slightly modified
because daily loads are a function of dam releases and nutrient concentration,
but concentration will be a function of lake inflows and or other lake processes.
 *Q*, *ma*, and *fa* terms were calculated based on total gaged inflow from
the 4 major tributaries to the lake and laily loads at the dam were calculated 
from the discrete daily concentration at the discharge point of the lake and 
corresponding reported daily discharge from the dam.
Flow-normalized loads were estimated similar to WRTDS by setting flow-based 
covariates on each day of the year equal to each of the historical values for
that day of the year over the study period [@hirsch_weighted_2010]. The 
flow-normalized estimate was calculated as the mean of all the predictions for 
each day considering all possible flow values.
Standard deviations and credible intervals were obtained by 
drawing samples from the multivariate normal posterior distribution of the 
fitted GAM [@woodConfidenceIntervalsGeneralized2006; 
@marraCoveragePropertiesConfidence2012; @mcdowell_implications_2021].
Uncertainty in loads were reported as 90% credible intervals developed by 
drawing 1000 realizations of parameter estimates from the multivariate normal 
posterior distribution of the model parameters. GAM performance was evaluated 
using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
and average Nash-Suttclifee Efficiency (NSE), r^2^ and percent bias (PBIAS) 
metrics across folds were calculated for each model.

## Linking Estuary Water Quality to Hydrology and Nutrient Loads

To test if changes in freshwater inflow and nutrient loading had explanatory
effect on changes in estuary water quality a series of GAM models were fit at
each site relating parameter concentration to temporal trends, inflow, and 
nutrient loads [@murphyNutrientImprovementsChesapeake2022]:

\begin{equation}\label{eq:5}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

\begin{equation}\label{eq:6}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

\begin{equation}\label{eq:7}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

where $\mu$ is the conditional expected response (nutrient concentration), *g()* 
is the log link, and response variable was modeled as Gamma distributed with 
mean $\mu$ and scale $\lambda$. *f~1~(ddate)* is decimal date smoothed with a 
thin-plate regression spline, *f~2~(yday)* is the numeric day of year smoothed 
with a cyclic cubic regression spline, *f~3~(Q)* is mean daily inflow (the 
combined measurements from Lavaca River and Lake Texana) and *f~4~(Load)* is 
the total NO~3~ or TP watershed load. The set of models specified for each water
quality response are in Table \ref{rab:estgammodels}.

```{r estgammodels, echo=FALSE}
df <- tibble(
  `Water Quality Response Parameter` = c(rep("TP", 3), rep("NO\\textsubscript{\\emph{x}}", 3),
               rep("Chlorophyll-\\emph{a}", 3), rep("Dissolved Oxygen", 3),
               rep("TKN", 2)),
  Model = c("Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow", "Flow+Load",
            "Temporal", "Flow"),
  `Model Terms` = c("s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(NO\\textsubscript{3} Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3} Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)",
                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3}  Load)",
                    "s(ddate) + s(yday)",
                    "s(ddate) + s(yday) + s(Q)")
)

kbl(df, format = "latex",
    booktabs = TRUE,
    caption = "Set of GAM models specified for each water quality parameter response.",
    col.names = linebreak(c("Water Quality\nResponse Parameter",
                          "Model",
                          "Model Terms")),
    table.envir = "table", 
    position = "H", escape = FALSE) |> 
  collapse_rows(columns = 1, latex_hline = "major",
                valign = "middle")

```

Because streamflow and nutrient loads are tightly correlated, freshwater inflow
masks any potential signals from nutrient loads alone. Following the 
methodology implemented by @murphyNutrientImprovementsChesapeake2022,
both streamflow and nutrient loads were prepossessed to account for season
and flow. Instead of using raw freshwater inflow and nutrient loading values, 
these values were replaced by seasonally adjusted inflow and flow-adjusted 
nutrient loads by fitting a GAM relating season (day of year) to log transformed
daily freshwater inflow values:

\begin{equation}\label{eq:8}
g(\mu) = \alpha + f_1(yday),
\end{equation}

and a GAM relating log transformed NO~3~ or TP loads to log transformed daily
inflow:

\begin{equation}\label{eq:9}
g(\mu) = \alpha + f_1(log(Q)),
\end{equation}

where the response variables were modeled as normally distributed with an 
identity link function. Response residuals from the respective GAM models were 
used as *Q* and *Load* in Equation \ref{eq:6} and Equation \ref{eq:7}.

# Results

## Watershed Nutrient Loads

Lavaca NO~3~ r2 = 0.85 and 0.90 deviance explained
TP r2 = 0.27 and 0.33
Texana NO~3~ r2 = 0.75 and 0.81
TP = 0.32 and 0.38

```{r fig2, message=FALSE, warning=FALSE, fig.height=5, fig.width=5, fig.cap="Density plots of goodness-of-fit metrics from repeated 5-fold cross validation. Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics."}

## some functions to overide scales in facets via dewey dunnington
#https://dewey.dunnington.ca/post/2018/modifying-facet-scales-in-ggplot2/
scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
    stop("which must be an integer of length 1")
  }

  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
    stop("scale must be an x or y position scale")
  }

  structure(list(which = which, scale = scale), class = "scale_override")
}

CustomFacetWrap <- ggproto(
  "CustomFacetWrap", FacetWrap,
  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
    # make the initial x, y scales list
    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)

    if(is.null(params$scale_overrides)) return(scales)

    max_scale_x <- length(scales$x)
    max_scale_y <- length(scales$y)

    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
    for(scale_override in params$scale_overrides) {
      which <- scale_override$which
      scale <- scale_override$scale

      if("x" %in% scale$aesthetics) {
        if(!is.null(scales$x)) {
          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
          scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) {
          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
          scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }

    # return scales
    scales
  }
)

facet_wrap_custom <- function(..., scale_overrides = NULL) {
  # take advantage of the sanitizing that happens in facet_wrap
  facet_super <- facet_wrap(...)

  # sanitize scale overrides
  if(inherits(scale_overrides, "scale_override")) {
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) || 
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
  }

  facet_super$params$scale_overrides <- scale_overrides

  ggproto(NULL, CustomFacetWrap,
          shrink = facet_super$shrink,
          params = facet_super$params
  )
}

df <- tar_read(cv_no3_08164000, store = store) |> 
  mutate(site = "USGS-08164000",
         parameter = "NO<sub>3</sub>") |> 
  bind_rows(
    tar_read(cv_tp_08164000, store = store) |> 
      mutate(site = "USGS-08164000",
             parameter = "TP")
  ) |> 
  bind_rows(
    tar_read(cv_no3_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "NO<sub>3</sub>")
  ) |> 
  bind_rows(
    tar_read(cv_tp_texana, store = store) |> 
      mutate(site = "USGS-08164525",
             parameter = "TP")
  )

df_long <- df |> 
  ungroup() |> 
  select(NSE, r2, pbias, site, parameter) |> 
  pivot_longer(cols = c(NSE, r2, pbias),
               names_to = "metric")

## plots the density estimates of the repeated 5-fold cross-validation goodness-of-fit metric results,
## color indicates the tail probability calculated from the empirical cumulitive distribution of the goodness-of-fit metric values 
p1 <- ggplot(df, aes(y = site, x = NSE,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(150L, "pt"))) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-1,1),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "NSE", y = "") +
  theme_TWRI_print() +
  theme(axis.text = element_text(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(20L, "pt"))

p2 <- ggplot(df, aes(y = site, x = r2,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(150L, "pt"))) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(0,1),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "r<sup>2</sup>", y = "") +
  theme_TWRI_print() +
  theme(axis.text = element_text(size = 8),
        axis.title.x = element_markdown(),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(20L, "pt"))

p3 <- ggplot(df, aes(y = site, x = pbias,
               fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient",
                      calc_ecdf = TRUE,
                      n = 10000) +
  scale_fill_scico("Tail Probability", palette = "hawaii", direction = -1,
                   limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
  guides(fill = guide_colorbar(barwidth = unit(150L, "pt"))) +
  facet_wrap_custom(~parameter, scales = "free_x",
                    ncol = 2,
                    scale_overrides = list(
                      scale_override(1, scale_x_continuous(limits = c(-150,200),
                                                           expand = c(0,0))),
                      scale_override(2, scale_x_continuous(limits = c(-75,75),
                                                           expand = c(0,0)))
                    )) +
  labs(x = "PBIAS", y = "") +
  theme_TWRI_print() +
  theme(axis.text = element_text(size = 8),
        strip.text.x = element_markdown(size = 8),
        strip.background = element_rect(fill = "white"),
        panel.spacing.x = unit(20L, "pt"))

layout <- "
AAAA
AAAA
AAAA
BBBB
BBBB
BBBB
CCCC
CCCC
CCCC
DDDD
"

p1 / p2 / p3 / guide_area() + plot_layout(guides = 'collect',
                                          design = layout)

```

```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."}

load <- tar_read(daily_no3_08164000, store = store)
fn_load <- tar_read(daily_no3_08164000_fn, store = store)

load_lav_tp <- tar_read(daily_tp_08164000, store = store)
load_nav_tp <- tar_read(daily_tp_texana, store = store)

a <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual NO<sub>3</sub> Load [kg]", subtitle = "USGS-08164000") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load <- tar_read(daily_no3_texana, store = store)
fn_load <- tar_read(daily_no3_texana_fn, store = store)

b <- ggplot() +
  geom_point(data = load$annually, aes(year, NO3_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = NO3_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, NO3_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = NO3_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = NO3_Lower, ymax = NO3_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual NO<sub>3</sub> Load [kg]", subtitle = "USGS-08164525") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E7)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load <- tar_read(daily_tp_08164000, store = store)
fn_load <- tar_read(daily_tp_08164000_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

c <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "USGS-08164000") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load <- tar_read(daily_tp_texana, store = store) 
fn_load <- tar_read(daily_tp_texana_fn, store = store)

load$annually <- load$annually |> 
  filter(year >= 2005)

fn_load$annually <- fn_load$annually |> 
  filter(year >= 2005)

d <- ggplot() +
  geom_point(data = load$annually, aes(year, TP_Estimate,
                                   color = "Total Annual Load + 90% CI",
                                   shape = "Total Annual Load + 90% CI")) +
  geom_line(data = load$annually, aes(x = year, y = TP_Estimate,
                                  color = "Total Annual Load + 90% CI",
                                  linetype = "Total Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = load$annually, aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                                       color = "Total Annual Load + 90% CI")) +
  geom_point(data = fn_load$annually, aes(year, TP_Estimate,
                                 color = "Flow-Normalized Annual Load + 90% CI",
                                 shape = "Flow-Normalized Annual Load + 90% CI")) +
  geom_line(data = fn_load$annually,
            aes(x = year, y = TP_Estimate,
                color = "Flow-Normalized Annual Load + 90% CI",
                linetype = "Flow-Normalized Annual Load + 90% CI"),
            alpha = 0.5) +
  geom_linerange(data = fn_load$annually,
                 aes(x = year, ymin = TP_Lower, ymax = TP_Upper,
                     color = "Flow-Normalized Annual Load + 90% CI")) +
  labs(x = "", y = "Annual TP Load [kg]", subtitle = "USGS-08164525") +
  scale_shape_manual(name = "values",
                     values = c(21, 19)) +
  scale_color_manual(name = "values",
                     values = c("#7E1900", "#1A3399")) +
  scale_linetype_manual(name = "values",
                        values = c(1, 2)) +
  scale_y_continuous(trans = "pseudo_log",
                     breaks = c(0,1E1, 1E2, 1E3, 1E4, 1E5, 1E6, 1E7),
                     labels = scales::label_log()) +
  coord_cartesian(ylim = c(1000, 1E6)) +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        plot.subtitle = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

layout <- "
11112222
11112222
11112222
11112222
33334444
33334444
33334444
33334444
55555555
"

a + b + c + d + guide_area() + plot_layout(guides = 'collect', design = layout)
```

```{r fig4, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Comparison of delivered annual loads at USGS-08164000 and USGS-08164525."}
load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
load_nav_no3 <- tar_read(daily_no3_texana, store = store)

load_lav_no3$annually |> 
  mutate(site = "Lavaca River") -> lavaca_no3_annually

load_nav_no3$annually |> 
  mutate(site = "Navidad River") -> navidad_no3_annually

no3_annual <- bind_rows(lavaca_no3_annually, navidad_no3_annually) |> 
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))

b <- ggplot(no3_annual) +
  geom_col(aes(year, NO3_Estimate, fill = site), width = 0.8) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
                     breaks = c(2005, 2010, 2015, 2020)) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  labs(x = "", y = "Predicted Annual<br>NO<sub>3</sub>-N Load [kg]") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

prop <- no3_annual |> 
  mutate(proportion = NO3_Estimate/sum(NO3_Estimate))

c <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "Proportion of Annual<br>NO<sub>3</sub> Load") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

d <- tar_read(qdata, store = store) |> 
  filter(site_no %in% c("8164000", "lktexana_g")) |>
  filter(Date >= as.Date("2005-01-01")) |> 
  mutate(site = case_when(
    site_no == "8164000" ~ "Lavaca River",
    site_no == "lktexana_g" ~ "Navidad River"
  )) |> 
  mutate(year = year(Date)) |>

  mutate(Flow = as_units(Flow, "ft^3/s")) |> 
  mutate(Flow = set_units(Flow, "ft^3/day")) |> 
  mutate(Flow = set_units(Flow, "1E6gallons/day")) |> 
  group_by(year, site) |> 
  summarise(Flow = sum(Flow, na.rm = TRUE)) |> 
  mutate(Flow = drop_units(Flow)) |> 
  ggplot() +
  geom_col(aes(year, Flow, fill = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "Annual Discharge<br>[Millon Gallons]") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

load_lav_tp$annually |> 
  mutate(site = "Lavaca River") -> lavaca_tp_annually

load_nav_tp$annually |> 
  mutate(site = "Navidad River") -> navidad_tp_annually

tp_annual <- bind_rows(lavaca_tp_annually, navidad_tp_annually) |>
  filter(year >= 2005) |> 
  mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))

e <- ggplot(tp_annual) +
  geom_col(aes(year, TP_Estimate, fill = site), width = 0.8) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
                     breaks = c(2005, 2010, 2015, 2020)) +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  labs(x = "", y = "Predicted Annual<br>TP Load [kg]") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

prop <- tp_annual |> 
  mutate(proportion = TP_Estimate/sum(TP_Estimate))

f <- ggplot() +
  geom_col(data = prop, aes(year, proportion, fill  = site)) +
  scale_fill_carto_d(palette = "Vivid", direction = -1) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                     expand = expansion(mult = c(0.05, 0.05))) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "", y = "Proportion of Annual<br>TP Load") +
  theme_TWRI_print() +
  theme(axis.title.y = element_markdown(size = 8),
        axis.text = element_text(size = 8),
        panel.grid.major.x = element_line(color = "#d9d9d9",
                                          linetype = "dotted"),
        legend.title = element_blank(),
        legend.text = element_text(size = 8))

layout <- "
BE
BE
BE
BE
CF
CF
CF
CF
DD
DD
DD
DD
GG
"

b + c + d + e + f + guide_area() + 
  plot_layout(guides = 'collect', design  = layout)
```

## Linkages Between Water Quality and Watershed Flows and Loads

```{r}

dplyr::bind_rows(
    tar_read(tp_13563, store = store), 
    tar_read(tp_13383, store = store),
    tar_read(tp_13384, store = store),
    tar_read(no3_13563, store = store), 
    tar_read(no3_13383, store = store),
    tar_read(no3_13384, store = store),
    tar_read(chla_13563, store = store), 
    tar_read(chla_13383, store = store),
    tar_read(chla_13384, store = store),
    tar_read(do_13563, store = store), 
    tar_read(do_13383, store = store),
    tar_read(do_13384, store = store),
    tar_read(tkn_13563, store = store),
    tar_read(tkn_13383, store = store),
    tar_read(tkn_13384, store = store)
    ) |> 
  mutate(Parameter = case_when(
    Parameter == "Nitrite+Nitrate" ~ "NO\\textsubscript{\\emph{x}}",
    Parameter == "Chlorophyll-a" ~ "Chlorophyll-\\emph{a}",
    TRUE ~ as.character(Parameter)
  )) |> 
  mutate(Parameter = fct_relevel(Parameter, "TP", "NO\\textsubscript{\\emph{x}}", "Chlorophyll-\\emph{a}","TKN", "DO"),
         Model = fct_relevel(Model, "Temporal", "Flow", "Flow + Load")) |>
  tidyr::complete(Site, Model, nesting(Parameter)) |>
  mutate(model_prob = paste0(round(AICc,1), " (", round(weight,2), ")")) |> 
  mutate(model_prob = case_when(
    model_prob == "NA (NA)" ~ "-",
    model_prob != "NA (NA)" ~ model_prob
  )) |> 
  select(-c(AICc, weight)) |> 
  pivot_wider(names_from = Model, values_from = model_prob) |> 
  select(Parameter, Site, Temporal, Flow, `Flow + Load`) |> 
  arrange(Parameter, Site) |> 
  kbl(format = "latex",
    booktabs = TRUE,
    table.envir = "table", 
    position = "H", escape = FALSE,
    caption = "Model AIC\\textsubscript{c} values and associated model probabilities (in parenthesis). Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.") |> 
    collapse_rows(columns = 1,
                valign = "middle", latex_hline = "major") |> 
  column_spec(3, bold = c(rep(F, 7), T, F, T, T, F, T, T, F),
              italic = c(rep(F, 7), T, F, T, T, F, T, T, F)) |> 
  column_spec(4, bold = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T),
              italic = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T)) |> 
  column_spec(5, bold = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F),
              italic =c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F))

```

## Figures, Tables and Schemes

## Formatting of Mathematical Components

This is an example of an equation:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please punctuate 
equations as regular text.

<!-- If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph. -->

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

# Discussion

```{r tpcomp}
df <- tibble(
  Parameter = rep("TP", 5),
  `Reported Yield (kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})` = c("42.9 (34.4, 54.0)", "45.2", "42", "20.81-91.58", "28.9"),
  Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
  `Time Period` = c("2000-2020", "2012", "1977-2005", "2002", "1972-1993"),
  Reference = c("-", "@wiseSpatiallyReferencedModels2019", 
                "@omaniEstimationSedimentNutrient2014", "@rebichSourcesDeliveryNutrients2011",
                "@dunnTrendsNutrientInflows1996")
)

kbl(df, format = "latex",
    booktabs = TRUE,
    table.envir = "table", 
    position = "H", escape = FALSE,
    col.names = linebreak(c("Parameter", 
                            "Reported Yield\n(kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})",
                          "Approach",
                          "Time Period",
                          "Reference")))

```

Authors should discuss the results and how they can be interpreted in 
perspective of previous studies and of the working hypotheses. The findings and 
their implications should be discussed in the broadest context possible. Future 
research directions may also be highlighted.

# Conclusion

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.

### 2023-02-07 Schramm-Manuscript-2023/Schramm-Manuscript-2023.tex
first commit

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles
\documentclass[water,article,submit,oneauthor]{Definitions/mdpi}

% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'. The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}. This is especially recommended for submission to arXiv, where line numbers should be removed before posting. For preprints.org, the editorial staff will make this change immediately prior to posting.

%% Some pieces required from the pandoc template
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing,\gdef\@continuouspages{yes}} nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, %%nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, %oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology,\gdef\@ISSN{2813-0618}\gdef\@continuous pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal. Changing "submit" to "accept" before posting will remove line numbers.

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ }
\daterevised{ } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Assessing linkages between watershed nutrient loading and estuary
water quality in Lavaca Bay, Texas}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Assessing linkages between watershed nutrient loading and
estuary water quality in Lavaca Bay, Texas}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Michael
Schramm$^{1}$\href{https://orcid.org/0000-0003-1876-6592}
{\orcidicon}}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Michael Schramm}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
\AuthorCitation{Schramm, M.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Texas A\&M AgriLife Research - Texas Water Resources
Institute 1001 Holleman Dr.~E. College Station, TX
77840-2118; \href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}\\
}

% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
\simplesumm{A Simple summary goes here.}

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{A single paragraph of about 200 words maximum. For research
articles, abstracts should give a pertinent overview of the work. We
strongly encourage authors to use the following style of structured
abstracts, but without headings: 1) Background: Place the question
addressed in a broad context and highlight the purpose of the study; 2)
Methods: Describe briefly the main methods or treatments applied; 3)
Results: Summarize the article's main findings; and 4) Conclusion:
Indicate the main conclusions or interpretations. The abstract should be
an objective representation of the article, it must not contain results
which are not presented and substantiated in the main text and should
not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent
keywords specific to the article, yet reasonably common within the
subject discipline.).}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}

  \newcolumntype{d}{S[
    input-open-uncertainty=,
    input-close-uncertainty=,
    parse-numbers = false,
    table-align-text-pre=false,
    table-align-text-post=false
  ]}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Like many estuaries globally, estuaries along the Texas Gulf coast are
facing pressures from increasing population, increases in point source
and non-point source pollution and alterations to freshwater inflows
leading to increases in the occurrences and risks of algal blooms and
eutrophication
\citep{bricker_effects_2008, kennicuttWaterQualityGulf2017, bugica_water_2020}.
Recent studies indicate that estuary water quality dynamics in both
agriculturally dominated and urban watersheds within Texas are
displaying signals of conditions increasingly conducive to
eutrophication
\citep{wetzWaterQualityDynamics2016, wetz_exceptionally_2017, bugica_water_2020, chinPhytoplanktonBiomassCommunity2022}.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{study-area-and-data}{%
\subsection{Study Area and Data}\label{study-area-and-data}}

Lavaca Bay is a secondary bay in the Matagorda Bay system located on the
Texas Gulf coast, roughly halfway between the cities of Houston and
Corpus Christi (Figure \ref{fig:fig1}). Lavaca Bay is 190
km\textsuperscript{2} with the majority of freshwater inflow provided by
the Lavaca and Navidad River systems. The Garcitas-Arenosa, Placedo
Creek, and Cox Bay watersheds provide additional freshwater inflows. The
entire watershed land area for Lavaca Bay is 8,149
km\textsuperscript{2}. The Lavaca and Navidad River watersheds are a
combined 5,966 km\textsuperscript{2}, or approximately 73\% of the
entire Lavaca Bay watershed area. Discharge from the Navidad River is
regulated by Lake Texana which has been in operation since 1980. Lake
Texana provides 170,000 acre-feet of water storage and discharges into
the tidal section of the Navidad River which ultimately joins the tidal
section of the Lavaca River 15 km upstream of the confluence with the
Bay.

\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
\end{adjustwidth}
\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
\end{figure}

Daily discharges for the Lavaca River (USGS-08164000) were obtained from
the United States Geologic Survey (USGS) National Water Information
System using the \emph{dataRetrieval} R package
\citep{deciccoDataRetrievalPackagesDiscovering2022}. Gaged dailly
discharges from Lake Texana (USGS-0816425) were provided by the Texas
Water Development Board (TWDB) (April 21, 2022 email from R. Neupane,
TWDB).

Water quality sample data for both freshwater and estuary locations were
o btained from the Texas Commission on Environmental Quality (TCEQ)
Surface Water Quality Monitoring Information System. Data submitted
through the system are required to be collected under Quality Assurance
Project Plans and lab method procedures outlined by the TCEQ's
procedures manual. The QAPP and procedures manuals ensure the consistent
collection and laboratory methods are applied between samples collected
by different entities and under different projects. For freshwater
locations, total phosphorus (TP) and nitrate-nitrogen
(NO\textsubscript{3}) data were downloaded (Table \ref{tab:fwsummary}).
Unfortunately, insufficient data was available for assessment of total
nitrogen (TN) and total Kjeldahl nitrogen (TKN) loadings, so analysis
was restricted to TP and NO\textsubscript{3} loads. For estuary
locations, we obtained data for TP, Nitrite+Nitrate
(NO\emph{\textsubscript{x}}), TKN, chlorophyll-\emph{a}, and dissolved
oxygen (Table \ref{tab:estuarysummary}).

\begin{table}[H]

\caption{\label{tab:fwsummary}Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.}
\centering
\begin{tabular}[t]{llrrr}
\toprule
Station ID &   & Mean & SD & N\\
\midrule
USGS-08164000 & TP (mg/L) & \num{0.21} & \num{0.09} & 80\\
 & NO\textsubscript{3} (mg/L) & \num{0.18} & \num{0.24} & 74\\
 & Mean Daily Streamflow (cfs) & \num{332.78} & \num{1667.47} & 7671\\
USGS-08164525 & TP (mg/L) & \num{0.20} & \num{0.08} & 81\\
 & NO\textsubscript{3} (mg/L) & \num{0.29} & \num{0.26} & 62\\
 & Mean Daily Streamflow (cfs) & \num{666.14} & \num{2957.79} & 7671\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]

\caption{\label{tab:estuarysummary}Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.}
\centering
\begin{tabular}[t]{lllll}
\toprule
Station ID &   & Mean & SD & N\\
\midrule
 & TP (mg/L) & 0.11 & 0.05 & 47\\

 & NO\textsubscript{\emph{x}} (mg/L) & 0.07 & 0.15 & 51\\

 & TKN (mg/L) & 0.94 & 0.49 & 45\\

 & Chlorophyll-\emph{a} ($\mu$g/L) & 9.43 & 5.31 & 47\\

\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13383} & DO (mg/L) & 7.22 & 1.35 & 55\\
\cmidrule{1-5}
 & TP (mg/L) & 0.08 & 0.03 & 51\\

 & NO\textsubscript{\emph{x}} (mg/L) & 0.06 & 0.08 & 52\\

 & TKN (mg/L) & 0.76 & 0.40 & 48\\

 & Chlorophyll-\emph{a} ($\mu$g/L) & 8.22 & 6.44 & 46\\

\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13384} & DO (mg/L) & 7.51 & 1.32 & 54\\
\cmidrule{1-5}
 & TP (mg/L) & 0.13 & 0.06 & 50\\

 & NO\textsubscript{\emph{x}} (mg/L) & 0.09 & 0.13 & 53\\

 & TKN (mg/L) & 0.94 & 0.37 & 49\\

 & Chlorophyll-\emph{a} ($\mu$g/L) & 9.67 & 5.33 & 49\\

\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13563} & DO (mg/L) & 7.91 & 1.34 & 56\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{estimating-watershed-based-nutrient-loads}{%
\subsection{Estimating Watershed Based Nutrient
Loads}\label{estimating-watershed-based-nutrient-loads}}

Estimates of nutrient loads were developed using Generalized Additive
Models (GAMs) relating nutrient concentration to river discharge,
season, and time. Separate models were fit at each station for each
parameter and used to predict nutrient concentrations for each day in
the study period. GAMs can be specified in a functionally similar manner
to the commonly used LOADEST \citep{cohn_validity_1992} or WRTDS
\citep{hirsch_weighted_2010} regression models and have been shown to
produce reliable estimates of nutrient and sediment loadings
\citep{wangLoadEstimationUncertainties2011, kroonRiverLoadsSuspended2012, kuhnert_quantifying_2012, robson_prediction_2015-1, hagemannEstimatingNutrientOrganic2016, mcdowell_implications_2021, biagi_novel_2022}.
GAMs are a semiparametric extension of generalized linear models where
the linear predictor is represented as the sum of multiple unknown
smooth functions and parametric linear predictors
\citep{wood_fast_2011}. Although the underlying parameter estimation
procedure of GAMs is substantially different than WRTDS, both the
functional form and results are demonstrated to be similar
\citep{beckNumericalQualitativeContrasts2017}. The use of GAMs over
other regression-based approaches was (1) the ability to easily explore
and incorporate different model terms, (2) the ability to incorporate
non-linear smooth function without explicit apriori knowledge of the
expect shape, and (3) the ability to specify a link function that
relates the expected value of the response to the linear predictors and
allows use to avoid data transformations as much as possible.

GAMs were fit using the \emph{mgcv} package in R which makes available
multiple types of smooth functions with automatic smoothness selection
\citep{wood_fast_2011}. The general form of the model relating
NO\textsubscript{3} and TP concentration to streamflow, season, and time
was:

\begin{equation}\label{eq:1}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  \\
y &\sim \mathcal{N}(\mu,\,\sigma^{2}),
\end{aligned}
\end{equation}

where \(\mu\) is the conditional expected NO\textsubscript{3}-N or TP
concentration, \emph{g()} is the log-link, \(\alpha\) is the intercept,
\emph{f\textsubscript{n}()} are smoothing functions. \emph{y} is the
response variable (NO\textsubscript{3} or TP concentration) modeled as
normally distributed with mean \(\mu\) and standard deviation
\(\sigma\). \emph{ddate} is the date converted to decimal notation,
\emph{yday} is numeric day of year (1-366), and \emph{log1p(Q)} is the
natural log of mean daily streamflow plus 1.

Moving average (\emph{ma}) is an exponentially smoothed moving average
that attempts to incorporate the influence of prior streamflow events on
concentration at the current time period.
\citet{wangLoadEstimationUncertainties2011},
\citet{kuhnert_quantifying_2012} and \citet{zhang_improving_2017} refer
to this as averaged or smoothed discounted flow and demonstrated
improvements in nutrient loading models by including the term.
\citet{kuhnert_quantifying_2012} expresses MA as:

\begin{equation}\label{eq:2}
ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m,
\end{equation}

where \(\delta\) is the discount factor (here, set equal to 0.95),
\(\kappa_i\) is the cumulative flow (\emph{Q}) up to the \emph{i}th day.

Flow anomaly (\emph{fa}) is a unitless term that represents how wet or
dry the current time period is from a previous time period
\citep{vecchia_trends_2009, zhang_improving_2017}. Long-term flow
anomaly (\emph{ltfa}) is the streamflow over the previous year relative
to the entire period and calculated as described by
\citet{zhang_improving_2017}:

\begin{equation}\label{eq:3}
ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period} 
\end{equation}

and the short-term flow anomaly (\emph{stfa}) calculated as the current
day flow compared to the preceding 1-month streamflow:

\begin{equation}\label{eq:4}
stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t) 
\end{equation}

where \emph{x} are the averages of log-transformed streamflow over the
antecedent period (\emph{1-year}, \emph{1-month}, etc.) for time
\emph{t}. We used \emph{ltfa} in NO\textasciitilde3 models and
\emph{stfa} in TP models based on results from
\citet{zhang_improving_2017} demonstrating major improvements in
NO\textsubscript{x} regression models that incorporated \emph{ltfa} and
moderate improvements in TP regression models that incorporated
\emph{stfa}.

The calculation of model terms for the Lake Texana site were slightly
modified because daily loads are not a function of natural stream flow
processes alone, but of dam releases and nutrient concentrations at the
discharge point of the lake. \emph{Q}, \emph{ma}, and \emph{fa} terms
were calculated based on total gaged inflow from the 4 major tributaries
to the lake. Thin-plate regression splines were used for \emph{ddate},
\emph{log1p(Q)}, \emph{fa}, and \emph{ma}. A cyclic cubic regression
spline was used for \emph{yday} to ensure the ends of the spline match
(day 1 and day 366 are expected to match). First order penalties were
applied to the smooths of flow-based variables which penalize departures
from a flat function to help constrain extrapolations for high flow
measurements.

Left-censored data were not uncommon in this dataset. Several methods
are available to account for censored data. We transformed left-censored
nutrient concentrations to one-half the detection limit. Although this
simple approach can introduce bias
\citep{hornungEstimationAverageConcentration1990}, we considered it
acceptable because high concentrations and loadings are associated with
high-flow events and low-flow/low-concentration events will account for
a small proportion of total loadings \citep{mcdowell_implications_2021}.

Daily loads were estimated as the predicted concentration multiplied by
the daily streamflow. For the Lake Texana site, model terms were
slightly modified because daily loads are a function of dam releases and
nutrient concentration, but concentration will be a function of lake
inflows and or other lake processes. \emph{Q}, \emph{ma}, and \emph{fa}
terms were calculated based on total gaged inflow from the 4 major
tributaries to the lake and laily loads at the dam were calculated from
the discrete daily concentration at the discharge point of the lake and
corresponding reported daily discharge from the dam. Flow-normalized
loads were estimated similar to WRTDS by setting flow-based covariates
on each day of the year equal to each of the historical values for that
day of the year over the study period \citep{hirsch_weighted_2010}. The
flow-normalized estimate was calculated as the mean of all the
predictions for each day considering all possible flow values. Standard
deviations and credible intervals were obtained by drawing samples from
the multivariate normal posterior distribution of the fitted GAM
\citep{woodConfidenceIntervalsGeneralized2006, marraCoveragePropertiesConfidence2012, mcdowell_implications_2021}.
Uncertainty in loads were reported as 90\% credible intervals developed
by drawing 1000 realizations of parameter estimates from the
multivariate normal posterior distribution of the model parameters. GAM
performance was evaluated using repeated 5-fold cross validation
\citep{burmanComparativeStudyOrdinary1989} and average Nash-Suttclifee
Efficiency (NSE), r\textsuperscript{2} and percent bias (PBIAS) metrics
across folds were calculated for each model.

\hypertarget{linking-estuary-water-quality-to-hydrology-and-nutrient-loads}{%
\subsection{Linking Estuary Water Quality to Hydrology and Nutrient
Loads}\label{linking-estuary-water-quality-to-hydrology-and-nutrient-loads}}

To test if changes in freshwater inflow and nutrient loading had
explanatory effect on changes in estuary water quality a series of GAM
models were fit at each site relating parameter concentration to
temporal trends, inflow, and nutrient loads
\citep{murphyNutrientImprovementsChesapeake2022}:

\begin{equation}\label{eq:5}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

\begin{equation}\label{eq:6}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

\begin{equation}\label{eq:7}
\begin{aligned}
g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \\
y &\sim \Gamma(\mu,\lambda),
\end{aligned}
\end{equation}

where \(\mu\) is the conditional expected response (nutrient
concentration), \emph{g()} is the log link, and response variable was
modeled as Gamma distributed with mean \(\mu\) and scale \(\lambda\).
\emph{f\textsubscript{1}(ddate)} is decimal date smoothed with a
thin-plate regression spline, \emph{f\textsubscript{2}(yday)} is the
numeric day of year smoothed with a cyclic cubic regression spline,
\emph{f\textsubscript{3}(Q)} is mean daily inflow (the combined
measurements from Lavaca River and Lake Texana) and
\emph{f\textsubscript{4}(Load)} is the total NO\textsubscript{3} or TP
watershed load. The set of models specified for each water quality
response are in Table \ref{rab:estgammodels}.

\begin{table}[H]

\caption{\label{tab:estgammodels}Set of GAM models specified for each water quality parameter response.}
\centering
\begin{tabular}[t]{lll}
\toprule
\makecell[l]{Water Quality\\Response Parameter} & Model & Model Terms\\
\midrule
 & Temporal & s(ddate) + s(yday)\\

 & Flow & s(ddate) + s(yday) + s(Q)\\

\multirow{-3}{*}{\raggedright\arraybackslash TP} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(TP Load)\\
\cmidrule{1-3}
 & Temporal & s(ddate) + s(yday)\\

 & Flow & s(ddate) + s(yday) + s(Q)\\

\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{\emph{x}}} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(NO\textsubscript{3} Load)\\
\cmidrule{1-3}
 & Temporal & s(ddate) + s(yday)\\

 & Flow & s(ddate) + s(yday) + s(Q)\\

\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\emph{a}} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\textsubscript{3} Load)\\
\cmidrule{1-3}
 & Temporal & s(ddate) + s(yday)\\

 & Flow & s(ddate) + s(yday) + s(Q)\\

\multirow{-3}{*}{\raggedright\arraybackslash Dissolved Oxygen} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\textsubscript{3}  Load)\\
\cmidrule{1-3}
 & Temporal & s(ddate) + s(yday)\\

\multirow{-2}{*}{\raggedright\arraybackslash TKN} & Flow & s(ddate) + s(yday) + s(Q)\\
\bottomrule
\end{tabular}
\end{table}

Because streamflow and nutrient loads are tightly correlated, freshwater
inflow masks any potential signals from nutrient loads alone. Following
the methodology implemented by
\citet{murphyNutrientImprovementsChesapeake2022}, both streamflow and
nutrient loads were prepossessed to account for season and flow. Instead
of using raw freshwater inflow and nutrient loading values, these values
were replaced by seasonally adjusted inflow and flow-adjusted nutrient
loads by fitting a GAM relating season (day of year) to log transformed
daily freshwater inflow values:

\begin{equation}\label{eq:8}
g(\mu) = \alpha + f_1(yday),
\end{equation}

and a GAM relating log transformed NO\textsubscript{3} or TP loads to
log transformed daily inflow:

\begin{equation}\label{eq:9}
g(\mu) = \alpha + f_1(log(Q)),
\end{equation}

where the response variables were modeled as normally distributed with
an identity link function. Response residuals from the respective GAM
models were used as \emph{Q} and \emph{Load} in Equation \ref{eq:6} and
Equation \ref{eq:7}.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{watershed-nutrient-loads}{%
\subsection{Watershed Nutrient Loads}\label{watershed-nutrient-loads}}

Lavaca NO\textsubscript{3} r2 = 0.85 and 0.90 deviance explained TP r2 =
0.27 and 0.33 Texana NO\textsubscript{3} r2 = 0.75 and 0.81 TP = 0.32
and 0.38

\begin{figure}
\centering
\includegraphics{Schramm-Manuscript-2023_files/figure-latex/fig2-1.pdf}
\caption{Density plots of goodness-of-fit metrics from repeated 5-fold
cross validation. Color indicates the tail probability calcualted from
the empirical cumulative distribution of the goodness-of-fit metrics.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Schramm-Manuscript-2023_files/figure-latex/fig3-1.pdf}
\caption{Aggregated estimated annual and flow-normalized annual
NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Schramm-Manuscript-2023_files/figure-latex/fig4-1.pdf}
\caption{Comparison of delivered annual loads at USGS-08164000 and
USGS-08164525.}
\end{figure}

\hypertarget{linkages-between-water-quality-and-watershed-flows-and-loads}{%
\subsection{Linkages Between Water Quality and Watershed Flows and
Loads}\label{linkages-between-water-quality-and-watershed-flows-and-loads}}

\begin{verbatim}
## Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result. 
## Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result. 
## Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result.
\end{verbatim}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-1}Model AIC\textsubscript{c} values and associated model probabilities (in parenthesis). Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.}
\centering
\begin{tabular}[t]{ll>{}l>{}l>{}l}
\toprule
Parameter & Site & Temporal & Flow & Flow + Load\\
\midrule
 & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & \em{\textbf{-158.2 (0.72)}}\\

 & TCEQ-13384 & -194.4 (0.03) & \em{\textbf{-200.2 (0.49)}} & -200.2 (0.49)\\

\multirow{-3}{*}{\raggedright\arraybackslash TP} & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & \em{\textbf{-157.3 (0.59)}}\\
\cmidrule{1-5}
 & TCEQ-13383 & -218.9 (0) & \em{\textbf{-244.8 (0.5)}} & -244.8 (0.5)\\

 & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & \em{\textbf{-311.9 (0.52)}}\\

\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{\emph{x}}} & TCEQ-13563 & -175.1 (0) & \em{\textbf{-190.2 (0.5)}} & -190.2 (0.5)\\
\cmidrule{1-5}
 & TCEQ-13383 & 279.7 (0.18) & \em{\textbf{278.1 (0.41)}} & 278.1 (0.41)\\

 & TCEQ-13384 & \em{\textbf{268.2 (0.33)}} & 268.2 (0.33) & 268.2 (0.33)\\

\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\emph{a}} & TCEQ-13563 & 289.5 (0.08) & \em{\textbf{286.1 (0.46)}} & 286.1 (0.46)\\
\cmidrule{1-5}
 & TCEQ-13383 & \em{\textbf{42.2 (0.66)}} & 43.5 (0.34) & -\\

 & TCEQ-13384 & \em{\textbf{34.3 (0.57)}} & 34.8 (0.43) & -\\

\multirow{-3}{*}{\raggedright\arraybackslash TKN} & TCEQ-13563 & 31.1 (0.22) & \em{\textbf{28.7 (0.78)}} & -\\
\cmidrule{1-5}
 & TCEQ-13383 & \em{\textbf{146.4 (0.34)}} & 146.4 (0.34) & 146.5 (0.32)\\

 & TCEQ-13384 & \em{\textbf{135.9 (0.47)}} & 137 (0.27) & 137 (0.27)\\

\multirow{-3}{*}{\raggedright\arraybackslash DO} & TCEQ-13563 & 138.3 (0.25) & \em{\textbf{137.2 (0.43)}} & 137.8 (0.32)\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{figures-tables-and-schemes}{%
\subsection{Figures, Tables and
Schemes}\label{figures-tables-and-schemes}}

\hypertarget{formatting-of-mathematical-components}{%
\subsection{Formatting of Mathematical
Components}\label{formatting-of-mathematical-components}}

This is an example of an equation:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please
punctuate equations as regular text.

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\begin{table}[H]

\begin{tabular}[t]{lllll}
\toprule
Parameter & \makecell[c]{Reported Yield\\(kg$\cdot$km\textsuperscript{-2}$\cdot$year\textsuperscript{-1})} & Approach & Time Period & Reference\\
\midrule
TP & 42.9 (34.4, 54.0) & GAM & 2000-2020 & -\\
TP & 45.2 & SPARROW & 2012 & @wiseSpatiallyReferencedModels2019\\
TP & 42 & SWAT & 1977-2005 & @omaniEstimationSedimentNutrient2014\\
TP & 20.81-91.58 & SPARROW & 2002 & @rebichSourcesDeliveryNutrients2011\\
TP & 28.9 & LOADEST & 1972-1993 & @dunnTrendsNutrientInflows1996\\
\bottomrule
\end{tabular}
\end{table}

Authors should discuss the results and how they can be interpreted in
perspective of previous studies and of the working hypotheses. The
findings and their implications should be discussed in the broadest
context possible. Future research directions may also be highlighted.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\funding{This paper was funded by financial assistance provided by the
Coastal Zone Management Act of 1972, as amended, administered by the
National Oceanic and Atmospheric Administration (NOAA), Office for
Coastal Management, pursuant to NOAA Award No.~NA21NOS4190136. The views
expressed herein are those of the author(s) and do not necessarily
reflect the views of NOAA, the U.S. Department of Commerce, or any of
their subagencies.}

\dataavailability{Data and code are openly available in Zenodo at
\url{https://doi.org/10.5281/zenodo.7330754}.}

\acknowledgments{All sources of funding of the study should be
disclosed. Please clearly indicate grants that you have received in
support of your research work. Clearly state if you received funds for
covering the costs to publish in open access.}

\conflictsofinterest{The author declares no conflict of interest. The
founding sponsors had no role in the design of the study; in the
collection, analyses, or interpretation of data; in the writing of the
manuscript, an in the decision to publish the results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional

%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}

%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
\externalbibliography{yes}
\bibliography{mybibfile.bib}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\PublishersNote{}
\end{adjustwidth}

\end{document}

### 2023-02-07 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
add wide figures

@@
 library(units)
+library(gratia)
 
@@
                       ## figure chunk options
-                      dpi = 300)
+                      dpi = 300,
+                      fig.pos = "H",
+                      fig.align = "center",
+                      out.width = "100%")
+
+## custom hook for wide figures
+
+hook_plot_wide <- function(x, options) {
+  x_out <- knitr:::hook_plot_tex(x, options)
+  if(isTRUE(options$wideplot)) {
+    x_out <- gsub(x = x_out,
+                  pattern='\\begin{figure}[H]', 
+                  replacement='\\begin{figure}[H]\\begin{adjustwidth}{-\\extralength}{0cm}',
+                  fixed=TRUE)
+
+    x_out <- gsub(x = x_out,
+                  pattern='\\caption',
+                  replacement='\\end{adjustwidth}\\caption',
+                  fixed = TRUE)
+  }
+  x_out
+}
+knitr::knit_hooks$set(plot=hook_plot_wide)
 
@@
 
-```{r fig2, message=FALSE, warning=FALSE, fig.height=5, fig.width=5, fig.cap="Density plots of goodness-of-fit metrics from repeated 5-fold cross validation.
-Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics."}
+```{r fig2, message=FALSE, warning=FALSE, fig.height=3.5, fig.width=6.5, fig.pos="H", fig.cap="Density plots of goodness-of-fit metrics (NSE, r\\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
+Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics.", out.width="100%", wideplot=TRUE}
 
@@
                    limit = c(0, 0.5), breaks = c(0, 0.1, 0.2, 0.3, 0.4)) +
-  guides(fill = guide_colorbar(barwidth = unit(150L, "pt"))) +
+  guides(fill = guide_colorbar(barwidth = unit(100L, "pt"),
+                               title.position = "top")) +
   facet_wrap_custom(~parameter, scales = "free_x",
@@
   theme_TWRI_print() +
-  theme(axis.text = element_text(size = 8),
+  theme(axis.text = element_text(size = 6),
+        axis.title.x = element_markdown(size = 6),
         strip.text.x = element_markdown(size = 8),
         strip.background = element_rect(fill = "white"),
-        panel.spacing.x = unit(20L, "pt"))
+        panel.spacing.x = unit(20L, "pt"),
+        legend.direction = "horizontal",
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8))
 
@@
   theme_TWRI_print() +
-  theme(axis.text = element_text(size = 8),
-        axis.title.x = element_markdown(),
+  theme(axis.text = element_text(size = 6),
+        axis.title.x = element_markdown(size = 6),
         strip.text.x = element_markdown(size = 8),
         strip.background = element_rect(fill = "white"),
-        panel.spacing.x = unit(20L, "pt"))
+        panel.spacing.x = unit(20L, "pt"),
+        legend.direction = "horizontal",
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8))
 
@@
   theme_TWRI_print() +
-  theme(axis.text = element_text(size = 8),
+  theme(axis.text = element_text(size = 6),
+        axis.title.x = element_markdown(size = 6),
         strip.text.x = element_markdown(size = 8),
         strip.background = element_rect(fill = "white"),
-        panel.spacing.x = unit(20L, "pt"))
+        panel.spacing.x = unit(20L, "pt"),
+        legend.direction = "horizontal",
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        axis.text.y = element_blank())
 
 layout <- "
-AAAA
-AAAA
-AAAA
-BBBB
-BBBB
-BBBB
-CCCC
-CCCC
-CCCC
-DDDD
+AAAA####
+AAAA#DD#
+AAAA####
+BBBBCCCC
+BBBBCCCC
+BBBBCCCC
 "
@@
 
-```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."}
+```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525.", out.width="100%"}
 
@@
 
-```{r fig4, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Comparison of delivered annual loads at USGS-08164000 and USGS-08164525."}
+```{r fig4, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Comparison of delivered annual loads at USGS-08164000 and USGS-08164525.", out.width="100%"}
 load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
@@
     position = "H", escape = FALSE,
-    caption = "Model AIC\\textsubscript{c} values and associated model probabilities (in parenthesis).
+    caption = "Estuary GAM AIC\\textsubscript{c} values and associated model probabilities (in parenthesis).
 Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.") |>
@@
 
-## Figures, Tables and Schemes
-
-## Formatting of Mathematical Components
-
-This is an example of an equation:
-
-\begin{equation}
-a = 1,
-\end{equation}
-
-the text following an equation need not be a new paragraph. Please punctuate 
-equations as regular text.
-
-<!-- If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
--->
-
-This is the example 2 of equation:
-
-\begin{adjustwidth}{-\extralength}{0cm}
-\begin{equation}
-a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
-u + v + w + x + y + z
-\end{equation}
-\end{adjustwidth}
+```{r long-smooths, message=FALSE, warning=FALSE, fig.pos="H", fig.height=3.5, fig.pos="H", fig.width=6, fig.cap="Smoothed trend component for water quality paramaters obtained from Temporal estaury GAMs.", wideplot=list("0cm"), out.width="100%", wideplot=TRUE}
+draw_smoothed_years <- function(model1, model2, model3,
+                                ylab, subtitle) {
+
+  `TCEQ-13563` <- model1
+  `TCEQ-13383` <- model2
+  `TCEQ-13384` <- model3
+
+  comp <- gratia::compare_smooths(
+    `TCEQ-13563`,
+    `TCEQ-13383`,
+    `TCEQ-13384`,
+    smooths = "s(ddate)")
+
+  crit <- gratia:::coverage_normal(0.90)
+
+  comp |> 
+    unnest(data) |> 
+    mutate(lower_ci = est + (crit * se),
+           upper_ci = est - (crit * se),
+           label = if_else(ddate == max(ddate),
+                           as.character(model),
+                           NA_character_)) |> 
+    ggplot(aes(x = ddate, y = est, group = model)) +
+    geom_ribbon(aes(ymin = lower_ci,
+                    ymax = upper_ci,
+                    fill = model),
+                alpha = 0.15) +
+    geom_line(aes(color = model)) +
+    geom_hline(yintercept = 0, linetype = 2, alpha = 0.25) +
+    # geom_text_repel(aes(label = label,
+    #                     color = model), 
+    #                 nudge_x = 5,
+    #                 hjust = "left",
+    #                 size = 2) +
+    scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
+                       expand = expansion(mult = c(0.05,0.05))) +
+    colorspace::scale_fill_discrete_qualitative(name = "") +
+    colorspace::scale_color_discrete_qualitative(name = "") +
+    labs(x = "Year", y = ylab,
+         subtitle = subtitle) +
+    theme_TWRI_print() +
+    theme(axis.title.y = element_markdown(size = 8),
+          axis.title.x = element_markdown(size = 8),
+          plot.subtitle = element_markdown(size = 8),
+          legend.direction = "vertical")
+
+}
+
+p_tp <- draw_smoothed_years(tar_read(tp_lavaca_13563_temporal, store = store),
+                            tar_read(tp_lavaca_13383_temporal, store = store),
+                            tar_read(tp_lavaca_13384_temporal, store = store),
+                            ylab = "TP\nSmooth<br>Estimate [mg/L]", subtitle = "TP Trend")
+
+p_no3 <- draw_smoothed_years(tar_read(no3_lavaca_13563_temporal, store = store),
+                             tar_read(no3_lavaca_13383_temporal, store = store),
+                             tar_read(no3_lavaca_13384_temporal, store = store),
+                             ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]", subtitle = "NO*<sub>x</sub>* Trend")
+
+p_tkn <- draw_smoothed_years(tar_read(tkn_lavaca_13563_temporal, store = store),
+                             tar_read(tkn_lavaca_13383_temporal, store = store),
+                             tar_read(tkn_lavaca_13384_temporal, store = store),
+                             ylab = "TKN\nSmooth<br>Estimate [mg/L]", subtitle = "TKN Trend")
+
+p_chla <- draw_smoothed_years(tar_read(chla_lavaca_13563_temporal, store = store),
+                             tar_read(chla_lavaca_13383_temporal, store = store),
+                             tar_read(chla_lavaca_13384_temporal, store = store),
+                             ylab = "Chlorophyll-*a*\nSmooth<br>Estimate [µg/L]", subtitle = "Chlorophyll-*a* Trend")
+
+p_do <- draw_smoothed_years(tar_read(do_lavaca_13563_temporal, store = store),
+                             tar_read(do_lavaca_13383_temporal, store = store),
+                             tar_read(do_lavaca_13384_temporal, store = store),
+                             ylab = "DO\nSmooth<br>Estimate [mg/L]", subtitle = "DO Trend")
+
+design <- "
+112266
+334455
+"
+
+#(p_tp + p_no3) / (p_chla + p_tkn) / (p_do)
+p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
+                                                   design = design)
+```
 
@@
   `Time Period` = c("2000-2020", "2012", "1977-2005", "2002", "1972-1993"),
-  Reference = c("-", "@wiseSpatiallyReferencedModels2019", 
-                "@omaniEstimationSedimentNutrient2014", "@rebichSourcesDeliveryNutrients2011",
-                "@dunnTrendsNutrientInflows1996")
+  Reference = c("This work", "\\citet{wise_spatially_2019}", 
+                "\\citet{omaniEstimationSedimentNutrient2014}", 
+                "\\citet{rebich_sources_2011}",
+                "\\citet{dunnTrendsNutrientInflows1996}")
 )

### 2023-02-07 Schramm-Manuscript-2023/Schramm-Manuscript-2023.tex
add wide figures

@@
 
-\begin{figure}
-\centering
-\includegraphics{Schramm-Manuscript-2023_files/figure-latex/fig2-1.pdf}
-\caption{Density plots of goodness-of-fit metrics from repeated 5-fold
-cross validation. Color indicates the tail probability calcualted from
-the empirical cumulative distribution of the goodness-of-fit metrics.}
+\begin{figure}[H]\begin{adjustwidth}{-\extralength}{0cm}
+
+{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig2-1}
+
+}
+
+\end{adjustwidth}\caption[Density plots of goodness-of-fit metrics (NSE, r\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads]{Density plots of goodness-of-fit metrics (NSE, r\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
+Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics.}\label{fig:fig2}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{Schramm-Manuscript-2023_files/figure-latex/fig3-1.pdf}
-\caption{Aggregated estimated annual and flow-normalized annual
-NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig3-1}
+
+}
+
+\caption[Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525]{Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{Schramm-Manuscript-2023_files/figure-latex/fig4-1.pdf}
-\caption{Comparison of delivered annual loads at USGS-08164000 and
-USGS-08164525.}
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig4-1}
+
+}
+
+\caption[Comparison of delivered annual loads at USGS-08164000 and USGS-08164525]{Comparison of delivered annual loads at USGS-08164000 and USGS-08164525.}\label{fig:fig4}
 \end{figure}
@@
 
-\caption{\label{tab:unnamed-chunk-1}Model AIC\textsubscript{c} values and associated model probabilities (in parenthesis).
+\caption{\label{tab:unnamed-chunk-1}Estuary GAM AIC\textsubscript{c} values and associated model probabilities (in parenthesis).
 Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.}
@@
 
-\hypertarget{figures-tables-and-schemes}{%
-\subsection{Figures, Tables and
-Schemes}\label{figures-tables-and-schemes}}
+\begin{figure}[H]\begin{adjustwidth}{-\extralength}{0cm}
 
-\hypertarget{formatting-of-mathematical-components}{%
-\subsection{Formatting of Mathematical
-Components}\label{formatting-of-mathematical-components}}
+{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/long-smooths-1}
 
-This is an example of an equation:
+}
 
-\begin{equation}
-a = 1,
-\end{equation}
-
-the text following an equation need not be a new paragraph. Please
-punctuate equations as regular text.
-
-This is the example 2 of equation:
-
-\begin{adjustwidth}{-\extralength}{0cm}
-\begin{equation}
-a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
-u + v + w + x + y + z
-\end{equation}
-\end{adjustwidth}
+\end{adjustwidth}\caption[Smoothed trend component for water quality paramaters obtained from Temporal estaury GAMs]{Smoothed trend component for water quality paramaters obtained from Temporal estaury GAMs.}\label{fig:long-smooths}
+\end{figure}
 
@@
 \midrule
-TP & 42.9 (34.4, 54.0) & GAM & 2000-2020 & -\\
-TP & 45.2 & SPARROW & 2012 & @wiseSpatiallyReferencedModels2019\\
-TP & 42 & SWAT & 1977-2005 & @omaniEstimationSedimentNutrient2014\\
-TP & 20.81-91.58 & SPARROW & 2002 & @rebichSourcesDeliveryNutrients2011\\
-TP & 28.9 & LOADEST & 1972-1993 & @dunnTrendsNutrientInflows1996\\
+TP & 42.9 (34.4, 54.0) & GAM & 2000-2020 & This work\\
+TP & 45.2 & SPARROW & 2012 & \citet{wise_spatially_2019}\\
+TP & 42 & SWAT & 1977-2005 & \citet{omaniEstimationSedimentNutrient2014}\\
+TP & 20.81-91.58 & SPARROW & 2002 & \citet{rebich_sources_2011}\\
+TP & 28.9 & LOADEST & 1972-1993 & \citet{dunnTrendsNutrientInflows1996}\\
 \bottomrule

### 2023-02-09 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
results

@@
 Because streamflow and nutrient loads are tightly correlated, freshwater inflow
-masks any potential signals from nutrient loads alone. Following the 
+can mask signals from nutrient loads alone. Following the 
 methodology implemented by @murphyNutrientImprovementsChesapeake2022,
@@
 
-Lavaca NO~3~ r2 = 0.85 and 0.90 deviance explained
-TP r2 = 0.27 and 0.33
-Texana NO~3~ r2 = 0.75 and 0.81
-TP = 0.32 and 0.38
+Based on criteria provided by @moriasiHydrologicWaterQuality2015, GAMs ranged
+from "satisfactory" to "very good" based on median NSE, r^2^ and PBIAS 
+calculated from repeated 5-fold cross validation on predicted and measured
+nutrient loads. NO~3~ GAM models had median NSE values of 0.34 and 0.48 at
+USGS-08164000 and USGS-08164390 respectively. Median r^2^ values were 0.70
+(USGS-08164000) and 0.87 (USGS-08164525) and PBIAS values were 2.00
+(USGS-08164000) and 10.90 (USGS-08164525) for NO~3~ loads. Median NSE values were
+0.80 (USGS-08164000) and 0.91 (USGS-08164525) for TP loads. Median r^2^ values 
+for TP loads were 0.93 (USGS-08164000) and 0.99 (USGS-08164525) and median PBIAS 
+values were -7.20 (USGS-08164000) and -3.30 (USGS-08164525). Density plots of
+metrics show similar distribution of values between sites for the 
+same parameter, with the exception r^2^ values for NO~3~ loads where 
+USGS-08164000 showed a much larger variance in values compared to USGS-08164525
+(Figure \ref{fig:fig2}).In addition to higher average NSE and r^2^ values, 
+GAMs had smaller variance in metric values for TP compared to NO~3~.
 
@@
 
-```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525.", out.width="100%"}
-
+```{r echo=FALSE}
+## calculates min and max load values and years
+load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
+load_nav_no3 <- tar_read(daily_no3_texana, store = store)
+
+load_lav_no3$annually |> 
+  mutate(site = "Lavaca River") -> lavaca_no3_annually
+
+load_nav_no3$annually |> 
+  mutate(site = "Navidad River") -> navidad_no3_annually
+
+mean_no3 <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Estimate + navidad_no3_annually$NO3_Estimate ))
+mean_no3_low <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Lower + navidad_no3_annually$NO3_Lower ))
+mean_no3_upper <- scales::label_comma()(mean(lavaca_no3_annually$NO3_Upper + navidad_no3_annually$NO3_Upper ))
+
+no3 <- data.frame(lavaca = lavaca_no3_annually$NO3_Estimate,
+                  navidad = navidad_no3_annually$NO3_Estimate,
+                  year = lavaca_no3_annually$year)
+no3 <- no3 |> 
+  mutate(total = lavaca+navidad)
+
+min_no3 <- min(no3$total)
+min_no3_year <- no3[no3$total == min_no3,]$year
+min_no3 <- scales::label_comma()(min_no3)
+
+max_no3 <- max(no3$total)
+max_no3_year <- no3[no3$total == max_no3,]$year
+max_no3 <- scales::label_comma()(max_no3)
+
+no3_prop <- no3 |> 
+  mutate(prop = round(navidad/total, 2))
+
+no3_mean_prop <- scales::label_percent()(mean(no3_prop$prop))
+
+min_no3_prop <- scales::label_percent()(min(no3_prop$prop))
+
+load_lav_tp <- tar_read(daily_tp_08164000, store = store)
+load_nav_tp <- tar_read(daily_tp_texana, store = store)
+
+load_lav_tp$annually |> 
+  mutate(site = "Lavaca River") |> 
+  filter(year >= 2005)-> lavaca_tp_annually
+
+load_nav_tp$annually |> 
+  mutate(site = "Navidad River") |> 
+  filter(year >= 2005) -> navidad_tp_annually
+
+mean_tp <- scales::label_comma()(mean(lavaca_tp_annually$TP_Estimate + navidad_tp_annually$TP_Estimate ))
+mean_tp_low <- scales::label_comma()(mean(lavaca_tp_annually$TP_Lower + navidad_tp_annually$TP_Lower ))
+mean_tp_upper <- scales::label_comma()(mean(lavaca_tp_annually$TP_Upper + navidad_tp_annually$TP_Upper ))
+
+tp <- data.frame(lavaca = lavaca_tp_annually$TP_Estimate,
+                  navidad = navidad_tp_annually$TP_Estimate,
+                  year = lavaca_tp_annually$year)
+tp <- tp |> 
+  mutate(total = lavaca+navidad)
+
+min_tp <- min(tp$total)
+min_tp_year <- tp[tp$total == min_tp,]$year
+min_tp <- scales::label_comma()(min_tp)
+
+max_tp <- max(tp$total)
+max_tp_year <- tp[tp$total == max_tp,]$year
+max_tp <- scales::label_comma()(max_tp)
+
+tp_prop <- tp |> 
+  mutate(prop = round(navidad/total, 2))
+
+tp_mean_prop <- scales::label_percent()(mean(tp_prop$prop))
+
+min_tp_prop <- scales::label_percent()(min(tp_prop$prop))
+```
+
+Annual NO~3~ and TP loads show considerable variation, generally following
+patterns in discharge (Figure \ref{fig:fig3}, Figure \ref{fig:fig4}). 
+Flow-normalized TP loads at both sites and flow-normalized loads at 
+USGS-08164000 indicate watershed based loads have not changed much over time
+when accounting for changes in streamflow (Figure \ref{fig:fig3}). 
+Flow-normalized loads at USGS-08164000 show small variation over time with 
+some decreases in NO~3~ loads since 2013. Aggregated across both sites, the mean 
+annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
+`r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
+in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`.
+
+Total annual TP loads ranged from `r min_tp` kg in `r min_tp_year` to 
+`r max_tp` kg  in `r max_tp_year`. Mean annual TP loading from 2005 through 2020
+was `r mean_tp` kg (`r mean_tp_low` kg - `r mean_tp_upper` kg, 90% CI).
+On average, USGS-08164525 accounted for `r no3_mean_prop` of NO~3~ loads and 
+`r tp_mean_prop` of TP loads from 2005 through 2020. However, during periods of extreme 
+drought the Lavaca River (USGS-08164000) became the primary source of 
+nutrient loading in the watershed with the Navidad River only accounting for 
+`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 (Figure 
+\ref{fig:fig4}).
+
+```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.", out.width="100%"}
 load <- tar_read(daily_no3_08164000, store = store)
@@
 
-```{r}
+```{r fig5, message=FALSE, warning=FALSE, fig.height=3.5, fig.pos="H", fig.width=6, fig.cap="Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.", out.width="100%", wideplot=TRUE}
+draw_smoothed_years <- function(model1, model2, model3,
+                                ylab, subtitle) {
+
+  `TCEQ-13563` <- model1
+  `TCEQ-13383` <- model2
+  `TCEQ-13384` <- model3
+
+  comp <- gratia::compare_smooths(
+    `TCEQ-13563`,
+    `TCEQ-13383`,
+    `TCEQ-13384`,
+    smooths = "s(ddate)")
+
+  crit <- gratia:::coverage_normal(0.90)
+
+  comp |> 
+    unnest(data) |> 
+    mutate(lower_ci = est + (crit * se),
+           upper_ci = est - (crit * se),
+           label = if_else(ddate == max(ddate),
+                           as.character(model),
+                           NA_character_)) |> 
+    ggplot(aes(x = ddate, y = est, group = model)) +
+    geom_ribbon(aes(ymin = lower_ci,
+                    ymax = upper_ci,
+                    fill = model),
+                alpha = 0.15) +
+    geom_line(aes(color = model)) +
+    geom_hline(yintercept = 0, linetype = 2, alpha = 0.25) +
+    # geom_text_repel(aes(label = label,
+    #                     color = model), 
+    #                 nudge_x = 5,
+    #                 hjust = "left",
+    #                 size = 2) +
+    scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
+                       expand = expansion(mult = c(0.05,0.05))) +
+    colorspace::scale_fill_discrete_qualitative(name = "") +
+    colorspace::scale_color_discrete_qualitative(name = "") +
+    labs(x = "Year", y = ylab,
+         subtitle = subtitle) +
+    theme_TWRI_print() +
+    theme(axis.title.y = element_markdown(size = 8),
+          axis.title.x = element_markdown(size = 8),
+          plot.subtitle = element_markdown(size = 8),
+          legend.direction = "vertical",
+          legend.text = element_text(size = 8))
+
+}
+
+p_tp <- draw_smoothed_years(tar_read(tp_lavaca_13563_temporal, store = store),
+                            tar_read(tp_lavaca_13383_temporal, store = store),
+                            tar_read(tp_lavaca_13384_temporal, store = store),
+                            ylab = "TP Smooth<br>Estimate [mg/L]", 
+                            subtitle = "A: TP Trend")
+
+p_no3 <- draw_smoothed_years(tar_read(no3_lavaca_13563_temporal, store = store),
+                             tar_read(no3_lavaca_13383_temporal, store = store),
+                             tar_read(no3_lavaca_13384_temporal, store = store),
+                             ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]", 
+                             subtitle = "B: NO*<sub>x</sub>* Trend")
+
+p_tkn <- draw_smoothed_years(tar_read(tkn_lavaca_13563_temporal, store = store),
+                             tar_read(tkn_lavaca_13383_temporal, store = store),
+                             tar_read(tkn_lavaca_13384_temporal, store = store),
+                             ylab = "TKN Smooth<br>Estimate [mg/L]", 
+                             subtitle = "C: TKN Trend")
+
+p_chla <- draw_smoothed_years(tar_read(chla_lavaca_13563_temporal, store = store),
+                             tar_read(chla_lavaca_13383_temporal, store = store),
+                             tar_read(chla_lavaca_13384_temporal, store = store),
+                             ylab = "Chlorophyll-*a*\nSmooth<br>Estimate [µg/L]", 
+                             subtitle = "D: Chlorophyll-*a* Trend")
+
+p_do <- draw_smoothed_years(tar_read(do_lavaca_13563_temporal, store = store),
+                             tar_read(do_lavaca_13383_temporal, store = store),
+                             tar_read(do_lavaca_13384_temporal, store = store),
+                             ylab = "DO\nSmooth<br>Estimate [mg/L]", 
+                            subtitle = "E: DO Trend")
+
+design <- "
+112266
+334455
+"
+
+#(p_tp + p_no3) / (p_chla + p_tkn) / (p_do)
+p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
+                                                   design = design)
+```
+
+GAMs did not identify significant changes in TP or DO concentrations at any
+of the Lavaca Bay sites from 2005 through 2020 (Figure \ref{fig:fig5}). The
+upper-bay site, TCEQ-13563, had a linear increase in NO~x~ concentration and
+and decrease in chlorophyll-*a* from 2005 through 2014. The mid-bay site,
+TCEQ-13383, showed a periodic pattern in NO~x~ concentration that appeared
+similar to precipitation/inflow patterns, as well as a post 2011 increase
+in TKN concentrations. No significant long-term trends in concentrations 
+were identified by GAMs for the lower-bay TCEQ-13384 site.
+
+Freshwater inflow provided additional explanation for changes in TP and 
+NO~*x*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
+model probability values (Table \ref{tab:tab4}).
+TCEQ-13563, the site closest to the river outlet, was the only site that
+had improvements in the explanations of DO and TKN concentration with the inclusion 
+of inflow. Both TCEQ-13563 and TCEQ-13383, the mid-bay site, saw improvements 
+in explanations for variations in chlorophyll-*a* with the inclusion of freshwater
+inflow. The addition of nutrient loads (both TP and NO~3~) terms did not provide
+additional explanation for changes in chlorophyll-*a* or DO concentrations.
+Inclusion of TP loads provided additional explanation of TP concentrations at
+the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383. Inclusion of NO~3~
+loads only provided marginal improvements in the explanation of NO~*X*~ 
+concentration at the lower-bay TCEQ-13384 site.
+
+```{r tab4, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r long-smooths, message=FALSE, warning=FALSE, fig.pos="H", fig.height=3.5, fig.pos="H", fig.width=6, fig.cap="Smoothed trend component for water quality paramaters obtained from Temporal estaury GAMs.", wideplot=list("0cm"), out.width="100%", wideplot=TRUE}
-draw_smoothed_years <- function(model1, model2, model3,
-                                ylab, subtitle) {
-
-  `TCEQ-13563` <- model1
-  `TCEQ-13383` <- model2
-  `TCEQ-13384` <- model3
-
-  comp <- gratia::compare_smooths(
-    `TCEQ-13563`,
-    `TCEQ-13383`,
-    `TCEQ-13384`,
-    smooths = "s(ddate)")
+```{r fig6, message=FALSE, warning=FALSE, fig.height=3.5, fig.pos="H", fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO\\textsubscript{\\emph{x}}, chlorophyll-\\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.", out.width="100%", wideplot=TRUE}
+draw_smoothed_surface <- function(comp,
+                                  x = flw_res,
+                                  xlab,
+                                  ylab, 
+                                  subtitle) {
 
@@
     mutate(lower_ci = est + (crit * se),
-           upper_ci = est - (crit * se),
-           label = if_else(ddate == max(ddate),
+           upper_ci = est - (crit * se)) |>
+    group_by(model) |> 
+    mutate(label = if_else(
+      {{ x }} == max( {{ x }} ),
       as.character(model),
       NA_character_)) |> 
-    ggplot(aes(x = ddate, y = est, group = model)) +
+    ggplot(aes(x = {{ x }}, y = est, group = model)) +
     geom_ribbon(aes(ymin = lower_ci,
@@
     geom_line(aes(color = model)) +
-    geom_hline(yintercept = 0, linetype = 2, alpha = 0.25) +
+    geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
     # geom_text_repel(aes(label = label,
@@
     #                 size = 2) +
-    scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
-                       expand = expansion(mult = c(0.05,0.05))) +
+    scale_x_continuous(expand = expansion(mult = c(0.05,0.05))) +
     colorspace::scale_fill_discrete_qualitative(name = "") +
     colorspace::scale_color_discrete_qualitative(name = "") +
-    labs(x = "Year", y = ylab,
+    labs(x = xlab, y = ylab,
          subtitle = subtitle) +
@@
           plot.subtitle = element_markdown(size = 8),
-          legend.direction = "vertical")
+          legend.direction = "vertical",
+          legend.text = element_text(size = 8))
 
@@
 
-p_tp <- draw_smoothed_years(tar_read(tp_lavaca_13563_temporal, store = store),
-                            tar_read(tp_lavaca_13383_temporal, store = store),
-                            tar_read(tp_lavaca_13384_temporal, store = store),
-                            ylab = "TP\nSmooth<br>Estimate [mg/L]", subtitle = "TP Trend")
-
-p_no3 <- draw_smoothed_years(tar_read(no3_lavaca_13563_temporal, store = store),
-                             tar_read(no3_lavaca_13383_temporal, store = store),
-                             tar_read(no3_lavaca_13384_temporal, store = store),
-                             ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]", subtitle = "NO*<sub>x</sub>* Trend")
-
-p_tkn <- draw_smoothed_years(tar_read(tkn_lavaca_13563_temporal, store = store),
-                             tar_read(tkn_lavaca_13383_temporal, store = store),
-                             tar_read(tkn_lavaca_13384_temporal, store = store),
-                             ylab = "TKN\nSmooth<br>Estimate [mg/L]", subtitle = "TKN Trend")
-
-p_chla <- draw_smoothed_years(tar_read(chla_lavaca_13563_temporal, store = store),
-                             tar_read(chla_lavaca_13383_temporal, store = store),
-                             tar_read(chla_lavaca_13384_temporal, store = store),
-                             ylab = "Chlorophyll-*a*\nSmooth<br>Estimate [µg/L]", subtitle = "Chlorophyll-*a* Trend")
-
-p_do <- draw_smoothed_years(tar_read(do_lavaca_13563_temporal, store = store),
-                             tar_read(do_lavaca_13383_temporal, store = store),
-                             tar_read(do_lavaca_13384_temporal, store = store),
-                             ylab = "DO\nSmooth<br>Estimate [mg/L]", subtitle = "DO Trend")
+`TCEQ-13383` <- tar_read(tp_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `TCEQ-13383`,
+                      `TCEQ-13384` = tar_read(tp_lavaca_13384_flow, store = store),
+                      `TCEQ-13563` = tar_read(tp_lavaca_13563_flow, store = store),
+                      smooths = c("s(flw_res)"))
+p_tp <- draw_smoothed_surface(comp,
+                      x = flw_res,
+                      xlab = "log(Inflow) Residuals [cfs]",
+                      ylab =  "TP Smooth<br>Estimate [mg/L]",
+                      subtitle = "A: TP")
+
+`TCEQ-13383` <- tar_read(no3_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `TCEQ-13383`,
+                      `TCEQ-13384` = tar_read(no3_lavaca_13384_flow, store = store),
+                      `TCEQ-13563` = tar_read(no3_lavaca_13563_flow, store = store),
+                      smooths = c("s(flw_res)"))
+p_no3 <- draw_smoothed_surface(comp,
+                      x = flw_res,
+                      xlab = "log(Inflow) Residuals [cfs]",
+                      ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]",
+                      subtitle = "B: NO*<sub>x</sub>*")
+
+`TCEQ-13383` <- tar_read(chla_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `TCEQ-13383`,
+                      `TCEQ-13384` = tar_read(chla_lavaca_13384_flow, store = store),
+                      `TCEQ-13563` = tar_read(chla_lavaca_13563_flow, store = store),
+                      smooths = c("s(flw_res)"))
+p_chla <- draw_smoothed_surface(comp,
+                      x = flw_res,
+                      xlab = "log(Inflow) Residuals [cfs]",
+                      ylab = "Chlorophyll-*a* Smooth<br>Estimate [µg/L]",
+                      subtitle = "C: Chlorophyll-*a*")
+
+`TCEQ-13383` <- tar_read(tkn_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `TCEQ-13383`,
+                      `TCEQ-13384` = tar_read(tkn_lavaca_13384_flow, store = store),
+                      `TCEQ-13563` = tar_read(tkn_lavaca_13563_flow, store = store),
+                      smooths = c("s(flw_res)"))
+p_tkn <- draw_smoothed_surface(comp,
+                      x = flw_res,
+                      xlab = "log(Inflow) Residuals [cfs]",
+                      ylab = "TKN Smooth<br>Estimate [mg/L]",
+                      subtitle = "D: TKN")
+
+`TCEQ-13383` <- tar_read(do_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `TCEQ-13383`,
+                      `TCEQ-13384` = tar_read(do_lavaca_13384_flow, store = store),
+                      `TCEQ-13563` = tar_read(do_lavaca_13563_flow, store = store),
+                      smooths = c("s(flw_res)"))
+p_do <- draw_smoothed_surface(comp,
+                      x = flw_res,
+                      xlab = "log(Inflow) Residuals [cfs]",
+                      ylab = "DO Smooth<br>Estimate [mg/L]",
+                      subtitle = "E: DO")
 
@@
 
-#(p_tp + p_no3) / (p_chla + p_tkn) / (p_do)
 p_tp + p_no3 + p_chla + p_tkn + p_do + guide_area() + plot_layout(guides = "collect",
                                                    design = design)
+
 ```
 
+GAMs show that increases in freshwater inflow result in nearly linear increases 
+in TP and NO~*x*~ at all three sites (Figure \ref{fig:6}). At the upper-bay
+TCEQ-13563 site, GAMs showed increases in freshwater inflow initially 
+increased chlorophyll-*a* and DO concentration but concentrations leveled and
+potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
+nearly linear increased in chlorophyll-*a* concentration with increases in
+freshwater inflow. Freshwater flow did not have significant
+effects on chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.
+
+Increases in TP loads resulted in nearly linear increases of TP concentration at
+the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively (Figure \ref{fig:fig7}).
+The relative effect size appeared to much smaller than the effect of freshwater
+inflow alone. Increases in NO~3~ loads only showed an effect at the 
+lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
+provided only small improvements to the model (Table \ref{tab:tab4}). As noted
+above, nutrient loadings did not provide any explanation in changes in the
+remaining assessed water quality parameters.
+
+```{r fig7, message=FALSE, warning=FALSE, fig.height=2.25, fig.width=5.5, fig.pos="H", fig.cap="Estimated effects of nutrient load residuals on TP and NO\\textsubscript{\\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.", out.width="100%", wideplot=FALSE}
+
+`TCEQ-13383` <- tar_read(tp_lavaca_13383_full, store = store)
+
+comp <- compare_smooths(model = `TCEQ-13383`,
+                      `TCEQ-13384` = tar_read(tp_lavaca_13384_full, store = store),
+                      `TCEQ-13563` = tar_read(tp_lavaca_13563_full, store = store),
+                      smooths = c("s(TP_resid)"))
+p_tp <- draw_smoothed_surface(comp,
+                      x = TP_resid,
+                      xlab = "TP Load Residuals [kg]",
+                      ylab = "TP [mg/L]",
+                      subtitle = "A: TP")
+
+`TCEQ-13383` <- tar_read(no3_lavaca_13383_full, store = store)
+
+comp <- compare_smooths(model = `TCEQ-13383`,
+                      `TCEQ-13384` = tar_read(no3_lavaca_13384_full, store = store),
+                      `TCEQ-13563` = tar_read(no3_lavaca_13563_full, store = store),
+                      smooths = c("s(NO3_resid)"))
+p_no3 <- draw_smoothed_surface(comp,
+                      x = NO3_resid,
+                      xlab = "NO<sub>3</sub> Load Residuals [kg]",
+                      ylab = "NO*<sub>x</sub>* [mg/L]",
+                      subtitle = "B: NO*<sub>x</sub>*")
+
+design <- "
+11223
+"
+
+p_tp + p_no3 + guide_area() + plot_layout(guides = "collect",
+                                                   design = design)
+
+```
+
 # Discussion
@@
 ```{r tpcomp}
+
+load_lav_tp <- tar_read(daily_tp_08164000, store = store)
+load_nav_tp <- tar_read(daily_tp_texana, store = store)
+
+dat <- load_lav_tp$annually |> 
+  ungroup() |> 
+  filter(year >= 2005) |> 
+  summarise(annual = round(mean(TP_Estimate)/2116,1),
+            upper = round(mean(TP_Upper)/2116,1),
+            lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km
+
 df <- tibble(
   Parameter = rep("TP", 5),
-  `Reported Yield (kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})` = c("42.9 (34.4, 54.0)", "45.2", "42", "20.81-91.58", "28.9"),
+  `Reported Yield (kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})` = 
+    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")"), 
+      "45.2", "42", "20.81-91.58", "28.9"),
   Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
-  `Time Period` = c("2000-2020", "2012", "1977-2005", "2002", "1972-1993"),
+  `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
   Reference = c("This work", "\\citet{wise_spatially_2019}",
@@
 )
+
+df[[1,2]] <- paste0(df[[1,2]], footnote_marker_symbol(1))
+df[[4,2]] <- paste0(df[[4,2]], footnote_marker_symbol(2))
 
@@
                           "Time Period",
-                          "Reference")))
+                          "Reference")),
+    caption = "Mean estimates of annual TP yield in the Lavaca River watershed in published studies.") |>
+  add_footnote(c("Values represent the mean of annual point estimates, lower and upper 95% credible intervals.",
+                 "A single point estimate was not reported, these value represent the range depicted on the choropleth map provided in the report."),
+               notation = "symbol",
+               threeparttable = TRUE,
+               escape = TRUE)
 

### 2023-02-09 Schramm-Manuscript-2023/Schramm-Manuscript-2023.tex
results

@@
 Because streamflow and nutrient loads are tightly correlated, freshwater
-inflow masks any potential signals from nutrient loads alone. Following
-the methodology implemented by
+inflow can mask signals from nutrient loads alone. Following the
+methodology implemented by
 \citet{murphyNutrientImprovementsChesapeake2022}, both streamflow and
@@
 
-Lavaca NO\textsubscript{3} r2 = 0.85 and 0.90 deviance explained TP r2 =
-0.27 and 0.33 Texana NO\textsubscript{3} r2 = 0.75 and 0.81 TP = 0.32
-and 0.38
+Based on criteria provided by \citet{moriasiHydrologicWaterQuality2015},
+GAMs ranged from ``satisfactory'' to ``very good'' based on median NSE,
+r\textsuperscript{2} and PBIAS calculated from repeated 5-fold cross
+validation on predicted and measured nutrient loads. NO\textsubscript{3}
+GAM models had median NSE values of 0.34 and 0.48 at USGS-08164000 and
+USGS-08164390 respectively. Median r\textsuperscript{2} values were 0.70
+(USGS-08164000) and 0.87 (USGS-08164525) and PBIAS values were 2.00
+(USGS-08164000) and 10.90 (USGS-08164525) for NO\textsubscript{3} loads.
+Median NSE values were 0.80 (USGS-08164000) and 0.91 (USGS-08164525) for
+TP loads. Median r\textsuperscript{2} values for TP loads were 0.93
+(USGS-08164000) and 0.99 (USGS-08164525) and median PBIAS values were
+-7.20 (USGS-08164000) and -3.30 (USGS-08164525). Density plots of
+metrics show similar distribution of values between sites for the same
+parameter, with the exception r\textsuperscript{2} values for
+NO\textsubscript{3} loads where USGS-08164000 showed a much larger
+variance in values compared to USGS-08164525 (Figure \ref{fig:fig2}).In
+addition to higher average NSE and r\textsuperscript{2} values, GAMs had
+smaller variance in metric values for TP compared to
+NO\textsubscript{3}.
 
@@
 
+Annual NO\textsubscript{3} and TP loads show considerable variation,
+generally following patterns in discharge (Figure \ref{fig:fig3}, Figure
+\ref{fig:fig4}). Flow-normalized TP loads at both sites and
+flow-normalized loads at USGS-08164000 indicate watershed based loads
+have not changed much over time when accounting for changes in
+streamflow (Figure \ref{fig:fig3}). Flow-normalized loads at
+USGS-08164000 show small variation over time with some decreases in
+NO\textsubscript{3} loads since 2013. Aggregated across both sites, the
+mean annual NO\textsubscript{3} load 2005 through 2020 was 205,405 kg
+(126,867 kg - 341,569 kg, 90\% CI). Annual NO\textsubscript{3} loads
+ranged from 12,574 kg in 2011 to 794,510 kg in 2007.
+
+Total annual TP loads ranged from 7,839 kg in 2011 to 595,075 kg in
+2007. Mean annual TP loading from 2005 through 2020 was 182,673 kg
+(152,227 kg - 219,310 kg, 90\% CI). On average, USGS-08164525 accounted
+for 68\% of NO\textsubscript{3} loads and 59\% of TP loads from 2005
+through 2020. However, during periods of extreme drought the Lavaca
+River (USGS-08164000) became the primary source of nutrient loading in
+the watershed with the Navidad River only accounting for 15\% and 25\%
+of NO\textsubscript{3} and TP loads in 2011 (Figure \ref{fig:fig4}).
+
 \begin{figure}[H]
@@
 
-\caption[Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525]{Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
+\caption[Aggregated estimated annual and flow-normalized annual NO  extsubscript{3} and TP loads for USGS-08164000 and USGS-08164525]{Aggregated estimated annual and flow-normalized annual NO extsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
 \end{figure}
@@
 
-\begin{verbatim}
-## Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result. 
-## Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result. 
-## Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result.
-\end{verbatim}
+\begin{figure}[H]\begin{adjustwidth}{-\extralength}{0cm}
+
+{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig5-1}
+
+}
+
+\end{adjustwidth}\caption[Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs]{Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.}\label{fig:fig5}
+\end{figure}
+
+GAMs did not identify significant changes in TP or DO concentrations at
+any of the Lavaca Bay sites from 2005 through 2020 (Figure
+\ref{fig:fig5}). The upper-bay site, TCEQ-13563, had a linear increase
+in NO\textsubscript{x} concentration and and decrease in
+chlorophyll-\emph{a} from 2005 through 2014. The mid-bay site,
+TCEQ-13383, showed a periodic pattern in NO\textsubscript{x}
+concentration that appeared similar to precipitation/inflow patterns, as
+well as a post 2011 increase in TKN concentrations. No significant
+long-term trends in concentrations were identified by GAMs for the
+lower-bay TCEQ-13384 site.
+
+Freshwater inflow provided additional explanation for changes in TP and
+NO\textsubscript{\emph{x}} concentration at all of the Lavaca Bay sites
+according to AIC\textsubscript{c} and model probability values (Table
+\ref{tab:tab4}). TCEQ-13563, the site closest to the river outlet, was
+the only site that had improvements in the explanations of DO and TKN
+concentration with the inclusion of inflow. Both TCEQ-13563 and
+TCEQ-13383, the mid-bay site, saw improvements in explanations for
+variations in chlorophyll-\emph{a} with the inclusion of freshwater
+inflow. The addition of nutrient loads (both TP and NO\textsubscript{3})
+terms did not provide additional explanation for changes in
+chlorophyll-\emph{a} or DO concentrations. Inclusion of TP loads
+provided additional explanation of TP concentrations at the upper- and
+mid-bay sites, TCEQ-13563 and TCEQ-13383. Inclusion of
+NO\textsubscript{3} loads only provided marginal improvements in the
+explanation of NO\textsubscript{\emph{X}} concentration at the lower-bay
+TCEQ-13384 site.
 
@@
 
-\caption{\label{tab:unnamed-chunk-1}Estuary GAM AIC\textsubscript{c} values and associated model probabilities (in parenthesis).
+\caption{\label{tab:tab4}Estuary GAM AIC\textsubscript{c} values and associated model probabilities (in parenthesis).
 Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.}
@@
 
-{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/long-smooths-1}
+{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig6-1}
 
@@
 
-\end{adjustwidth}\caption[Smoothed trend component for water quality paramaters obtained from Temporal estaury GAMs]{Smoothed trend component for water quality paramaters obtained from Temporal estaury GAMs.}\label{fig:long-smooths}
+\end{adjustwidth}\caption[Estimated effects of mean daily inflow residuals on mean TP, NO\textsubscript{\emph{x}}, chlorophyll-\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs]{Estimated effects of mean daily inflow residuals on mean TP, NO\textsubscript{\emph{x}}, chlorophyll-\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.}\label{fig:fig6}
+\end{figure}
+
+GAMs show that increases in freshwater inflow result in nearly linear
+increases in TP and NO\textsubscript{\emph{x}} at all three sites
+(Figure \ref{fig:6}). At the upper-bay TCEQ-13563 site, GAMs showed
+increases in freshwater inflow initially increased chlorophyll-\emph{a}
+and DO concentration but concentrations leveled and potentially
+decreased at higher flows. The mid-bay TCEQ-13383 site showed a nearly
+linear increased in chlorophyll-\emph{a} concentration with increases in
+freshwater inflow. Freshwater flow did not have significant effects on
+chlorophyll-\emph{a}, TKN, or DO at the lower-bay TCEQ-13384 site.
+
+Increases in TP loads resulted in nearly linear increases of TP
+concentration at the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383
+respectively (Figure \ref{fig:fig7}). The relative effect size appeared
+to much smaller than the effect of freshwater inflow alone. Increases in
+NO\textsubscript{3} loads only showed an effect at the lower-bay
+TCEQ-13384 site. The effect was quite small compared to streamflow and
+provided only small improvements to the model (Table \ref{tab:tab4}). As
+noted above, nutrient loadings did not provide any explanation in
+changes in the remaining assessed water quality parameters.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig7-1}
+
+}
+
+\caption[Estimated effects of nutrient load residuals on TP and NO\textsubscript{\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs]{Estimated effects of nutrient load residuals on TP and NO\textsubscript{\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.}\label{fig:fig7}
 \end{figure}
@@
 
+\begin{threeparttable}
+\caption{\label{tab:tpcomp}Mean estimates of annual TP yield in the Lavaca River watershed in published studies.}
+\centering
 \begin{tabular}[t]{lllll}
@@
 \midrule
-TP & 42.9 (34.4, 54.0) & GAM & 2000-2020 & This work\\
-TP & 45.2 & SPARROW & 2012 & \citet{wise_spatially_2019}\\
+TP & 35.2 (28.8, 43.3)\textsuperscript{*} & GAM & 2005-2020 & This work\\
+TP & 45.2 & SPARROW & 2000-2014 & \citet{wise_spatially_2019}\\
 TP & 42 & SWAT & 1977-2005 & \citet{omaniEstimationSedimentNutrient2014}\\
-TP & 20.81-91.58 & SPARROW & 2002 & \citet{rebich_sources_2011}\\
+TP & 20.81-91.58\textsuperscript{\dag} & SPARROW & 1980-2002 & \citet{rebich_sources_2011}\\
 TP & 28.9 & LOADEST & 1972-1993 & \citet{dunnTrendsNutrientInflows1996}\\
@@
 \end{tabular}
+\begin{tablenotes}
+\small
+\item [*] Values represent the mean of annual point estimates, lower and upper 95\% credible intervals.
+\item [\dag] A single point estimate was not reported, these value represent the range depicted on the choropleth map provided in the report.
+\end{tablenotes}
+\end{threeparttable}
 \end{table}

### 2023-02-10 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
Intro kinda done

@@
 # front matter
-simplesummary: |
-  A Simple summary goes here.
 abstract: |
@@
 keywords: |
-  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
-  to the article, yet reasonably common within the subject discipline.).
+  estuary; nutrient loading; water quality; Texas
 acknowledgement: |
-  All sources of funding of the study should be disclosed. Please clearly 
-  indicate grants that you have received in support of your research work. 
-  Clearly state if you received funds for covering the costs to publish in open 
-  access.
+  Thank you to Mike Wetz (Harte Institute, Texas A&M Corpus Christi),
+  Chad Kinsfather and Partick Brzozowski (Lavaca-Navidad River Authority),
+  Brian Koch (Texas State Soil and Water Conservation Board), Bill Balboa 
+  (Matagorda Bay Foundation) and the Lavaca Bay Foundation for supporting 
+  development of this project and providing valuable feedback. 
+  Thank you to Ram Neupane (Texas Water Development Board) for providing data.
 funding: |
@@
 conflictsofinterest: |
-  The author declares no conflict of interest. The founding sponsors had no 
+  The author declares no conflict of interest. The project sponsors had no 
   role in the design of the study; in the collection, analyses, or 
-  interpretation of data; in the writing of the manuscript, an in the decision 
+  interpretation of data; in the writing of the manuscript, or in the decision 
   to publish the results.
@@
                       dpi = 300,
-                      fig.pos = "H",
                       fig.align = "center",
@@
     x_out <- gsub(x = x_out,
-                  pattern='\\begin{figure}[H]', 
-                  replacement='\\begin{figure}[H]\\begin{adjustwidth}{-\\extralength}{0cm}',
+                  pattern='\\begin{figure}', 
+                  replacement='\\begin{figure}\\begin{adjustwidth}{-\\extralength}{0cm}',
                   fixed=TRUE)
@@
 
-Like many estuaries globally, estuaries along the Texas Gulf coast are facing 
-pressures from increasing population, increases in point source 
-and non-point source pollution and alterations to freshwater inflows leading to
-increases in the occurrences and risks of algal blooms and eutrophication [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; 
-@bugica_water_2020]. Recent studies indicate that estuary water quality
-dynamics in both agriculturally dominated and urban watersheds within Texas are
-displaying signals of conditions increasingly conducive to eutrophication 
-[@wetzWaterQualityDynamics2016; @wetz_exceptionally_2017; 
-@bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022].
+Like many coastal areas globally, coastal watersheds along the Texas Gulf coast 
+are facing pressures from increasing population, increases in point source 
+and non-point source pollution and alterations to freshwater flows that alter
+water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. Despite these increasing pressures, national scale 
+assessments have classified coastal estuaries in Texas as moderate or lower for
+exhibiting eutrophic conditions [@bricker_effects_2008]. However, a suite of 
+recent studies indicate that estuary water quality dynamics in both 
+agriculturally dominated and urban watersheds within Texas are in fact 
+expressing conditions that are increasingly conducive to algal blooms and 
+eutrophication [@wetzWaterQualityDynamics2016; @wetz_exceptionally_2017; 
+@bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022]. With identification
+of localized areas of estuary water quality concern along the Texas coast 
+[@bugica_water_2020], localized studies are being prioritized to better 
+inform management actions.
+
+This project provides an assessment of nutrient loading and water quality
+responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
+Matagorda Bay system located roughly halfway between Houston, Texas and Corpus
+Christi, Texas. Lavaca Bay has faced substantial challenges associated with
+legacy contamination but general water quality parameters such as dissolved oxygen 
+(DO), nutrients, and biological parameters have been well within state
+water quality standards. However, long-term declines of benthic 
+fauna abundance, biomass, and diversity in Lavaca Bay that are 
+linked to reductions in freshwater inflows and changes in estuary salinity 
+[@beserespollackLongtermTrendsResponse2011; @palmerImpactsDroughtsLow2015;
+@montagnaAssessmentRelationshipFreshwater2020] and are a concern to local 
+stakeholders. @bugica_water_2020 identified monotonic increases in total 
+phosphorus (TP), orthophosphate, total Kjeldahl nitrogen (TKN), and
+chlorophyll-*a* at sites within Lavaca Bay. Although no long-term changes in 
+DO concentrations were identified, the trends in nutrient concentrations are 
+concerning due to the role of nitrogen as a limiting factor for primary 
+production in many Texas estuaries [@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; @paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the
+ramifications that major changes in nitrogen loadings could have for 
+productivity and eutrophication in Lavaca Bay.
+
+There are ongoing efforts between local, state, and federal agencies to 
+address water quality impairments in the freshwater portions of the 
+Lavaca Bay watershed [@jainTechnicalSupportDocument2021; @schramm_lavaca_2018; @bertholdDirectMailingEducation2021]. However, at a statewide scale, these
+approaches have shown limited success and emphasize a need for improved efforts
+at assessing and linking management actions with downstream water quality to
+better identify  and replicate effective management actions across the state
+[@schrammTotalMaximumDaily2022]. The identification and communication of 
+changes and trends in water quality is complicated by the fact that trends are 
+often non-linear and confounded by precipitation and runoff that 
+hinder traditional analysis [@wazniakLinkingWaterQuality2007; 
+@lloydMethodsDetectingChange2014]. To provide
+actionable information for resource managers, water quality conditions must be 
+evaluated relative to changes in natural environmental drivers to 
+better understand  and manage potential anthropogenic effects.
+
+Here something some we present methodology and results for estimating nutrient
+loading- and assessing linkages in estuary water quality parameters with watershed derived
+nutrient loads....
 
@@
 ## Study Area and Data
+
+\begin{figure}
+\begin{adjustwidth}{-\extralength}{0cm}
+\centering
+\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
+\end{adjustwidth}
+\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
+\end{figure}
 
@@
 
-\begin{figure}[H]
-\begin{adjustwidth}{-\extralength}{0cm}
-\centering
-\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
-\end{adjustwidth}
-\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
-\end{figure}
-
-Daily discharges for the Lavaca River (USGS-08164000) were obtained from the 
-United States Geologic Survey (USGS) National Water Information System using the 
-*dataRetrieval* R package [@deciccoDataRetrievalPackagesDiscovering2022]. Gaged
-dailly discharges from Lake Texana (USGS-0816425) were provided by the Texas
-Water Development Board (TWDB) (April 21, 2022 email from R. Neupane, TWDB).
-
-Water quality sample data for both freshwater and estuary locations were o
-btained from the Texas Commission on Environmental Quality (TCEQ) Surface Water 
+**Need to add a short description of land use**
+
+Daily discharges for the Lavaca River (USGS-08164000, Figure \ref{fig:fig1}) 
+were obtained from the United States Geologic Survey (USGS) National Water 
+Information System using the *dataRetrieval* R package 
+[@deciccoDataRetrievalPackagesDiscovering2022]. Gaged daily discharges from 
+the outlet of Lake Texana on the Navidad River (USGS-0816425) were provided by 
+the Texas Water Development Board (TWDB) 
+(April 21, 2022 email from R. Neupane, TWDB).
+
+Water quality sample data for the two freshwater and three estuary locations were 
+obtained from the Texas Commission on Environmental Quality (TCEQ) Surface Water 
 Quality Monitoring Information System. Data submitted through the system 
@@
 different projects. 
-For freshwater locations, total phosphorus (TP) and nitrate-nitrogen (NO~3~) 
-data were downloaded (Table \ref{tab:fwsummary}). Unfortunately, insufficient data was available for 
-assessment of total nitrogen (TN) and total Kjeldahl nitrogen (TKN) loadings, 
-so analysis was restricted to TP and NO~3~ loads. For estuary locations, we 
-obtained data for TP, Nitrite+Nitrate (NO*~x~*), TKN, chlorophyll-*a*, and
-dissolved oxygen (Table \ref{tab:estuarysummary}).
+All sites had varying lengths of and availability of data.
+For freshwater locations, TP from January 2000 through December 2020 and 
+nitrate-nitrogen (NO~3~) data from January 2005 through December 2020 were 
+downloaded (Table \ref{tab:fwsummary}). Less than 5-years of total
+nitrogen and TKN concentration data were available at the freshwater sites and
+deemed insufficient to develop load estimation models (**CITE**). 
+The three estuary sites included an upper Lavaca Bay site near the outlet of the
+Lavaca River system (TCEQ-13563), a mid-Lavaca Bay site (TCEQ-13383), and the 
+lower Lavaca Bay site near the mouth of the Bay (TCEQ-13384). 
+For estuary locations, we obtained data for TP, Nitrite+Nitrate (NO*~x~*), TKN, 
+chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
+(Table \ref{tab:estuarysummary}).
 
@@
 
-Estimates of nutrient loads were developed using Generalized Additive Models
-(GAMs) relating nutrient concentration to river discharge, season, and time.
+Estimates of NO~3~ and TP loads at the Lavaca River (USGS-08164000) and the 
+outlet of Lake Texana on the Navidad River (USGS-08164250) were developed using 
+Generalized Additive Models (GAMs) relating nutrient concentration to river 
+discharge, season, and time.
 Separate models were fit at each station for each parameter and used to predict
@@
 smoothed discounted flow and demonstrated improvements in nutrient loading 
-models by including the term. @kuhnert_quantifying_2012 expresses MA as:
+models by including the term. @kuhnert_quantifying_2012 expresses *ma* as:
 
@@
 Based on criteria provided by @moriasiHydrologicWaterQuality2015, GAMs ranged
-from "satisfactory" to "very good" based on median NSE, r^2^ and PBIAS 
-calculated from repeated 5-fold cross validation on predicted and measured
+from "satisfactory" to "very good" using median NSE, r^2^ and PBIAS 
+calculated using repeated 5-fold cross validation on predicted and measured
 nutrient loads. NO~3~ GAM models had median NSE values of 0.34 and 0.48 at
@@
 
-```{r fig2, message=FALSE, warning=FALSE, fig.height=3.5, fig.width=6.5, fig.pos="H", fig.cap="Density plots of goodness-of-fit metrics (NSE, r\\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
+```{r fig2, message=FALSE, warning=FALSE, fig.height=3.5, fig.width=6.5, fig.cap="Density plots of goodness-of-fit metrics (NSE, r\\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
 Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics.", out.width="100%", wideplot=TRUE}
@@
 
-```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.", out.width="100%"}
+```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.", out.width="100%"}
 load <- tar_read(daily_no3_08164000, store = store)
@@
 
-```{r fig4, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.pos="H", fig.cap="Comparison of delivered annual loads at USGS-08164000 and USGS-08164525.", out.width="100%"}
+```{r fig4, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Comparison of delivered annual loads at USGS-08164000 and USGS-08164525.", out.width="100%"}
 load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
@@
 
-```{r fig5, message=FALSE, warning=FALSE, fig.height=3.5, fig.pos="H", fig.width=6, fig.cap="Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.", out.width="100%", wideplot=TRUE}
+```{r fig5, message=FALSE, warning=FALSE, fig.height=3.5, fig.width=6, fig.cap="Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.", out.width="100%", wideplot=TRUE}
 draw_smoothed_years <- function(model1, model2, model3,
@@
 
-```{r fig6, message=FALSE, warning=FALSE, fig.height=3.5, fig.pos="H", fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO\\textsubscript{\\emph{x}}, chlorophyll-\\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.", out.width="100%", wideplot=TRUE}
+```{r fig6, message=FALSE, warning=FALSE, fig.height=3.5, fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO\\textsubscript{\\emph{x}}, chlorophyll-\\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.", out.width="100%", wideplot=TRUE}
 draw_smoothed_surface <- function(comp,
@@
 
-```{r fig7, message=FALSE, warning=FALSE, fig.height=2.25, fig.width=5.5, fig.pos="H", fig.cap="Estimated effects of nutrient load residuals on TP and NO\\textsubscript{\\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.", out.width="100%", wideplot=FALSE}
+```{r fig7, message=FALSE, warning=FALSE, fig.height=2.25, fig.width=5.5, fig.cap="Estimated effects of nutrient load residuals on TP and NO\\textsubscript{\\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.", out.width="100%", wideplot=FALSE}
 
@@
 # Discussion
+
+TP and NO~3~ loadings from the Lavaca Bay watershed showed high inter-annual
+variability tied with changes in discharge. There is little evidence for changes
+in flow-normalized TP loads in either rivers. There is some evidence of recent
+decreases in flow-normalized NO~3~ loads in the Lavaca River. Although there is
+no work directly correlating water quality planning and implementation efforts
+in the watershed to water quality outcomes, efforts to increase agricultural 
+producer participation in the watershed have been ongoing since 2016
+[@schramm_lavaca_2018; @bertholdDirectMailingEducation2021]. The decrease in 
+flow-normalized NO~3~ loads could be a reflection of those collective efforts 
+but further data collection and research is required to support that statement.
+
+Other studies have published estimates of TP loading estimates for the Lavaca
+River (USGS-08164000) over various time periods (Table \ref{tab:tpcomp}).
+Converted to average annual yield, the estimates of TP loading in this study
+are within the ranges of previous studies [@dunnTrendsNutrientInflows1996; @rebich_sources_2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019], with the TP 
+estimates in @dunnTrendsNutrientInflows1996 being notably lower. Given that none
+of the studies identify substantially sized trends in TP, it is is possible that 
+the period used in @dunnTrendsNutrientInflows1996 was drier on average than the
+other studies. The SPARROW models used in @rebich_sources_2011 and 
+@wise_spatially_2019 utilize a version of LOADEST in the underlying load
+estimation procedure, so a difference due to methodology alone is unlikely.
 
@@
 
-Authors should discuss the results and how they can be interpreted in 
-perspective of previous studies and of the working hypotheses. The findings and 
-their implications should be discussed in the broadest context possible. Future 
-research directions may also be highlighted.
+Cross-validation of the GAM loading models indicated that GAMs performed well on
+average at predicting daily nutrient loading values. The variance in scores was
+very high indicating subsets of values were problematic at characterizing functional
+relationships between nutrients and predictors. All of the water
+quality data for the two river sites in the TCEQ database were ambient water
+quality data, collected to be representative of typical conditions. This results
+in few data collected at the highest portions of the flow duration curve
+under which the majority of loadings take place. Supplementary flow-biased 
+monitoring targeting storm- or high-flow conditions is recommended here to
+improve the precision of GAM predictions [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017].
+
+The non-linear temporal water quality trends identified using GAMs differed
+slightly from the trends identified by @bugicaWaterQualityTrends2020. This is
+not unexpected due to the different time periods, different methodology, and
+generally small slopes identified for most of the significant water quality
+parameters in @bugica_water_2020. The trend in DO and 
+cholorophyll-*a* concentrations are stable in comparison to other Texas
+estuaries that might be facing larger demands for freshwater diversions, higher
+population growth, and more intense agricultural production 
+[@wetzWaterQualityDynamics2016; @bugica_water_2020]. The trend of 
+increasing NO~*x*~ concentration at the upper-bay TCEQ-13563 site and recent 
+increases in TKN concentration at the mid-bay TCEQ-13383
+site are concerning due to the nitrogen limitation identified in many Texas 
+estuaries [@gardnerNitrogenFixationDissimilatory2006; 
+@houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; 
+@paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] 
+and the relatively low ambient concentrations observed in Lavaca Bay.
+
+The strong positive effect of freshwater inflow on NO~*x*~, TKN, and TP are
+suggestive of nonpoint watershed sources, consistent with watershed uses, and
+consistent with other studies relating freshwater inflow with nutrient 
+concentrations in Lavaca Bay and other estuaries [@russell_effect_2006; 
+@caffreyHighNutrientPulses2007; @peierlsNonmonotonicResponsesPhytoplankton2012; 
+@palmerImpactsDroughtsLow2015; @ciraPhytoplanktonDynamicsLowInflow2021]. Inflow 
+had a non-linear relationship with TKN at the two upstream sites, with TKN 
+increasing as freshwater inflow transitioned from low to moderate levels. 
+At higher freshwater inflows, the effect was attenuated, possibly indicating a 
+flushing effect at higher freshwater inflow. No relationship between TKN and 
+freshwater inflow were observed at TCEQ-13384 located in the lower reach of 
+Lavaca Bay. Tidal flushing from Matagorda Bay might be responsible for 
+decreasing TKN concentrations and limiting the effects of freshwater inflow in 
+lower reaches of Lavaca Bay. @russell_effect_2006 suggested the processing of 
+organic loads in the upper portions of Lavaca Bay reduces the transport of 
+nutrients into the lower reaches of the Bay.
+
+Freshwater inflow also had a strong positive effect on chlorophyll-*a* at 
+the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed decreases
+in chlorophyll-*a* at the highest freshwater inflow volumes. Freshwater flushing
+or increases in turbidity are associated with decreases in chlorophyll-*a* in
+other estuaries [@peierlsNonmonotonicResponsesPhytoplankton2012; 
+@cloernPhytoplanktonPrimaryProduction2014]. No relationships between inorganic
+nitrogen or TP loadings with chlorophyll-*a* were observed. Due to the lack
+of TKN loading information, no assessment between organic nitrogen loads and
+chlorophyll-*a* were possible.
+
+Although other studies have identified complex relationships between estuary 
+nutrient concentrations, nutrient loading and chlorophyll-*a* concentrations in 
+Texas estuaries [@ornolfsdottirNutrientPulsingRegulator2004; 
+@doradoUnderstandingInteractionsFreshwater2015; @ciraPhytoplanktonDynamicsLowInflow2021; @tominackVariabilityPhytoplanktonBiomass2022], 
+this study specifically used flow-adjusted freshwater derived nutrient loads to 
+parse out contributions from changes in nutrient loadings while accounting for 
+variations in load due to flow. Loading GAMs indicated no evidence of changes in 
+flow-normalized TP loads in either river, and no changes in flow-normalized NO~3~
+loads in the Navidad River. The small changes in flow-normalized NO~3~ loads in
+the Lavaca River are probably masked under most conditions by discharge from the
+Navidad River. Given the relatively little variation in flow-normalized loads,
+it can be expected that they would contribute little to the variance in downstream
+water quality.
+
+GAMs did not identify responses in DO concentration to inflows or nutrient loads.
+The seasonality term in the temporal GAM models explained a substantial amount of DO
+variation at all of the sites. Responses of estuary metabolic processes and 
+resulting DO concentrations can be quite complicated and often locally specific [@caffreyFactorsControllingNet2004]. While the lack of total nitrogen or TKN 
+loading data hinders interpretation, the large seasonal effect on DO suggests 
+physical factors play an important role and should be included in future models.
+Prior work suggests that there is some evidence the Lavaca Bay may not be 
+limited by nutrients alone, with high turbidity or nutrient processing in 
+upper portions of the Bay or intertidal river limiting production 
+[@russell_effect_2006].
 

### 2023-02-10 Schramm-Manuscript-2023/Schramm-Manuscript-2023.tex
Intro kinda done

@@
 % Simple summary
-\simplesumm{A Simple summary goes here.}
 
@@
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent
-keywords specific to the article, yet reasonably common within the
-subject discipline.).}
+\keyword{estuary; nutrient loading; water quality; Texas}
 
@@
 
-Like many estuaries globally, estuaries along the Texas Gulf coast are
-facing pressures from increasing population, increases in point source
-and non-point source pollution and alterations to freshwater inflows
-leading to increases in the occurrences and risks of algal blooms and
-eutrophication
+Like many coastal areas globally, coastal watersheds along the Texas
+Gulf coast are facing pressures from increasing population, increases in
+point source and non-point source pollution and alterations to
+freshwater flows that alter water quality in downstream estuaries
 \citep{bricker_effects_2008, kennicuttWaterQualityGulf2017, bugica_water_2020}.
-Recent studies indicate that estuary water quality dynamics in both
-agriculturally dominated and urban watersheds within Texas are
-displaying signals of conditions increasingly conducive to
-eutrophication
+Despite these increasing pressures, national scale assessments have
+classified coastal estuaries in Texas as moderate or lower for
+exhibiting eutrophic conditions \citep{bricker_effects_2008}. However, a
+suite of recent studies indicate that estuary water quality dynamics in
+both agriculturally dominated and urban watersheds within Texas are in
+fact expressing conditions that are increasingly conducive to algal
+blooms and eutrophication
 \citep{wetzWaterQualityDynamics2016, wetz_exceptionally_2017, bugica_water_2020, chinPhytoplanktonBiomassCommunity2022}.
+With identification of localized areas of estuary water quality concern
+along the Texas coast \citep{bugica_water_2020}, localized studies are
+being prioritized to better inform management actions.
+
+This project provides an assessment of nutrient loading and water
+quality responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in
+the larger Matagorda Bay system located roughly halfway between Houston,
+Texas and Corpus Christi, Texas. Lavaca Bay has faced substantial
+challenges associated with legacy contamination but general water
+quality parameters such as dissolved oxygen (DO), nutrients, and
+biological parameters have been well within state water quality
+standards. However, long-term declines of benthic fauna abundance,
+biomass, and diversity in Lavaca Bay that are linked to reductions in
+freshwater inflows and changes in estuary salinity
+\citep{beserespollackLongtermTrendsResponse2011, palmerImpactsDroughtsLow2015, montagnaAssessmentRelationshipFreshwater2020}
+and are a concern to local stakeholders. \citet{bugica_water_2020}
+identified monotonic increases in total phosphorus (TP), orthophosphate,
+total Kjeldahl nitrogen (TKN), and chlorophyll-\emph{a} at sites within
+Lavaca Bay. Although no long-term changes in DO concentrations were
+identified, the trends in nutrient concentrations are concerning due to
+the role of nitrogen as a limiting factor for primary production in many
+Texas estuaries
+\citep{gardnerNitrogenFixationDissimilatory2006, houTransformationFateNitrate2012, doradoUnderstandingInteractionsFreshwater2015, paudelRelationshipSuspendedSolids2019, wetz_exceptionally_2017}
+and the ramifications that major changes in nitrogen loadings could have
+for productivity and eutrophication in Lavaca Bay.
+
+There are ongoing efforts between local, state, and federal agencies to
+address water quality impairments in the freshwater portions of the
+Lavaca Bay watershed
+\citep{jainTechnicalSupportDocument2021, schramm_lavaca_2018, bertholdDirectMailingEducation2021}.
+However, at a statewide scale, these approaches have shown limited
+success and emphasize a need for improved efforts at assessing and
+linking management actions with downstream water quality to better
+identify and replicate effective management actions across the state
+\citep{schrammTotalMaximumDaily2022}. The identification and
+communication of changes and trends in water quality is complicated by
+the fact that trends are often non-linear and confounded by
+precipitation and runoff that hinder traditional analysis
+\citep{wazniakLinkingWaterQuality2007, lloydMethodsDetectingChange2014}.
+To provide actionable information for resource managers, water quality
+conditions must be evaluated relative to changes in natural
+environmental drivers to better understand and manage potential
+anthropogenic effects.
+
+Here something some we present methodology and results for estimating
+nutrient loading- and assessing linkages in estuary water quality
+parameters with watershed derived nutrient loads\ldots.
 
@@
 \subsection{Study Area and Data}\label{study-area-and-data}}
+
+\begin{figure}
+\begin{adjustwidth}{-\extralength}{0cm}
+\centering
+\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
+\end{adjustwidth}
+\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
+\end{figure}
 
@@
 
-\begin{figure}[H]
-\begin{adjustwidth}{-\extralength}{0cm}
-\centering
-\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
-\end{adjustwidth}
-\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
-\end{figure}
-
-Daily discharges for the Lavaca River (USGS-08164000) were obtained from
-the United States Geologic Survey (USGS) National Water Information
-System using the \emph{dataRetrieval} R package
-\citep{deciccoDataRetrievalPackagesDiscovering2022}. Gaged dailly
-discharges from Lake Texana (USGS-0816425) were provided by the Texas
-Water Development Board (TWDB) (April 21, 2022 email from R. Neupane,
-TWDB).
-
-Water quality sample data for both freshwater and estuary locations were
-o btained from the Texas Commission on Environmental Quality (TCEQ)
-Surface Water Quality Monitoring Information System. Data submitted
-through the system are required to be collected under Quality Assurance
-Project Plans and lab method procedures outlined by the TCEQ's
+\textbf{Need to add a short description of land use}
+
+Daily discharges for the Lavaca River (USGS-08164000, Figure
+\ref{fig:fig1}) were obtained from the United States Geologic Survey
+(USGS) National Water Information System using the \emph{dataRetrieval}
+R package \citep{deciccoDataRetrievalPackagesDiscovering2022}. Gaged
+daily discharges from the outlet of Lake Texana on the Navidad River
+(USGS-0816425) were provided by the Texas Water Development Board (TWDB)
+(April 21, 2022 email from R. Neupane, TWDB).
+
+Water quality sample data for the two freshwater and three estuary
+locations were obtained from the Texas Commission on Environmental
+Quality (TCEQ) Surface Water Quality Monitoring Information System. Data
+submitted through the system are required to be collected under Quality
+Assurance Project Plans and lab method procedures outlined by the TCEQ's
 procedures manual. The QAPP and procedures manuals ensure the consistent
 collection and laboratory methods are applied between samples collected
-by different entities and under different projects. For freshwater
-locations, total phosphorus (TP) and nitrate-nitrogen
-(NO\textsubscript{3}) data were downloaded (Table \ref{tab:fwsummary}).
-Unfortunately, insufficient data was available for assessment of total
-nitrogen (TN) and total Kjeldahl nitrogen (TKN) loadings, so analysis
-was restricted to TP and NO\textsubscript{3} loads. For estuary
-locations, we obtained data for TP, Nitrite+Nitrate
-(NO\emph{\textsubscript{x}}), TKN, chlorophyll-\emph{a}, and dissolved
-oxygen (Table \ref{tab:estuarysummary}).
+by different entities and under different projects. All sites had
+varying lengths of and availability of data. For freshwater locations,
+TP from January 2000 through December 2020 and nitrate-nitrogen
+(NO\textsubscript{3}) data from January 2005 through December 2020 were
+downloaded (Table \ref{tab:fwsummary}). Less than 5-years of total
+nitrogen and TKN concentration data were available at the freshwater
+sites and deemed insufficient to develop load estimation models
+(\textbf{CITE}). The three estuary sites included an upper Lavaca Bay
+site near the outlet of the Lavaca River system (TCEQ-13563), a
+mid-Lavaca Bay site (TCEQ-13383), and the lower Lavaca Bay site near the
+mouth of the Bay (TCEQ-13384). For estuary locations, we obtained data
+for TP, Nitrite+Nitrate (NO\emph{\textsubscript{x}}), TKN,
+chlorophyll-\emph{a}, and DO concentrations from January 2005 through
+December 2020 (Table \ref{tab:estuarysummary}).
 
@@
 
-Estimates of nutrient loads were developed using Generalized Additive
-Models (GAMs) relating nutrient concentration to river discharge,
-season, and time. Separate models were fit at each station for each
-parameter and used to predict nutrient concentrations for each day in
-the study period. GAMs can be specified in a functionally similar manner
-to the commonly used LOADEST \citep{cohn_validity_1992} or WRTDS
-\citep{hirsch_weighted_2010} regression models and have been shown to
-produce reliable estimates of nutrient and sediment loadings
+Estimates of NO\textsubscript{3} and TP loads at the Lavaca River
+(USGS-08164000) and the outlet of Lake Texana on the Navidad River
+(USGS-08164250) were developed using Generalized Additive Models (GAMs)
+relating nutrient concentration to river discharge, season, and time.
+Separate models were fit at each station for each parameter and used to
+predict nutrient concentrations for each day in the study period. GAMs
+can be specified in a functionally similar manner to the commonly used
+LOADEST \citep{cohn_validity_1992} or WRTDS \citep{hirsch_weighted_2010}
+regression models and have been shown to produce reliable estimates of
+nutrient and sediment loadings
 \citep{wangLoadEstimationUncertainties2011, kroonRiverLoadsSuspended2012, kuhnert_quantifying_2012, robson_prediction_2015-1, hagemannEstimatingNutrientOrganic2016, mcdowell_implications_2021, biagi_novel_2022}.
@@
 improvements in nutrient loading models by including the term.
-\citet{kuhnert_quantifying_2012} expresses MA as:
+\citet{kuhnert_quantifying_2012} expresses \emph{ma} as:
 
@@
 Based on criteria provided by \citet{moriasiHydrologicWaterQuality2015},
-GAMs ranged from ``satisfactory'' to ``very good'' based on median NSE,
-r\textsuperscript{2} and PBIAS calculated from repeated 5-fold cross
+GAMs ranged from ``satisfactory'' to ``very good'' using median NSE,
+r\textsuperscript{2} and PBIAS calculated using repeated 5-fold cross
 validation on predicted and measured nutrient loads. NO\textsubscript{3}
@@
 
-\begin{figure}[H]\begin{adjustwidth}{-\extralength}{0cm}
+\begin{figure}\begin{adjustwidth}{-\extralength}{0cm}
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 
@@
 
-\caption[Aggregated estimated annual and flow-normalized annual NO  extsubscript{3} and TP loads for USGS-08164000 and USGS-08164525]{Aggregated estimated annual and flow-normalized annual NO extsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
+\caption[Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525]{Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
 \end{figure}
 
-\begin{figure}[H]
+\begin{figure}
 
@@
 \section{Discussion}\label{discussion}}
+
+TP and NO\textsubscript{3} loadings from the Lavaca Bay watershed showed
+high inter-annual variability tied with changes in discharge. There is
+little evidence for changes in flow-normalized TP loads in either
+rivers. There is some evidence of recent decreases in flow-normalized
+NO\textsubscript{3} loads in the Lavaca River. Although there is no work
+directly correlating water quality planning and implementation efforts
+in the watershed to water quality outcomes, efforts to increase
+agricultural producer participation in the watershed have been ongoing
+since 2016
+\citep{schramm_lavaca_2018, bertholdDirectMailingEducation2021}. The
+decrease in flow-normalized NO\textsubscript{3} loads could be a
+reflection of those collective efforts but further data collection and
+research is required to support that statement.
+
+Other studies have published estimates of TP loading estimates for the
+Lavaca River (USGS-08164000) over various time periods (Table
+\ref{tab:tpcomp}). Converted to average annual yield, the estimates of
+TP loading in this study are within the ranges of previous studies
+\citep{dunnTrendsNutrientInflows1996, rebich_sources_2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019},
+with the TP estimates in \citet{dunnTrendsNutrientInflows1996} being
+notably lower. Given that none of the studies identify substantially
+sized trends in TP, it is is possible that the period used in
+\citet{dunnTrendsNutrientInflows1996} was drier on average than the
+other studies. The SPARROW models used in \citet{rebich_sources_2011}
+and \citet{wise_spatially_2019} utilize a version of LOADEST in the
+underlying load estimation procedure, so a difference due to methodology
+alone is unlikely.
 
@@
 
-Authors should discuss the results and how they can be interpreted in
-perspective of previous studies and of the working hypotheses. The
-findings and their implications should be discussed in the broadest
-context possible. Future research directions may also be highlighted.
+Cross-validation of the GAM loading models indicated that GAMs performed
+well on average at predicting daily nutrient loading values. The
+variance in scores was very high indicating subsets of values were
+problematic at characterizing functional relationships between nutrients
+and predictors. All of the water quality data for the two river sites in
+the TCEQ database were ambient water quality data, collected to be
+representative of typical conditions. This results in few data collected
+at the highest portions of the flow duration curve under which the
+majority of loadings take place. Supplementary flow-biased monitoring
+targeting storm- or high-flow conditions is recommended here to improve
+the precision of GAM predictions
+\citep{horowitzEvaluationSedimentRating2003, snelderEstimationCatchmentNutrient2017}.
+
+The non-linear temporal water quality trends identified using GAMs
+differed slightly from the trends identified by
+\citet{bugicaWaterQualityTrends2020}. This is not unexpected due to the
+different time periods, different methodology, and generally small
+slopes identified for most of the significant water quality parameters
+in \citet{bugica_water_2020}. The trend in DO and cholorophyll-\emph{a}
+concentrations are stable in comparison to other Texas estuaries that
+might be facing larger demands for freshwater diversions, higher
+population growth, and more intense agricultural production
+\citep{wetzWaterQualityDynamics2016, bugica_water_2020}. The trend of
+increasing NO\textsubscript{\emph{x}} concentration at the upper-bay
+TCEQ-13563 site and recent increases in TKN concentration at the mid-bay
+TCEQ-13383 site are concerning due to the nitrogen limitation identified
+in many Texas estuaries
+\citep{gardnerNitrogenFixationDissimilatory2006, houTransformationFateNitrate2012, doradoUnderstandingInteractionsFreshwater2015, paudelRelationshipSuspendedSolids2019, wetz_exceptionally_2017}
+and the relatively low ambient concentrations observed in Lavaca Bay.
+
+The strong positive effect of freshwater inflow on
+NO\textsubscript{\emph{x}}, TKN, and TP are suggestive of nonpoint
+watershed sources, consistent with watershed uses, and consistent with
+other studies relating freshwater inflow with nutrient concentrations in
+Lavaca Bay and other estuaries
+\citep{russell_effect_2006, caffreyHighNutrientPulses2007, peierlsNonmonotonicResponsesPhytoplankton2012, palmerImpactsDroughtsLow2015, ciraPhytoplanktonDynamicsLowInflow2021}.
+Inflow had a non-linear relationship with TKN at the two upstream sites,
+with TKN increasing as freshwater inflow transitioned from low to
+moderate levels. At higher freshwater inflows, the effect was
+attenuated, possibly indicating a flushing effect at higher freshwater
+inflow. No relationship between TKN and freshwater inflow were observed
+at TCEQ-13384 located in the lower reach of Lavaca Bay. Tidal flushing
+from Matagorda Bay might be responsible for decreasing TKN
+concentrations and limiting the effects of freshwater inflow in lower
+reaches of Lavaca Bay. \citet{russell_effect_2006} suggested the
+processing of organic loads in the upper portions of Lavaca Bay reduces
+the transport of nutrients into the lower reaches of the Bay.
+
+Freshwater inflow also had a strong positive effect on
+chlorophyll-\emph{a} at the upper- and mid-bay sites. The upper-bay
+site, TCEQ-13563, showed decreases in chlorophyll-\emph{a} at the
+highest freshwater inflow volumes. Freshwater flushing or increases in
+turbidity are associated with decreases in chlorophyll-\emph{a} in other
+estuaries
+\citep{peierlsNonmonotonicResponsesPhytoplankton2012, cloernPhytoplanktonPrimaryProduction2014}.
+No relationships between inorganic nitrogen or TP loadings with
+chlorophyll-\emph{a} were observed. Due to the lack of TKN loading
+information, no assessment between organic nitrogen loads and
+chlorophyll-\emph{a} were possible.
+
+Although other studies have identified complex relationships between
+estuary nutrient concentrations, nutrient loading and
+chlorophyll-\emph{a} concentrations in Texas estuaries
+\citep{ornolfsdottirNutrientPulsingRegulator2004, doradoUnderstandingInteractionsFreshwater2015, ciraPhytoplanktonDynamicsLowInflow2021, tominackVariabilityPhytoplanktonBiomass2022},
+this study specifically used flow-adjusted freshwater derived nutrient
+loads to parse out contributions from changes in nutrient loadings while
+accounting for variations in load due to flow. Loading GAMs indicated no
+evidence of changes in flow-normalized TP loads in either river, and no
+changes in flow-normalized NO\textsubscript{3} loads in the Navidad
+River. The small changes in flow-normalized NO\textsubscript{3} loads in
+the Lavaca River are probably masked under most conditions by discharge
+from the Navidad River. Given the relatively little variation in
+flow-normalized loads, it can be expected that they would contribute
+little to the variance in downstream water quality.
+
+GAMs did not identify responses in DO concentration to inflows or
+nutrient loads. The seasonality term in the temporal GAM models
+explained a substantial amount of DO variation at all of the sites.
+Responses of estuary metabolic processes and resulting DO concentrations
+can be quite complicated and often locally specific
+\citep{caffreyFactorsControllingNet2004}. While the lack of total
+nitrogen or TKN loading data hinders interpretation, the large seasonal
+effect on DO suggests physical factors play an important role and should
+be included in future models. Prior work suggests that there is some
+evidence the Lavaca Bay may not be limited by nutrients alone, with high
+turbidity or nutrient processing in upper portions of the Bay or
+intertidal river limiting production \citep{russell_effect_2006}.
 
@@
 
-\acknowledgments{All sources of funding of the study should be
-disclosed. Please clearly indicate grants that you have received in
-support of your research work. Clearly state if you received funds for
-covering the costs to publish in open access.}
+\acknowledgments{Thank you to Mike Wetz (Harte Institute, Texas A\&M
+Corpus Christi), Chad Kinsfather and Partick Brzozowski (Lavaca-Navidad
+River Authority), Brian Koch (Texas State Soil and Water Conservation
+Board), Bill Balboa (Matagorda Bay Foundation) and the Lavaca Bay
+Foundation for supporting development of this project and providing
+valuable feedback. Thank you to Ram Neupane (Texas Water Development
+Board) for providing data.}
 
 \conflictsofinterest{The author declares no conflict of interest. The
-founding sponsors had no role in the design of the study; in the
+project sponsors had no role in the design of the study; in the
 collection, analyses, or interpretation of data; in the writing of the
-manuscript, an in the decision to publish the results.}
+manuscript, or in the decision to publish the results.}
 

### 2023-02-11 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
spell checks

@@
 abstract: |
-  A single paragraph of about 200 words maximum. For research articles, 
-  abstracts should give a pertinent overview of the work. We strongly encourage
-  authors to use the following style of structured abstracts, but without 
-  headings: 1) Background: Place the question addressed in a broad context and
-  highlight the purpose of the study; 2) Methods: Describe briefly the main
-  methods or treatments applied; 3) Results: Summarize the article's main 
-  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
-  The abstract should be an objective representation of the article, it must not 
-  contain results which are not presented and substantiated in the main text and 
-  should not exaggerate the main conclusions.
+  Lavaca Bay is a small secondary embayment on the Texas coast that is
+  displaying early signals of water quality degradation. This
+  study uses Generalized Additive Models to assesses both watershed nutrient loading 
+  and responses in estuary water quality to nutrient loading and streamflow.
+  While delivered nutrient loads varied greatly, flow-normalized nutrient loads indicated
+  limited evidence of changes in nutrient runoff in the watershed since 2005. 
+  As a consequence, there wasn't strong evidence for an effect of nutrient loading
+  (controlled for flow) on dissolved oxygen or chlorophyll-*a* concentrations.
+  There was evidence total phosphorus (TP) loading had a strong positive
+  effect on TP concentration in the Bay, particularly at sites closest to the
+  river mouth. Seasonally-adjusted streamflow had strong effects on nutrient
+  concentration and chlorophyll-*a* concentrations at most sites. Although
+  temporal trends did not identify significant estuary water quality degredation 
+  in dissolved oxygen or chlorophyll-*a* concentrations, site specific trends
+  in both organic and inorganic nitrogen are concerning for the long-term
+  health of the estuary. This study provides a baseline assessment of the
+  linkages between watershed nutrient loading and water quality responses in 
+  Lavaca Bay. Some of the data gaps highlighted in the study are being addressed
+  in ongoing projects to provide improved information for resource managers.
 # back matter
@@
 acknowledgement: |
-  Thank you to Mike Wetz (Harte Institute, Texas A&M Corpus Christi),
-  Chad Kinsfather and Partick Brzozowski (Lavaca-Navidad River Authority),
-  Brian Koch (Texas State Soil and Water Conservation Board), Bill Balboa 
-  (Matagorda Bay Foundation) and the Lavaca Bay Foundation for supporting 
+  The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
+  Texas A&M Corpus Christi), Chad Kinsfather and Partick Brzozowski 
+  (Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water 
+  Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason Pinchbeck 
+  (Texas General Land Office) and the Lavaca Bay Foundation for supporting 
   development of this project and providing valuable feedback.
-  Thank you to Ram Neupane (Texas Water Development Board) for providing data.
 funding: |
-  This paper was funded by financial assistance provided
-  by the Coastal Zone Management Act of 1972, as amended, administered by the
-  National Oceanic and Atmospheric Administration (NOAA), Office for Coastal
-  Management, pursuant to NOAA Award No. NA21NOS4190136. The views
-  expressed herein are those of the author(s) and do not necessarily 
-  reflect the views of NOAA, the U.S. Department of Commerce, or any of their 
-  subagencies.
+  This project was funded in part by a Texas Coastal Management Program grant approved by the Texas Land
+  Commissioner, providing financial assistance under the Coastal Zone Management Act of 1972, as amended,
+  awarded by the National Oceanic and Atmospheric Administration (NOAA), Office for Coastal Management,
+  pursuant to NOAA Award No. NA21NOS4190136. The views expressed herein are those of the author(s) and do not
+  necessarily reflect the views of NOAA, the U.S. Department of Commerce, or any of their subagencies
 dataavailability: |
@@
 are facing pressures from increasing population, increases in point source 
-and non-point source pollution and alterations to freshwater flows that alter
+and non-point source pollution and alterations to freshwater flows that affect
 water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. Despite these increasing pressures, national scale 
@@
 water quality standards. However, long-term declines of benthic 
-fauna abundance, biomass, and diversity in Lavaca Bay that are 
+fauna abundance, biomass, and diversity in Lavaca Bay primarily
 linked to reductions in freshwater inflows and changes in estuary salinity 
 [@beserespollackLongtermTrendsResponse2011; @palmerImpactsDroughtsLow2015;
-@montagnaAssessmentRelationshipFreshwater2020] and are a concern to local 
+@montagnaAssessmentRelationshipFreshwater2020] are a concern to local 
 stakeholders. @bugica_water_2020 identified monotonic increases in total 
 phosphorus (TP), orthophosphate, total Kjeldahl nitrogen (TKN), and
-chlorophyll-*a* at sites within Lavaca Bay. Although no long-term changes in 
-DO concentrations were identified, the trends in nutrient concentrations are 
+chlorophyll-*a* at sites within Lavaca Bay. Although long-term changes in 
+DO concentrations were not identified, the trends in nutrient concentrations are 
 concerning due to the role of nitrogen as a limiting factor for primary 
 production in many Texas estuaries [@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; @paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the
-ramifications that major changes in nitrogen loadings could have for 
+ramifications that changes in nitrogen loadings could have for 
 productivity and eutrophication in Lavaca Bay.
@@
 better understand  and manage potential anthropogenic effects.
-
-Here something some we present methodology and results for estimating nutrient
-loading- and assessing linkages in estuary water quality parameters with watershed derived
-nutrient loads....
+This study utilizes semiparemetric methods to develop estimates of delivered 
+and flow-normalized nutrient loads and assess changes in loads delivered to 
+Lavaca Bay. The study also assesses the response of water quality parameters in
+Lavaca Bay over time and in response to freshwater inflow controlled
+for seasonality and to watershed nutrient loads that are controlled for
+environmentally driven variation.
 
@@
 
+```{=latex}
 \begin{figure}
-\begin{adjustwidth}{-\extralength}{0cm}
-\centering
-\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
-\end{adjustwidth}
-\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
+\includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
+\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig:fig1}}
 \end{figure}  
-
-Lavaca Bay is a secondary bay in the Matagorda Bay system located on the 
-Texas Gulf coast, roughly halfway between the cities of Houston and Corpus 
-Christi (Figure  \ref{fig:fig1}). Lavaca Bay is 190 km^2^ with the majority
-of freshwater inflow provided by the Lavaca and Navidad River systems. The 
-Garcitas-Arenosa, Placedo Creek, and Cox Bay watersheds provide additional 
-freshwater inflows. The entire watershed land area for Lavaca Bay is 8,149 km^2^. 
-The Lavaca and Navidad River watersheds are a combined 5,966 km^2^, or 
-approximately 73% of the entire Lavaca Bay watershed area. Discharge from the 
-Navidad River is regulated by Lake Texana which has been in operation since 1980.
-Lake Texana provides 170,000 acre-feet of water storage and discharges into the 
-tidal section of the Navidad River which ultimately joins the tidal section of 
-the Lavaca River 15 km upstream of the confluence with the Bay.
-
-**Need to add a short description of land use**
+```
+
+Lavaca Bay is 190 km^2^ with the majority of freshwater inflow provided by the 
+Lavaca and Navidad River systems (Figure \ref{fig:fig1}). The Garcitas-Arenosa, 
+Placedo Creek, and Cox Bay watersheds provide additional freshwater inflows. 
+The entire watershed land area is 8,149 km^2^ and primarily rural. Watershed 
+land cover and land use is 50% grazed pasture and rangeland, 20% cultivated 
+cropland (primarily rows crops such as corn, cotton, and sorghum), and 5% 
+suburban/urban. Pasture and rangeland is concentrated in the Lavaca River 
+watershed, while cultivated crops are generally located along the eastern 
+tributaries of the Navidad river. The Lavaca and Navidad River watersheds are a 
+combined 5,966 km^2^, or approximately 73% of the entire Lavaca Bay watershed 
+area. Discharge from the Navidad River is regulated by Lake Texana which has 
+been in operation since 1980. Lake Texana provides 0.210 km^3^ of water 
+storage and discharges into the tidal section of the Navidad River which 
+ultimately joins the tidal section of the Lavaca River 15 km upstream of the 
+confluence with the Lavaca Bay.
 
@@
 nitrogen and TKN concentration data were available at the freshwater sites and
-deemed insufficient to develop load estimation models (**CITE**). 
+deemed insufficient to develop load estimation models 
+[@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017]. 
 The three estuary sites included an upper Lavaca Bay site near the outlet of the
@@
 Estimates of NO~3~ and TP loads at the Lavaca River (USGS-08164000) and the 
-outlet of Lake Texana on the Navidad River (USGS-08164250) were developed using 
+outlet of Lake Texana on the Navidad River (USGS-08164525) were developed using 
 Generalized Additive Models (GAMs) relating nutrient concentration to river 
@@
 incorporate different model terms, (2) the ability to incorporate non-linear
-smooth function without explicit apriori knowledge of the expect shape, and (3)
+smooth function without explicit a priori knowledge of the expect shape, and (3)
 the ability to specify a link function that relates the expected value of the
-response to the linear predictors and allows use to avoid data transformations
-as much as possible.
+response to the linear predictors and avoid data transformations.
 
@@
 types of smooth functions with automatic smoothness selection [@wood_fast_2011]. 
-The general form of the model relating NO~3~ and TP concentration to streamflow,
-season, and time was:
+The general form of the model related NO~3~ or TP concentration to a long term
+tend, season, streamflow, and two different antecedent discharge terms:
 
@@
 
-where $\mu$ is the conditional expected NO~3~-N or TP concentration, *g()* is the 
+where $\mu$ is the conditional expected NO~3~ or TP concentration, *g()* is the 
 log-link, $\alpha$ is the intercept, *f~n~()* are smoothing functions. *y* is the 
@@
 to incorporate the influence of prior streamflow events on concentration at the 
-current time period. @wangLoadEstimationUncertainties2011, 
-@kuhnert_quantifying_2012 and @zhang_improving_2017 refer to this as averaged or 
-smoothed discounted flow and demonstrated improvements in nutrient loading 
-models by including the term. @kuhnert_quantifying_2012 expresses *ma* as:
+current time period [@wangLoadEstimationUncertainties2011;  
+@kuhnert_quantifying_2012; @zhang_improving_2017], and 
+expressed in @kuhnert_quantifying_2012 as:
 
@@
 where *x* are the averages of log-transformed streamflow over the antecedent 
-period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3 models 
+period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3~ models 
 and *stfa* in TP models based on results from @zhang_improving_2017 
@@
 Daily loads were estimated as the predicted concentration multiplied by the 
-daily streamflow. For the Lake Texana site, model terms were slightly modified
-because daily loads are a function of dam releases and nutrient concentration,
-but concentration will be a function of lake inflows and or other lake processes.
- *Q*, *ma*, and *fa* terms were calculated based on total gaged inflow from
-the 4 major tributaries to the lake and laily loads at the dam were calculated 
+daily streamflow. For the Navidad River (USGS-08164525) site, daily loads at the 
+dam were calculated 
 from the discrete daily concentration at the discharge point of the lake and 
@@
 using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
-and average Nash-Suttclifee Efficiency (NSE), r^2^ and percent bias (PBIAS) 
+and average Nash-Sutcliffe Efficiency (NSE), r^2^ and percent bias (PBIAS) 
 metrics across folds were calculated for each model.
@@
 with a cyclic cubic regression spline, *f~3~(Q)* is mean daily inflow (the 
-combined measurements from Lavaca River and Lake Texana) and *f~4~(Load)* is 
+combined measurements from Lavaca River and Navidad River) and *f~4~(Load)* is 
 the total NO~3~ or TP watershed load. The set of models specified for each water
-quality response are in Table \ref{rab:estgammodels}.
+quality response are in Table \ref{tab:estgammodels}.
 
@@
 both streamflow and nutrient loads were prepossessed to account for season
-and flow. Instead of using raw freshwater inflow and nutrient loading values, 
-these values were replaced by seasonally adjusted inflow and flow-adjusted 
-nutrient loads by fitting a GAM relating season (day of year) to log transformed
-daily freshwater inflow values:
+and flow. Freshwater inflow and nutrient loads were replaced by 
+seasonally adjusted log transformed inflow and flow-adjusted 
+log transformed nutrient loads obtained by fitting a GAM relating season 
+(day of year) to log transformed daily freshwater inflow values:
 
@@
 used as *Q* and *Load* in Equation \ref{eq:6} and Equation \ref{eq:7}.
+
+<!---
+Need to add model selection criteria!
+--->
 
@@
 USGS-08164000 showed a much larger variance in values compared to USGS-08164525
-(Figure \ref{fig:fig2}).In addition to higher average NSE and r^2^ values, 
-GAMs had smaller variance in metric values for TP compared to NO~3~.
+(Figure \ref{fig:fig2}). TP GAMS had higher average NSE and r^2^ values and 
+lower variance in metric values compared to NO~3~.
 
@@
   scale_fill_carto_d(palette = "Vivid", direction = -1) +
-  labs(x = "", y = "Predicted Annual<br>NO<sub>3</sub>-N Load [kg]") +
+  labs(x = "", y = "Predicted Annual<br>NO<sub>3</sub> Load [kg]") +
   theme_TWRI_print() +
@@
                       x = flw_res,
-                      xlab = "log(Inflow) Residuals [cfs]",
+                      xlab = "Residuals log(Inflow) [cfs]",
                       ylab =  "TP Smooth<br>Estimate [mg/L]",
@@
                       x = flw_res,
-                      xlab = "log(Inflow) Residuals [cfs]",
+                      xlab = "Residuals log(Inflow) [cfs]",
                       ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]",
@@
                       x = flw_res,
-                      xlab = "log(Inflow) Residuals [cfs]",
+                      xlab = "Residuals log(Inflow) [cfs]",
                       ylab = "Chlorophyll-*a* Smooth<br>Estimate [µg/L]",
@@
                       x = flw_res,
-                      xlab = "log(Inflow) Residuals [cfs]",
+                      xlab = "Residuals log(Inflow) [cfs]",
                       ylab = "TKN Smooth<br>Estimate [mg/L]",
@@
                       x = flw_res,
-                      xlab = "log(Inflow) Residuals [cfs]",
+                      xlab = "Residuals log(Inflow) [cfs]",
                       ylab = "DO Smooth<br>Estimate [mg/L]",
@@
 
-GAMs show that increases in freshwater inflow result in nearly linear increases 
-in TP and NO~*x*~ at all three sites (Figure \ref{fig:6}). At the upper-bay
+GAMs showed increases in freshwater inflow resulted in nearly linear increases 
+in TP and NO~*x*~ concentration at all three sites (Figure \ref{fig:fig6}). At the upper-bay
 TCEQ-13563 site, GAMs showed increases in freshwater inflow initially 
@@
 potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
-nearly linear increased in chlorophyll-*a* concentration with increases in
+nearly linear increased in chlorophyll-*a* concentration in response to increases
 freshwater inflow. Freshwater flow did not have significant
@@
 
-Increases in TP loads resulted in nearly linear increases of TP concentration at
+Increased TP loads resulted in nearly linear increases of TP concentration at
 the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively (Figure \ref{fig:fig7}).
 The relative effect size appeared to much smaller than the effect of freshwater
-inflow alone. Increases in NO~3~ loads only showed an effect at the 
+inflow alone. Increased NO~3~ loads only showed an effect at the 
 lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
@@
                       x = TP_resid,
-                      xlab = "TP Load Residuals [kg]",
+                      xlab = "Residuals log(TP) Load [kg]",
                       ylab = "TP [mg/L]",
@@
                       x = NO3_resid,
-                      xlab = "NO<sub>3</sub> Load Residuals [kg]",
+                      xlab = "Residuals log(NO<sub>3</sub>) Load [kg]",
                       ylab = "NO*<sub>x</sub>* [mg/L]",
@@
 The non-linear temporal water quality trends identified using GAMs differed
-slightly from the trends identified by @bugicaWaterQualityTrends2020. This is
+slightly from the trends identified by @bugica_water_2020. This is
 not unexpected due to the different time periods, different methodology, and
 generally small slopes identified for most of the significant water quality
-parameters in @bugica_water_2020. The trend in DO and 
+parameters in prior work. The trend in DO and 
 cholorophyll-*a* concentrations are stable in comparison to other Texas
@@
 physical factors play an important role and should be included in future models.
-Prior work suggests that there is some evidence the Lavaca Bay may not be 
-limited by nutrients alone, with high turbidity or nutrient processing in 
-upper portions of the Bay or intertidal river limiting production 
-[@russell_effect_2006].
+Prior work suggests that Lavaca Bay may not be limited by nutrients alone, 
+with high turbidity or nutrient processing in upper portions of the Bay or 
+intertidal river limiting production [@russell_effect_2006].
 
@@
 
-This section is not mandatory, but can be added to the manuscript if the
-discussion is unusually long or complex.
+GAM models appear to provide reliable estimates of nutrient loads in the Lavaca
+Bay watershed. However, additional flow-biased data collection efforts would
+decrease the prediction variance and probably improve accuracy at critical 
+high flow events. Ongoing projects will fill data gaps for total nitrogen and 
+TKN loading. This study, consistent with others along the Texas coast, found
+strong effects of freshwater flow on nutrient and chlorophyll-*a* 
+concentrations. DO concentrations, dominated by seasonal effects, did not show
+strong direct responses to freshwater flow. Small variance in flow-adjusted
+nutrient loads indicates that (1) there have been limited changes in non point
+sources of nutrients and (2) that those changes have limited effect on 
+water quality in Lavaca Bay to date. Although the study did not identify
+strong responses to changes in nutrient loading, this does provide a baseline
+assessment for future water quality management activities in the watershed.

### 2023-02-11 Schramm-Manuscript-2023/Schramm-Manuscript-2023.tex
spell checks

@@
 % Abstract (Do not insert blank lines, i.e. \\)
-\abstract{A single paragraph of about 200 words maximum. For research
-articles, abstracts should give a pertinent overview of the work. We
-strongly encourage authors to use the following style of structured
-abstracts, but without headings: 1) Background: Place the question
-addressed in a broad context and highlight the purpose of the study; 2)
-Methods: Describe briefly the main methods or treatments applied; 3)
-Results: Summarize the article's main findings; and 4) Conclusion:
-Indicate the main conclusions or interpretations. The abstract should be
-an objective representation of the article, it must not contain results
-which are not presented and substantiated in the main text and should
-not exaggerate the main conclusions.}
+\abstract{Lavaca Bay is a small secondary embayment on the Texas coast
+that is displaying early signals of water quality degradation. This
+study uses Generalized Additive Models to assesses both watershed
+nutrient loading and responses in estuary water quality to nutrient
+loading and streamflow. While delivered nutrient loads varied greatly,
+flow-normalized nutrient loads indicated limited evidence of changes in
+nutrient runoff in the watershed since 2005. As a consequence, there
+wasn't strong evidence for an effect of nutrient loading (controlled for
+flow) on dissolved oxygen or chlorophyll-\emph{a} concentrations. There
+was evidence total phosphorus (TP) loading had a strong positive effect
+on TP concentration in the Bay, particularly at sites closest to the
+river mouth. Seasonally-adjusted streamflow had strong effects on
+nutrient concentration and chlorophyll-\emph{a} concentrations at most
+sites. Although temporal trends did not identify significant estuary
+water quality degredation in dissolved oxygen or chlorophyll-\emph{a}
+concentrations, site specific trends in both organic and inorganic
+nitrogen are concerning for the long-term health of the estuary. This
+study provides a baseline assessment of the linkages between watershed
+nutrient loading and water quality responses in Lavaca Bay. Some of the
+data gaps highlighted in the study are being addressed in ongoing
+projects to provide improved information for resource managers.}
 
@@
 point source and non-point source pollution and alterations to
-freshwater flows that alter water quality in downstream estuaries
+freshwater flows that affect water quality in downstream estuaries
 \citep{bricker_effects_2008, kennicuttWaterQualityGulf2017, bugica_water_2020}.
@@
 standards. However, long-term declines of benthic fauna abundance,
-biomass, and diversity in Lavaca Bay that are linked to reductions in
+biomass, and diversity in Lavaca Bay primarily linked to reductions in
 freshwater inflows and changes in estuary salinity
 \citep{beserespollackLongtermTrendsResponse2011, palmerImpactsDroughtsLow2015, montagnaAssessmentRelationshipFreshwater2020}
-and are a concern to local stakeholders. \citet{bugica_water_2020}
+are a concern to local stakeholders. \citet{bugica_water_2020}
 identified monotonic increases in total phosphorus (TP), orthophosphate,
 total Kjeldahl nitrogen (TKN), and chlorophyll-\emph{a} at sites within
-Lavaca Bay. Although no long-term changes in DO concentrations were
+Lavaca Bay. Although long-term changes in DO concentrations were not
 identified, the trends in nutrient concentrations are concerning due to
@@
 \citep{gardnerNitrogenFixationDissimilatory2006, houTransformationFateNitrate2012, doradoUnderstandingInteractionsFreshwater2015, paudelRelationshipSuspendedSolids2019, wetz_exceptionally_2017}
-and the ramifications that major changes in nitrogen loadings could have
-for productivity and eutrophication in Lavaca Bay.
+and the ramifications that changes in nitrogen loadings could have for
+productivity and eutrophication in Lavaca Bay.
 
@@
 environmental drivers to better understand and manage potential
-anthropogenic effects.
-
-Here something some we present methodology and results for estimating
-nutrient loading- and assessing linkages in estuary water quality
-parameters with watershed derived nutrient loads\ldots.
+anthropogenic effects. This study utilizes semiparemetric methods to
+develop estimates of delivered and flow-normalized nutrient loads and
+assess changes in loads delivered to Lavaca Bay. The study also assesses
+the response of water quality parameters in Lavaca Bay over time and in
+response to freshwater inflow controlled for seasonality and to
+watershed nutrient loads that are controlled for environmentally driven
+variation.
 
@@
 \begin{figure}
-\begin{adjustwidth}{-\extralength}{0cm}
-\centering
-\includegraphics[width=15.5cm]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
-\end{adjustwidth}
-\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig1}}
+\includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
+\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig:fig1}}
 \end{figure}
 
-Lavaca Bay is a secondary bay in the Matagorda Bay system located on the
-Texas Gulf coast, roughly halfway between the cities of Houston and
-Corpus Christi (Figure \ref{fig:fig1}). Lavaca Bay is 190
-km\textsuperscript{2} with the majority of freshwater inflow provided by
-the Lavaca and Navidad River systems. The Garcitas-Arenosa, Placedo
-Creek, and Cox Bay watersheds provide additional freshwater inflows. The
-entire watershed land area for Lavaca Bay is 8,149
-km\textsuperscript{2}. The Lavaca and Navidad River watersheds are a
-combined 5,966 km\textsuperscript{2}, or approximately 73\% of the
-entire Lavaca Bay watershed area. Discharge from the Navidad River is
-regulated by Lake Texana which has been in operation since 1980. Lake
-Texana provides 170,000 acre-feet of water storage and discharges into
-the tidal section of the Navidad River which ultimately joins the tidal
-section of the Lavaca River 15 km upstream of the confluence with the
-Bay.
-
-\textbf{Need to add a short description of land use}
+Lavaca Bay is 190 km\textsuperscript{2} with the majority of freshwater
+inflow provided by the Lavaca and Navidad River systems (Figure
+\ref{fig:fig1}). The Garcitas-Arenosa, Placedo Creek, and Cox Bay
+watersheds provide additional freshwater inflows. The entire watershed
+land area is 8,149 km\textsuperscript{2} and primarily rural. Watershed
+land cover and land use is 50\% grazed pasture and rangeland, 20\%
+cultivated cropland (primarily rows crops such as corn, cotton, and
+sorghum), and 5\% suburban/urban. Pasture and rangeland is concentrated
+in the Lavaca River watershed, while cultivated crops are generally
+located along the eastern tributaries of the Navidad river. The Lavaca
+and Navidad River watersheds are a combined 5,966 km\textsuperscript{2},
+or approximately 73\% of the entire Lavaca Bay watershed area. Discharge
+from the Navidad River is regulated by Lake Texana which has been in
+operation since 1980. Lake Texana provides 0.210 km\textsuperscript{3}
+of water storage and discharges into the tidal section of the Navidad
+River which ultimately joins the tidal section of the Lavaca River 15 km
+upstream of the confluence with the Lavaca Bay.
 
@@
 sites and deemed insufficient to develop load estimation models
-(\textbf{CITE}). The three estuary sites included an upper Lavaca Bay
-site near the outlet of the Lavaca River system (TCEQ-13563), a
-mid-Lavaca Bay site (TCEQ-13383), and the lower Lavaca Bay site near the
-mouth of the Bay (TCEQ-13384). For estuary locations, we obtained data
-for TP, Nitrite+Nitrate (NO\emph{\textsubscript{x}}), TKN,
-chlorophyll-\emph{a}, and DO concentrations from January 2005 through
-December 2020 (Table \ref{tab:estuarysummary}).
+\citep{horowitzEvaluationSedimentRating2003, snelderEstimationCatchmentNutrient2017}.
+The three estuary sites included an upper Lavaca Bay site near the
+outlet of the Lavaca River system (TCEQ-13563), a mid-Lavaca Bay site
+(TCEQ-13383), and the lower Lavaca Bay site near the mouth of the Bay
+(TCEQ-13384). For estuary locations, we obtained data for TP,
+Nitrite+Nitrate (NO\emph{\textsubscript{x}}), TKN, chlorophyll-\emph{a},
+and DO concentrations from January 2005 through December 2020 (Table
+\ref{tab:estuarysummary}).
 
@@
 (USGS-08164000) and the outlet of Lake Texana on the Navidad River
-(USGS-08164250) were developed using Generalized Additive Models (GAMs)
+(USGS-08164525) were developed using Generalized Additive Models (GAMs)
 relating nutrient concentration to river discharge, season, and time.
@@
 and incorporate different model terms, (2) the ability to incorporate
-non-linear smooth function without explicit apriori knowledge of the
+non-linear smooth function without explicit a priori knowledge of the
 expect shape, and (3) the ability to specify a link function that
 relates the expected value of the response to the linear predictors and
-allows use to avoid data transformations as much as possible.
+avoid data transformations.
 
@@
 multiple types of smooth functions with automatic smoothness selection
-\citep{wood_fast_2011}. The general form of the model relating
-NO\textsubscript{3} and TP concentration to streamflow, season, and time
-was:
+\citep{wood_fast_2011}. The general form of the model related
+NO\textsubscript{3} or TP concentration to a long term tend, season,
+streamflow, and two different antecedent discharge terms:
 
@@
 
-where \(\mu\) is the conditional expected NO\textsubscript{3}-N or TP
+where \(\mu\) is the conditional expected NO\textsubscript{3} or TP
 concentration, \emph{g()} is the log-link, \(\alpha\) is the intercept,
@@
 that attempts to incorporate the influence of prior streamflow events on
-concentration at the current time period.
-\citet{wangLoadEstimationUncertainties2011},
-\citet{kuhnert_quantifying_2012} and \citet{zhang_improving_2017} refer
-to this as averaged or smoothed discounted flow and demonstrated
-improvements in nutrient loading models by including the term.
-\citet{kuhnert_quantifying_2012} expresses \emph{ma} as:
+concentration at the current time period
+\citep{wangLoadEstimationUncertainties2011, kuhnert_quantifying_2012, zhang_improving_2017},
+and expressed in \citet{kuhnert_quantifying_2012} as:
 
@@
 antecedent period (\emph{1-year}, \emph{1-month}, etc.) for time
-\emph{t}. We used \emph{ltfa} in NO\textasciitilde3 models and
+\emph{t}. We used \emph{ltfa} in NO\textsubscript{3} models and
 \emph{stfa} in TP models based on results from
@@
 Daily loads were estimated as the predicted concentration multiplied by
-the daily streamflow. For the Lake Texana site, model terms were
-slightly modified because daily loads are a function of dam releases and
-nutrient concentration, but concentration will be a function of lake
-inflows and or other lake processes. \emph{Q}, \emph{ma}, and \emph{fa}
-terms were calculated based on total gaged inflow from the 4 major
-tributaries to the lake and laily loads at the dam were calculated from
-the discrete daily concentration at the discharge point of the lake and
-corresponding reported daily discharge from the dam. Flow-normalized
-loads were estimated similar to WRTDS by setting flow-based covariates
-on each day of the year equal to each of the historical values for that
-day of the year over the study period \citep{hirsch_weighted_2010}. The
-flow-normalized estimate was calculated as the mean of all the
-predictions for each day considering all possible flow values. Standard
-deviations and credible intervals were obtained by drawing samples from
-the multivariate normal posterior distribution of the fitted GAM
+the daily streamflow. For the Navidad River (USGS-08164525) site, daily
+loads at the dam were calculated from the discrete daily concentration
+at the discharge point of the lake and corresponding reported daily
+discharge from the dam. Flow-normalized loads were estimated similar to
+WRTDS by setting flow-based covariates on each day of the year equal to
+each of the historical values for that day of the year over the study
+period \citep{hirsch_weighted_2010}. The flow-normalized estimate was
+calculated as the mean of all the predictions for each day considering
+all possible flow values. Standard deviations and credible intervals
+were obtained by drawing samples from the multivariate normal posterior
+distribution of the fitted GAM
 \citep{woodConfidenceIntervalsGeneralized2006, marraCoveragePropertiesConfidence2012, mcdowell_implications_2021}.
@@
 performance was evaluated using repeated 5-fold cross validation
-\citep{burmanComparativeStudyOrdinary1989} and average Nash-Suttclifee
+\citep{burmanComparativeStudyOrdinary1989} and average Nash-Sutcliffe
 Efficiency (NSE), r\textsuperscript{2} and percent bias (PBIAS) metrics
@@
 \emph{f\textsubscript{3}(Q)} is mean daily inflow (the combined
-measurements from Lavaca River and Lake Texana) and
+measurements from Lavaca River and Navidad River) and
 \emph{f\textsubscript{4}(Load)} is the total NO\textsubscript{3} or TP
 watershed load. The set of models specified for each water quality
-response are in Table \ref{rab:estgammodels}.
+response are in Table \ref{tab:estgammodels}.
 
@@
 \citet{murphyNutrientImprovementsChesapeake2022}, both streamflow and
-nutrient loads were prepossessed to account for season and flow. Instead
-of using raw freshwater inflow and nutrient loading values, these values
-were replaced by seasonally adjusted inflow and flow-adjusted nutrient
-loads by fitting a GAM relating season (day of year) to log transformed
-daily freshwater inflow values:
+nutrient loads were prepossessed to account for season and flow.
+Freshwater inflow and nutrient loads were replaced by seasonally
+adjusted log transformed inflow and flow-adjusted log transformed
+nutrient loads obtained by fitting a GAM relating season (day of year)
+to log transformed daily freshwater inflow values:
 
@@
 NO\textsubscript{3} loads where USGS-08164000 showed a much larger
-variance in values compared to USGS-08164525 (Figure \ref{fig:fig2}).In
-addition to higher average NSE and r\textsuperscript{2} values, GAMs had
-smaller variance in metric values for TP compared to
-NO\textsubscript{3}.
+variance in values compared to USGS-08164525 (Figure \ref{fig:fig2}). TP
+GAMS had higher average NSE and r\textsuperscript{2} values and lower
+variance in metric values compared to NO\textsubscript{3}.
 
@@
 
-GAMs show that increases in freshwater inflow result in nearly linear
-increases in TP and NO\textsubscript{\emph{x}} at all three sites
-(Figure \ref{fig:6}). At the upper-bay TCEQ-13563 site, GAMs showed
-increases in freshwater inflow initially increased chlorophyll-\emph{a}
-and DO concentration but concentrations leveled and potentially
-decreased at higher flows. The mid-bay TCEQ-13383 site showed a nearly
-linear increased in chlorophyll-\emph{a} concentration with increases in
-freshwater inflow. Freshwater flow did not have significant effects on
-chlorophyll-\emph{a}, TKN, or DO at the lower-bay TCEQ-13384 site.
-
-Increases in TP loads resulted in nearly linear increases of TP
+GAMs showed increases in freshwater inflow resulted in nearly linear
+increases in TP and NO\textsubscript{\emph{x}} concentration at all
+three sites (Figure \ref{fig:fig6}). At the upper-bay TCEQ-13563 site,
+GAMs showed increases in freshwater inflow initially increased
+chlorophyll-\emph{a} and DO concentration but concentrations leveled and
+potentially decreased at higher flows. The mid-bay TCEQ-13383 site
+showed a nearly linear increased in chlorophyll-\emph{a} concentration
+in response to increases freshwater inflow. Freshwater flow did not have
+significant effects on chlorophyll-\emph{a}, TKN, or DO at the lower-bay
+TCEQ-13384 site.
+
+Increased TP loads resulted in nearly linear increases of TP
 concentration at the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383
 respectively (Figure \ref{fig:fig7}). The relative effect size appeared
-to much smaller than the effect of freshwater inflow alone. Increases in
+to much smaller than the effect of freshwater inflow alone. Increased
 NO\textsubscript{3} loads only showed an effect at the lower-bay
@@
 differed slightly from the trends identified by
-\citet{bugicaWaterQualityTrends2020}. This is not unexpected due to the
-different time periods, different methodology, and generally small
-slopes identified for most of the significant water quality parameters
-in \citet{bugica_water_2020}. The trend in DO and cholorophyll-\emph{a}
-concentrations are stable in comparison to other Texas estuaries that
-might be facing larger demands for freshwater diversions, higher
-population growth, and more intense agricultural production
+\citet{bugica_water_2020}. This is not unexpected due to the different
+time periods, different methodology, and generally small slopes
+identified for most of the significant water quality parameters in prior
+work. The trend in DO and cholorophyll-\emph{a} concentrations are
+stable in comparison to other Texas estuaries that might be facing
+larger demands for freshwater diversions, higher population growth, and
+more intense agricultural production
 \citep{wetzWaterQualityDynamics2016, bugica_water_2020}. The trend of
@@
 effect on DO suggests physical factors play an important role and should
-be included in future models. Prior work suggests that there is some
-evidence the Lavaca Bay may not be limited by nutrients alone, with high
-turbidity or nutrient processing in upper portions of the Bay or
-intertidal river limiting production \citep{russell_effect_2006}.
+be included in future models. Prior work suggests that Lavaca Bay may
+not be limited by nutrients alone, with high turbidity or nutrient
+processing in upper portions of the Bay or intertidal river limiting
+production \citep{russell_effect_2006}.
 
@@
 
-This section is not mandatory, but can be added to the manuscript if the
-discussion is unusually long or complex.
+GAM models appear to provide reliable estimates of nutrient loads in the
+Lavaca Bay watershed. However, additional flow-biased data collection
+efforts would decrease the prediction variance and probably improve
+accuracy at critical high flow events. Ongoing projects will fill data
+gaps for total nitrogen and TKN loading. This study, consistent with
+others along the Texas coast, found strong effects of freshwater flow on
+nutrient and chlorophyll-\emph{a} concentrations. DO concentrations,
+dominated by seasonal effects, did not show strong direct responses to
+freshwater flow. Small variance in flow-adjusted nutrient loads
+indicates that (1) there have been limited changes in non point sources
+of nutrients and (2) that those changes have limited effect on water
+quality in Lavaca Bay to date. Although the study did not identify
+strong responses to changes in nutrient loading, this does provide a
+baseline assessment for future water quality management activities in
+the watershed.
 
@@
 
-\funding{This paper was funded by financial assistance provided by the
-Coastal Zone Management Act of 1972, as amended, administered by the
-National Oceanic and Atmospheric Administration (NOAA), Office for
-Coastal Management, pursuant to NOAA Award No.~NA21NOS4190136. The views
-expressed herein are those of the author(s) and do not necessarily
-reflect the views of NOAA, the U.S. Department of Commerce, or any of
-their subagencies.}
+\funding{This project was funded in part by a Texas Coastal Management
+Program grant approved by the Texas Land Commissioner, providing
+financial assistance under the Coastal Zone Management Act of 1972, as
+amended, awarded by the National Oceanic and Atmospheric Administration
+(NOAA), Office for Coastal Management, pursuant to NOAA Award
+No.~NA21NOS4190136. The views expressed herein are those of the
+author(s) and do not necessarily reflect the views of NOAA, the U.S.
+Department of Commerce, or any of their subagencies}
 
@@
 
-\acknowledgments{Thank you to Mike Wetz (Harte Institute, Texas A\&M
-Corpus Christi), Chad Kinsfather and Partick Brzozowski (Lavaca-Navidad
-River Authority), Brian Koch (Texas State Soil and Water Conservation
-Board), Bill Balboa (Matagorda Bay Foundation) and the Lavaca Bay
-Foundation for supporting development of this project and providing
-valuable feedback. Thank you to Ram Neupane (Texas Water Development
-Board) for providing data.}
+\acknowledgments{The author extends thanks to Dr.~Mike Wetz (Harte
+Research Institute, Texas A\&M Corpus Christi), Chad Kinsfather and
+Partick Brzozowski (Lavaca-Navidad River Authority), Brian Koch (Texas
+State Soil and Water Conservation Board), Bill Balboa (Matagorda Bay
+Foundation), Jason Pinchbeck (Texas General Land Office) and the Lavaca
+Bay Foundation for supporting development of this project and providing
+valuable feedback.}
 

### 2023-02-11 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
edits. Add IT approach to methods, add lagged responses in discussion.

@@
 
-<!---
-Need to add model selection criteria!
---->
+This study used an information theoretic approach to evaluate 
+if nutrient loads and/or freshwater inflows provided evidence of effects on water quality concentrations in Lavaca Bay.
+Model probabilities were calculated and compared using the AIC~c~ scores between each group of temporal, flow, and flow+load models [@burnhamAICModelSelection2011].
+Improvements in model probabilities provide evidence that the terms explain additional variation in the response variable.
+If model probabilities were tied, there wasn't evidence the more complicated model explains additional variation in water quality.
 
@@
 
-Based on criteria provided by @moriasiHydrologicWaterQuality2015, GAMs ranged
-from "satisfactory" to "very good" using median NSE, r^2^ and PBIAS 
-calculated using repeated 5-fold cross validation on predicted and measured
-nutrient loads. NO~3~ GAM models had median NSE values of 0.34 and 0.48 at
-USGS-08164000 and USGS-08164390 respectively. Median r^2^ values were 0.70
-(USGS-08164000) and 0.87 (USGS-08164525) and PBIAS values were 2.00
-(USGS-08164000) and 10.90 (USGS-08164525) for NO~3~ loads. Median NSE values were
-0.80 (USGS-08164000) and 0.91 (USGS-08164525) for TP loads. Median r^2^ values 
-for TP loads were 0.93 (USGS-08164000) and 0.99 (USGS-08164525) and median PBIAS 
-values were -7.20 (USGS-08164000) and -3.30 (USGS-08164525). Density plots of
+Using evaluation criteria from @moriasiHydrologicWaterQuality2015, predictive 
+performance of nutrient loading GAMs ranged from "satisfactory" to
+"very good" based on median NSE, r^2^, and PBIAS metrics calculated using 5-fold 
+cross validation. Median NSE values were 0.34 (Lavaca River) and 
+0.48 (Navidad River) for NO~3~ loads and 0.81 (Lavaca River) and 0.91 (Navidad 
+River) for TP loads. Median r^2^ values were 0.70
+(Lavaca River) and 0.87 (Navidad River) for NO~3~ loads and 0.93 (Lavaca River) 
+and 0.99 (Navidad River) fopr TP loads. Median PBIAS values were 2.00
+(Lavaca River) and 10.90 (Navidad River) for NO~3~ loads and -7.20 
+(Lavaca River) and -3.30 (Navidad River) for TP loads. Density plots of
 metrics show similar distribution of values between sites for the 
@@
 
-Annual NO~3~ and TP loads show considerable variation, generally following
+Predicted annual NO~3~ and TP loads show considerable variation, generally following
 patterns in discharge (Figure \ref{fig:fig3}, Figure \ref{fig:fig4}). 
-Flow-normalized TP loads at both sites and flow-normalized loads at 
-USGS-08164000 indicate watershed based loads have not changed much over time
-when accounting for changes in streamflow (Figure \ref{fig:fig3}). 
+Flow-normalized TP loads at both sites and flow-normalized NO~3~ loads at 
+USGS-08164000 indicate watershed based loads did not change much over time
+when accounting for variation driven by streamflow (Figure \ref{fig:fig3}). 
 Flow-normalized loads at USGS-08164000 show small variation over time with 
-some decreases in NO~3~ loads since 2013. Aggregated across both sites, the mean 
+some decreases in NO~3~ loads since 2013.
+
+Aggregated across both sites, the mean 
 annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
 `r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
-in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`.
-
-Total annual TP loads ranged from `r min_tp` kg in `r min_tp_year` to 
+in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`. Total annual TP loads 
+ranged from `r min_tp` kg in `r min_tp_year` to 
 `r max_tp` kg  in `r max_tp_year`. Mean annual TP loading from 2005 through 2020
@@
 
-Other studies have published estimates of TP loading estimates for the Lavaca
-River (USGS-08164000) over various time periods (Table \ref{tab:tpcomp}).
-Converted to average annual yield, the estimates of TP loading in this study
-are within the ranges of previous studies [@dunnTrendsNutrientInflows1996; @rebich_sources_2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019], with the TP 
-estimates in @dunnTrendsNutrientInflows1996 being notably lower. Given that none
+Converted to average annual yield, the estimates of annual TP loads at
+Lavaca River (USGS-08164000) are within the ranges in previous published studies
+(Table \ref{tab:tpcomp}; [@dunnTrendsNutrientInflows1996; @rebich_sources_2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]). It isn't obvious
+why TP estimates in @dunnTrendsNutrientInflows1996 were notably lower. Given that none
 of the studies identify substantially sized trends in TP, it is is possible that 
@@
 very high indicating subsets of values were problematic at characterizing functional
-relationships between nutrients and predictors. All of the water
-quality data for the two river sites in the TCEQ database were ambient water
-quality data, collected to be representative of typical conditions. This results
-in few data collected at the highest portions of the flow duration curve
-under which the majority of loadings take place. Supplementary flow-biased 
+relationships between nutrients and predictors. Because all of the water quality
+data for these two locations in the TCEQ databases were ambient water quality
+data, collected to be representative of typical flow conditions, there were few
+data at the highest portions of the flow-duration curve. It was beyond the scope
+of the current study to evaluate the subsets of cross-validation data and scores.
+However, the cross-validation procedure is indicative that more robust sampling
+would be beneficial for reducing prediction variance. Supplementary flow-biased 
 monitoring targeting storm- or high-flow conditions is recommended here to
@@
 cholorophyll-*a* concentrations are stable in comparison to other Texas
-estuaries that might be facing larger demands for freshwater diversions, higher
+estuaries that are facing larger demands for freshwater diversions, higher
 population growth, and more intense agricultural production 
@@
 The strong positive effect of freshwater inflow on NO~*x*~, TKN, and TP are
-suggestive of nonpoint watershed sources, consistent with watershed uses, and
-consistent with other studies relating freshwater inflow with nutrient 
+suggestive of nonpoint watershed sources, consistent with watershed uses and
+with other studies relating freshwater inflow with nutrient 
 concentrations in Lavaca Bay and other estuaries [@russell_effect_2006; 
@@
 freshwater inflow were observed at TCEQ-13384 located in the lower reach of 
-Lavaca Bay. Tidal flushing from Matagorda Bay might be responsible for 
-decreasing TKN concentrations and limiting the effects of freshwater inflow in 
-lower reaches of Lavaca Bay. @russell_effect_2006 suggested the processing of 
+Lavaca Bay. Tidal flushing from Matagorda Bay could be responsible for 
+diluting TKN and acting as a control on the effects of freshwater inflow in 
+lower reaches of Lavaca Bay. @russell_effect_2006 also suggested the processing of 
 organic loads in the upper portions of Lavaca Bay reduces the transport of 
@@
 
-Freshwater inflow also had a strong positive effect on chlorophyll-*a* at 
+Freshwater inflow had a strong positive effect on chlorophyll-*a* at 
 the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed decreases
@@
 the Lavaca River are probably masked under most conditions by discharge from the
-Navidad River. Given the relatively little variation in flow-normalized loads,
+Navidad River. Given the relatively small variation in flow-normalized loads,
 it can be expected that they would contribute little to the variance in downstream
@@
 with high turbidity or nutrient processing in upper portions of the Bay or 
-intertidal river limiting production [@russell_effect_2006].
+intertidal river limiting production [@russell_effect_2006]. Finally, it
+is reasonable to assume that fluctuations in DO may not occur immediately in 
+response to nutrient pulses or freshwater inflow. Work has has shown that various
+water quality parameters may have lagged effects lasting days or even months following
+storms and large discharge events [@mooney_watershed_2012; @wetzExtremeFutureEstuaries2013; @bukaveckas_influence_2020; @walkerTimescalesMagnitudeWater2021]. 
+However, our work only evaluates responses to
+loading and inflows occuring the day of water quality observations.
 
@@
 Bay watershed. However, additional flow-biased data collection efforts would
-decrease the prediction variance and probably improve accuracy at critical 
+decrease the prediction variance and improve accuracy at critical 
 high flow events. Ongoing projects will fill data gaps for total nitrogen and 
@@
 strong direct responses to freshwater flow. Small variance in flow-adjusted
-nutrient loads indicates that (1) there have been limited changes in non point
-sources of nutrients and (2) that those changes have limited effect on 
-water quality in Lavaca Bay to date. Although the study did not identify
-strong responses to changes in nutrient loading, this does provide a baseline
-assessment for future water quality management activities in the watershed.
+nutrient loads indicates that (1) there have been limited changes in non-point
+sources of nutrients and (2) that there isn't strong evidence that those small 
+changes have had effects on chlorophyll-*a* or dissolved oxygen in Lavaca Bay. 
+Although the study did not identify strong responses to changes in nutrient 
+loading, this does provide a baseline assessment for future water quality 
+management activities in the watershed.

### 2023-02-11 Schramm-Manuscript-2023/Schramm-Manuscript-2023.tex
edits. Add IT approach to methods, add lagged responses in discussion.

@@
 
+This study used an information theoretic approach to evaluate if
+nutrient loads and/or freshwater inflows provided evidence of effects on
+water quality concentrations in Lavaca Bay. Model probabilities were
+calculated and compared using the AIC\textsubscript{c} scores between
+each group of temporal, flow, and flow+load models
+\citep{burnhamAICModelSelection2011}. Improvements in model
+probabilities provide evidence that the terms explain additional
+variation in the response variable. If model probabilities were tied,
+there wasn't evidence the more complicated model explains additional
+variation in water quality.
+
 \hypertarget{results}{%
@@
 
-Based on criteria provided by \citet{moriasiHydrologicWaterQuality2015},
-GAMs ranged from ``satisfactory'' to ``very good'' using median NSE,
-r\textsuperscript{2} and PBIAS calculated using repeated 5-fold cross
-validation on predicted and measured nutrient loads. NO\textsubscript{3}
-GAM models had median NSE values of 0.34 and 0.48 at USGS-08164000 and
-USGS-08164390 respectively. Median r\textsuperscript{2} values were 0.70
-(USGS-08164000) and 0.87 (USGS-08164525) and PBIAS values were 2.00
-(USGS-08164000) and 10.90 (USGS-08164525) for NO\textsubscript{3} loads.
-Median NSE values were 0.80 (USGS-08164000) and 0.91 (USGS-08164525) for
-TP loads. Median r\textsuperscript{2} values for TP loads were 0.93
-(USGS-08164000) and 0.99 (USGS-08164525) and median PBIAS values were
--7.20 (USGS-08164000) and -3.30 (USGS-08164525). Density plots of
+Using evaluation criteria from
+\citet{moriasiHydrologicWaterQuality2015}, predictive performance of
+nutrient loading GAMs ranged from ``satisfactory'' to ``very good''
+based on median NSE, r\textsuperscript{2}, and PBIAS metrics calculated
+using 5-fold cross validation. Median NSE values were 0.34 (Lavaca
+River) and 0.48 (Navidad River) for NO\textsubscript{3} loads and 0.81
+(Lavaca River) and 0.91 (Navidad River) for TP loads. Median
+r\textsuperscript{2} values were 0.70 (Lavaca River) and 0.87 (Navidad
+River) for NO\textsubscript{3} loads and 0.93 (Lavaca River) and 0.99
+(Navidad River) fopr TP loads. Median PBIAS values were 2.00 (Lavaca
+River) and 10.90 (Navidad River) for NO\textsubscript{3} loads and -7.20
+(Lavaca River) and -3.30 (Navidad River) for TP loads. Density plots of
 metrics show similar distribution of values between sites for the same
@@
 
-Annual NO\textsubscript{3} and TP loads show considerable variation,
-generally following patterns in discharge (Figure \ref{fig:fig3}, Figure
-\ref{fig:fig4}). Flow-normalized TP loads at both sites and
-flow-normalized loads at USGS-08164000 indicate watershed based loads
-have not changed much over time when accounting for changes in
-streamflow (Figure \ref{fig:fig3}). Flow-normalized loads at
-USGS-08164000 show small variation over time with some decreases in
-NO\textsubscript{3} loads since 2013. Aggregated across both sites, the
-mean annual NO\textsubscript{3} load 2005 through 2020 was 205,405 kg
-(126,867 kg - 341,569 kg, 90\% CI). Annual NO\textsubscript{3} loads
-ranged from 12,574 kg in 2011 to 794,510 kg in 2007.
-
-Total annual TP loads ranged from 7,839 kg in 2011 to 595,075 kg in
-2007. Mean annual TP loading from 2005 through 2020 was 182,673 kg
-(152,227 kg - 219,310 kg, 90\% CI). On average, USGS-08164525 accounted
-for 68\% of NO\textsubscript{3} loads and 59\% of TP loads from 2005
-through 2020. However, during periods of extreme drought the Lavaca
-River (USGS-08164000) became the primary source of nutrient loading in
-the watershed with the Navidad River only accounting for 15\% and 25\%
-of NO\textsubscript{3} and TP loads in 2011 (Figure \ref{fig:fig4}).
+Predicted annual NO\textsubscript{3} and TP loads show considerable
+variation, generally following patterns in discharge (Figure
+\ref{fig:fig3}, Figure \ref{fig:fig4}). Flow-normalized TP loads at both
+sites and flow-normalized NO\textsubscript{3} loads at USGS-08164000
+indicate watershed based loads did not change much over time when
+accounting for variation driven by streamflow (Figure \ref{fig:fig3}).
+Flow-normalized loads at USGS-08164000 show small variation over time
+with some decreases in NO\textsubscript{3} loads since 2013.
+
+Aggregated across both sites, the mean annual NO\textsubscript{3} load
+2005 through 2020 was 205,405 kg (126,867 kg - 341,569 kg, 90\% CI).
+Annual NO\textsubscript{3} loads ranged from 12,574 kg in 2011 to
+794,510 kg in 2007. Total annual TP loads ranged from 7,839 kg in 2011
+to 595,075 kg in 2007. Mean annual TP loading from 2005 through 2020 was
+182,673 kg (152,227 kg - 219,310 kg, 90\% CI). On average, USGS-08164525
+accounted for 68\% of NO\textsubscript{3} loads and 59\% of TP loads
+from 2005 through 2020. However, during periods of extreme drought the
+Lavaca River (USGS-08164000) became the primary source of nutrient
+loading in the watershed with the Navidad River only accounting for 15\%
+and 25\% of NO\textsubscript{3} and TP loads in 2011 (Figure
+\ref{fig:fig4}).
 
@@
 
-Other studies have published estimates of TP loading estimates for the
-Lavaca River (USGS-08164000) over various time periods (Table
-\ref{tab:tpcomp}). Converted to average annual yield, the estimates of
-TP loading in this study are within the ranges of previous studies
-\citep{dunnTrendsNutrientInflows1996, rebich_sources_2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019},
-with the TP estimates in \citet{dunnTrendsNutrientInflows1996} being
-notably lower. Given that none of the studies identify substantially
-sized trends in TP, it is is possible that the period used in
-\citet{dunnTrendsNutrientInflows1996} was drier on average than the
-other studies. The SPARROW models used in \citet{rebich_sources_2011}
-and \citet{wise_spatially_2019} utilize a version of LOADEST in the
-underlying load estimation procedure, so a difference due to methodology
-alone is unlikely.
+Converted to average annual yield, the estimates of annual TP loads at
+Lavaca River (USGS-08164000) are within the ranges in previous published
+studies (Table \ref{tab:tpcomp};
+\citep{dunnTrendsNutrientInflows1996, rebich_sources_2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019}).
+It isn't obvious why TP estimates in
+\citet{dunnTrendsNutrientInflows1996} were notably lower. Given that
+none of the studies identify substantially sized trends in TP, it is is
+possible that the period used in \citet{dunnTrendsNutrientInflows1996}
+was drier on average than the other studies. The SPARROW models used in
+\citet{rebich_sources_2011} and \citet{wise_spatially_2019} utilize a
+version of LOADEST in the underlying load estimation procedure, so a
+difference due to methodology alone is unlikely.
 
@@
 problematic at characterizing functional relationships between nutrients
-and predictors. All of the water quality data for the two river sites in
-the TCEQ database were ambient water quality data, collected to be
-representative of typical conditions. This results in few data collected
-at the highest portions of the flow duration curve under which the
-majority of loadings take place. Supplementary flow-biased monitoring
+and predictors. Because all of the water quality data for these two
+locations in the TCEQ databases were ambient water quality data,
+collected to be representative of typical flow conditions, there were
+few data at the highest portions of the flow-duration curve. It was
+beyond the scope of the current study to evaluate the subsets of
+cross-validation data and scores. However, the cross-validation
+procedure is indicative that more robust sampling would be beneficial
+for reducing prediction variance. Supplementary flow-biased monitoring
 targeting storm- or high-flow conditions is recommended here to improve
@@
 work. The trend in DO and cholorophyll-\emph{a} concentrations are
-stable in comparison to other Texas estuaries that might be facing
-larger demands for freshwater diversions, higher population growth, and
-more intense agricultural production
+stable in comparison to other Texas estuaries that are facing larger
+demands for freshwater diversions, higher population growth, and more
+intense agricultural production
 \citep{wetzWaterQualityDynamics2016, bugica_water_2020}. The trend of
@@
 NO\textsubscript{\emph{x}}, TKN, and TP are suggestive of nonpoint
-watershed sources, consistent with watershed uses, and consistent with
-other studies relating freshwater inflow with nutrient concentrations in
-Lavaca Bay and other estuaries
+watershed sources, consistent with watershed uses and with other studies
+relating freshwater inflow with nutrient concentrations in Lavaca Bay
+and other estuaries
 \citep{russell_effect_2006, caffreyHighNutrientPulses2007, peierlsNonmonotonicResponsesPhytoplankton2012, palmerImpactsDroughtsLow2015, ciraPhytoplanktonDynamicsLowInflow2021}.
@@
 at TCEQ-13384 located in the lower reach of Lavaca Bay. Tidal flushing
-from Matagorda Bay might be responsible for decreasing TKN
-concentrations and limiting the effects of freshwater inflow in lower
-reaches of Lavaca Bay. \citet{russell_effect_2006} suggested the
-processing of organic loads in the upper portions of Lavaca Bay reduces
-the transport of nutrients into the lower reaches of the Bay.
-
-Freshwater inflow also had a strong positive effect on
-chlorophyll-\emph{a} at the upper- and mid-bay sites. The upper-bay
-site, TCEQ-13563, showed decreases in chlorophyll-\emph{a} at the
-highest freshwater inflow volumes. Freshwater flushing or increases in
-turbidity are associated with decreases in chlorophyll-\emph{a} in other
-estuaries
+from Matagorda Bay could be responsible for diluting TKN and acting as a
+control on the effects of freshwater inflow in lower reaches of Lavaca
+Bay. \citet{russell_effect_2006} also suggested the processing of
+organic loads in the upper portions of Lavaca Bay reduces the transport
+of nutrients into the lower reaches of the Bay.
+
+Freshwater inflow had a strong positive effect on chlorophyll-\emph{a}
+at the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed
+decreases in chlorophyll-\emph{a} at the highest freshwater inflow
+volumes. Freshwater flushing or increases in turbidity are associated
+with decreases in chlorophyll-\emph{a} in other estuaries
 \citep{peierlsNonmonotonicResponsesPhytoplankton2012, cloernPhytoplanktonPrimaryProduction2014}.
@@
 the Lavaca River are probably masked under most conditions by discharge
-from the Navidad River. Given the relatively little variation in
+from the Navidad River. Given the relatively small variation in
 flow-normalized loads, it can be expected that they would contribute
@@
 processing in upper portions of the Bay or intertidal river limiting
-production \citep{russell_effect_2006}.
+production \citep{russell_effect_2006}. Finally, it is reasonable to
+assume that fluctuations in DO may not occur immediately in response to
+nutrient pulses or freshwater inflow. Work has has shown that various
+water quality parameters may have lagged effects lasting days or even
+months following storms and large discharge events
+\citep{mooney_watershed_2012, wetzExtremeFutureEstuaries2013, bukaveckas_influence_2020, walkerTimescalesMagnitudeWater2021}.
+However, our work only evaluates responses to loading and inflows
+occuring the day of water quality observations.
 
@@
 Lavaca Bay watershed. However, additional flow-biased data collection
-efforts would decrease the prediction variance and probably improve
-accuracy at critical high flow events. Ongoing projects will fill data
-gaps for total nitrogen and TKN loading. This study, consistent with
-others along the Texas coast, found strong effects of freshwater flow on
-nutrient and chlorophyll-\emph{a} concentrations. DO concentrations,
-dominated by seasonal effects, did not show strong direct responses to
-freshwater flow. Small variance in flow-adjusted nutrient loads
-indicates that (1) there have been limited changes in non point sources
-of nutrients and (2) that those changes have limited effect on water
-quality in Lavaca Bay to date. Although the study did not identify
-strong responses to changes in nutrient loading, this does provide a
-baseline assessment for future water quality management activities in
-the watershed.
+efforts would decrease the prediction variance and improve accuracy at
+critical high flow events. Ongoing projects will fill data gaps for
+total nitrogen and TKN loading. This study, consistent with others along
+the Texas coast, found strong effects of freshwater flow on nutrient and
+chlorophyll-\emph{a} concentrations. DO concentrations, dominated by
+seasonal effects, did not show strong direct responses to freshwater
+flow. Small variance in flow-adjusted nutrient loads indicates that (1)
+there have been limited changes in non-point sources of nutrients and
+(2) that there isn't strong evidence that those small changes have had
+effects on chlorophyll-\emph{a} or dissolved oxygen in Lavaca Bay.
+Although the study did not identify strong responses to changes in
+nutrient loading, this does provide a baseline assessment for future
+water quality management activities in the watershed.
 

### 2023-02-20 Schramm_2023_02_PLOS/Schramm_2023_02_PLOS.qmd
format plos paper

--- 2023-02-11 Schramm-Manuscript-2023/Schramm-Manuscript-2023.Rmd
@@
 ---
-title: Assessing linkages between watershed nutrient loading and estuary water quality in Lavaca Bay, Texas
+title: "Assessing linkages between watershed nutrient loading and water quality in a subtropical estuary with semiparametric models."
 author: 
   - name: Michael Schramm
-    affil: 1
-    orcid: 0000-0003-1876-6592
-affiliation:
-  - num: 1
-    address: |
-      Texas A&M AgriLife Research - 
-      Texas Water Resources Institute
-      1001 Holleman Dr. E. 
-      College Station, TX 77840-2118
     email: michael.schramm@ag.tamu.edu
-# author citation list in chicago format
-authorcitation: |
-  Schramm, M.
-correspondence: |
-  michael.schramm@ag.tamu.edu
-# specify the mdpi cls and location of file
-# as of 2021 the template files are located in
-# a subdirectory and should be retained there 
-# since file locations are hard coded in the cls file.
-cls: Definitions/mdpi
-# class options
-classoption:
-  - water #first option is the journal name
-  - article #second option is the type of manuscript
-  - submit #do not change
-  - oneauthor #change to oneauthor for single author papers
-  #- pdftex #remove if compiling with XeLaTeX
-# front matter
-abstract: |
-  Lavaca Bay is a small secondary embayment on the Texas coast that is
-  displaying early signals of water quality degradation. This
-  study uses Generalized Additive Models to assesses both watershed nutrient loading 
-  and responses in estuary water quality to nutrient loading and streamflow.
-  While delivered nutrient loads varied greatly, flow-normalized nutrient loads indicated
-  limited evidence of changes in nutrient runoff in the watershed since 2005. 
-  As a consequence, there wasn't strong evidence for an effect of nutrient loading
-  (controlled for flow) on dissolved oxygen or chlorophyll-*a* concentrations.
-  There was evidence total phosphorus (TP) loading had a strong positive
-  effect on TP concentration in the Bay, particularly at sites closest to the
-  river mouth. Seasonally-adjusted streamflow had strong effects on nutrient
-  concentration and chlorophyll-*a* concentrations at most sites. Although
-  temporal trends did not identify significant estuary water quality degredation 
-  in dissolved oxygen or chlorophyll-*a* concentrations, site specific trends
-  in both organic and inorganic nitrogen are concerning for the long-term
-  health of the estuary. This study provides a baseline assessment of the
-  linkages between watershed nutrient loading and water quality responses in 
-  Lavaca Bay. Some of the data gaps highlighted in the study are being addressed
-  in ongoing projects to provide improved information for resource managers.
-# back matter
-keywords: |
-  estuary; nutrient loading; water quality; Texas
-acknowledgement: |
-  The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
-  Texas A&M Corpus Christi), Chad Kinsfather and Partick Brzozowski 
-  (Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water 
-  Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason Pinchbeck 
-  (Texas General Land Office) and the Lavaca Bay Foundation for supporting 
-  development of this project and providing valuable feedback.
-funding: |
-  This project was funded in part by a Texas Coastal Management Program grant approved by the Texas Land
-  Commissioner, providing financial assistance under the Coastal Zone Management Act of 1972, as amended,
-  awarded by the National Oceanic and Atmospheric Administration (NOAA), Office for Coastal Management,
-  pursuant to NOAA Award No. NA21NOS4190136. The views expressed herein are those of the author(s) and do not
-  necessarily reflect the views of NOAA, the U.S. Department of Commerce, or any of their subagencies
-dataavailability: |
-  Data and code are openly available in Zenodo at 
-  https://doi.org/10.5281/zenodo.7330754.
-conflictsofinterest: |
-  The author declares no conflict of interest. The project sponsors had no 
-  role in the design of the study; in the collection, analyses, or 
-  interpretation of data; in the writing of the manuscript, or in the decision 
-  to publish the results.
+    corresponding: true
+    affiliations:
+      - ref: agrilife
+affiliations:
+  - id: agrilife
+    name: Texas A&M AgriLife Research, Texas Water Resources Institute
+    city: College Station
+    state: Texas
+    country: United States of America
+format: 
+  docx:
+    number-sections: false
+    reference-doc: format.docx
+crossref:
+  fig-prefix: Fig   # (default is "Figure")
+  tbl-prefix: Table    # (default is "Table")
+  title-delim: .     # (default is ":")
+  ref-hyperlink: false # (default is true)
 bibliography: mybibfile.bib
-endnotes: false
-output: 
-  rticles::mdpi_article:
-    latex_engine: xelatex #pdflatex or XeLaTeX allowed
+csl: https://www.zotero.org/styles/plos?source=1
+filters: 
+  - authors-affiliations.lua
+keep-md: true
 ---
+
+<!--
+TODO:
+- find right csl format
+- fix cross refs
+- fix equations
+- bold captions(?)
+- fix symbols, subscripts, superscripts in tables
+ List corresponding author’s initials in 
+parentheses after the email address.
+make intial field? for corresponding author?
+- use imagemagick to convert png to tiff?
+-->
 
@@
 library(targets)
-#library(unitar)
 library(knitr)
@@
 library(gratia)
+library(gt)
+
+ragg_tiff <- function(..., res = 300) {
+  ragg::agg_tiff(..., res = res, units = "in")
+}
 
 ## this sets our default code chunk options
-knitr::opts_chunk$set(dev = "cairo_pdf",
-                      echo = FALSE,
+knitr::opts_chunk$set(echo = FALSE,
                       ## figure chunk options
+                      dev = "CairoTIFF",
                       dpi = 300,
@@
 
-## custom hook for wide figures
-
-hook_plot_wide <- function(x, options) {
-  x_out <- knitr:::hook_plot_tex(x, options)
-  if(isTRUE(options$wideplot)) {
-    x_out <- gsub(x = x_out,
-                  pattern='\\begin{figure}', 
-                  replacement='\\begin{figure}\\begin{adjustwidth}{-\\extralength}{0cm}',
-                  fixed=TRUE)
-
-    x_out <- gsub(x = x_out,
-                  pattern='\\caption',
-                  replacement='\\end{adjustwidth}\\caption',
-                  fixed = TRUE)
-  }
-  x_out
-}
-knitr::knit_hooks$set(plot=hook_plot_wide)
-
 store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
@@
 
+# Abstract
+Lavaca Bay is a small secondary embayment on the Texas coast that is
+displaying early signals of water quality degradation. This
+study applied a semiparametric approach to assess both watershed nutrient loads 
+and the responses in estuary water quality to nutrient loading and streamflow. 
+Cross-validation indicated that, despite data constraints, semiparametric models 
+performed well at nutrient load prediction. Based on these models, delivered
+annual nutrient loads varied substantially from year to year. In contrast,
+minimal changes in calculated flow-normalized loads indicate that changes in 
+nutrient loading were driven by natural variation in precipitation and runoff as
+opposed to changes in discharge or nonpoint source management efforts.
+Models did not identify significant long-term changes within Lavaca Bay for dissolved 
+oxygen or chlorophyll-*a*. However, site specific long-term increases in both
+organic and inorganic nitrogen are concerning. Further analysis found freshwater
+inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within
+Lavaca Bay but there was no evidence that changes in nutrient loading explained
+variation in dissolved oxygen or chlorophyll-*a* concentrations. In addition to
+providing a baseline assessment of watershed nutrient loading and water quality
+responses in the Lavaca Bay watershed, this study provides methodological
+support for the use of semiparametric methods in load regression models and 
+estuary assessments.
+
 # Introduction
 
-Like many coastal areas globally, coastal watersheds along the Texas Gulf coast 
+Like many coastal areas globally, the coastal watersheds along the Texas Gulf coast 
 are facing pressures from increasing population, increases in point source 
-and non-point source pollution and alterations to freshwater flows that affect
+and non-point source pollution and alterations to freshwater flows that degrade
 water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. Despite these increasing pressures, national scale 
@@
 Matagorda Bay system located roughly halfway between Houston, Texas and Corpus
-Christi, Texas. Lavaca Bay has faced substantial challenges associated with
+Christi, Texas. Lavaca Bay faces substantial challenges associated with
 legacy contamination but general water quality parameters such as dissolved oxygen 
@@
 @montagnaAssessmentRelationshipFreshwater2020] are a concern to local 
-stakeholders. @bugica_water_2020 identified monotonic increases in total 
+stakeholders. Recent water quality assessments identified monotonic increases in total 
 phosphorus (TP), orthophosphate, total Kjeldahl nitrogen (TKN), and
-chlorophyll-*a* at sites within Lavaca Bay. Although long-term changes in 
-DO concentrations were not identified, the trends in nutrient concentrations are 
-concerning due to the role of nitrogen as a limiting factor for primary 
-production in many Texas estuaries [@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; @paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the
+chlorophyll-*a* at sites within Lavaca Bay [@bugica_water_2020]. Although 
+long-term changes in DO concentrations were not identified, the trends in 
+nutrient concentrations are concerning due to the role of nitrogen as a 
+limiting factor for primary production in many Texas estuaries 
+[@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; @paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the
 ramifications that changes in nitrogen loadings could have for 
@@
 
-```{=latex}
-\begin{figure}
-\includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
-\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig:fig1}}
-\end{figure}  
-```
-
 Lavaca Bay is 190 km^2^ with the majority of freshwater inflow provided by the 
-Lavaca and Navidad River systems (Figure \ref{fig:fig1}). The Garcitas-Arenosa, 
+Lavaca and Navidad River systems (@fig-fig1). The Garcitas-Arenosa, 
 Placedo Creek, and Cox Bay watersheds provide additional freshwater inflows. 
@@
 
-Daily discharges for the Lavaca River (USGS-08164000, Figure \ref{fig:fig1}) 
+```{r}
+#| label: fig-fig1
+#| fig-width: 5.2
+#| fig-height: 6.75
+#| out-width: "5.2in"
+#| fig-cap: "Map of Lavaca Bay and the contribution watershed.
+The freshwater sites are the most downstream freshwater stream locations with water quality and streamflow data used for nutrient load models.
+Water quality concentration data at the three Lavaca Bay sites were used to assess relationships between freshwater flows, loads and estuary water quality."
+
+knitr::include_graphics("Figures/fig1-1.tiff")
+
+```
+
+Daily discharges for the Lavaca River (USGS-08164000, @fig-fig1) 
 were obtained from the United States Geologic Survey (USGS) National Water 
@@
 nitrate-nitrogen (NO~3~) data from January 2005 through December 2020 were 
-downloaded (Table \ref{tab:fwsummary}). Less than 5-years of total
+downloaded (@tbl-table1). Less than 5-years of total
 nitrogen and TKN concentration data were available at the freshwater sites and
@@
 chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
-(Table \ref{tab:estuarysummary}).
-
-```{r fwsummary}
+(@tbl-table2).
+
+```{r}
+#| label: tbl-table1
+#| tbl-cap: "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020."
+#| warning: false
 df <- tar_read(model_data, store = store)
@@
            site_no == "usgs08164525") |> 
-  rename(`NO\\textsubscript{3} (mg/L)` = original_NO3,
+  rename(`NO3 (mg/L)` = original_NO3,
          `TP (mg/L)` = original_TP,
@@
 
-datasummary(`Station ID` * (`TP (mg/L)` + `NO\\textsubscript{3} (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
+datasummary(`Station ID` * (`TP (mg/L)` + `NO3 (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
                Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-            data = df, output = "kableExtra",
-            booktabs = TRUE, table.envir = "table", 
-            position = "H", escape = FALSE,
-            caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.")
+            data = df, output = "kableExtra", format = "pipe") 
 ```
 
-```{r estuarysummary}
+```{r}
+#| label: tbl-table2
+#| warning: false
+#| tbl-cap: "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020."
 df <- tar_read(estuary_model_data, store = store)
@@
            case_when(
-             parameter_code == "70953" ~ "Chlorophyll-\\emph{a} ($\\mu$g/L)",
+             parameter_code == "70953" ~ "Chlorophyll-a (ug/L)",
              parameter_code == "00625" ~ "TKN (mg/L)",
-             parameter_code == "00630" ~ "NO\\textsubscript{\\emph{x}} (mg/L)",
+             parameter_code == "00630" ~ "NOx (mg/L)",
              parameter_code == "00665" ~ "TP (mg/L)",
@@
  modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
-                            `NO\\textsubscript{\\emph{x}} (mg/L)` +
+                            `NOx (mg/L)` +
                             `TKN (mg/L)` +
-                            `Chlorophyll-\\emph{a} ($\\mu$g/L)` +
+                            `Chlorophyll-a (ug/L)` +
                             `DO (mg/L)`) ~
                             Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-                          data = df, output = "dataframe") |> 
-   mutate(`Station ID` = case_when(
-     `Station ID` == "" ~ NA_character_,
-     `Station ID` != "" ~ `Station ID`
-     )) |>
-   tidyr::fill(`Station ID`, .direction = "down") |> 
-   kbl(format = "latex", booktabs = TRUE, table.envir = "table", 
-       position = "H", escape = FALSE,
-       caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.") |> 
-   collapse_rows(columns = 1, latex_hline = "major",
-                 valign = "middle")
+                          data = df, output = "kableExtra", format = "pipe") #|>
+  #  flextable::autofit()
+   # mutate(`Station ID` = case_when(
+   #   `Station ID` == "" ~ NA_character_,
+   #   `Station ID` != "" ~ `Station ID`
+   #   )) |>
+   # tidyr::fill(`Station ID`, .direction = "down") |> 
+   # gt::gt() |> 
+   # gt::as_word()
 
@@
 the underlying parameter estimation procedure of GAMs is substantially different 
-than WRTDS, both the functional form and results are demonstrated to be similar [@beckNumericalQualitativeContrasts2017]. The use of GAMs over other 
-regression-based approaches was (1) the ability to easily explore and 
-incorporate different model terms, (2) the ability to incorporate non-linear
-smooth function without explicit a priori knowledge of the expect shape, and (3)
-the ability to specify a link function that relates the expected value of the
-response to the linear predictors and avoid data transformations.
+than WRTDS, both the functional form and results are demonstrated to be similar [@beckNumericalQualitativeContrasts2017]. GAMs were chosen over other regression-based
+approached for use in this study due to; (1) the ability to easily explore and 
+incorporate different model terms; (2) the incorporation of non-liner smooth
+functions that do not require explicit a priori knowledge of the expected
+shape; and (3) inclusion of a link function that related the expected value of
+the response to linear predictors thus avoiding unneeded data transformations
+and bias corrections.
 
@@
 
-\begin{equation}\label{eq:1}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  \\
-y &\sim \mathcal{N}(\mu,\,\sigma^{2}),
-\end{aligned}
-\end{equation}
-
-where $\mu$ is the conditional expected NO~3~ or TP concentration, *g()* is the 
-log-link, $\alpha$ is the intercept, *f~n~()* are smoothing functions. *y* is the 
+$$
+g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  
+y \sim \mathcal{N}(\mu,\,\sigma^{2})
+$${#eq-eq1}
+
+where *μ* is the conditional expected NO~3~ or TP concentration, *g()* is the 
+log-link, *α* is the intercept, *f~n~()* are smoothing functions. *y* is the 
 response variable (NO~3~ or TP concentration) modeled as normally distributed 
-with mean $\mu$ and standard deviation $\sigma$. *ddate* is the date converted to 
+with mean *μ* and standard deviation *σ*. *ddate* is the date converted to 
 decimal notation, *yday* is numeric day of year (1-366), and *log1p(Q)* is the 
@@
 current time period [@wangLoadEstimationUncertainties2011;  
-@kuhnert_quantifying_2012; @zhang_improving_2017], and 
-expressed in @kuhnert_quantifying_2012 as:
-
-\begin{equation}\label{eq:2}
-ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m,
-\end{equation}
-
-where $\delta$ is the discount factor (here, set equal to 0.95), $\kappa_i$ 
+@kuhnert_quantifying_2012; @zhang_improving_2017]:
+
+$$
+ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m
+$${#eq-eq2}
+
+where *δ* is the discount factor (here, set equal to 0.95), κ~i~
 is the cumulative flow (*Q*) up to the *i*th day.
@@
 @zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
-the previous year relative to the entire period and calculated as described 
-by @zhang_improving_2017:
-
-\begin{equation}\label{eq:3}
+the previous year relative to the entire period and calculated as described as
+[@zhang_improving_2017]:
+
+$$
 ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}
-\end{equation}
+$${#eq-eq3}
 
@@
 
-\begin{equation}\label{eq:4}
+$$
 stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t)
-\end{equation}
+$${#eq-eq4}
 
@@
 period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3~ models 
-and *stfa* in TP models based on results from @zhang_improving_2017 
+and *stfa* in TP models based on previous work  
 demonstrating major improvements in NO~x~ regression models that incorporated 
 *ltfa* and moderate improvements in TP regression models that incorporated 
-*stfa*.
+*stfa* [@zhang_improving_2017].
 
@@
 @marraCoveragePropertiesConfidence2012; @mcdowell_implications_2021].
-Uncertainty in loads were reported as 90% credible intervals developed by 
+Uncertainty in loads were calculated as 90% credible intervals estimated by 
 drawing 1000 realizations of parameter estimates from the multivariate normal 
@@
 
-\begin{equation}\label{eq:5}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) \\
-y &\sim \Gamma(\mu,\lambda),
-\end{aligned}
-\end{equation}
-
-\begin{equation}\label{eq:6}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) \\
-y &\sim \Gamma(\mu,\lambda),
-\end{aligned}
-\end{equation}
-
-\begin{equation}\label{eq:7}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \\
-y &\sim \Gamma(\mu,\lambda),
-\end{aligned}
-\end{equation}
-
-where $\mu$ is the conditional expected response (nutrient concentration), *g()* 
+$$
+g(\mu) = \alpha + f_1(ddate) + f_2(yday) y \sim \Gamma(\mu,\lambda)
+$${#eq-eq5}
+
+$$
+g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(Q) y \sim \Gamma(\mu,\lambda)
+$${#eq-eq6}
+
+$$
+g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) y \sim \Gamma(\mu,\lambda)
+$${#eq-eq7}
+
+where *μ* is the conditional expected response (nutrient concentration), *g()* 
 is the log link, and response variable was modeled as Gamma distributed with 
-mean $\mu$ and scale $\lambda$. *f~1~(ddate)* is decimal date smoothed with a 
+mean *μ* and scale *λ*. *f~1~(ddate)* is decimal date smoothed with a 
 thin-plate regression spline, *f~2~(yday)* is the numeric day of year smoothed 
@@
 the total NO~3~ or TP watershed load. The set of models specified for each water
-quality response are in Table \ref{tab:estgammodels}.
-
-```{r estgammodels, echo=FALSE}
+quality response are in @tbl-table3.
+
+```{r}
+#| label: tbl-table3
+#| tbl-cap: "Set of GAM models specified for each water quality parameter response."
+#| warning: false
 df <- tibble(
-  `Water Quality Response Parameter` = c(rep("TP", 3), rep("NO\\textsubscript{\\emph{x}}", 3),
-               rep("Chlorophyll-\\emph{a}", 3), rep("Dissolved Oxygen", 3),
+  `Water Quality Response Parameter` = c(rep("TP", 3), rep("NOx", 3),
+               rep("Chlorophyll-a", 3), rep("Dissolved Oxygen", 3),
                rep("TKN", 2)),
@@
                     "s(ddate) + s(yday) + s(Q)",
-                    "s(ddate) + s(yday) + s(Q) + s(NO\\textsubscript{3} Load)",
+                    "s(ddate) + s(yday) + s(Q) + s(NO3 Load)",
                     "s(ddate) + s(yday)",
                     "s(ddate) + s(yday) + s(Q)",
-                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3} Load)",
+                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO3 Load)",
                     "s(ddate) + s(yday)",
                     "s(ddate) + s(yday) + s(Q)",
-                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3}  Load)",
+                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO3  Load)",
                     "s(ddate) + s(yday)",
@@
 
-kbl(df, format = "latex",
-    booktabs = TRUE,
-    caption = "Set of GAM models specified for each water quality parameter response.",
-    col.names = linebreak(c("Water Quality\nResponse Parameter",
-                          "Model",
-                          "Model Terms")),
-    table.envir = "table", 
-    position = "H", escape = FALSE) |> 
-  collapse_rows(columns = 1, latex_hline = "major",
-                valign = "middle")
+kbl(df, format = "pipe")
 
@@
 
-\begin{equation}\label{eq:8}
+$$
 g(\mu) = \alpha + f_1(yday),
-\end{equation}
+$${#eq-eq8}
 
@@
 
-\begin{equation}\label{eq:9}
+$$
 g(\mu) = \alpha + f_1(log(Q)),
-\end{equation}
+$${#eq-eq9}
 
@@
 identity link function. Response residuals from the respective GAM models were 
-used as *Q* and *Load* in Equation \ref{eq:6} and Equation \ref{eq:7}.
+used as *Q* and *Load* in @eq-eq6 and @eq-eq7.
 
@@
 (Lavaca River) and 0.87 (Navidad River) for NO~3~ loads and 0.93 (Lavaca River) 
-and 0.99 (Navidad River) fopr TP loads. Median PBIAS values were 2.00
+and 0.99 (Navidad River) for TP loads. Median PBIAS values were 2.00
 (Lavaca River) and 10.90 (Navidad River) for NO~3~ loads and -7.20 
@@
 same parameter, with the exception r^2^ values for NO~3~ loads where 
-USGS-08164000 showed a much larger variance in values compared to USGS-08164525
-(Figure \ref{fig:fig2}). TP GAMS had higher average NSE and r^2^ values and 
+Lavaca River had a much larger variance in values compared to the Navidad River
+(@fig-fig2). TP GAMS had higher average NSE and r^2^ values and 
 lower variance in metric values compared to NO~3~.
 
-```{r fig2, message=FALSE, warning=FALSE, fig.height=3.5, fig.width=6.5, fig.cap="Density plots of goodness-of-fit metrics (NSE, r\\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
-Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics.", out.width="100%", wideplot=TRUE}
+```{r}
+#| label: fig-fig2
+#| warning: false
+#| message: false
+#| fig-width: 5.2
+#| fig-asp: 0.65
+#| out-width: "5.2in"
+#| fig-cap: "Density plots of goodness-of-fit metrics (NSE, r^2^, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
+Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics."
 
@@
   mutate(site = "USGS-08164000",
-         parameter = "NO<sub>3</sub>") |> 
+         parameter = "NO <sub>3</sub>") |> 
   bind_rows(
@@
       mutate(site = "USGS-08164525",
-             parameter = "NO<sub>3</sub>")
+             parameter = "NO <sub>3</sub>")
   ) |> 
@@
              parameter = "TP")
-  )
+  ) |> 
+  mutate(site = case_when(
+    site == "USGS-08164525" ~ "Navidad",
+    site == "USGS-08164000" ~ "Lavaca"
+  ))
 
@@
                       scale_override(1, scale_x_continuous(limits = c(-1,1),
-                                                           expand = c(0,0))),
+                                                           expand = c(0,0),
+                                                           breaks = c(-1,0,1))),
                       scale_override(2, scale_x_continuous(limits = c(0,1),
-                                                           expand = c(0,0)))
+                                                           expand = c(0,0),
+                                                           breaks = c(0,0.5,1)))
                     )) +
   labs(x = "NSE", y = "") +
-  theme_TWRI_print() +
-  theme(axis.text = element_text(size = 6),
-        axis.title.x = element_markdown(size = 6),
+  theme_TWRI_print(base_family = "Arial") +
+  theme(axis.text = element_text(size = 8),
+        axis.title.x = element_markdown(size = 8),
         strip.text.x = element_markdown(size = 8),
         strip.background = element_rect(fill = "white"),
-        panel.spacing.x = unit(20L, "pt"),
+        panel.spacing.x = unit(15L, "pt"),
         legend.direction = "horizontal",
         legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8))
+        legend.title = element_text(size = 8),
+        plot.margin = unit(c(1L,0L,1L,0L), "pt"))
 
@@
                       scale_override(1, scale_x_continuous(limits = c(0,1),
-                                                           expand = c(0,0))),
+                                                           expand = c(0,0),
+                                                           breaks = c(0,0.5,1))),
                       scale_override(2, scale_x_continuous(limits = c(0,1),
-                                                           expand = c(0,0)))
+                                                           expand = c(0,0),
+                                                           breaks = c(0,0.5,1)))
                     )) +
   labs(x = "r<sup>2</sup>", y = "") +
-  theme_TWRI_print() +
-  theme(axis.text = element_text(size = 6),
-        axis.title.x = element_markdown(size = 6),
+  theme_TWRI_print(base_family = "Arial") +
+  theme(axis.text = element_text(size = 8),
+        axis.title.x = element_markdown(size = 8),
         strip.text.x = element_markdown(size = 8),
         strip.background = element_rect(fill = "white"),
-        panel.spacing.x = unit(20L, "pt"),
+        panel.spacing.x = unit(15L, "pt"),
         legend.direction = "horizontal",
         legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8))
+        legend.title = element_text(size = 8),
+        plot.margin = unit(c(0L,0L,1L,0L), "pt"))
 
@@
   labs(x = "PBIAS", y = "") +
-  theme_TWRI_print() +
-  theme(axis.text = element_text(size = 6),
-        axis.title.x = element_markdown(size = 6),
+  theme_TWRI_print(base_family = "Arial") +
+  theme(axis.text = element_text(size = 8),
+        axis.title.x = element_markdown(size = 8),
         strip.text.x = element_markdown(size = 8),
         strip.background = element_rect(fill = "white"),
-        panel.spacing.x = unit(20L, "pt"),
+        panel.spacing.x = unit(15L, "pt"),
         legend.direction = "horizontal",
@@
         legend.title = element_text(size = 8),
-        axis.text.y = element_blank())
+        axis.text.y = element_blank(),
+        plot.margin = unit(c(0L,0.25,1L,0L), "pt"))
 
@@
 Predicted annual NO~3~ and TP loads show considerable variation, generally following
-patterns in discharge (Figure \ref{fig:fig3}, Figure \ref{fig:fig4}). 
-Flow-normalized TP loads at both sites and flow-normalized NO~3~ loads at 
-USGS-08164000 indicate watershed based loads did not change much over time
-when accounting for variation driven by streamflow (Figure \ref{fig:fig3}). 
-Flow-normalized loads at USGS-08164000 show small variation over time with 
+patterns in discharge (@fig-fig3; @fig-fig4). 
+Flow-normalized TP loads at both sites and flow-normalized NO~3~ loads in the Lavaca River
+indicated watershed based loads did not change much over time
+when accounting for variation driven by streamflow (@fig-fig3). 
+Flow-normalized loads in the Lavaca River showed small variation over time with 
 some decreases in NO~3~ loads since 2013.
 
-Aggregated across both sites, the mean 
-annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
-`r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
-in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`. Total annual TP loads 
-ranged from `r min_tp` kg in `r min_tp_year` to 
-`r max_tp` kg  in `r max_tp_year`. Mean annual TP loading from 2005 through 2020
-was `r mean_tp` kg (`r mean_tp_low` kg - `r mean_tp_upper` kg, 90% CI).
-On average, USGS-08164525 accounted for `r no3_mean_prop` of NO~3~ loads and 
-`r tp_mean_prop` of TP loads from 2005 through 2020. However, during periods of extreme 
-drought the Lavaca River (USGS-08164000) became the primary source of 
-nutrient loading in the watershed with the Navidad River only accounting for 
-`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 (Figure 
-\ref{fig:fig4}).
-
-```{r fig3, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.", out.width="100%"}
+```{r}
+#| label: fig-fig3
+#| error: false
+#| warning: false
+#| message: false
+#| fig-width: 5.2
+#| fig-asp: 0.7
+#| out-width: "5.2in"
+#| fig-cap: "Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."
 load <- tar_read(daily_no3_08164000, store = store)
@@
                      color = "Flow-Normalized Annual Load + 90% CI")) +
-  labs(x = "", y = "Annual NO<sub>3</sub> Load [kg]", subtitle = "USGS-08164000") +
+  labs(x = "", 
+       y = expression(Annual~NO[3]~Load~"[kg]"), 
+       subtitle = "Lavaca River") +
   scale_shape_manual(name = "values",
@@
   coord_cartesian(ylim = c(1000, 1E7)) +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
-        plot.subtitle = element_text(size = 8),
+  theme_TWRI_print(base_family = "sans") +
+  theme(axis.title.y = element_text(size = 8),
+        axis.text.x = element_blank(),
+        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "bold"),
         panel.grid.major.x = element_line(color = "#d9d9d9",
@@
         legend.title = element_blank(),
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 8),
+        plot.margin = unit(c(0L,3L,0L,0L), "pt"))
 
@@
                      color = "Flow-Normalized Annual Load + 90% CI")) +
-  labs(x = "", y = "Annual NO<sub>3</sub> Load [kg]", subtitle = "USGS-08164525") +
+  labs(x = "", 
+       y = expression(Annual~NO[3]~Load~"[kg]"), 
+       subtitle = "Navidad River") +
   scale_shape_manual(name = "values",
@@
   coord_cartesian(ylim = c(1000, 1E7)) +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
-        plot.subtitle = element_text(size = 8),
+  theme_TWRI_print(base_family = "sans") +
+  theme(axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.text.x = element_blank(),
+        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "bold"),
         panel.grid.major.x = element_line(color = "#d9d9d9",
@@
         legend.title = element_blank(),
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 8),
+        plot.margin = unit(c(0L,0L,0L,3L), "pt"))
 
@@
                      color = "Flow-Normalized Annual Load + 90% CI")) +
-  labs(x = "", y = "Annual TP Load [kg]", subtitle = "USGS-08164000") +
+  labs(x = "", y = "Annual TP Load [kg]", subtitle = "Lavaca River") +
   scale_shape_manual(name = "values",
@@
   coord_cartesian(ylim = c(1000, 1E6)) +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
-        plot.subtitle = element_text(size = 8),
+  theme_TWRI_print(base_family = "sans") +
+  theme(axis.title.y = element_text(size = 8),
         panel.grid.major.x = element_line(color = "#d9d9d9",
@@
         legend.title = element_blank(),
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 8),
+        plot.margin = unit(c(0L,3L,0L,0L), "pt"),
+        plot.subtitle = element_blank())
 
@@
                      color = "Flow-Normalized Annual Load + 90% CI")) +
-  labs(x = "", y = "Annual TP Load [kg]", subtitle = "USGS-08164525") +
+  labs(x = "", y = "Annual TP Load [kg]", subtitle = "Navidad River") +
   scale_shape_manual(name = "values",
@@
   coord_cartesian(ylim = c(1000, 1E6)) +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
-        plot.subtitle = element_text(size = 8),
+  theme_TWRI_print(base_family = "sans") +
+  theme(axis.title.y = element_blank(),
+        axis.text.y = element_blank(),
         panel.grid.major.x = element_line(color = "#d9d9d9",
@@
         legend.title = element_blank(),
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 8),
+        plot.margin = unit(c(0L,1L,0L,3L), "pt"),
+        plot.subtitle = element_blank())
 
@@
 
-```{r fig4, error=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.cap="Comparison of delivered annual loads at USGS-08164000 and USGS-08164525.", out.width="100%"}
+Aggregated across both sites, the mean 
+annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
+`r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
+in `r min_no3_year` to `r max_no3` kg in `r max_no3_year`. Total annual TP loads 
+ranged from `r min_tp` kg in `r min_tp_year` to 
+`r max_tp` kg  in `r max_tp_year`. Mean annual TP loading from 2005 through 2020
+was `r mean_tp` kg (`r mean_tp_low` kg - `r mean_tp_upper` kg, 90% CI).
+On average, the Navidad River accounted for `r no3_mean_prop` of NO~3~ loads and 
+`r tp_mean_prop` of TP loads from 2005 through 2020. However, during periods of extreme 
+drought the Lavaca River became the primary source of 
+nutrient loading in the watershed with the Navidad River only accounting for 
+`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 (@fig-fig4).
+
+```{r}
+#| label: fig-fig4
+#| error: false
+#| message: false
+#| warning: false
+#| fig-width: 3
+#| fig-asp: 2
+#| fig-cap: "Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers."
+#| out-width: "3in"
 load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
@@
 
-b <- ggplot(no3_annual) +
-  geom_col(aes(year, NO3_Estimate, fill = site), width = 0.8) +
-  scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
-                     breaks = c(2005, 2010, 2015, 2020)) +
-  scale_y_continuous(labels = scales::comma) +
-  scale_fill_carto_d(palette = "Vivid", direction = -1) +
-  labs(x = "", y = "Predicted Annual<br>NO<sub>3</sub> Load [kg]") +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
-        axis.text = element_text(size = 8),
-        panel.grid.major.x = element_line(color = "#d9d9d9",
-                                          linetype = "dotted"),
-        legend.title = element_blank(),
-        legend.text = element_text(size = 8))
+# b <- ggplot(no3_annual) +
+#   geom_col(aes(year, NO3_Estimate, fill = site), width = 0.8) +
+#   scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
+#                      breaks = c(2005, 2010, 2015, 2020)) +
+#   scale_y_continuous(labels = scales::comma) +
+#   scale_fill_carto_d(palette = "Vivid", direction = -1) +
+#   labs(x = "", 
+#        y = "Predicted Annual<br>NO <sub>3</sub> Load [kg]") +
+#   theme_TWRI_print(base_family = "Arial") +
+#   theme(axis.title.y = element_text(size = 8),
+#         axis.text = element_text(size = 8),
+#         panel.grid.major.x = element_line(color = "#d9d9d9",
+#                                           linetype = "dotted"),
+#         legend.title = element_blank(),
+#         legend.text = element_text(size = 8))
 
@@
 
+ylab1 <- expression(Annual~NO[3]~Load)
+ylab2 <- expression("[Proportion]")
 c <- ggplot() +
@@
   scale_y_continuous(expand = c(0,0)) +
-  labs(x = "", y = "Proportion of Annual<br>NO<sub>3</sub> Load") +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
+  labs(x = "",
+       y = "")  +
+  coord_cartesian(clip = "off") +
+  annotation_custom(grid::textGrob(ylab1,
+                                   x = unit(-0.23, "npc"),
+                                   rot = 90,
+                                   gp = grid::gpar(fontfamily = "sans",
+                                             fontsize = 8))) +
+  annotation_custom(grid::textGrob(ylab2,
+                                   x = unit(-0.18, "npc"),
+                                   rot = 90,
+                                   gp = grid::gpar(fontfamily = "sans",
+                                                   fontsize = 8))) +
+  theme_TWRI_print(base_family = "sans") +
+  theme(axis.title.y = element_text(size = 8),
         axis.text = element_text(size = 8),
@@
 
+ylab1 <- expression("Annual Discharge")
+ylab2 <- expression("[Million Gallons]")
 d <- tar_read(qdata, store = store) |>
@@
   scale_y_continuous(expand = c(0,0)) +
-  labs(x = "", y = "Annual Discharge<br>[Millon Gallons]") +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
+  labs(x = "",
+       y = "") +
+  coord_cartesian(clip = "off") +
+  annotation_custom(grid::textGrob(ylab1,
+                                   x = unit(-0.25, "npc"),
+                                   rot = 90,
+                                   gp = grid::gpar(fontfamily = "sans",
+                                             fontsize = 8))) +
+  annotation_custom(grid::textGrob(ylab2,
+                                   x = unit(-0.20, "npc"),
+                                   rot = 90,
+                                   gp = grid::gpar(fontfamily = "sans",
+                                                   fontsize = 8))) +
+  theme_TWRI_print(base_family = "Arial") +
+  theme(axis.title.y = element_text(size = 8),
         axis.text = element_text(size = 8),
@@
         legend.title = element_blank(),
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 8),
+        legend.position = "none",
+        plot.margin = unit(c(1L,1L,1L,2L), "pt"))
 
@@
   mutate(x = as.Date(paste0(year, "-01-01"), "%Y-%m-%d"))
-
-e <- ggplot(tp_annual) +
-  geom_col(aes(year, TP_Estimate, fill = site), width = 0.8) +
-  scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
-                     breaks = c(2005, 2010, 2015, 2020)) +
-  scale_y_continuous(labels = scales::comma) +
-  scale_fill_carto_d(palette = "Vivid", direction = -1) +
-  labs(x = "", y = "Predicted Annual<br>TP Load [kg]") +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
-        axis.text = element_text(size = 8),
-        panel.grid.major.x = element_line(color = "#d9d9d9",
-                                          linetype = "dotted"),
-        legend.title = element_blank(),
-        legend.text = element_text(size = 8))
-
+# 
+# e <- ggplot(tp_annual) +
+#   geom_col(aes(year, TP_Estimate, fill = site), width = 0.8) +
+#   scale_x_continuous(expand = expansion(mult = c(0.05, 0.05)), 
+#                      breaks = c(2005, 2010, 2015, 2020)) +
+#   scale_y_continuous(labels = scales::comma) +
+#   scale_fill_carto_d(palette = "Vivid", direction = -1) +
+#   labs(x = "", y = "Predicted Annual<br>TP Load [kg]") +
+#   theme_TWRI_print(base_family = "Arial") +
+#   theme(axis.title.y = element_text(size = 8),
+#         axis.text = element_text(size = 8),
+#         panel.grid.major.x = element_line(color = "#d9d9d9",
+#                                           linetype = "dotted"),
+#         legend.title = element_blank(),
+#         legend.text = element_text(size = 8))
+# 
+# 
 prop <- tp_annual |>
@@
 
+ylab1 <- expression("Annual TP Load")
+ylab2 <- expression("[Proportion]")
 f <- ggplot() +
@@
   scale_y_continuous(expand = c(0,0)) +
-  labs(x = "", y = "Proportion of Annual<br>TP Load") +
-  theme_TWRI_print() +
-  theme(axis.title.y = element_markdown(size = 8),
+  labs(x = "",
+       y = "") +
+  coord_cartesian(clip = "off") +
+  annotation_custom(grid::textGrob(ylab1,
+                                   x = unit(-0.23, "npc"),
+                                   rot = 90,
+                                   gp = grid::gpar(fontfamily = "sans",
+                                             fontsize = 8))) +
+  annotation_custom(grid::textGrob(ylab2,
+                                   x = unit(-0.18, "npc"),
+                                   rot = 90,
+                                   gp = grid::gpar(fontfamily = "sans",
+                                                   fontsize = 8))) +
+  theme_TWRI_print(base_family = "Arial") +
+  theme(axis.title.y = element_text(size = 8),
         axis.text = element_text(size = 8),
@@
         legend.title = element_blank(),
-        legend.text = element_text(size = 8))
+        legend.text = element_text(size = 8),
+        legend.position = "none")
 
 layout <- "
-BE
-BE
-BE
-BE
-CF
-CF
-CF
-CF
+CC
+CC
+CC
+CC
+FF
+FF
+FF
+FF
 DD
@@
 
-b + c + d + e + f + guide_area() + 
+c + d + f + guide_area() +
   plot_layout(guides = 'collect', design  = layout)
+
 ```
@@
 
-```{r fig5, message=FALSE, warning=FALSE, fig.height=3.5, fig.width=6, fig.cap="Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.", out.width="100%", wideplot=TRUE}
+GAMs did not identify significant changes in TP or DO concentrations at any
+of the Lavaca Bay sites from 2005 through 2020 (@fig-fig5). The
+upper-bay site, TCEQ-13563, had a linear increase in NO~x~ concentration and
+and decrease in chlorophyll-*a* from 2005 through 2014. The mid-bay site,
+TCEQ-13383, showed a periodic pattern in NO~x~ concentration that appeared
+similar to precipitation/inflow patterns, as well as a post 2011 increase
+in TKN concentrations. No significant long-term trends in concentrations 
+were identified by GAMs for the lower-bay TCEQ-13384 site.
+
+```{r}
+#| label: fig-fig5
+#| message: false
+#| warning: false
+#| fig-width: 5.2
+#| fig-asp: 0.7
+#| fig-cap: "Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs."
+#| out-width: "5.2in"
 draw_smoothed_years <- function(model1, model2, model3,
@@
   `TCEQ-13384` <- model3
+
+  newdata <- data_slice(model1, ddate = evenly(2005:2020, n = 200))
 
@@
     `TCEQ-13384`,
-    smooths = "s(ddate)")
+    smooths = "s(ddate)",
+    data = newdata)
 
@@
          subtitle = subtitle) +
-    theme_TWRI_print() +
-    theme(axis.title.y = element_markdown(size = 8),
-          axis.title.x = element_markdown(size = 8),
-          plot.subtitle = element_markdown(size = 8),
+    theme_TWRI_print(base_family = "sans") +
+    theme(axis.title.y = element_text(size = 8),
+          axis.title.x = element_text(size = 8),
+          axis.text.x = element_text(size = 8),
+          axis.text.y = element_text(size = 8),
+          plot.subtitle = element_text(size = 8),
           legend.direction = "vertical",
@@
                             tar_read(tp_lavaca_13384_temporal, store = store),
-                            ylab = "TP Smooth<br>Estimate [mg/L]", 
-                            subtitle = "A: TP Trend")
+                            ylab = "TP [mg/L]", 
+                            subtitle = "A")
 
@@
                              tar_read(no3_lavaca_13384_temporal, store = store),
-                             ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]", 
-                             subtitle = "B: NO*<sub>x</sub>* Trend")
+                             ylab = "NOx [mg/L]", 
+                             subtitle = "B")
 
@@
                              tar_read(tkn_lavaca_13384_temporal, store = store),
-                             ylab = "TKN Smooth<br>Estimate [mg/L]", 
-                             subtitle = "C: TKN Trend")
+                             ylab = "TKN [mg/L]", 
+                             subtitle = "C")
 
@@
                              tar_read(chla_lavaca_13384_temporal, store = store),
-                             ylab = "Chlorophyll-*a*\nSmooth<br>Estimate [µg/L]", 
-                             subtitle = "D: Chlorophyll-*a* Trend")
+                             ylab = expression("Chlorophyll-"~italic(a)~"[μg/L]"), 
+                             subtitle = "D")
 
@@
                              tar_read(do_lavaca_13384_temporal, store = store),
-                             ylab = "DO\nSmooth<br>Estimate [mg/L]", 
-                            subtitle = "E: DO Trend")
+                             ylab = "DO [mg/L]", 
+                            subtitle = "E")
 
@@
 
-GAMs did not identify significant changes in TP or DO concentrations at any
-of the Lavaca Bay sites from 2005 through 2020 (Figure \ref{fig:fig5}). The
-upper-bay site, TCEQ-13563, had a linear increase in NO~x~ concentration and
-and decrease in chlorophyll-*a* from 2005 through 2014. The mid-bay site,
-TCEQ-13383, showed a periodic pattern in NO~x~ concentration that appeared
-similar to precipitation/inflow patterns, as well as a post 2011 increase
-in TKN concentrations. No significant long-term trends in concentrations 
-were identified by GAMs for the lower-bay TCEQ-13384 site.
-
 Freshwater inflow provided additional explanation for changes in TP and 
 NO~*x*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
-model probability values (Table \ref{tab:tab4}).
+model probability values (@tbl-table4).
 TCEQ-13563, the site closest to the river outlet, was the only site that
@@
 
-```{r tab4, echo=FALSE, message=FALSE, warning=FALSE}
+```{r}
+#| label: tbl-table4
+#| message: false
+#| warning: false
+#| tbl-cap: "Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
+Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis."
 
@@
   mutate(Parameter = case_when(
-    Parameter == "Nitrite+Nitrate" ~ "NO\\textsubscript{\\emph{x}}",
-    Parameter == "Chlorophyll-a" ~ "Chlorophyll-\\emph{a}",
+    Parameter == "Nitrite+Nitrate" ~ "NO~*x*~",
+    Parameter == "Chlorophyll-a" ~ "Chlorophyll-*a*",
     TRUE ~ as.character(Parameter)
   )) |> 
-  mutate(Parameter = fct_relevel(Parameter, "TP", "NO\\textsubscript{\\emph{x}}", "Chlorophyll-\\emph{a}","TKN", "DO"),
+  mutate(Parameter = fct_relevel(Parameter, "TP", "NO~*x*~", "Chlorophyll-*a*","TKN", "DO"),
          Model = fct_relevel(Model, "Temporal", "Flow", "Flow + Load")) |>
@@
   arrange(Parameter, Site) |> 
-  kbl(format = "latex",
-    booktabs = TRUE,
-    table.envir = "table", 
-    position = "H", escape = FALSE,
-    caption = "Estuary GAM AIC\\textsubscript{c} values and associated model probabilities (in parenthesis).
-Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.") |>
-    collapse_rows(columns = 1,
-                valign = "middle", latex_hline = "major") |> 
-  column_spec(3, bold = c(rep(F, 7), T, F, T, T, F, T, T, F),
-              italic = c(rep(F, 7), T, F, T, T, F, T, T, F)) |> 
-  column_spec(4, bold = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T),
-              italic = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T)) |> 
-  column_spec(5, bold = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F),
-              italic =c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F))
+  kbl(format = "pipe")
 
@@
 
-```{r fig6, message=FALSE, warning=FALSE, fig.height=3.5, fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO\\textsubscript{\\emph{x}}, chlorophyll-\\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.", out.width="100%", wideplot=TRUE}
+GAMs showed increases in freshwater inflow resulted in nearly linear increases 
+in TP and NO~*x*~ concentration at all three sites (@fig-fig6). At the upper-bay
+TCEQ-13563 site, GAMs showed increases in freshwater inflow initially 
+increased chlorophyll-*a* and DO concentration but concentrations leveled and
+potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
+nearly linear increased in chlorophyll-*a* concentration in response to increases
+freshwater inflow. Freshwater flow did not have significant
+effects on chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.
+
+```{r}
+#| label: fig-fig6
+#| message: false
+#| warning: false
+#| fig-width: 5.2
+#| fig-asp: 0.7
+#| out-width: "5.2in"
+#| fig-cap: "Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."
 draw_smoothed_surface <- function(comp,
@@
          subtitle = subtitle) +
-    theme_TWRI_print() +
+    theme_TWRI_print(base_family = "sans") +
     theme(axis.title.y = element_markdown(size = 8),
@@
 
-GAMs showed increases in freshwater inflow resulted in nearly linear increases 
-in TP and NO~*x*~ concentration at all three sites (Figure \ref{fig:fig6}). At the upper-bay
-TCEQ-13563 site, GAMs showed increases in freshwater inflow initially 
-increased chlorophyll-*a* and DO concentration but concentrations leveled and
-potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
-nearly linear increased in chlorophyll-*a* concentration in response to increases
-freshwater inflow. Freshwater flow did not have significant
-effects on chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.
-
 Increased TP loads resulted in nearly linear increases of TP concentration at
-the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively (Figure \ref{fig:fig7}).
+the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively (@fig-fig7).
 The relative effect size appeared to much smaller than the effect of freshwater
@@
 lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
-provided only small improvements to the model (Table \ref{tab:tab4}). As noted
+provided only small improvements to the model (@tbl-table4). As noted
 above, nutrient loadings did not provide any explanation in changes in the
@@
 
-```{r fig7, message=FALSE, warning=FALSE, fig.height=2.25, fig.width=5.5, fig.cap="Estimated effects of nutrient load residuals on TP and NO\\textsubscript{\\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.", out.width="100%", wideplot=FALSE}
+```{r}
+#| label: fig-fig7
+#| message: false
+#| warning: false
+#| fig-width: 5.2
+#| fig-asp: 0.45
+#| out-width: "5.2in"
+#| fig-cap: "Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs."
 
@@
 
-Converted to average annual yield, the estimates of annual TP loads at
-Lavaca River (USGS-08164000) are within the ranges in previous published studies
-(Table \ref{tab:tpcomp}; [@dunnTrendsNutrientInflows1996; @rebich_sources_2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]). It isn't obvious
+Converted to average annual yield, the estimates of annual TP loads for the
+Lavaca River are within the ranges in previous published studies
+(@tbl-table5; [@dunnTrendsNutrientInflows1996; @rebich_sources_2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]). It isn't obvious
 why TP estimates in @dunnTrendsNutrientInflows1996 were notably lower. Given that none
@@
 
-```{r tpcomp}
+```{r}
+#| label: tbl-table5
+#| warning: false
+#| message: false
+#| tbl-cap: "Comparisons of previously published estimates of mean annual TP yield at the Lavaca River site."
+#| results: asis
 
@@
   Parameter = rep("TP", 5),
-  `Reported Yield (kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})` = 
-    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")"), 
-      "45.2", "42", "20.81-91.58", "28.9"),
+  "Reported Yield (kg ⋅ km^-2^ ⋅ year^-1^)" = 
+    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")^a^"), 
+      "45.2", "42", "20.81-91.58^b^", "28.9"),
   Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
   `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
-  Reference = c("This work", "\\citet{wise_spatially_2019}",
-                "\\citet{omaniEstimationSedimentNutrient2014}",
-                "\\citet{rebich_sources_2011}",
-                "\\citet{dunnTrendsNutrientInflows1996}")
+  Reference = c("This work", "@wise_spatially_2019",
+                "@omaniEstimationSedimentNutrient2014",
+                "@rebich_sources_2011",
+                "@dunnTrendsNutrientInflows1996")
 )
 
-df[[1,2]] <- paste0(df[[1,2]], footnote_marker_symbol(1))
-df[[4,2]] <- paste0(df[[4,2]], footnote_marker_symbol(2))
-
-kbl(df, format = "latex",
-    booktabs = TRUE,
-    table.envir = "table",
-    position = "H", escape = FALSE,
-    col.names = linebreak(c("Parameter",
-                            "Reported Yield\n(kg$\\cdot$km\\textsuperscript{-2}$\\cdot$year\\textsuperscript{-1})",
-                          "Approach",
-                          "Time Period",
-                          "Reference")),
-    caption = "Mean estimates of annual TP yield in the Lavaca River watershed in published studies.") |>
-  add_footnote(c("Values represent the mean of annual point estimates, lower and upper 95% credible intervals.",
-                 "A single point estimate was not reported, these value represent the range depicted on the choropleth map provided in the report."),
-               notation = "symbol",
-               threeparttable = TRUE,
-               escape = TRUE)
+x <- kable(df, format = "pipe")
+
+cat(x, 
+    "^a^Mean of the annual point estimates and the lower and upper 95% credible intervals.", 
+    "^b^Represents a binned value range from a choropleth map.",
+    sep = "\n")
 
@@
 The non-linear temporal water quality trends identified using GAMs differed
-slightly from the trends identified by @bugica_water_2020. This is
+slightly from previously identified trends [@bugica_water_2020]. This is
 not unexpected due to the different time periods, different methodology, and
@@
 diluting TKN and acting as a control on the effects of freshwater inflow in 
-lower reaches of Lavaca Bay. @russell_effect_2006 also suggested the processing of 
-organic loads in the upper portions of Lavaca Bay reduces the transport of 
-nutrients into the lower reaches of the Bay.
+lower reaches of Lavaca Bay. Previous work suggests the processing of 
+organic loads in the upper portions of Lavaca Bay mgiht reduce the transport of 
+nutrients into the lower reaches of the Bay [@russell_effect_2006].
 
@@
 management activities in the watershed.
+
+# Acknowledgements
+
+The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
+Texas A&M Corpus Christi), Chad Kinsfather and Patrick Brzozowski 
+(Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water 
+Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason Pinchbeck 
+(Texas General Land Office) and the Lavaca Bay Foundation for supporting 
+development of this project and providing valuable project feedback.
+
+# References

### 2023-02-20 map.R

@@ -1,0 +1,331 @@
+library(targets)
+library(tidyverse)
+library(ragg)
+library(magick)
+library(sf)
+library(ggrepel)
+library(patchwork)
+library(dataRetrieval)
+#library(FedData)
+library(ggspatial)
+library(rcartocolor)
+library(twriTemplates)
+
+download_usgs_sites <- function() {
+  site_numbers <- c("08164000", #lavaca
+                    "08164525")
+  sites <- readNWISsite(site_numbers)
+  sites
+}
+
+download_tceq_sites <-function() {
+  site_numbers <- c("TCEQMAIN-13563",
+                    "TCEQMAIN-13383",
+                    "TCEQMAIN-13384")
+  sites <- whatWQPsites(siteid = site_numbers)
+  sites
+}
+
+
+## usgs sites
+df_sites <- download_usgs_sites() |> 
+  filter(site_no == "08164000" | site_no == "08164525")
+
+df_sites <- st_as_sf(df_sites, coords = c("dec_long_va", "dec_lat_va"),
+                     crs = 4326)  
+
+## tceq sites
+tceq_sites <- download_tceq_sites()
+tceq_sites <- st_as_sf(tceq_sites, coords = c("LongitudeMeasure", "LatitudeMeasure"),
+                       crs = 4326)
+
+tceq_sites$MonitoringLocationIdentifier <- stringr::str_replace_all(tceq_sites$MonitoringLocationIdentifier, "MAIN", "")
+
+## read rivers spatial data
+rivers <- read_sf("data/Spatial/lavaca.gpkg",
+                  layer = "rivers")
+
+## read lake texana polygon
+waterbody <- read_sf("data/Spatial/lavaca.gpkg",
+                     layer = "waterbody")
+
+## read watershed polygons
+lavaca_ws <- read_sf("data/Spatial/lavaca.gpkg",
+                     layer = "lavaca_ws")
+
+navidad_ws <- read_sf("data/Spatial/lavaca.gpkg",
+                      layer = "navidad_ws")
+ds_ws <- read_sf("data/Spatial/lavaca.gpkg",
+                 layer = "ds_ws")
+
+## read counties polygons
+counties <- read_sf("data/Spatial/lavaca.gpkg",
+                    layer = "counties")
+
+## combine watersheds into one sf
+ws <- bind_rows(lavaca_ws, navidad_ws, ds_ws)
+
[…]

### 2023-04-20 Schramm_2023_02_PLOS/Schramm_2023_02_PLOS.qmd
Plos draft rejected :sad:

@@
                       ## figure chunk options
-                      dev = "CairoTIFF",
-                      dpi = 300,
+                      dev = "CairoPS",
+                      dpi = 600,
                       fig.align = "center",
@@
 
-p1 / p2 / p3 / guide_area() + plot_layout(guides = 'collect',
+p4 <- p1 / p2 / p3 / guide_area() + plot_layout(guides = 'collect',
                                           design = layout)
+
+p4
 

### 2023-04-20 figs-eps.R

@@ -1,0 +1,944 @@
+## manuscript figs
+
+
+library(targets)
+library(knitr)
+library(tidyverse)
+library(kableExtra)
+library(patchwork)
+library(ggrepel)
+library(rcartocolor)
+library(twriTemplates)
+library(ggspatial)
+library(ragg)
+library(flextable)
+library(modelsummary)
+library(scico)
+library(ggridges)
+library(ggtext)
+library(lubridate)
+library(units)
+library(gratia)
+library(gt)
+library(extrafont)
+library(systemfonts)
+
+store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
+# Fig 2 -----------------------------------------------------------------------
+
+## some functions to overide scales in facets via dewey dunnington
+#https://dewey.dunnington.ca/post/2018/modifying-facet-scales-in-ggplot2/
+scale_override <- function(which, scale) {
+  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
+    stop("which must be an integer of length 1")
+  }
+  
+  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
+    stop("scale must be an x or y position scale")
+  }
+  
+  structure(list(which = which, scale = scale), class = "scale_override")
+}
+
+CustomFacetWrap <- ggproto(
+  "CustomFacetWrap", FacetWrap,
+  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
+    # make the initial x, y scales list
+    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)
+    
+    if(is.null(params$scale_overrides)) return(scales)
+    
+    max_scale_x <- length(scales$x)
+    max_scale_y <- length(scales$y)
+    
+    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
+    for(scale_override in params$scale_overrides) {
+      which <- scale_override$which
+      scale <- scale_override$scale
+      
+      if("x" %in% scale$aesthetics) {
+        if(!is.null(scales$x)) {
+          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
+          scales$x[[which]] <- scale$clone()
+        }
+      } else if("y" %in% scale$aesthetics) {
+        if(!is.null(scales$y)) {
+          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
[…]

### 2023-04-20 map.R

@@ -321,11 +321,10 @@ p3 <-p1 + inset_element(p2, left = 0.7, bottom = 0.7, right = 1, top = 1,
                    align_to = "full")
 
 
-ragg::agg_png("map.png", width = 5.2, height = 6.75, units = "in", res = 300)
+ragg::agg_tiff("map.tif", width = 5.2, height = 6.75, units = "in", res = 600,
+               compression = "lzw")
 p3
 dev.off()
 
 
-img <- magick::image_read("map.png")
-img
-magick::image_write(img, path = "map.tiff", format = "tiff")
+

### 2023-04-20 map_final_report.R

@@ -1,0 +1,333 @@
+library(targets)
+library(tidyverse)
+library(ragg)
+library(magick)
+library(sf)
+library(ggrepel)
+library(patchwork)
+library(dataRetrieval)
+#library(FedData)
+library(ggspatial)
+library(rcartocolor)
+library(twriTemplates)
+
+download_usgs_sites <- function() {
+  site_numbers <- c("08164000", #lavaca
+                    "08164390", #navidad
+                    "08164450", #sandy
+                    "08164503", #w mustang
+                    "08164504", #e mustang
+                    "08164525")
+  sites <- readNWISsite(site_numbers)
+  sites
+}
+
+download_tceq_sites <-function() {
+  site_numbers <- c("TCEQMAIN-13563",
+                    "TCEQMAIN-13383",
+                    "TCEQMAIN-13384")
+  sites <- whatWQPsites(siteid = site_numbers)
+  sites
+}
+
+
+## usgs sites
+df_sites <- download_usgs_sites()
+
+df_sites <- st_as_sf(df_sites, coords = c("dec_long_va", "dec_lat_va"),
+                     crs = 4326)  
+
+## tceq sites
+tceq_sites <- download_tceq_sites()
+tceq_sites <- st_as_sf(tceq_sites, coords = c("LongitudeMeasure", "LatitudeMeasure"),
+                       crs = 4326)
+
+tceq_sites$MonitoringLocationIdentifier <- stringr::str_replace_all(tceq_sites$MonitoringLocationIdentifier, "MAIN", "")
+
+## read rivers spatial data
+rivers <- read_sf("data/Spatial/lavaca.gpkg",
+                  layer = "rivers")
+
+## read lake texana polygon
+waterbody <- read_sf("data/Spatial/lavaca.gpkg",
+                     layer = "waterbody")
+
+## read watershed polygons
+lavaca_ws <- read_sf("data/Spatial/lavaca.gpkg",
+                     layer = "lavaca_ws")
+
+navidad_ws <- read_sf("data/Spatial/lavaca.gpkg",
+                      layer = "navidad_ws")
+ds_ws <- read_sf("data/Spatial/lavaca.gpkg",
+                 layer = "ds_ws")
+
+## read counties polygons
+counties <- read_sf("data/Spatial/lavaca.gpkg",
+                    layer = "counties")
+
+## combine watersheds into one sf
+ws <- bind_rows(lavaca_ws, navidad_ws, ds_ws)
+
[…]

### 2023-05-20 Schramm-2023-04-AS/Lavaca-Manuscript-AS.qmd
init springer nature tex submission for AS

--- 2023-04-20 Schramm_2023_02_PLOS/Schramm_2023_02_PLOS.qmd
@@
 title: "Assessing linkages between watershed nutrient loading and water quality in a subtropical estuary with semiparametric models."
+format:
+  wordcount-docx:
+    toc: false
+    number-sections: false
+    reference-doc: format.docx
 author: 
@@
     country: United States of America
-format: 
-  docx:
-    number-sections: false
-    reference-doc: format.docx
 crossref:
@@
   ref-hyperlink: false # (default is true)
-bibliography: mybibfile.bib
-csl: https://www.zotero.org/styles/plos?source=1
+  eq-prefix: Eq
+bibliography: mybib.yaml
+csl: https://www.zotero.org/styles/springer-basic-author-date-no-et-al?source=1
 filters: 
@@
 keep-md: true
+cite-method: citeproc
+link-bibliography: false
 ---
-
-<!--
-TODO:
-- find right csl format
-- fix cross refs
-- fix equations
-- bold captions(?)
-- fix symbols, subscripts, superscripts in tables
- List corresponding author’s initials in 
-parentheses after the email address.
-make intial field? for corresponding author?
-- use imagemagick to convert png to tiff?
--->
 
@@
 library(gt)
+library(officer)
 
@@
                       ## figure chunk options
-                      dev = "CairoPS",
-                      dpi = 600,
+                      dev = "ragg_tiff",
+                      dpi = 300,
                       fig.align = "center",
+                      fig.ext = "tiff",
                       out.width = "100%")
@@
 nutrient loading were driven by natural variation in precipitation and runoff as
-opposed to changes in discharge or nonpoint source management efforts.
-Models did not identify significant long-term changes within Lavaca Bay for dissolved 
-oxygen or chlorophyll-*a*. However, site specific long-term increases in both
-organic and inorganic nitrogen are concerning. Further analysis found freshwater
-inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within
-Lavaca Bay but there was no evidence that changes in nutrient loading explained
-variation in dissolved oxygen or chlorophyll-*a* concentrations. In addition to
-providing a baseline assessment of watershed nutrient loading and water quality
-responses in the Lavaca Bay watershed, this study provides methodological
-support for the use of semiparametric methods in load regression models and 
-estuary assessments.
+opposed to changes in nonpoint sources.
+Estuary water quality models did not identify significant long-term changes 
+within Lavaca Bay for dissolved oxygen or chlorophyll-*a*. However, site specific 
+long-term increases in both organic and inorganic nitrogen are concerning. 
+Further analysis found freshwater inflow was a strong driver of nutrient and 
+chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that 
+changes in nutrient loading explained variation in dissolved oxygen or 
+chlorophyll-*a* concentrations. In addition to providing a baseline assessment 
+of watershed nutrient loading and water quality responses in the Lavaca Bay 
+watershed, this study provides methodological support for the use of 
+semiparametric methods in load regression models and estuary assessments.
 
@@
 and non-point source pollution and alterations to freshwater flows that degrade
-water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. Despite these increasing pressures, national scale 
-assessments have classified coastal estuaries in Texas as moderate or lower for
-exhibiting eutrophic conditions [@bricker_effects_2008]. However, a suite of 
-recent studies indicate that estuary water quality dynamics in both 
-agriculturally dominated and urban watersheds within Texas are in fact 
-expressing conditions that are increasingly conducive to algal blooms and 
-eutrophication [@wetzWaterQualityDynamics2016; @wetz_exceptionally_2017; 
-@bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022]. With identification
-of localized areas of estuary water quality concern along the Texas coast 
-[@bugica_water_2020], localized studies are being prioritized to better 
-inform management actions.
-
-This project provides an assessment of nutrient loading and water quality
+water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. 
+Despite these escalating pressures, national scale assessments have classified 
+coastal estuaries in Texas as moderate or low risk for eutrophic conditions 
+[@bricker_effects_2008]. However, a suite of recent studies indicates that 
+estuary water quality dynamics in both agriculturally dominated and urban 
+watersheds within Texas are in fact expressing conditions that are increasingly 
+conducive to algal blooms and eutrophication [@wetzWaterQualityDynamics2016; 
+@wetz_exceptionally_2017; @bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022]. 
+With identification of localized areas of estuary water quality concern along 
+the Texas coast [@bugica_water_2020], localized studies are being prioritized 
+to better inform management actions.
+
+This project aims to provide an assessment of nutrient loading and water quality
 responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
@@
 (DO), nutrients, and biological parameters have been well within state
-water quality standards. However, long-term declines of benthic 
-fauna abundance, biomass, and diversity in Lavaca Bay primarily
+water quality standards. More recently identified long-term declines 
+in abundance, biomass, and diversity of benthic fauna in Lavaca Bay have been 
 linked to reductions in freshwater inflows and changes in estuary salinity 
 [@beserespollackLongtermTrendsResponse2011; @palmerImpactsDroughtsLow2015;
-@montagnaAssessmentRelationshipFreshwater2020] are a concern to local 
-stakeholders. Recent water quality assessments identified monotonic increases in total 
-phosphorus (TP), orthophosphate, total Kjeldahl nitrogen (TKN), and
+@montagnaAssessmentRelationshipFreshwater2020] and are a concern to local 
+stakeholders. Water quality assessments identified monotonic increases 
+in total phosphorus (TP), orthophosphate, total Kjeldahl nitrogen (TKN), and
 chlorophyll-*a* at sites within Lavaca Bay [@bugica_water_2020]. Although 
@@
 limiting factor for primary production in many Texas estuaries 
-[@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012; @doradoUnderstandingInteractionsFreshwater2015; @paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the
+[@gardnerNitrogenFixationDissimilatory2006; @houTransformationFateNitrate2012;
+@doradoUnderstandingInteractionsFreshwater2015; 
+@paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the 
 ramifications that changes in nitrogen loadings could have for 
@@
 address water quality impairments in the freshwater portions of the 
-Lavaca Bay watershed [@jainTechnicalSupportDocument2021; @schramm_lavaca_2018; @bertholdDirectMailingEducation2021]. However, at a statewide scale, these
-approaches have shown limited success and emphasize a need for improved efforts
-at assessing and linking management actions with downstream water quality to
-better identify  and replicate effective management actions across the state
-[@schrammTotalMaximumDaily2022]. The identification and communication of 
-changes and trends in water quality is complicated by the fact that trends are 
-often non-linear and confounded by precipitation and runoff that 
-hinder traditional analysis [@wazniakLinkingWaterQuality2007; 
-@lloydMethodsDetectingChange2014]. To provide
-actionable information for resource managers, water quality conditions must be 
-evaluated relative to changes in natural environmental drivers to 
-better understand  and manage potential anthropogenic effects.
-This study utilizes semiparemetric methods to develop estimates of delivered 
+Lavaca Bay watershed [@jainTechnicalSupportDocument2021; 
+@schrammLavacaRiverWatershed2018; @bertholdDirectMailingEducation2021]. 
+However, at a statewide scale, these approaches have shown limited success and 
+emphasize a need for improved efforts at assessing and linking management 
+actions with downstream water quality to identify and replicate effective 
+management actions across the state [@schrammTotalMaximumDaily2022]. The 
+identification and communication of changes and trends in water quality is 
+complicated by the fact that trends are often non-linear and confounded by 
+precipitation and runoff that hinder traditional analysis 
+[@wazniakLinkingWaterQuality2007; @lloydMethodsDetectingChange2014]. 
+The development and application of statistical methods such as Weighted
+Regressions on Time, Discharge and Season [WRTDS, @hirschWeightedRegressionsTime2010] 
+and Generalized Additive Models [GAMs, @woodFastStableRestricted2011] has 
+provided effective tools for researchers to quantify and
+communicate non-linear changes in river and estuary pollutant loadings.
+
+WRTDS calculates a time series of in stream concentrations or loads 
+(daily, monthly, or annually) and flow-normalized estimates of concentrations and loads
+using locally weighted regresssion for unique combinations of time, discharge, and season. 
+WRTDS has been widely used to assess and identify trends in riverine nutrients 
+[@oelsner_recent_2019; @stackpooleLongTermMississippi2021],
+chlorides [@stetsIncreasingChlorideRivers2018], and other pollutants of concern 
+[@shodaWaterqualityTrendsRivers2019]. WRTDS has also been succesfully adapted to 
+assess trends in estuarine water quality concentrations [@beckFourDecadesWater2018].
+
+While WRTDS is a statistical approach developed specifically for water quality applications, 
+GAMs are a broadly applicable statistical method. GAMs are a semiparametric 
+extension of generalized linear models where the linear predictor is represented
+as the sum of multiple unknown smooth functions and parametric linear predictors
+[@woodFastStableRestricted2011]. Although the underlying parameter estimation 
+procedure of GAMs is substantially different than WRTDS, both the functional 
+form and results have been demonstrated to be similar 
+[@beckNumericalQualitativeContrasts2017]. Water quality applications of GAMs
+include river and catchment nutrient concentration and load models 
+[@wangLoadEstimationUncertainties2011;@kroonRiverLoadsSuspended2012;
+@kuhnertQuantifyingTotalSuspended2012;@robson_prediction_2015-1;
+@hagemannEstimatingNutrientOrganic2016;@mcdowell_implications_2021;
+@biagi_novel_2022]. GAMs can also be used to identify non-linear temporal trends
+(including flow-normalized trends) in pollutant concentrations and loads 
+[@beckNumericalQualitativeContrasts2017; @murphyGeneralizedAdditiveModel2019]. 
+Recently GAMs have also been used to link water quality responses in 
+receiving water bodies to changes in nutrient inputs 
+[@murphyNutrientImprovementsChesapeake2022]. @beckNumericalQualitativeContrasts2017
+provides a substantial discussion on the differences (and similarities) between 
+GAMs and WRTDS for water quality applications.
+
+To provide actionable information for resource managers in Lavaca Bay, 
+water quality conditions must be evaluated relative to changes in natural 
+environmental drivers to better understand and manage potential human impacts.
+This study utilizes GAMs to develop estimates of delivered 
 and flow-normalized nutrient loads and assess changes in loads delivered to 
-Lavaca Bay. The study also assesses the response of water quality parameters in
+Lavaca Bay. GAMs were chosen over other regression-based
+approached for use in this study due to; (1) the ability to easily explore and 
+incorporate different model terms; (2) the incorporation of non-liner smooth
+functions that do not require explicit a priori knowledge of the expected
+shape; and (3) inclusion of a link function that related the expected value of
+the response to linear predictors thus avoiding unneeded data transformations
+and bias corrections. The study also assesses the response of water quality parameters in
 Lavaca Bay over time and in response to freshwater inflow controlled
@@
 nitrate-nitrogen (NO~3~) data from January 2005 through December 2020 were 
-downloaded (@tbl-table1). Less than 5-years of total
+downloaded (`#r run_reference("table1")`). Less than 5-years of total
 nitrogen and TKN concentration data were available at the freshwater sites and
@@
 ```{r}
-#| label: tbl-table1
-#| tbl-cap: "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020."
-#| warning: false
+
 df <- tar_read(model_data, store = store)
@@
                Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-            data = df, output = "kableExtra", format = "pipe") 
+            data = df, output = "kableExtra", format = "simple")
+
+# datasummary(`Station ID` * (`TP (mg/L)` + `NO3 (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
+#                Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
+#             data = df, output = "flextable") |> 
+#   set_caption(caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.",
+#               autonum = run_autonum(seq_id = "tab", bkm = "table1")) |> 
+#   style(pr_t = fp_text_default(font.size = 9), part = "body") |>
+#   style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
+#   set_table_properties(layout = "autofit") |>
+#   padding(padding.top = 0, padding.bottom = 0, part = "body")
 ```
@@
 ```{r}
-#| label: tbl-table2
 #| warning: false
-#| tbl-cap: "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020."
+
 df <- tar_read(estuary_model_data, store = store)
@@
                             Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-                          data = df, output = "kableExtra", format = "pipe") #|>
+                          data = df, output = "flextable") |> 
+  set_caption(caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.",
+              autonum = run_autonum(seq_id = "tab", bkm = "table2")) |> 
+  style(pr_t = fp_text_default(font.size = 9), part = "body") |>
+  style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
+  set_table_properties(layout = "autofit") |>
+  padding(padding.top = 0, padding.bottom = 0, part = "body")
+
+ # modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
+ #                            `NOx (mg/L)` +
+ #                            `TKN (mg/L)` +
+ #                            `Chlorophyll-a (ug/L)` +
+ #                            `DO (mg/L)`) ~
+ #                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
+ #                          data = df, output = "kableExtra", format = "pipe") #|>
   #  flextable::autofit()
@@
 outlet of Lake Texana on the Navidad River (USGS-08164525) were developed using 
-Generalized Additive Models (GAMs) relating nutrient concentration to river 
-discharge, season, and time.
+GAMs relating nutrient concentration to river discharge, season, and time.
 Separate models were fit at each station for each parameter and used to predict
-nutrient concentrations for each day in the study period. GAMs can be specified
-in a functionally similar manner to the commonly used LOADEST 
-[@cohn_validity_1992] or WRTDS 
-[@hirsch_weighted_2010] regression models and have been shown to
-produce reliable estimates of nutrient and sediment loadings 
-[@wangLoadEstimationUncertainties2011; @kroonRiverLoadsSuspended2012; 
-@kuhnert_quantifying_2012; @robson_prediction_2015-1; 
-@hagemannEstimatingNutrientOrganic2016; @mcdowell_implications_2021; 
-@biagi_novel_2022]. GAMs are a semiparametric extension of generalized linear 
-models where the linear predictor is represented as the sum of multiple unknown 
-smooth functions and parametric linear predictors [@wood_fast_2011]. Although 
-the underlying parameter estimation procedure of GAMs is substantially different 
-than WRTDS, both the functional form and results are demonstrated to be similar [@beckNumericalQualitativeContrasts2017]. GAMs were chosen over other regression-based
-approached for use in this study due to; (1) the ability to easily explore and 
-incorporate different model terms; (2) the incorporation of non-liner smooth
-functions that do not require explicit a priori knowledge of the expected
-shape; and (3) inclusion of a link function that related the expected value of
-the response to linear predictors thus avoiding unneeded data transformations
-and bias corrections.
-
-GAMs were fit using the *mgcv* package in R which makes available multiple 
-types of smooth functions with automatic smoothness selection [@wood_fast_2011]. 
+nutrient concentrations for each day in the study period. GAMs were fit using 
+the *mgcv* package in R which makes available multiple 
+types of smooth functions with automatic smoothness selection [@woodFastStableRestricted2011]. 
 The general form of the model related NO~3~ or TP concentration to a long term
-tend, season, streamflow, and two different antecedent discharge terms:
+tend, season, streamflow, and two different antecedent discharge terms, 
+shown in [@eq-eq1]:
 
 $$
-g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  
-y \sim \mathcal{N}(\mu,\,\sigma^{2})
+\begin{align*}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  \\
+y &\sim \mathcal{N}(\mu,\,\sigma^{2})
+\end{align*}
 $${#eq-eq1}
@@
 current time period [@wangLoadEstimationUncertainties2011;  
-@kuhnert_quantifying_2012; @zhang_improving_2017]:
+@kuhnertQuantifyingTotalSuspended2012; @zhang_improving_2017], using [@eq-eq2]:
 
 $$
-ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m
+\begin{align*}
+ma(\delta) &= d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1},\\
+\kappa_{i} &=\sum_{m=1}^{i}\hat{Q}_m,
+\end{align*}
 $${#eq-eq2}
@@
 Flow anomaly (*fa*) is a unitless term that represents how wet or dry the 
-current time period is from a previous time period [@vecchia_trends_2009; 
+current time period is from a previous time period [@vecchiaTrendsConcentrationsUse2009; 
 @zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
-the previous year relative to the entire period and calculated as described as
-[@zhang_improving_2017]:
+the previous year relative to the entire period [@zhang_improving_2017] and 
+calculated using [@eq-eq3]:
 
@@
 and the short-term flow anomaly (*stfa*) calculated as the current day flow 
-compared to the preceding 1-month streamflow:
+compared to the preceding 1-month streamflow using [@eq-eq4]:
 
@@
 period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3~ models 
-and *stfa* in TP models based on previous work  
-demonstrating major improvements in NO~x~ regression models that incorporated 
+and *stfa* in TP models based on previous work demonstrating major improvements 
+in NO~x~ regression models that incorporated 
 *ltfa* and moderate improvements in TP regression models that incorporated 
@@
 
-The calculation of model terms for the Lake Texana site were slightly modified
+The calculation of model terms for the Lake Texana site were modified
 because daily loads are not a function of natural stream flow processes alone,
@@
 covariates on each day of the year equal to each of the historical values for
-that day of the year over the study period [@hirsch_weighted_2010]. The 
+that day of the year over the study period [@hirschWeightedRegressionsTime2010]. The 
 flow-normalized estimate was calculated as the mean of all the predictions for 
@@
 using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
-and average Nash-Sutcliffe Efficiency (NSE), r^2^ and percent bias (PBIAS) 
-metrics across folds were calculated for each model.
+and average Nash-Sutcliffe Efficiency (NSE), Pearson sample correlation (*r*) 
+and percent bias (PBIAS) metrics across folds were calculated for each model.
 
@@
 effect on changes in estuary water quality a series of GAM models were fit at
-each site relating parameter concentration to temporal trends, inflow, and 
-nutrient loads [@murphyNutrientImprovementsChesapeake2022]:
+each site relating parameter concentration to temporal trends ([@eq-eq5]), 
+temporal trends and inflow (@eq-eq6), and temporal trends, inflow, and
+nutrient loads (@eq-eq7):
 
 $$
-g(\mu) = \alpha + f_1(ddate) + f_2(yday) y \sim \Gamma(\mu,\lambda)
+\begin{align*}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday)\\
+y &\sim \Gamma(\mu,\lambda)
+\end{align*}
 $${#eq-eq5}
@@
 $$
-g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(Q) y \sim \Gamma(\mu,\lambda)
+\begin{align*}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) \\
+y &\sim \Gamma(\mu,\lambda)
+\end{align*}
 $${#eq-eq6}
@@
 $$
-g(\mu) = \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) y \sim \Gamma(\mu,\lambda)
+\begin{align*}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \\
+y &\sim \Gamma(\mu,\lambda)
+\end{align*}
 $${#eq-eq7}
@@
 ```{r}
-#| label: tbl-table3
-#| tbl-cap: "Set of GAM models specified for each water quality parameter response."
 #| warning: false
@@
 
-kbl(df, format = "pipe")
+# kbl(df, format = "pipe")
+
+flextable::flextable(df) |> 
+  set_caption(caption = "Set of GAM models specified for each water quality parameter response.",
+              autonum = run_autonum(seq_id = "tab", bkm = "table3")) |> 
+  style(pr_t = fp_text_default(font.size = 9), part = "body") |>
+  style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
+  set_table_properties(layout = "autofit") |>
+  padding(padding.top = 0, padding.bottom = 0, part = "body")
 
@@
 log transformed nutrient loads obtained by fitting a GAM relating season 
-(day of year) to log transformed daily freshwater inflow values:
-
-$$
-g(\mu) = \alpha + f_1(yday),
-$${#eq-eq8}
-
+(day of year) to log transformed daily freshwater inflow values
 and a GAM relating log transformed NO~3~ or TP loads to log transformed daily
-inflow:
-
-$$
-g(\mu) = \alpha + f_1(log(Q)),
-$${#eq-eq9}
-
-where the response variables were modeled as normally distributed with an 
-identity link function. Response residuals from the respective GAM models were 
+inflow.Response residuals from the respective GAM models were 
 used as *Q* and *Load* in @eq-eq6 and @eq-eq7.
@@
 This study used an information theoretic approach to evaluate 
-if nutrient loads and/or freshwater inflows provided evidence of effects on water quality concentrations in Lavaca Bay.
-Model probabilities were calculated and compared using the AIC~c~ scores between each group of temporal, flow, and flow+load models [@burnhamAICModelSelection2011].
-Improvements in model probabilities provide evidence that the terms explain additional variation in the response variable.
-If model probabilities were tied, there wasn't evidence the more complicated model explains additional variation in water quality.
+if nutrient loads and/or freshwater inflows provided evidence of effects on 
+water quality concentrations in Lavaca Bay. Model probabilities were calculated 
+and compared using the AIC~c~ scores between each group of temporal, flow, and 
+flow+load models [@burnhamAICModelSelection2011]. Improvements in model 
+probabilities provide evidence that the terms explain additional variation in 
+the response variable. If model probabilities were tied, there wasn't evidence 
+the more complicated model explains additional variation in water quality.
 
@@
 
-Using evaluation criteria from @moriasiHydrologicWaterQuality2015, predictive 
-performance of nutrient loading GAMs ranged from "satisfactory" to
-"very good" based on median NSE, r^2^, and PBIAS metrics calculated using 5-fold 
-cross validation. Median NSE values were 0.34 (Lavaca River) and 
-0.48 (Navidad River) for NO~3~ loads and 0.81 (Lavaca River) and 0.91 (Navidad 
-River) for TP loads. Median r^2^ values were 0.70
-(Lavaca River) and 0.87 (Navidad River) for NO~3~ loads and 0.93 (Lavaca River) 
-and 0.99 (Navidad River) for TP loads. Median PBIAS values were 2.00
-(Lavaca River) and 10.90 (Navidad River) for NO~3~ loads and -7.20 
-(Lavaca River) and -3.30 (Navidad River) for TP loads. Density plots of
+Using evaluation criteria recommended by @moriasiHydrologicWaterQuality2015, 
+predictive performance of nutrient loading GAMs ranged from "satisfactory" to
+"very good" based on median NSE, *r*, and PBIAS metrics calculated using 5-fold 
+cross validation. Median goodness-of-fit metrics for NO~3~ models in the Lavaca 
+River were 0.34 NSE, 0.70 *r*, and 2.00 PBIAS. Navidad River NO~3~ models 
+appeared to perform slightly better with 0.48 NSE and 0.87 *r* but with higher 
+bias at 10.90 PBIAS. Generally, TP models performed better than NO~3~ models. 
+Median goodness-of-fit metrics for TP in the Lavaca River were 0.81 NSE, 
+0.93 *r*, and -7.20 PBIAS. Navidad River TP models has similar performance 
+with 0.91 NSE, 0.99 *r*, and -3.30 PBIAS. Density plots of
 metrics show similar distribution of values between sites for the 
-same parameter, with the exception r^2^ values for NO~3~ loads where 
+same parameter, with the exception *r* values for NO~3~ loads where 
 Lavaca River had a much larger variance in values compared to the Navidad River
-(@fig-fig2). TP GAMS had higher average NSE and r^2^ values and 
+(@fig-fig2). TP GAMS had higher average NSE and *r* values and 
 lower variance in metric values compared to NO~3~.
@@
 #| message: false
-#| fig-width: 5.2
+#| fig-width: 6.85
 #| fig-asp: 0.65
-#| out-width: "5.2in"
-#| fig-cap: "Density plots of goodness-of-fit metrics (NSE, r^2^, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
+#| out-width: "6.85in"
+#| fig-cap: "Density plots of goodness-of-fit metrics (NSE, *r*, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
 Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics."
@@
                     )) +
-  labs(x = "r<sup>2</sup>", y = "") +
+  labs(x = "*r*", y = "") +
   theme_TWRI_print(base_family = "Arial") +
@@
 #| message: false
-#| fig-width: 5.2
+#| fig-width: 6.85
 #| fig-asp: 0.7
-#| out-width: "5.2in"
+#| out-width: "6.85in"
 #| fig-cap: "Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."
@@
 #| warning: false
-#| fig-width: 3
+#| fig-width: 3.3
 #| fig-asp: 2
 #| fig-cap: "Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers."
-#| out-width: "3in"
+#| out-width: "3.3in"
 load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
@@
 #| warning: false
-#| fig-width: 5.2
+#| fig-width: 6.85
 #| fig-asp: 0.7
 #| fig-cap: "Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs."
-#| out-width: "5.2in"
+#| out-width: "6.85in"
 draw_smoothed_years <- function(model1, model2, model3,
@@
 ```{r}
-#| label: tbl-table4
 #| message: false
 #| warning: false
-#| tbl-cap: "Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
-Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis."
 
@@
   mutate(Parameter = case_when(
-    Parameter == "Nitrite+Nitrate" ~ "NO~*x*~",
+    Parameter == "Nitrite+Nitrate" ~ "NOx",
     Parameter == "Chlorophyll-a" ~ "Chlorophyll-*a*",
@@
   )) |>
-  mutate(Parameter = fct_relevel(Parameter, "TP", "NO~*x*~", "Chlorophyll-*a*","TKN", "DO"),
+  mutate(Parameter = fct_relevel(Parameter, "TP", "NOx", "Chlorophyll-*a*","TKN", "DO"),
          Model = fct_relevel(Model, "Temporal", "Flow", "Flow + Load")) |>
@@
   mutate(model_prob = case_when(
-    model_prob == "NA (NA)" ~ "-",
+    model_prob == "NA (NA)" ~ "NA",
     model_prob != "NA (NA)" ~ model_prob
@@
   arrange(Parameter, Site) |>
-  kbl(format = "pipe")
+  gt() |> 
+  fmt_markdown(columns = starts_with("Par"), rows = starts_with("Chl"), md_engine = "markdown") |>
+  tab_header(title = gt::md("Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
+Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis."))
+  #kbl(format = "pipe")
+  # flextable() |>   
+  # set_caption(caption = "Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
+Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.",
+  #             autonum = run_autonum(seq_id = "tab", bkm = "table4")) |> 
+  # style(pr_t = fp_text_default(font.size = 9), part = "body") |>
+  # style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
+  # set_table_properties(layout = "autofit") |>
+  # padding(padding.top = 0, padding.bottom = 0, part = "body")
 
@@
 
-GAMs showed increases in freshwater inflow resulted in nearly linear increases 
+GAMs showed that increases in freshwater inflow resulted in nearly linear increases 
 in TP and NO~*x*~ concentration at all three sites (@fig-fig6). At the upper-bay
-TCEQ-13563 site, GAMs showed increases in freshwater inflow initially 
+TCEQ-13563 site, GAMs showed that increases in freshwater inflow initially 
 increased chlorophyll-*a* and DO concentration but concentrations leveled and
@@
 #| warning: false
-#| fig-width: 5.2
+#| fig-width: 6.85
 #| fig-asp: 0.7
-#| out-width: "5.2in"
+#| out-width: "6.85in"
 #| fig-cap: "Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."
@@
 #| warning: false
-#| fig-width: 5.2
+#| fig-width: 6.85
 #| fig-asp: 0.45
-#| out-width: "5.2in"
+#| out-width: "6.85in"
 #| fig-cap: "Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs."
@@
 
+**resume here**
+
 TP and NO~3~ loadings from the Lavaca Bay watershed showed high inter-annual
@@
 producer participation in the watershed have been ongoing since 2016
-[@schramm_lavaca_2018; @bertholdDirectMailingEducation2021]. The decrease in 
-flow-normalized NO~3~ loads could be a reflection of those collective efforts 
-but further data collection and research is required to support that statement.
+[@schrammLavacaRiverWatershed2018; @bertholdDirectMailingEducation2021]. The 
+decrease in flow-normalized NO~3~ loads could be a reflection of those 
+collective efforts but further data collection and research is required to 
+support that statement.
 
@@
 Lavaca River are within the ranges in previous published studies
-(@tbl-table5; [@dunnTrendsNutrientInflows1996; @rebich_sources_2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]). It isn't obvious
+(@tbl-table5; [@dunnTrendsNutrientInflows1996; @rebichSourcesDeliveryNutrients2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]). It isn't obvious
 why TP estimates in @dunnTrendsNutrientInflows1996 were notably lower. Given that none
@@
 the period used in @dunnTrendsNutrientInflows1996 was drier on average than the
-other studies. The SPARROW models used in @rebich_sources_2011 and 
+other studies. The SPARROW models used in @rebichSourcesDeliveryNutrients2011 and 
 @wise_spatially_2019 utilize a version of LOADEST in the underlying load
@@
 
-load_lav_tp <- tar_read(daily_tp_08164000, store = store)
-load_nav_tp <- tar_read(daily_tp_texana, store = store)
-
-dat <- load_lav_tp$annually |> 
-  ungroup() |> 
-  filter(year >= 2005) |> 
-  summarise(annual = round(mean(TP_Estimate)/2116,1),
-            upper = round(mean(TP_Upper)/2116,1),
-            lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km
-
-df <- tibble(
-  Parameter = rep("TP", 5),
-  "Reported Yield (kg ⋅ km^-2^ ⋅ year^-1^)" = 
-    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")^a^"), 
-      "45.2", "42", "20.81-91.58^b^", "28.9"),
-  Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
-  `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
-  Reference = c("This work", "@wise_spatially_2019",
-                "@omaniEstimationSedimentNutrient2014",
-                "@rebich_sources_2011",
-                "@dunnTrendsNutrientInflows1996")
-)
-
-x <- kable(df, format = "pipe")
-
-cat(x, 
-    "^a^Mean of the annual point estimates and the lower and upper 95% credible intervals.", 
-    "^b^Represents a binned value range from a choropleth map.",
-    sep = "\n")
+# load_lav_tp <- tar_read(daily_tp_08164000, store = store)
+# load_nav_tp <- tar_read(daily_tp_texana, store = store)
+# 
+# dat <- load_lav_tp$annually |> 
+#   ungroup() |> 
+#   filter(year >= 2005) |> 
+#   summarise(annual = round(mean(TP_Estimate)/2116,1),
+#             upper = round(mean(TP_Upper)/2116,1),
+#             lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km
+# 
+# df <- tibble(
+#   Parameter = rep("TP", 5),
+#   "Reported Yield (kg ⋅ km^-2^ ⋅ year^-1^)" = 
+#     c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")^a^"), 
+#       "45.2", "42", "20.81-91.58^b^", "28.9"),
+#   Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
+#   `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
+#   Reference = c("This work", "@wise_spatially_2019",
+#                 "@omaniEstimationSedimentNutrient2014",
+#                 "@rebichSourcesDeliveryNutrients2011",
+#                 "@dunnTrendsNutrientInflows1996")
+# )
+# 
+# 
+# x <- kable(df, format = "pipe")
+# 
+# cat(x, 
+#     "^a^Mean of the annual point estimates and the lower and upper 95% credible intervals.", 
+#     "^b^Represents a binned value range from a choropleth map.",
+#     sep = "\n")
 
@@
 @caffreyHighNutrientPulses2007; @peierlsNonmonotonicResponsesPhytoplankton2012; 
-@palmerImpactsDroughtsLow2015; @ciraPhytoplanktonDynamicsLowInflow2021]. Inflow 
+@palmerImpactsDroughtsLow2015; @ciraPhytoplanktonDynamicsLowinflow2021]. Inflow 
 had a non-linear relationship with TKN at the two upstream sites, with TKN 
@@
 lower reaches of Lavaca Bay. Previous work suggests the processing of 
-organic loads in the upper portions of Lavaca Bay mgiht reduce the transport of 
+organic loads in the upper portions of Lavaca Bay might reduce the transport of 
 nutrients into the lower reaches of the Bay [@russell_effect_2006].
@@
 Texas estuaries [@ornolfsdottirNutrientPulsingRegulator2004; 
-@doradoUnderstandingInteractionsFreshwater2015; @ciraPhytoplanktonDynamicsLowInflow2021; @tominackVariabilityPhytoplanktonBiomass2022], 
+@doradoUnderstandingInteractionsFreshwater2015; @ciraPhytoplanktonDynamicsLowinflow2021; @tominackVariabilityPhytoplanktonBiomass2022], 
 this study specifically used flow-adjusted freshwater derived nutrient loads to 
@@
 water quality parameters may have lagged effects lasting days or even months following
-storms and large discharge events [@mooney_watershed_2012; @wetzExtremeFutureEstuaries2013; @bukaveckas_influence_2020; @walkerTimescalesMagnitudeWater2021]. 
+storms and large discharge events [@mooneyWatershedExportEvents2012a; @wetzExtremeFutureEstuaries2013; @bukaveckasInfluenceStormEvents2020; @walkerTimescalesMagnitudeWater2021]. 
 However, our work only evaluates responses to
@@
 # Acknowledgements
-
 The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
-Texas A&M Corpus Christi), Chad Kinsfather and Patrick Brzozowski 
+Texas A&M Corpus Christi), Chad Kinsfather and Partick Brzozowski 
 (Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water 
@@
 (Texas General Land Office) and the Lavaca Bay Foundation for supporting 
-development of this project and providing valuable project feedback.
+development of this project and providing valuable feedback.
+
+# Funding
+
+This project was funded in part by a Texas Coastal Management Program grant 
+approved by the Texas Land Commissioner, providing financial assistance under 
+the Coastal Zone Management Act of 1972, as amended, awarded by the National 
+Oceanic and Atmospheric Administration (NOAA), Office for Coastal Management,
+pursuant to NOAA Award No. NA21NOS4190136. The views expressed herein are those 
+of the author(s) and do not necessarily reflect the views of NOAA, the U.S. 
+Department of Commerce, or any of their subagencies.
+
+# Data Availability
+
+Reproducible code and datasets generated during this study are available in
+the Zenodo repository, https://doi.org/10.5281/zenodo.733075.
 

### 2023-05-20 Schramm-2023-05-AS/Schramm-2023-05-AS.Rmd
init springer nature tex submission for AS

--- 2023-05-20 Schramm-2023-04-AS/Lavaca-Manuscript-AS.qmd
@@
 ---
-title: "Assessing linkages between watershed nutrient loading and water quality in a subtropical estuary with semiparametric models."
-format:
-  wordcount-docx:
-    toc: false
-    number-sections: false
-    reference-doc: format.docx
-author: 
-  - name: Michael Schramm
+# Supported options: 
+#   sn-nature:       Style for submissions to Nature Portfolio journals
+#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
+#   sn-mathphys:     Math and Physical Sciences Reference Style
+#   sn-aps:          American Physical Society (APS) Reference Style
+#   sn-vancouver:    Vancouver Reference Style
+#   sn-apa:          APA Reference Style 
+#   sn-chicago:      Chicago-based Humanities Reference Style
+#   default:         Default
+
+classoptions: 
+  - sn-basic      
+  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
+  - referee       # Optional: Use double line spacing 
+  - lineno        # Optional: Add line numbers
+  # - iicol         # Optional: Double column layour
+
+title: Assessing linkages between watershed nutrient loading and water quality in a subtropical estuary with semiparametric models.
+titlerunning: Watershed nutrient loading and water quality in a subtropical estuary
+
+authors: 
+  - firstname: Michael
+    lastname: Schramm
     email: michael.schramm@ag.tamu.edu
-    corresponding: true
+    affiliation: [1]
+    corresponding: TRUE
+
 affiliations:
-      - ref: agrilife
-affiliations:
-  - id: agrilife
-    name: Texas A&M AgriLife Research, Texas Water Resources Institute
+  - number: 1
+    corresponding: TRUE
+    info:
+      orgdiv: Texas Water Resources Institute
+      orgname: Texas A&M AgriLife Research
+    address:
+        street: 1001 Holleman Dr. E.
         city: College Station
-    state: Texas
-    country: United States of America
-crossref:
-  fig-prefix: Fig   # (default is "Figure")
-  tbl-prefix: Table    # (default is "Table")
-  title-delim: .     # (default is ":")
-  ref-hyperlink: false # (default is true)
-  eq-prefix: Eq
-bibliography: mybib.yaml
-csl: https://www.zotero.org/styles/springer-basic-author-date-no-et-al?source=1
-filters: 
-  - authors-affiliations.lua
-keep-md: true
-cite-method: citeproc
-link-bibliography: false
+        postcode: 77840
+        state: TX
+        country: United States
+
+keywords:
+  - key
+  - dictionary
+  - word
+
+# Sample for unstructured abstract 
+abstract: |
+  Lavaca Bay is a small secondary embayment on the Texas coast that is displaying early signals of water quality degradation.
+This study applied a semiparametric approach to assess both watershed nutrient loads and the responses in estuary water quality to nutrient loading and streamflow.
+Cross-validation indicated that, despite data constraints, semiparametric models performed well at nutrient load prediction.
+Based on these models, delivered annual nutrient loads varied substantially from year to year.
+In contrast, minimal changes in calculated flow-normalized loads indicate that changes in nutrient loading were driven by natural variation in precipitation and runoff as opposed to changes in nonpoint sources.
+Estuary water quality models did not identify significant long-term changes within Lavaca Bay for dissolved oxygen or chlorophyll-*a*.
+However, site specific long-term increases in both organic and inorganic nitrogen are concerning.
+Further analysis found freshwater inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that changes in nutrient loading explained variation in dissolved oxygen or chlorophyll-*a* concentrations.
+In addition to providing a baseline assessment of watershed nutrient loading and water quality responses in the Lavaca Bay watershed, this study provides methodological support for the use of semiparametric methods in load regression models and estuary assessments.
+
+bibliography: bibliography.bib
+output: rticles::springer_article
 ---
 
-```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
+```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(echo = FALSE,
+                      out.extra = "")  # This forces knitr to label all figures.
+
 library(targets)
@@
 library(ragg)
-library(flextable)
 library(modelsummary)
@@
 library(gt)
-library(officer)
-
-ragg_tiff <- function(..., res = 300) {
-  ragg::agg_tiff(..., res = res, units = "in")
-}
 
@@
                       ## figure chunk options
-                      dev = "ragg_tiff",
+                      dev = "ragg_png",
                       dpi = 300,
                       fig.align = "center",
-                      fig.ext = "tiff",
                       out.width = "100%")
@@
 
-# Abstract
-Lavaca Bay is a small secondary embayment on the Texas coast that is
-displaying early signals of water quality degradation. This
-study applied a semiparametric approach to assess both watershed nutrient loads 
-and the responses in estuary water quality to nutrient loading and streamflow. 
-Cross-validation indicated that, despite data constraints, semiparametric models 
-performed well at nutrient load prediction. Based on these models, delivered
-annual nutrient loads varied substantially from year to year. In contrast,
-minimal changes in calculated flow-normalized loads indicate that changes in 
-nutrient loading were driven by natural variation in precipitation and runoff as
-opposed to changes in nonpoint sources.
-Estuary water quality models did not identify significant long-term changes 
-within Lavaca Bay for dissolved oxygen or chlorophyll-*a*. However, site specific 
-long-term increases in both organic and inorganic nitrogen are concerning. 
-Further analysis found freshwater inflow was a strong driver of nutrient and 
-chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that 
-changes in nutrient loading explained variation in dissolved oxygen or 
-chlorophyll-*a* concentrations. In addition to providing a baseline assessment 
-of watershed nutrient loading and water quality responses in the Lavaca Bay 
-watershed, this study provides methodological support for the use of 
-semiparametric methods in load regression models and estuary assessments.
-
-# Introduction
+# Introduction {#sec1}
 
@@
 
-# Materials and Methods
-
-## Study Area and Data
-
+# Methods {#sec2}
+
+## Location and Data
 Lavaca Bay is 190 km^2^ with the majority of freshwater inflow provided by the 
-Lavaca and Navidad River systems (@fig-fig1). The Garcitas-Arenosa, 
+Lavaca and Navidad River systems (Figure\ \ref{fig:fig1}). The Garcitas-Arenosa, 
 Placedo Creek, and Cox Bay watersheds provide additional freshwater inflows. 
@@
 
-```{r}
-#| label: fig-fig1
-#| fig-width: 5.2
-#| fig-height: 6.75
-#| out-width: "5.2in"
-#| fig-cap: "Map of Lavaca Bay and the contribution watershed.
+```{r fig1, fig.width=6.75, fig.width=5.2, out.width="5.2in", fig.cap="Map of Lavaca Bay and the contribution watershed.
 The freshwater sites are the most downstream freshwater stream locations with water quality and streamflow data used for nutrient load models.
-Water quality concentration data at the three Lavaca Bay sites were used to assess relationships between freshwater flows, loads and estuary water quality."
-
-knitr::include_graphics("Figures/fig1-1.tiff")
+Water quality concentration data at the three Lavaca Bay sites were used to assess relationships between freshwater flows, loads and estuary water quality."}
+
+knitr::include_graphics("../fig1.png")
 
@@
 
-Daily discharges for the Lavaca River (USGS-08164000, @fig-fig1) 
+Daily discharges for the Lavaca River (USGS-08164000, Figure\ \ref{fig:fig1}) 
 were obtained from the United States Geologic Survey (USGS) National Water 
@@
 nitrate-nitrogen (NO~3~) data from January 2005 through December 2020 were 
-downloaded (`#r run_reference("table1")`). Less than 5-years of total
+downloaded (Table\ \ref{tab:table1}). Less than 5-years of total
 nitrogen and TKN concentration data were available at the freshwater sites and
@@
 chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
-(@tbl-table2).
-
-```{r}
+(Table\ \ref{tab:table2}).
+
+```{r table1}
 
@@
                Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-            data = df, output = "kableExtra", format = "simple")
-
-# datasummary(`Station ID` * (`TP (mg/L)` + `NO3 (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
-#                Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-#             data = df, output = "flextable") |> 
-#   set_caption(caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.",
-#               autonum = run_autonum(seq_id = "tab", bkm = "table1")) |> 
-#   style(pr_t = fp_text_default(font.size = 9), part = "body") |>
-#   style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
-#   set_table_properties(layout = "autofit") |>
-#   padding(padding.top = 0, padding.bottom = 0, part = "body")
+            data = df, output = "kableExtra", format = "latex",
+            caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.")
+
 ```
 
-```{r}
+```{r table2}
 #| warning: false
@@
                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-                          data = df, output = "flextable") |> 
-  set_caption(caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.",
-              autonum = run_autonum(seq_id = "tab", bkm = "table2")) |> 
-  style(pr_t = fp_text_default(font.size = 9), part = "body") |>
-  style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
-  set_table_properties(layout = "autofit") |>
-  padding(padding.top = 0, padding.bottom = 0, part = "body")
-
- # modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
- #                            `NOx (mg/L)` +
- #                            `TKN (mg/L)` +
- #                            `Chlorophyll-a (ug/L)` +
- #                            `DO (mg/L)`) ~
- #                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
- #                          data = df, output = "kableExtra", format = "pipe") #|>
-  #  flextable::autofit()
-   # mutate(`Station ID` = case_when(
-   #   `Station ID` == "" ~ NA_character_,
-   #   `Station ID` != "" ~ `Station ID`
-   #   )) |>
-   # tidyr::fill(`Station ID`, .direction = "down") |> 
-   # gt::gt() |> 
-   # gt::as_word()
-
+                         data = df, output = "kableExtra", format = "latex",
+                         caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.")
 ```
@@
 tend, season, streamflow, and two different antecedent discharge terms, 
-shown in [@eq-eq1]:
-
-$$
-\begin{align*}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  \\
+shown in Eq. \ref{eq1}:
+
+\begin{align}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa), \nonumber \\
 y &\sim \mathcal{N}(\mu,\,\sigma^{2})
-\end{align*}
-$${#eq-eq1}
-
-where *μ* is the conditional expected NO~3~ or TP concentration, *g()* is the 
-log-link, *α* is the intercept, *f~n~()* are smoothing functions. *y* is the 
+\label{eq1}
+\end{align}
+
+where $\mu$ is the conditional expected NO~3~ or TP concentration, $g()$ is the 
+log-link, $\alpha$ is the intercept, $f_n()$ are smoothing functions. $y$ is the 
 response variable (NO~3~ or TP concentration) modeled as normally distributed 
-with mean *μ* and standard deviation *σ*. *ddate* is the date converted to 
+with mean $\mu$ and standard deviation $\sigma$. *ddate* is the date converted to 
 decimal notation, *yday* is numeric day of year (1-366), and *log1p(Q)* is the 
@@
 current time period [@wangLoadEstimationUncertainties2011;  
-@kuhnertQuantifyingTotalSuspended2012; @zhang_improving_2017], using [@eq-eq2]:
-
-$$
-\begin{align*}
-ma(\delta) &= d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1},\\
-\kappa_{i} &=\sum_{m=1}^{i}\hat{Q}_m,
-\end{align*}
-$${#eq-eq2}
-
-where *δ* is the discount factor (here, set equal to 0.95), κ~i~
+@kuhnertQuantifyingTotalSuspended2012; @zhang_improving_2017], using Eq. \ref{eq2}:
+
+\begin{align}
+ma(\delta) &= d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}, \nonumber \\
+\kappa_{i} &=\sum_{m=1}^{i}\hat{Q}_m
+\label{eq2}
+\end{align}
+
+where $\delta$ is the discount factor (here, set equal to 0.95), $\kappa_i$
 is the cumulative flow (*Q*) up to the *i*th day.
@@
 the previous year relative to the entire period [@zhang_improving_2017] and 
-calculated using [@eq-eq3]:
-
-$$
+calculated using Eq. \ref{eq3}:
+
+\begin{equation}
 ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}
-$${#eq-eq3}
+\label{eq3}
+\end{equation}
 
 and the short-term flow anomaly (*stfa*) calculated as the current day flow 
-compared to the preceding 1-month streamflow using [@eq-eq4]:
-
-$$
+compared to the preceding 1-month streamflow using \ref{eq4}:
+
+\begin{equation}
 stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t)
-$${#eq-eq4}
-
-where *x* are the averages of log-transformed streamflow over the antecedent 
+\label{eq4}
+\end{equation}
+
+where $x$ are the averages of log-transformed streamflow over the antecedent 
 period (*1-year*, *1-month*, etc.) for time *t*.  We used *ltfa* in NO~3~ models 
@@
 nutrient loads (@eq-eq7):
-
-$$
-\begin{align*}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday)\\
-y &\sim \Gamma(\mu,\lambda)
-\end{align*}
-$${#eq-eq5}
-
-$$
-\begin{align*}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) \\
-y &\sim \Gamma(\mu,\lambda)
-\end{align*}
-$${#eq-eq6}
-
-$$
-\begin{align*}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \\
-y &\sim \Gamma(\mu,\lambda)
-\end{align*}
-$${#eq-eq7}
-
-where *μ* is the conditional expected response (nutrient concentration), *g()* 
+\begin{align}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday), \label{eq5} \\
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q), \label{eq6} \\
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \label{eq7}
+\end{align}
+
+where $\mu$ is the conditional expected response (nutrient concentration), $g()$ 
 is the log link, and response variable was modeled as Gamma distributed with 
-mean *μ* and scale *λ*. *f~1~(ddate)* is decimal date smoothed with a 
-thin-plate regression spline, *f~2~(yday)* is the numeric day of year smoothed 
-with a cyclic cubic regression spline, *f~3~(Q)* is mean daily inflow (the 
-combined measurements from Lavaca River and Navidad River) and *f~4~(Load)* is 
+mean $\mu$ and scale $\lambda$. $f_1(ddate)$ is decimal date smoothed with a 
+thin-plate regression spline, $f_2(yday)$ is the numeric day of year smoothed 
+with a cyclic cubic regression spline, $f_3(Q)$ is mean daily inflow (the 
+combined measurements from Lavaca River and Navidad River) and $f_4(Load)$ is 
 the total NO~3~ or TP watershed load. The set of models specified for each water
-quality response are in @tbl-table3.
-
-```{r}
+quality response are in Table\ \ref{tab:table3}.
+
+```{r table3}
 #| warning: false
 df <- tibble(
-  `Water Quality Response Parameter` = c(rep("TP", 3), rep("NOx", 3),
-               rep("Chlorophyll-a", 3), rep("Dissolved Oxygen", 3),
+  `Parameter` = c(rep("TP", 3), rep("NO\\textsubscript{x}", 3),
+               rep("Chlorophyll-\\textit{a}", 3), rep("Dissolved Oxygen", 3),
                rep("TKN", 2)),
@@
                     "s(ddate) + s(yday) + s(Q)",
-                    "s(ddate) + s(yday) + s(Q) + s(NO3 Load)",
+                    "s(ddate) + s(yday) + s(Q) + s(NO\\textsubscript{3} Load)",
                     "s(ddate) + s(yday)",
                     "s(ddate) + s(yday) + s(Q)",
-                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO3 Load)",
+                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3} Load)",
                     "s(ddate) + s(yday)",
                     "s(ddate) + s(yday) + s(Q)",
-                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO3  Load)",
+                    "s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\\textsubscript{3}  Load)",
                     "s(ddate) + s(yday)",
@@
 
-# kbl(df, format = "pipe")
-
-flextable::flextable(df) |> 
-  set_caption(caption = "Set of GAM models specified for each water quality parameter response.",
-              autonum = run_autonum(seq_id = "tab", bkm = "table3")) |> 
-  style(pr_t = fp_text_default(font.size = 9), part = "body") |>
-  style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
-  set_table_properties(layout = "autofit") |>
-  padding(padding.top = 0, padding.bottom = 0, part = "body")
+kbl(df, format = "latex",
+    caption = "Set of GAM models specified for each water quality parameter response.",
+    booktabs = TRUE,
+    escape = FALSE) |> 
+  collapse_rows(columns = 1, latex_hline = "major")
+
+# flextable::flextable(df) |> 
+#   set_caption(caption = "Set of GAM models specified for each water quality parameter response.",
+#               autonum = run_autonum(seq_id = "tab", bkm = "table3")) |> 
+#   style(pr_t = fp_text_default(font.size = 9), part = "body") |>
+#   style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
+#   set_table_properties(layout = "autofit") |>
+#   padding(padding.top = 0, padding.bottom = 0, part = "body")
 
@@
 inflow.Response residuals from the respective GAM models were 
-used as *Q* and *Load* in @eq-eq6 and @eq-eq7.
+used as *Q* and *Load* in Eq.\ \ref{eq6} and Eq.\ \ref{eq7}.
 
@@
 
-# Results
+# Results {#sec3}
 
@@
 Lavaca River had a much larger variance in values compared to the Navidad River
-(@fig-fig2). TP GAMS had higher average NSE and *r* values and 
+(Figure\ \ref{fig:fig2}). TP GAMS had higher average NSE and *r* values and 
 lower variance in metric values compared to NO~3~.
 
-```{r}
-#| label: fig-fig2
-#| warning: false
-#| message: false
-#| fig-width: 6.85
-#| fig-asp: 0.65
-#| out-width: "6.85in"
-#| fig-cap: "Density plots of goodness-of-fit metrics (NSE, *r*, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
-Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics."
+```{r fig2, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.65, fig.cap="Density plots of goodness-of-fit metrics (NSE, \\textit{r}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
+Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics."}
 
@@
 Predicted annual NO~3~ and TP loads show considerable variation, generally following
-patterns in discharge (@fig-fig3; @fig-fig4). 
+patterns in discharge (Figures\ \ref{fig:fig3}, \ref{fig:fig4}). 
 Flow-normalized TP loads at both sites and flow-normalized NO~3~ loads in the Lavaca River
 indicated watershed based loads did not change much over time
-when accounting for variation driven by streamflow (@fig-fig3). 
+when accounting for variation driven by streamflow (Figure\ \ref{fig:fig3}). 
 Flow-normalized loads in the Lavaca River showed small variation over time with 
@@
 
-```{r}
-#| label: fig-fig3
-#| error: false
-#| warning: false
-#| message: false
-#| fig-width: 6.85
-#| fig-asp: 0.7
-#| out-width: "6.85in"
-#| fig-cap: "Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."
+```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."}
+
 load <- tar_read(daily_no3_08164000, store = store)
@@
 nutrient loading in the watershed with the Navidad River only accounting for 
-`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 (@fig-fig4).
-
-```{r}
-#| label: fig-fig4
-#| error: false
-#| message: false
-#| warning: false
-#| fig-width: 3.3
-#| fig-asp: 2
-#| fig-cap: "Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers."
-#| out-width: "3.3in"
+`r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 
+(Figure\ \ref{tab:fig4}).
+
+```{r fig4, error=FALSE, message=FALSE, warning=FALSE, fig.width=3.3, fig.asp=2, fig.cap="Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers.", out.width="0.5\\linewidth"}
+
 load_lav_no3 <- tar_read(daily_no3_08164000, store = store)
@@
 GAMs did not identify significant changes in TP or DO concentrations at any
-of the Lavaca Bay sites from 2005 through 2020 (@fig-fig5). The
+of the Lavaca Bay sites from 2005 through 2020 (Figure\ \ref{fig:fig5}). The
 upper-bay site, TCEQ-13563, had a linear increase in NO~x~ concentration and
@@
 
-```{r}
-#| label: fig-fig5
-#| message: false
-#| warning: false
-#| fig-width: 6.85
-#| fig-asp: 0.7
-#| fig-cap: "Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs."
-#| out-width: "6.85in"
+```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs."}
+
 draw_smoothed_years <- function(model1, model2, model3,
@@
 NO~*x*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
-model probability values (@tbl-table4).
+model probability values (Table\ \ref{tab:table4}).
 TCEQ-13563, the site closest to the river outlet, was the only site that
@@
 
-```{r}
-#| message: false
-#| warning: false
+```{r table4, message=FALSE, warning=FALSE}
 
@@
   arrange(Parameter, Site) |>
-  gt() |> 
-  fmt_markdown(columns = starts_with("Par"), rows = starts_with("Chl"), md_engine = "markdown") |>
-  tab_header(title = gt::md("Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
-Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis."))
-  #kbl(format = "pipe")
+  kbl(format = "latex",
+      caption = "Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
+Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.",
+      booktabs = TRUE)
   # flextable() |>   
@@
 GAMs showed that increases in freshwater inflow resulted in nearly linear increases 
-in TP and NO~*x*~ concentration at all three sites (@fig-fig6). At the upper-bay
-TCEQ-13563 site, GAMs showed that increases in freshwater inflow initially 
+in TP and NO~*x*~ concentration at all three sites (Figure\ \ref{fig:fig6}). 
+At the upper-bay TCEQ-13563 site, GAMs showed that increases in freshwater inflow initially 
 increased chlorophyll-*a* and DO concentration but concentrations leveled and
@@
 
-```{r}
-#| label: fig-fig6
-#| message: false
-#| warning: false
-#| fig-width: 6.85
-#| fig-asp: 0.7
-#| out-width: "6.85in"
-#| fig-cap: "Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."
+```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."}
+
 draw_smoothed_surface <- function(comp,
@@
 Increased TP loads resulted in nearly linear increases of TP concentration at
-the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively (@fig-fig7).
+the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively 
+(Figure\ \ref{fig:fig7}).
 The relative effect size appeared to much smaller than the effect of freshwater
@@
 
-```{r}
-#| label: fig-fig7
-#| message: false
-#| warning: false
-#| fig-width: 6.85
-#| fig-asp: 0.45
-#| out-width: "6.85in"
-#| fig-cap: "Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs."
+```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.45, fig.cap="Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs."}
 
@@
 
-```{r}
-#| label: tbl-table5
-#| warning: false
-#| message: false
-#| tbl-cap: "Comparisons of previously published estimates of mean annual TP yield at the Lavaca River site."
-#| results: asis
-
-# load_lav_tp <- tar_read(daily_tp_08164000, store = store)
-# load_nav_tp <- tar_read(daily_tp_texana, store = store)
-# 
-# dat <- load_lav_tp$annually |> 
-#   ungroup() |> 
-#   filter(year >= 2005) |> 
-#   summarise(annual = round(mean(TP_Estimate)/2116,1),
-#             upper = round(mean(TP_Upper)/2116,1),
-#             lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km
-# 
-# df <- tibble(
+```{r table5, warning=FALSE, message=FALSE}
+
+load_lav_tp <- tar_read(daily_tp_08164000, store = store)
+load_nav_tp <- tar_read(daily_tp_texana, store = store)
+
+dat <- load_lav_tp$annually |>
+  ungroup() |>
+  filter(year >= 2005) |>
+  summarise(annual = round(mean(TP_Estimate)/2116,1),
+            upper = round(mean(TP_Upper)/2116,1),
+            lower = round(mean(TP_Lower)/2116,1)) ## 2116 sq km
+
+df <- tibble(
  # Parameter = rep("TP", 5),
-#   "Reported Yield (kg ⋅ km^-2^ ⋅ year^-1^)" = 
-#     c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")^a^"), 
-#       "45.2", "42", "20.81-91.58^b^", "28.9"),
-#   Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
-#   `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
-#   Reference = c("This work", "@wise_spatially_2019",
-#                 "@omaniEstimationSedimentNutrient2014",
-#                 "@rebichSourcesDeliveryNutrients2011",
-#                 "@dunnTrendsNutrientInflows1996")
-# )
-# 
-# 
-# x <- kable(df, format = "pipe")
-# 
-# cat(x, 
-#     "^a^Mean of the annual point estimates and the lower and upper 95% credible intervals.", 
-#     "^b^Represents a binned value range from a choropleth map.",
-#     sep = "\n")
+  "Reported Yield (kg$\\cdot$km\\textsuperscript{2}$\\cdot$year\\textsuperscript{-1})" =
+    c(paste0(dat[[1,1]]," (", dat[[1,3]], ", ", dat[[1,2]], ")\\textsuperscript{a}"),
+      "45.2", "42", "20.81-91.58\\textsuperscript{b}", "28.9"),
+  Approach = c("GAM", "SPARROW", "SWAT", "SPARROW", "LOADEST"),
+  `Time Period` = c("2005-2020", "2000-2014", "1977-2005", "1980-2002", "1972-1993"),
+  Reference = c("This work", "\\cite{wise_spatially_2019}",
+                "\\cite{omaniEstimationSedimentNutrient2014}",
+                "\\cite{rebichSourcesDeliveryNutrients2011}",
+                "\\cite{dunnTrendsNutrientInflows1996}")
+)
+
+kbl(df, format = "latex", booktabs = TRUE, escape = FALSE,
+    caption = "Comparisons of previously published estimates of mean annual TP yield at the Lavaca River site.") |>
+  footnote(alphabet = c("Mean of the annual point estimates and the lower and upper 95% credible intervals.",
+                        "Represents a binned value range from a choropleth map."))
 
@@
 
-# Acknowledgements
+\backmatter
+
+\bmhead{Acknowledgments}
 The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
@@
 
-# Funding
-
+# Declarations {.unnumbered}
+
+\bmhead{Funding}
 This project was funded in part by a Texas Coastal Management Program grant 
@@
 
-# Data Availability
-
+\bmhead{Data Availability}
 Reproducible code and datasets generated during this study are available in
 the Zenodo repository, https://doi.org/10.5281/zenodo.733075.
-
-# References

### 2023-05-20 Schramm-2023-05-AS/Schramm-2023-05-AS.tex
init springer nature tex submission for AS

--- 2023-02-11 Schramm-Manuscript-2023/Schramm-Manuscript-2023.tex
@@
-%  LaTeX support: latex@mdpi.com
-%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
-
-%=================================================================
-% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles
-\documentclass[water,article,submit,oneauthor]{Definitions/mdpi}
-
-% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
-The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
-This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
-For preprints.org, the editorial staff will make this change immediately prior to posting.
-
-%% Some pieces required from the pandoc template
-\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
-\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}
-
-%--------------------
-% Class Options:
-%--------------------
-%----------
-% journal
-%----------
-% Choose between the following MDPI journals:
-% acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing,\gdef\@continuouspages{yes}} nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, %%nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, %oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology,\gdef\@ISSN{2813-0618}\gdef\@continuous pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
-% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal. Changing "submit" to "accept" before posting will remove line numbers.
-
-%---------
-% article
-%---------
-% The default type of manuscript is "article", but can be replaced by:
-% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
-% supfile = supplementary materials
-
-%----------
-% submit
-%----------
-% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
-This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
-Also, line numbering will be removed.
-Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
-
-%------------------
-% moreauthors
-%------------------
-% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
-
-%---------
-% pdftex
-%---------
-% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.
-
-%=================================================================
-% MDPI internal commands - do not modify
-\firstpage{1}
-\makeatletter
-\setcounter{page}{\@firstpage}
-\makeatother
-\pubvolume{1}
-\issuenum{1}
-\articlenumber{0}
-\pubyear{2023}
-\copyrightyear{2023}
-%\externaleditor{Academic Editor: Firstname Lastname}
-\datereceived{ }
-\daterevised{ } % Comment out if no revised date
-\dateaccepted{ }
-\datepublished{ }
-%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
-%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
-\hreflink{https://doi.org/} % If needed use \linebreak
-%\doinum{}
-%\pdfoutput=1 % Uncommented for upload to arXiv.org
-
-%=================================================================
-% Add packages and commands here.
-The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
-% cleveref: load \crefname definitions after \begin{document}
-
-%=================================================================
-% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
-%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
-
-%=================================================================
-% Full title of the paper (Capitalized)
-\Title{Assessing linkages between watershed nutrient loading and estuary
-water quality in Lavaca Bay, Texas}
-
-% MDPI internal command: Title for citation in the left column
-\TitleCitation{Assessing linkages between watershed nutrient loading and
-estuary water quality in Lavaca Bay, Texas}
-
-% Author Orchid ID: enter ID or remove command
-%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
-%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-
-% Authors, for the paper (add full first names)
-\Author{Michael
-Schramm$^{1}$\href{https://orcid.org/0000-0003-1876-6592}
-{\orcidicon}}
-
-%\longauthorlist{yes}
-
-% MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Michael Schramm}
-
-% MDPI internal command: Authors, for citation in the left column
-%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
-% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
-\AuthorCitation{Schramm, M.}
-
-% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{%
-$^{1}$ \quad Texas A\&M AgriLife Research - Texas Water Resources
-Institute 1001 Holleman Dr.~E. College Station, TX
-77840-2118; \href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}\\
-}
-
-% Contact information of the corresponding author
-\corres{Correspondence: \href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}
-
-% Current address and/or shared authorship
-
-% The commands \thirdnote{} till \eighthnote{} are available for further notes
-
-% Simple summary
-
-%\conference{} % An extended version of a conference paper
-
-% Abstract (Do not insert blank lines, i.e. \\)
-\abstract{Lavaca Bay is a small secondary embayment on the Texas coast
-that is displaying early signals of water quality degradation. This
-study uses Generalized Additive Models to assesses both watershed
-nutrient loading and responses in estuary water quality to nutrient
-loading and streamflow. While delivered nutrient loads varied greatly,
-flow-normalized nutrient loads indicated limited evidence of changes in
-nutrient runoff in the watershed since 2005. As a consequence, there
-wasn't strong evidence for an effect of nutrient loading (controlled for
-flow) on dissolved oxygen or chlorophyll-\emph{a} concentrations. There
-was evidence total phosphorus (TP) loading had a strong positive effect
-on TP concentration in the Bay, particularly at sites closest to the
-river mouth. Seasonally-adjusted streamflow had strong effects on
-nutrient concentration and chlorophyll-\emph{a} concentrations at most
-sites. Although temporal trends did not identify significant estuary
-water quality degredation in dissolved oxygen or chlorophyll-\emph{a}
-concentrations, site specific trends in both organic and inorganic
-nitrogen are concerning for the long-term health of the estuary. This
-study provides a baseline assessment of the linkages between watershed
-nutrient loading and water quality responses in Lavaca Bay. Some of the
-data gaps highlighted in the study are being addressed in ongoing
-projects to provide improved information for resource managers.}
-
-% Keywords
-\keyword{estuary; nutrient loading; water quality; Texas}
-
-% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
-%\PACS{J0101}
-%\MSC{}
-%\JEL{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences
-%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data
-%\dataset{DOI number or link to the deposited data set if the data set is published separately.
-If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors.
-In this case, please submit the data set as a supplement.}
-%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Encyclopedia
-%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Advances in Respiratory Medicine
-%\addhighlights{yes}
-%\renewcommand{\addhighlights}{%
-
-%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars.
-Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it.
-Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
-%\textbf{What are the main findings?}
-% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
-% \item First bullet.
-% \item Second bullet.
-% \end{itemize}\vspace{3pt}
-%\textbf{What is the implication of the main finding?}
-% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
-% \item First bullet.
-% \item Second bullet.
-% \end{itemize}
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+%Version 2.1 April 2023
+% See section 11 of the User Manual for version history
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                                                 %%
+%% Please do not use \input{...} to include other tex files.       %%
+%% Submit your LaTeX manuscript as one .tex document.              %%
+%%                                                                 %%
+%% All additional figures and files should be attached             %%
+%% separately and not embedded in the \TeX\ document itself.       %%
+%%                                                                 %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\documentclass[sn-basic,referee,lineno,pdflatex]{sn-jnl}
+
+%%%% Standard Packages
+%%<additional latex packages if required can be included here>
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
+%%%%
+
+%%%%%=============================================================================%%%%
+%%%%  Remarks: This template is provided to aid authors with the preparation
+%%%%  of original research articles intended for submission to journals published
+%%%%  by Springer Nature. The guidance has been prepared in partnership with
+%%%%  production teams to conform to Springer Nature technical requirements.
+%%%%  Editorial and presentation requirements differ among journal portfolios and
+%%%%  research disciplines. You may find sections in this template are irrelevant
+%%%%  to your work and are empowered to omit any such section if allowed by the
+%%%%  journal you intend to submit to. The submission guidelines and policies
+%%%%  of the journal take precedence. A detailed User Manual is available in the
+%%%%  template package for technical guidance.
+%%%%%=============================================================================%%%%
 
@@
 
+\raggedbottom
+
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 \begin{document}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Like many coastal areas globally, coastal watersheds along the Texas
+\title[Watershed nutrient loading and water quality in a subtropical
+estuary]{Assessing linkages between watershed nutrient loading and water
+quality in a subtropical estuary with semiparametric models.}
+
+%%=============================================================%%
+%% Prefix	-> \pfx{Dr}
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% NatureName	-> \tanm{Poet Laureate} -> Title after name
+%% Degrees	-> \dgr{MSc, PhD}
+%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
+%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
+%%=============================================================%%
+
+\author*[1]{\fnm{Michael} \sur{Schramm} }\email{\href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}
+
+  \affil*[1]{\orgdiv{Texas Water Resources Institute}, \orgname{Texas
+A\&M AgriLife Research}, \orgaddress{\city{College
+Station}, \country{United
+States}, \postcode{77840}, \state{TX}, \street{1001 Holleman Dr.~E.}}}
+
+\abstract{Lavaca Bay is a small secondary embayment on the Texas coast
+that is displaying early signals of water quality degradation. This
+study applied a semiparametric approach to assess both watershed
+nutrient loads and the responses in estuary water quality to nutrient
+loading and streamflow. Cross-validation indicated that, despite data
+constraints, semiparametric models performed well at nutrient load
+prediction. Based on these models, delivered annual nutrient loads
+varied substantially from year to year. In contrast, minimal changes in
+calculated flow-normalized loads indicate that changes in nutrient
+loading were driven by natural variation in precipitation and runoff as
+opposed to changes in nonpoint sources. Estuary water quality models did
+not identify significant long-term changes within Lavaca Bay for
+dissolved oxygen or chlorophyll-\emph{a}. However, site specific
+long-term increases in both organic and inorganic nitrogen are
+concerning. Further analysis found freshwater inflow was a strong driver
+of nutrient and chlorophyll-\emph{a} concentrations within Lavaca Bay
+but there was no evidence that changes in nutrient loading explained
+variation in dissolved oxygen or chlorophyll-\emph{a} concentrations. In
+addition to providing a baseline assessment of watershed nutrient
+loading and water quality responses in the Lavaca Bay watershed, this
+study provides methodological support for the use of semiparametric
+methods in load regression models and estuary assessments.}
+
+\keywords{key, dictionary, word}
+
+\maketitle
+
+\hypertarget{sec1}{%
+\section{Introduction}\label{sec1}}
+
+Like many coastal areas globally, the coastal watersheds along the Texas
 Gulf coast are facing pressures from increasing population, increases in
 point source and non-point source pollution and alterations to
-freshwater flows that affect water quality in downstream estuaries
+freshwater flows that degrade water quality in downstream estuaries
 \citep{bricker_effects_2008, kennicuttWaterQualityGulf2017, bugica_water_2020}.
-Despite these increasing pressures, national scale assessments have
-classified coastal estuaries in Texas as moderate or lower for
-exhibiting eutrophic conditions \citep{bricker_effects_2008}. However, a
-suite of recent studies indicate that estuary water quality dynamics in
-both agriculturally dominated and urban watersheds within Texas are in
-fact expressing conditions that are increasingly conducive to algal
-blooms and eutrophication
+Despite these escalating pressures, national scale assessments have
+classified coastal estuaries in Texas as moderate or low risk for
+eutrophic conditions \citep{bricker_effects_2008}. However, a suite of
+recent studies indicates that estuary water quality dynamics in both
+agriculturally dominated and urban watersheds within Texas are in fact
+expressing conditions that are increasingly conducive to algal blooms
+and eutrophication
 \citep{wetzWaterQualityDynamics2016, wetz_exceptionally_2017, bugica_water_2020, chinPhytoplanktonBiomassCommunity2022}.
@@
 
-This project provides an assessment of nutrient loading and water
+This project aims to provide an assessment of nutrient loading and water
 quality responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in
 the larger Matagorda Bay system located roughly halfway between Houston,
-Texas and Corpus Christi, Texas. Lavaca Bay has faced substantial
-challenges associated with legacy contamination but general water
-quality parameters such as dissolved oxygen (DO), nutrients, and
-biological parameters have been well within state water quality
-standards. However, long-term declines of benthic fauna abundance,
-biomass, and diversity in Lavaca Bay primarily linked to reductions in
-freshwater inflows and changes in estuary salinity
+Texas and Corpus Christi, Texas. Lavaca Bay faces substantial challenges
+associated with legacy contamination but general water quality
+parameters such as dissolved oxygen (DO), nutrients, and biological
+parameters have been well within state water quality standards. More
+recently identified long-term declines in abundance, biomass, and
+diversity of benthic fauna in Lavaca Bay have been linked to reductions
+in freshwater inflows and changes in estuary salinity
 \citep{beserespollackLongtermTrendsResponse2011, palmerImpactsDroughtsLow2015, montagnaAssessmentRelationshipFreshwater2020}
-are a concern to local stakeholders. \citet{bugica_water_2020}
+and are a concern to local stakeholders. Water quality assessments
 identified monotonic increases in total phosphorus (TP), orthophosphate,
 total Kjeldahl nitrogen (TKN), and chlorophyll-\emph{a} at sites within
-Lavaca Bay. Although long-term changes in DO concentrations were not
-identified, the trends in nutrient concentrations are concerning due to
-the role of nitrogen as a limiting factor for primary production in many
-Texas estuaries
+Lavaca Bay \citep{bugica_water_2020}. Although long-term changes in DO
+concentrations were not identified, the trends in nutrient
+concentrations are concerning due to the role of nitrogen as a limiting
+factor for primary production in many Texas estuaries
 \citep{gardnerNitrogenFixationDissimilatory2006, houTransformationFateNitrate2012, doradoUnderstandingInteractionsFreshwater2015, paudelRelationshipSuspendedSolids2019, wetz_exceptionally_2017}
@@
 Lavaca Bay watershed
-\citep{jainTechnicalSupportDocument2021, schramm_lavaca_2018, bertholdDirectMailingEducation2021}.
+\citep{jainTechnicalSupportDocument2021, schrammLavacaRiverWatershed2018, bertholdDirectMailingEducation2021}.
 However, at a statewide scale, these approaches have shown limited
 success and emphasize a need for improved efforts at assessing and
-linking management actions with downstream water quality to better
-identify and replicate effective management actions across the state
+linking management actions with downstream water quality to identify and
+replicate effective management actions across the state
 \citep{schrammTotalMaximumDaily2022}. The identification and
@@
 \citep{wazniakLinkingWaterQuality2007, lloydMethodsDetectingChange2014}.
-To provide actionable information for resource managers, water quality
-conditions must be evaluated relative to changes in natural
-environmental drivers to better understand and manage potential
-anthropogenic effects. This study utilizes semiparemetric methods to
-develop estimates of delivered and flow-normalized nutrient loads and
-assess changes in loads delivered to Lavaca Bay. The study also assesses
-the response of water quality parameters in Lavaca Bay over time and in
-response to freshwater inflow controlled for seasonality and to
-watershed nutrient loads that are controlled for environmentally driven
-variation.
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-\hypertarget{study-area-and-data}{%
-\subsection{Study Area and Data}\label{study-area-and-data}}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig1-1.png}
-\caption{Map of the Lavaca Bay watershed, location of USGS gages where nutrient loads were calculated, and location of estuary water quality sampling sites.\label{fig:fig1}}
-\end{figure}
+The development and application of statistical methods such as Weighted
+Regressions on Time, Discharge and Season
+\citep[WRTDS,][]{hirschWeightedRegressionsTime2010} and Generalized
+Additive Models \citep[GAMs,][]{woodFastStableRestricted2011} has
+provided effective tools for researchers to quantify and communicate
+non-linear changes in river and estuary pollutant loadings.
+
+WRTDS calculates a time series of in stream concentrations or loads
+(daily, monthly, or annually) and flow-normalized estimates of
+concentrations and loads using locally weighted regresssion for unique
+combinations of time, discharge, and season. WRTDS has been widely used
+to assess and identify trends in riverine nutrients
+\citep{oelsner_recent_2019, stackpooleLongTermMississippi2021},
+chlorides \citep{stetsIncreasingChlorideRivers2018}, and other
+pollutants of concern \citep{shodaWaterqualityTrendsRivers2019}. WRTDS
+has also been succesfully adapted to assess trends in estuarine water
+quality concentrations \citep{beckFourDecadesWater2018}.
+
+While WRTDS is a statistical approach developed specifically for water
+quality applications, GAMs are a broadly applicable statistical method.
+GAMs are a semiparametric extension of generalized linear models where
+the linear predictor is represented as the sum of multiple unknown
+smooth functions and parametric linear predictors
+\citep{woodFastStableRestricted2011}. Although the underlying parameter
+estimation procedure of GAMs is substantially different than WRTDS, both
+the functional form and results have been demonstrated to be similar
+\citep{beckNumericalQualitativeContrasts2017}. Water quality
+applications of GAMs include river and catchment nutrient concentration
+and load models
+\citep{wangLoadEstimationUncertainties2011, kroonRiverLoadsSuspended2012, kuhnertQuantifyingTotalSuspended2012, robson_prediction_2015-1, hagemannEstimatingNutrientOrganic2016, mcdowell_implications_2021, biagi_novel_2022}.
+GAMs can also be used to identify non-linear temporal trends (including
+flow-normalized trends) in pollutant concentrations and loads
+\citep{beckNumericalQualitativeContrasts2017, murphyGeneralizedAdditiveModel2019}.
+Recently GAMs have also been used to link water quality responses in
+receiving water bodies to changes in nutrient inputs
+\citep{murphyNutrientImprovementsChesapeake2022}.
+\citet{beckNumericalQualitativeContrasts2017} provides a substantial
+discussion on the differences (and similarities) between GAMs and WRTDS
+for water quality applications.
+
+To provide actionable information for resource managers in Lavaca Bay,
+water quality conditions must be evaluated relative to changes in
+natural environmental drivers to better understand and manage potential
+human impacts. This study utilizes GAMs to develop estimates of
+delivered and flow-normalized nutrient loads and assess changes in loads
+delivered to Lavaca Bay. GAMs were chosen over other regression-based
+approached for use in this study due to; (1) the ability to easily
+explore and incorporate different model terms; (2) the incorporation of
+non-liner smooth functions that do not require explicit a priori
+knowledge of the expected shape; and (3) inclusion of a link function
+that related the expected value of the response to linear predictors
+thus avoiding unneeded data transformations and bias corrections. The
+study also assesses the response of water quality parameters in Lavaca
+Bay over time and in response to freshwater inflow controlled for
+seasonality and to watershed nutrient loads that are controlled for
+environmentally driven variation.
+
+\hypertarget{sec2}{%
+\section{Methods}\label{sec2}}
+
+\hypertarget{location-and-data}{%
+\subsection{Location and Data}\label{location-and-data}}
 
 Lavaca Bay is 190 km\textsuperscript{2} with the majority of freshwater
-inflow provided by the Lavaca and Navidad River systems (Figure
-\ref{fig:fig1}). The Garcitas-Arenosa, Placedo Creek, and Cox Bay
-watersheds provide additional freshwater inflows. The entire watershed
-land area is 8,149 km\textsuperscript{2} and primarily rural. Watershed
-land cover and land use is 50\% grazed pasture and rangeland, 20\%
-cultivated cropland (primarily rows crops such as corn, cotton, and
+inflow provided by the Lavaca and Navidad River systems
+(Figure~\ref{fig:fig1}). The Garcitas-Arenosa, Placedo Creek, and Cox
+Bay watersheds provide additional freshwater inflows. The entire
+watershed land area is 8,149 km\textsuperscript{2} and primarily rural.
+Watershed land cover and land use is 50\% grazed pasture and rangeland,
+20\% cultivated cropland (primarily rows crops such as corn, cotton, and
 sorghum), and 5\% suburban/urban. Pasture and rangeland is concentrated
@@
 
-Daily discharges for the Lavaca River (USGS-08164000, Figure
-\ref{fig:fig1}) were obtained from the United States Geologic Survey
-(USGS) National Water Information System using the \emph{dataRetrieval}
-R package \citep{deciccoDataRetrievalPackagesDiscovering2022}. Gaged
-daily discharges from the outlet of Lake Texana on the Navidad River
+\begin{figure}
+
+{\centering \includegraphics[width=5.2in,]{../fig1}
+
+}
+
+\caption{Map of Lavaca Bay and the contribution watershed.
+The freshwater sites are the most downstream freshwater stream locations with water quality and streamflow data used for nutrient load models.
+Water quality concentration data at the three Lavaca Bay sites were used to assess relationships between freshwater flows, loads and estuary water quality.}\label{fig:fig1}
+\end{figure}
+
+Daily discharges for the Lavaca River (USGS-08164000,
+Figure~\ref{fig:fig1}) were obtained from the United States Geologic
+Survey (USGS) National Water Information System using the
+\emph{dataRetrieval} R package
+\citep{deciccoDataRetrievalPackagesDiscovering2022}. Gaged daily
+discharges from the outlet of Lake Texana on the Navidad River
 (USGS-0816425) were provided by the Texas Water Development Board (TWDB)
@@
 (NO\textsubscript{3}) data from January 2005 through December 2020 were
-downloaded (Table \ref{tab:fwsummary}). Less than 5-years of total
-nitrogen and TKN concentration data were available at the freshwater
-sites and deemed insufficient to develop load estimation models
+downloaded (Table~\ref{tab:table1}). Less than 5-years of total nitrogen
+and TKN concentration data were available at the freshwater sites and
+deemed insufficient to develop load estimation models
 \citep{horowitzEvaluationSedimentRating2003, snelderEstimationCatchmentNutrient2017}.
@@
 Nitrite+Nitrate (NO\emph{\textsubscript{x}}), TKN, chlorophyll-\emph{a},
-and DO concentrations from January 2005 through December 2020 (Table
-\ref{tab:estuarysummary}).
-
-\begin{table}[H]
-
-\caption{\label{tab:fwsummary}Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.}
+and DO concentrations from January 2005 through December 2020
+(Table~\ref{tab:table2}).
+
+\begin{table}
+
+\caption{\label{tab:table1}Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.}
 \centering
@@
 USGS-08164000 & TP (mg/L) & \num{0.21} & \num{0.09} & 80\\
- & NO\textsubscript{3} (mg/L) & \num{0.18} & \num{0.24} & 74\\
+ & NO3 (mg/L) & \num{0.18} & \num{0.24} & 74\\
  & Mean Daily Streamflow (cfs) & \num{332.78} & \num{1667.47} & 7671\\
 USGS-08164525 & TP (mg/L) & \num{0.20} & \num{0.08} & 81\\
- & NO\textsubscript{3} (mg/L) & \num{0.29} & \num{0.26} & 62\\
+ & NO3 (mg/L) & \num{0.29} & \num{0.26} & 62\\
  & Mean Daily Streamflow (cfs) & \num{666.14} & \num{2957.79} & 7671\\
@@
 
-\begin{table}[H]
-
-\caption{\label{tab:estuarysummary}Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.}
+\begin{table}
+
+\caption{\label{tab:table2}Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.}
 \centering
-\begin{tabular}[t]{lllll}
+\begin{tabular}[t]{llrrr}
 \toprule
@@
 \midrule
- & TP (mg/L) & 0.11 & 0.05 & 47\\
-
- & NO\textsubscript{\emph{x}} (mg/L) & 0.07 & 0.15 & 51\\
-
- & TKN (mg/L) & 0.94 & 0.49 & 45\\
-
- & Chlorophyll-\emph{a} ($\mu$g/L) & 9.43 & 5.31 & 47\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13383} & DO (mg/L) & 7.22 & 1.35 & 55\\
-\cmidrule{1-5}
- & TP (mg/L) & 0.08 & 0.03 & 51\\
-
- & NO\textsubscript{\emph{x}} (mg/L) & 0.06 & 0.08 & 52\\
-
- & TKN (mg/L) & 0.76 & 0.40 & 48\\
-
- & Chlorophyll-\emph{a} ($\mu$g/L) & 8.22 & 6.44 & 46\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13384} & DO (mg/L) & 7.51 & 1.32 & 54\\
-\cmidrule{1-5}
- & TP (mg/L) & 0.13 & 0.06 & 50\\
-
- & NO\textsubscript{\emph{x}} (mg/L) & 0.09 & 0.13 & 53\\
-
- & TKN (mg/L) & 0.94 & 0.37 & 49\\
-
- & Chlorophyll-\emph{a} ($\mu$g/L) & 9.67 & 5.33 & 49\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13563} & DO (mg/L) & 7.91 & 1.34 & 56\\
+TCEQ-13383 & TP (mg/L) & \num{0.11} & \num{0.05} & 47\\
+ & NOx (mg/L) & \num{0.07} & \num{0.15} & 51\\
+ & TKN (mg/L) & \num{0.94} & \num{0.49} & 45\\
+ & Chlorophyll-a (ug/L) & \num{9.43} & \num{5.31} & 47\\
+ & DO (mg/L) & \num{7.22} & \num{1.35} & 55\\
+TCEQ-13384 & TP (mg/L) & \num{0.08} & \num{0.03} & 51\\
+ & NOx (mg/L) & \num{0.06} & \num{0.08} & 52\\
+ & TKN (mg/L) & \num{0.76} & \num{0.40} & 48\\
+ & Chlorophyll-a (ug/L) & \num{8.22} & \num{6.44} & 46\\
+ & DO (mg/L) & \num{7.51} & \num{1.32} & 54\\
+TCEQ-13563 & TP (mg/L) & \num{0.13} & \num{0.06} & 50\\
+ & NOx (mg/L) & \num{0.09} & \num{0.13} & 53\\
+ & TKN (mg/L) & \num{0.94} & \num{0.37} & 49\\
+ & Chlorophyll-a (ug/L) & \num{9.67} & \num{5.33} & 49\\
+ & DO (mg/L) & \num{7.91} & \num{1.34} & 56\\
 \bottomrule
@@
 (USGS-08164000) and the outlet of Lake Texana on the Navidad River
-(USGS-08164525) were developed using Generalized Additive Models (GAMs)
-relating nutrient concentration to river discharge, season, and time.
-Separate models were fit at each station for each parameter and used to
-predict nutrient concentrations for each day in the study period. GAMs
-can be specified in a functionally similar manner to the commonly used
-LOADEST \citep{cohn_validity_1992} or WRTDS \citep{hirsch_weighted_2010}
-regression models and have been shown to produce reliable estimates of
-nutrient and sediment loadings
-\citep{wangLoadEstimationUncertainties2011, kroonRiverLoadsSuspended2012, kuhnert_quantifying_2012, robson_prediction_2015-1, hagemannEstimatingNutrientOrganic2016, mcdowell_implications_2021, biagi_novel_2022}.
-GAMs are a semiparametric extension of generalized linear models where
-the linear predictor is represented as the sum of multiple unknown
-smooth functions and parametric linear predictors
-\citep{wood_fast_2011}. Although the underlying parameter estimation
-procedure of GAMs is substantially different than WRTDS, both the
-functional form and results are demonstrated to be similar
-\citep{beckNumericalQualitativeContrasts2017}. The use of GAMs over
-other regression-based approaches was (1) the ability to easily explore
-and incorporate different model terms, (2) the ability to incorporate
-non-linear smooth function without explicit a priori knowledge of the
-expect shape, and (3) the ability to specify a link function that
-relates the expected value of the response to the linear predictors and
-avoid data transformations.
-
-GAMs were fit using the \emph{mgcv} package in R which makes available
-multiple types of smooth functions with automatic smoothness selection
-\citep{wood_fast_2011}. The general form of the model related
-NO\textsubscript{3} or TP concentration to a long term tend, season,
-streamflow, and two different antecedent discharge terms:
-
-\begin{equation}\label{eq:1}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa)  \\
-y &\sim \mathcal{N}(\mu,\,\sigma^{2}),
-\end{aligned}
-\end{equation}
+(USGS-08164525) were developed using GAMs relating nutrient
+concentration to river discharge, season, and time. Separate models were
+fit at each station for each parameter and used to predict nutrient
+concentrations for each day in the study period. GAMs were fit using the
+\emph{mgcv} package in R which makes available multiple types of smooth
+functions with automatic smoothness selection
+\citep{woodFastStableRestricted2011}. The general form of the model
+related NO\textsubscript{3} or TP concentration to a long term tend,
+season, streamflow, and two different antecedent discharge terms, shown
+in Eq. \ref{eq1}:
+
+\begin{align}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa), \nonumber \\
+y &\sim \mathcal{N}(\mu,\,\sigma^{2})
+\label{eq1}
+\end{align}
 
 where \(\mu\) is the conditional expected NO\textsubscript{3} or TP
-concentration, \emph{g()} is the log-link, \(\alpha\) is the intercept,
-\emph{f\textsubscript{n}()} are smoothing functions. \emph{y} is the
-response variable (NO\textsubscript{3} or TP concentration) modeled as
-normally distributed with mean \(\mu\) and standard deviation
-\(\sigma\). \emph{ddate} is the date converted to decimal notation,
-\emph{yday} is numeric day of year (1-366), and \emph{log1p(Q)} is the
-natural log of mean daily streamflow plus 1.
+concentration, \(g()\) is the log-link, \(\alpha\) is the intercept,
+\(f_n()\) are smoothing functions. \(y\) is the response variable
+(NO\textsubscript{3} or TP concentration) modeled as normally
+distributed with mean \(\mu\) and standard deviation \(\sigma\).
+\emph{ddate} is the date converted to decimal notation, \emph{yday} is
+numeric day of year (1-366), and \emph{log1p(Q)} is the natural log of
+mean daily streamflow plus 1.
 
@@
 concentration at the current time period
-\citep{wangLoadEstimationUncertainties2011, kuhnert_quantifying_2012, zhang_improving_2017},
-and expressed in \citet{kuhnert_quantifying_2012} as:
-
-\begin{equation}\label{eq:2}
-ma(\delta) = d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}\quad\text{and}\quad \kappa_{i}=\sum_{m=1}^{i}\hat{Q}_m,
-\end{equation}
+\citep{wangLoadEstimationUncertainties2011, kuhnertQuantifyingTotalSuspended2012, zhang_improving_2017},
+using Eq. \ref{eq2}:
+
+\begin{align}
+ma(\delta) &= d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}, \nonumber \\
+\kappa_{i} &=\sum_{m=1}^{i}\hat{Q}_m
+\label{eq2}
+\end{align}
 
@@
 dry the current time period is from a previous time period
-\citep{vecchia_trends_2009, zhang_improving_2017}. Long-term flow
-anomaly (\emph{ltfa}) is the streamflow over the previous year relative
-to the entire period and calculated as described by
-\citet{zhang_improving_2017}:
-
-\begin{equation}\label{eq:3}
+\citep{vecchiaTrendsConcentrationsUse2009, zhang_improving_2017}.
+Long-term flow anomaly (\emph{ltfa}) is the streamflow over the previous
+year relative to the entire period \citep{zhang_improving_2017} and
+calculated using Eq. \ref{eq3}:
+
+\begin{equation}
 ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}
+\label{eq3}
 \end{equation}
@@
 and the short-term flow anomaly (\emph{stfa}) calculated as the current
-day flow compared to the preceding 1-month streamflow:
-
-\begin{equation}\label{eq:4}
+day flow compared to the preceding 1-month streamflow using \ref{eq4}:
+
+\begin{equation}
 stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t)
+\label{eq4}
 \end{equation}
 
-where \emph{x} are the averages of log-transformed streamflow over the
+where \(x\) are the averages of log-transformed streamflow over the
 antecedent period (\emph{1-year}, \emph{1-month}, etc.) for time
 \emph{t}. We used \emph{ltfa} in NO\textsubscript{3} models and
-\emph{stfa} in TP models based on results from
-\citet{zhang_improving_2017} demonstrating major improvements in
-NO\textsubscript{x} regression models that incorporated \emph{ltfa} and
-moderate improvements in TP regression models that incorporated
-\emph{stfa}.
-
-The calculation of model terms for the Lake Texana site were slightly
-modified because daily loads are not a function of natural stream flow
-processes alone, but of dam releases and nutrient concentrations at the
-discharge point of the lake. \emph{Q}, \emph{ma}, and \emph{fa} terms
-were calculated based on total gaged inflow from the 4 major tributaries
-to the lake. Thin-plate regression splines were used for \emph{ddate},
+\emph{stfa} in TP models based on previous work demonstrating major
+improvements in NO\textsubscript{x} regression models that incorporated
+\emph{ltfa} and moderate improvements in TP regression models that
+incorporated \emph{stfa} \citep{zhang_improving_2017}.
+
+The calculation of model terms for the Lake Texana site were modified
+because daily loads are not a function of natural stream flow processes
+alone, but of dam releases and nutrient concentrations at the discharge
+point of the lake. \emph{Q}, \emph{ma}, and \emph{fa} terms were
+calculated based on total gaged inflow from the 4 major tributaries to
+the lake. Thin-plate regression splines were used for \emph{ddate},
 \emph{log1p(Q)}, \emph{fa}, and \emph{ma}. A cyclic cubic regression
@@
 each of the historical values for that day of the year over the study
-period \citep{hirsch_weighted_2010}. The flow-normalized estimate was
-calculated as the mean of all the predictions for each day considering
-all possible flow values. Standard deviations and credible intervals
-were obtained by drawing samples from the multivariate normal posterior
-distribution of the fitted GAM
+period \citep{hirschWeightedRegressionsTime2010}. The flow-normalized
+estimate was calculated as the mean of all the predictions for each day
+considering all possible flow values. Standard deviations and credible
+intervals were obtained by drawing samples from the multivariate normal
+posterior distribution of the fitted GAM
 \citep{woodConfidenceIntervalsGeneralized2006, marraCoveragePropertiesConfidence2012, mcdowell_implications_2021}.
-Uncertainty in loads were reported as 90\% credible intervals developed
-by drawing 1000 realizations of parameter estimates from the
+Uncertainty in loads were calculated as 90\% credible intervals
+estimated by drawing 1000 realizations of parameter estimates from the
 multivariate normal posterior distribution of the model parameters. GAM
@@
 \citep{burmanComparativeStudyOrdinary1989} and average Nash-Sutcliffe
-Efficiency (NSE), r\textsuperscript{2} and percent bias (PBIAS) metrics
-across folds were calculated for each model.
+Efficiency (NSE), Pearson sample correlation (\emph{r}) and percent bias
+(PBIAS) metrics across folds were calculated for each model.
 
@@
 models were fit at each site relating parameter concentration to
-temporal trends, inflow, and nutrient loads
-\citep{murphyNutrientImprovementsChesapeake2022}:
-
-\begin{equation}\label{eq:5}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) \\
-y &\sim \Gamma(\mu,\lambda),
-\end{aligned}
-\end{equation}
-
-\begin{equation}\label{eq:6}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) \\
-y &\sim \Gamma(\mu,\lambda),
-\end{aligned}
-\end{equation}
-
-\begin{equation}\label{eq:7}
-\begin{aligned}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \\
-y &\sim \Gamma(\mu,\lambda),
-\end{aligned}
-\end{equation}
+temporal trends (\citep{eq-eq5}), temporal trends and inflow
+(\citet{eq-eq6}), and temporal trends, inflow, and nutrient loads
+(\citet{eq-eq7}): \begin{align}
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday), \label{eq5} \\
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q), \label{eq6} \\
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(Q) + f_4(Load) \label{eq7}
+\end{align}
 
 where \(\mu\) is the conditional expected response (nutrient
-concentration), \emph{g()} is the log link, and response variable was
+concentration), \(g()\) is the log link, and response variable was
 modeled as Gamma distributed with mean \(\mu\) and scale \(\lambda\).
-\emph{f\textsubscript{1}(ddate)} is decimal date smoothed with a
-thin-plate regression spline, \emph{f\textsubscript{2}(yday)} is the
-numeric day of year smoothed with a cyclic cubic regression spline,
-\emph{f\textsubscript{3}(Q)} is mean daily inflow (the combined
-measurements from Lavaca River and Navidad River) and
-\emph{f\textsubscript{4}(Load)} is the total NO\textsubscript{3} or TP
-watershed load. The set of models specified for each water quality
-response are in Table \ref{tab:estgammodels}.
-
-\begin{table}[H]
-
-\caption{\label{tab:estgammodels}Set of GAM models specified for each water quality parameter response.}
+\(f_1(ddate)\) is decimal date smoothed with a thin-plate regression
+spline, \(f_2(yday)\) is the numeric day of year smoothed with a cyclic
+cubic regression spline, \(f_3(Q)\) is mean daily inflow (the combined
+measurements from Lavaca River and Navidad River) and \(f_4(Load)\) is
+the total NO\textsubscript{3} or TP watershed load. The set of models
+specified for each water quality response are in Table~\ref{tab:table3}.
+
+\begin{table}
+
+\caption{\label{tab:table3}Set of GAM models specified for each water quality parameter response.}
 \centering
@@
 \toprule
-\makecell[l]{Water Quality\\Response Parameter} & Model & Model Terms\\
+Parameter & Model & Model Terms\\
 \midrule
@@
 
-\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{\emph{x}}} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(NO\textsubscript{3} Load)\\
+\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(NO\textsubscript{3} Load)\\
 \cmidrule{1-3}
@@
 
-\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\emph{a}} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\textsubscript{3} Load)\\
+\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & Flow+Load & s(ddate) + s(yday) + s(Q) + s(TP Load) + s(NO\textsubscript{3} Load)\\
 \cmidrule{1-3}
@@
 nutrient loads obtained by fitting a GAM relating season (day of year)
-to log transformed daily freshwater inflow values:
-
-\begin{equation}\label{eq:8}
-g(\mu) = \alpha + f_1(yday),
-\end{equation}
-
-and a GAM relating log transformed NO\textsubscript{3} or TP loads to
-log transformed daily inflow:
-
-\begin{equation}\label{eq:9}
-g(\mu) = \alpha + f_1(log(Q)),
-\end{equation}
-
-where the response variables were modeled as normally distributed with
-an identity link function. Response residuals from the respective GAM
-models were used as \emph{Q} and \emph{Load} in Equation \ref{eq:6} and
-Equation \ref{eq:7}.
+to log transformed daily freshwater inflow values and a GAM relating log
+transformed NO\textsubscript{3} or TP loads to log transformed daily
+inflow.Response residuals from the respective GAM models were used as
+\emph{Q} and \emph{Load} in Eq.~\ref{eq6} and Eq.~\ref{eq7}.
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{sec3}{%
+\section{Results}\label{sec3}}
 
@@
 
-Using evaluation criteria from
+Using evaluation criteria recommended by
 \citet{moriasiHydrologicWaterQuality2015}, predictive performance of
 nutrient loading GAMs ranged from ``satisfactory'' to ``very good''
-based on median NSE, r\textsuperscript{2}, and PBIAS metrics calculated
-using 5-fold cross validation. Median NSE values were 0.34 (Lavaca
-River) and 0.48 (Navidad River) for NO\textsubscript{3} loads and 0.81
-(Lavaca River) and 0.91 (Navidad River) for TP loads. Median
-r\textsuperscript{2} values were 0.70 (Lavaca River) and 0.87 (Navidad
-River) for NO\textsubscript{3} loads and 0.93 (Lavaca River) and 0.99
-(Navidad River) fopr TP loads. Median PBIAS values were 2.00 (Lavaca
-River) and 10.90 (Navidad River) for NO\textsubscript{3} loads and -7.20
-(Lavaca River) and -3.30 (Navidad River) for TP loads. Density plots of
-metrics show similar distribution of values between sites for the same
-parameter, with the exception r\textsuperscript{2} values for
-NO\textsubscript{3} loads where USGS-08164000 showed a much larger
-variance in values compared to USGS-08164525 (Figure \ref{fig:fig2}). TP
-GAMS had higher average NSE and r\textsuperscript{2} values and lower
-variance in metric values compared to NO\textsubscript{3}.
-
-\begin{figure}\begin{adjustwidth}{-\extralength}{0cm}
-
-{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig2-1}
+based on median NSE, \emph{r}, and PBIAS metrics calculated using 5-fold
+cross validation. Median goodness-of-fit metrics for NO\textsubscript{3}
+models in the Lavaca River were 0.34 NSE, 0.70 \emph{r}, and 2.00 PBIAS.
+Navidad River NO\textsubscript{3} models appeared to perform slightly
+better with 0.48 NSE and 0.87 \emph{r} but with higher bias at 10.90
+PBIAS. Generally, TP models performed better than NO\textsubscript{3}
+models. Median goodness-of-fit metrics for TP in the Lavaca River were
+0.81 NSE, 0.93 \emph{r}, and -7.20 PBIAS. Navidad River TP models has
+similar performance with 0.91 NSE, 0.99 \emph{r}, and -3.30 PBIAS.
+Density plots of metrics show similar distribution of values between
+sites for the same parameter, with the exception \emph{r} values for
+NO\textsubscript{3} loads where Lavaca River had a much larger variance
+in values compared to the Navidad River (Figure~\ref{fig:fig2}). TP GAMS
+had higher average NSE and \emph{r} values and lower variance in metric
+values compared to NO\textsubscript{3}.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig2-1}
 
@@
 
-\end{adjustwidth}\caption[Density plots of goodness-of-fit metrics (NSE, r\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads]{Density plots of goodness-of-fit metrics (NSE, r\textsuperscript{2}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
-Color indicates the tail probability calcualted from the empirical cumulative distribution of the goodness-of-fit metrics.}\label{fig:fig2}
+\caption{Density plots of goodness-of-fit metrics (NSE, \textit{r}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
+Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics.}\label{fig:fig2}
 \end{figure}
@@
 Predicted annual NO\textsubscript{3} and TP loads show considerable
-variation, generally following patterns in discharge (Figure
-\ref{fig:fig3}, Figure \ref{fig:fig4}). Flow-normalized TP loads at both
-sites and flow-normalized NO\textsubscript{3} loads at USGS-08164000
-indicate watershed based loads did not change much over time when
-accounting for variation driven by streamflow (Figure \ref{fig:fig3}).
-Flow-normalized loads at USGS-08164000 show small variation over time
-with some decreases in NO\textsubscript{3} loads since 2013.
+variation, generally following patterns in discharge
+(Figures~\ref{fig:fig3}, \ref{fig:fig4}). Flow-normalized TP loads at
+both sites and flow-normalized NO\textsubscript{3} loads in the Lavaca
+River indicated watershed based loads did not change much over time when
+accounting for variation driven by streamflow (Figure~\ref{fig:fig3}).
+Flow-normalized loads in the Lavaca River showed small variation over
+time with some decreases in NO\textsubscript{3} loads since 2013.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig3-1}
+
+}
+
+\caption{Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
+\end{figure}
 
@@
 to 595,075 kg in 2007. Mean annual TP loading from 2005 through 2020 was
-182,673 kg (152,227 kg - 219,310 kg, 90\% CI). On average, USGS-08164525
-accounted for 68\% of NO\textsubscript{3} loads and 59\% of TP loads
-from 2005 through 2020. However, during periods of extreme drought the
-Lavaca River (USGS-08164000) became the primary source of nutrient
-loading in the watershed with the Navidad River only accounting for 15\%
-and 25\% of NO\textsubscript{3} and TP loads in 2011 (Figure
-\ref{fig:fig4}).
+182,673 kg (152,227 kg - 219,310 kg, 90\% CI). On average, the Navidad
+River accounted for 68\% of NO\textsubscript{3} loads and 59\% of TP
+loads from 2005 through 2020. However, during periods of extreme drought
+the Lavaca River became the primary source of nutrient loading in the
+watershed with the Navidad River only accounting for 15\% and 25\% of
+NO\textsubscript{3} and TP loads in 2011 (Figure~\ref{tab:fig4}).
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig3-1}
+{\centering \includegraphics[width=0.5\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig4-1}
 
@@
 
-\caption[Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525]{Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
-\end{figure}
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig4-1}
-
-}
-
-\caption[Comparison of delivered annual loads at USGS-08164000 and USGS-08164525]{Comparison of delivered annual loads at USGS-08164000 and USGS-08164525.}\label{fig:fig4}
+\caption{Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers.}\label{fig:fig4}
 \end{figure}
@@
 
-\begin{figure}\begin{adjustwidth}{-\extralength}{0cm}
-
-{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig5-1}
-
-}
-
-\end{adjustwidth}\caption[Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs]{Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.}\label{fig:fig5}
-\end{figure}
-
 GAMs did not identify significant changes in TP or DO concentrations at
-any of the Lavaca Bay sites from 2005 through 2020 (Figure
-\ref{fig:fig5}). The upper-bay site, TCEQ-13563, had a linear increase
-in NO\textsubscript{x} concentration and and decrease in
+any of the Lavaca Bay sites from 2005 through 2020
+(Figure~\ref{fig:fig5}). The upper-bay site, TCEQ-13563, had a linear
+increase in NO\textsubscript{x} concentration and and decrease in
 chlorophyll-\emph{a} from 2005 through 2014. The mid-bay site,
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig5-1}
+
+}
+
+\caption{Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.}\label{fig:fig5}
+\end{figure}
+
 Freshwater inflow provided additional explanation for changes in TP and
 NO\textsubscript{\emph{x}} concentration at all of the Lavaca Bay sites
-according to AIC\textsubscript{c} and model probability values (Table
-\ref{tab:tab4}). TCEQ-13563, the site closest to the river outlet, was
-the only site that had improvements in the explanations of DO and TKN
-concentration with the inclusion of inflow. Both TCEQ-13563 and
-TCEQ-13383, the mid-bay site, saw improvements in explanations for
+according to AIC\textsubscript{c} and model probability values
+(Table~\ref{tab:table4}). TCEQ-13563, the site closest to the river
+outlet, was the only site that had improvements in the explanations of
+DO and TKN concentration with the inclusion of inflow. Both TCEQ-13563
+and TCEQ-13383, the mid-bay site, saw improvements in explanations for
 variations in chlorophyll-\emph{a} with the inclusion of freshwater
@@
 
-\begin{table}[H]
-
-\caption{\label{tab:tab4}Estuary GAM AIC\textsubscript{c} values and associated model probabilities (in parenthesis).
+\begin{table}
+
+\caption{\label{tab:table4}Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
 Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.}
 \centering
-\begin{tabular}[t]{ll>{}l>{}l>{}l}
+\begin{tabular}[t]{lllll}
 \toprule
@@
 \midrule
- & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & \em{\textbf{-158.2 (0.72)}}\\
-
- & TCEQ-13384 & -194.4 (0.03) & \em{\textbf{-200.2 (0.49)}} & -200.2 (0.49)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash TP} & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & \em{\textbf{-157.3 (0.59)}}\\
-\cmidrule{1-5}
- & TCEQ-13383 & -218.9 (0) & \em{\textbf{-244.8 (0.5)}} & -244.8 (0.5)\\
-
- & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & \em{\textbf{-311.9 (0.52)}}\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{\emph{x}}} & TCEQ-13563 & -175.1 (0) & \em{\textbf{-190.2 (0.5)}} & -190.2 (0.5)\\
-\cmidrule{1-5}
- & TCEQ-13383 & 279.7 (0.18) & \em{\textbf{278.1 (0.41)}} & 278.1 (0.41)\\
-
- & TCEQ-13384 & \em{\textbf{268.2 (0.33)}} & 268.2 (0.33) & 268.2 (0.33)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\emph{a}} & TCEQ-13563 & 289.5 (0.08) & \em{\textbf{286.1 (0.46)}} & 286.1 (0.46)\\
-\cmidrule{1-5}
- & TCEQ-13383 & \em{\textbf{42.2 (0.66)}} & 43.5 (0.34) & -\\
-
- & TCEQ-13384 & \em{\textbf{34.3 (0.57)}} & 34.8 (0.43) & -\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash TKN} & TCEQ-13563 & 31.1 (0.22) & \em{\textbf{28.7 (0.78)}} & -\\
-\cmidrule{1-5}
- & TCEQ-13383 & \em{\textbf{146.4 (0.34)}} & 146.4 (0.34) & 146.5 (0.32)\\
-
- & TCEQ-13384 & \em{\textbf{135.9 (0.47)}} & 137 (0.27) & 137 (0.27)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash DO} & TCEQ-13563 & 138.3 (0.25) & \em{\textbf{137.2 (0.43)}} & 137.8 (0.32)\\
+TP & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & -158.2 (0.72)\\
+TP & TCEQ-13384 & -194.4 (0.03) & -200.2 (0.49) & -200.2 (0.49)\\
+TP & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & -157.3 (0.59)\\
+NOx & TCEQ-13383 & -218.9 (0) & -244.8 (0.5) & -244.8 (0.5)\\
+NOx & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & -311.9 (0.52)\\
+\addlinespace
+NOx & TCEQ-13563 & -175.1 (0) & -190.2 (0.5) & -190.2 (0.5)\\
+Chlorophyll-*a* & TCEQ-13383 & 279.7 (0.18) & 278.1 (0.41) & 278.1 (0.41)\\
+Chlorophyll-*a* & TCEQ-13384 & 268.2 (0.33) & 268.2 (0.33) & 268.2 (0.33)\\
+Chlorophyll-*a* & TCEQ-13563 & 289.5 (0.08) & 286.1 (0.46) & 286.1 (0.46)\\
+TKN & TCEQ-13383 & 42.2 (0.66) & 43.5 (0.34) & NA\\
+\addlinespace
+TKN & TCEQ-13384 & 34.3 (0.57) & 34.8 (0.43) & NA\\
+TKN & TCEQ-13563 & 31.1 (0.22) & 28.7 (0.78) & NA\\
+DO & TCEQ-13383 & 146.4 (0.34) & 146.4 (0.34) & 146.5 (0.32)\\
+DO & TCEQ-13384 & 135.9 (0.47) & 137 (0.27) & 137 (0.27)\\
+DO & TCEQ-13563 & 138.3 (0.25) & 137.2 (0.43) & 137.8 (0.32)\\
 \bottomrule
@@
 
-\begin{figure}\begin{adjustwidth}{-\extralength}{0cm}
-
-{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig6-1}
+GAMs showed that increases in freshwater inflow resulted in nearly
+linear increases in TP and NO\textsubscript{\emph{x}} concentration at
+all three sites (Figure~\ref{fig:fig6}). At the upper-bay TCEQ-13563
+site, GAMs showed that increases in freshwater inflow initially
+increased chlorophyll-\emph{a} and DO concentration but concentrations
+leveled and potentially decreased at higher flows. The mid-bay
+TCEQ-13383 site showed a nearly linear increased in chlorophyll-\emph{a}
+concentration in response to increases freshwater inflow. Freshwater
+flow did not have significant effects on chlorophyll-\emph{a}, TKN, or
+DO at the lower-bay TCEQ-13384 site.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig6-1}
 
@@
 
-\end{adjustwidth}\caption[Estimated effects of mean daily inflow residuals on mean TP, NO\textsubscript{\emph{x}}, chlorophyll-\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs]{Estimated effects of mean daily inflow residuals on mean TP, NO\textsubscript{\emph{x}}, chlorophyll-\emph{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.}\label{fig:fig6}
+\caption{Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.}\label{fig:fig6}
 \end{figure}
-
-GAMs showed increases in freshwater inflow resulted in nearly linear
-increases in TP and NO\textsubscript{\emph{x}} concentration at all
-three sites (Figure \ref{fig:fig6}). At the upper-bay TCEQ-13563 site,
-GAMs showed increases in freshwater inflow initially increased
-chlorophyll-\emph{a} and DO concentration but concentrations leveled and
-potentially decreased at higher flows. The mid-bay TCEQ-13383 site
-showed a nearly linear increased in chlorophyll-\emph{a} concentration
-in response to increases freshwater inflow. Freshwater flow did not have
-significant effects on chlorophyll-\emph{a}, TKN, or DO at the lower-bay
-TCEQ-13384 site.
 
@@
 concentration at the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383
-respectively (Figure \ref{fig:fig7}). The relative effect size appeared
+respectively (Figure~\ref{fig:fig7}). The relative effect size appeared
 to much smaller than the effect of freshwater inflow alone. Increased
@@
 TCEQ-13384 site. The effect was quite small compared to streamflow and
-provided only small improvements to the model (Table \ref{tab:tab4}). As
+provided only small improvements to the model (\citet{tbl-table4}). As
 noted above, nutrient loadings did not provide any explanation in
@@
 
-{\centering \includegraphics[width=1\linewidth]{Schramm-Manuscript-2023_files/figure-latex/fig7-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig7-1}
 
@@
 
-\caption[Estimated effects of nutrient load residuals on TP and NO\textsubscript{\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs]{Estimated effects of nutrient load residuals on TP and NO\textsubscript{\emph{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.}\label{fig:fig7}
+\caption{Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs.}\label{fig:fig7}
 \end{figure}
@@
 \section{Discussion}\label{discussion}}
+
+\textbf{resume here}
 
@@
 since 2016
-\citep{schramm_lavaca_2018, bertholdDirectMailingEducation2021}. The
-decrease in flow-normalized NO\textsubscript{3} loads could be a
+\citep{schrammLavacaRiverWatershed2018, bertholdDirectMailingEducation2021}.
+The decrease in flow-normalized NO\textsubscript{3} loads could be a
 reflection of those collective efforts but further data collection and
@@
 
-Converted to average annual yield, the estimates of annual TP loads at
-Lavaca River (USGS-08164000) are within the ranges in previous published
-studies (Table \ref{tab:tpcomp};
-\citep{dunnTrendsNutrientInflows1996, rebich_sources_2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019}).
+Converted to average annual yield, the estimates of annual TP loads for
+the Lavaca River are within the ranges in previous published studies
+(\citet{tbl-table5};
+\citep{dunnTrendsNutrientInflows1996, rebichSourcesDeliveryNutrients2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019}).
 It isn't obvious why TP estimates in
@@
 was drier on average than the other studies. The SPARROW models used in
-\citet{rebich_sources_2011} and \citet{wise_spatially_2019} utilize a
-version of LOADEST in the underlying load estimation procedure, so a
-difference due to methodology alone is unlikely.
-
-\begin{table}[H]
-
-\begin{threeparttable}
-\caption{\label{tab:tpcomp}Mean estimates of annual TP yield in the Lavaca River watershed in published studies.}
+\citet{rebichSourcesDeliveryNutrients2011} and
+\citet{wise_spatially_2019} utilize a version of LOADEST in the
+underlying load estimation procedure, so a difference due to methodology
+alone is unlikely.
+
+\begin{table}
+
+\caption{\label{tab:table5}Comparisons of previously published estimates of mean annual TP yield at the Lavaca River site.}
 \centering
-\begin{tabular}[t]{lllll}
+\begin{tabular}[t]{llll}
 \toprule
-Parameter & \makecell[c]{Reported Yield\\(kg$\cdot$km\textsuperscript{-2}$\cdot$year\textsuperscript{-1})} & Approach & Time Period & Reference\\
+Reported Yield (kg$\cdot$km\textsuperscript{2}$\cdot$year\textsuperscript{-1}) & Approach & Time Period & Reference\\
 \midrule
-TP & 35.2 (28.8, 43.3)\textsuperscript{*} & GAM & 2005-2020 & This work\\
-TP & 45.2 & SPARROW & 2000-2014 & \citet{wise_spatially_2019}\\
-TP & 42 & SWAT & 1977-2005 & \citet{omaniEstimationSedimentNutrient2014}\\
-TP & 20.81-91.58\textsuperscript{\dag} & SPARROW & 1980-2002 & \citet{rebich_sources_2011}\\
-TP & 28.9 & LOADEST & 1972-1993 & \citet{dunnTrendsNutrientInflows1996}\\
+35.2 (28.8, 43.3)\textsuperscript{a} & GAM & 2005-2020 & This work\\
+45.2 & SPARROW & 2000-2014 & \cite{wise_spatially_2019}\\
+42 & SWAT & 1977-2005 & \cite{omaniEstimationSedimentNutrient2014}\\
+20.81-91.58\textsuperscript{b} & SPARROW & 1980-2002 & \cite{rebichSourcesDeliveryNutrients2011}\\
+28.9 & LOADEST & 1972-1993 & \cite{dunnTrendsNutrientInflows1996}\\
 \bottomrule
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{a} Mean of the annual point estimates and the lower and upper 95\% credible intervals.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{b} Represents a binned value range from a choropleth map.}\\
 \end{tabular}
-\begin{tablenotes}
-\small
-\item [*] Values represent the mean of annual point estimates, lower and upper 95\% credible intervals.
-\item [\dag] A single point estimate was not reported, these value represent the range depicted on the choropleth map provided in the report.
-\end{tablenotes}
-\end{threeparttable}
 \end{table}
@@
 The non-linear temporal water quality trends identified using GAMs
-differed slightly from the trends identified by
-\citet{bugica_water_2020}. This is not unexpected due to the different
+differed slightly from previously identified trends
+\citep{bugica_water_2020}. This is not unexpected due to the different
 time periods, different methodology, and generally small slopes
@@
 and other estuaries
-\citep{russell_effect_2006, caffreyHighNutrientPulses2007, peierlsNonmonotonicResponsesPhytoplankton2012, palmerImpactsDroughtsLow2015, ciraPhytoplanktonDynamicsLowInflow2021}.
+\citep{russell_effect_2006, caffreyHighNutrientPulses2007, peierlsNonmonotonicResponsesPhytoplankton2012, palmerImpactsDroughtsLow2015, ciraPhytoplanktonDynamicsLowinflow2021}.
 Inflow had a non-linear relationship with TKN at the two upstream sites,
@@
 control on the effects of freshwater inflow in lower reaches of Lavaca
-Bay. \citet{russell_effect_2006} also suggested the processing of
-organic loads in the upper portions of Lavaca Bay reduces the transport
-of nutrients into the lower reaches of the Bay.
+Bay. Previous work suggests the processing of organic loads in the upper
+portions of Lavaca Bay might reduce the transport of nutrients into the
+lower reaches of the Bay \citep{russell_effect_2006}.
 
@@
 chlorophyll-\emph{a} concentrations in Texas estuaries
-\citep{ornolfsdottirNutrientPulsingRegulator2004, doradoUnderstandingInteractionsFreshwater2015, ciraPhytoplanktonDynamicsLowInflow2021, tominackVariabilityPhytoplanktonBiomass2022},
+\citep{ornolfsdottirNutrientPulsingRegulator2004, doradoUnderstandingInteractionsFreshwater2015, ciraPhytoplanktonDynamicsLowinflow2021, tominackVariabilityPhytoplanktonBiomass2022},
 this study specifically used flow-adjusted freshwater derived nutrient
@@
 months following storms and large discharge events
-\citep{mooney_watershed_2012, wetzExtremeFutureEstuaries2013, bukaveckas_influence_2020, walkerTimescalesMagnitudeWater2021}.
+\citep{mooneyWatershedExportEvents2012a, wetzExtremeFutureEstuaries2013, bukaveckasInfluenceStormEvents2020, walkerTimescalesMagnitudeWater2021}.
 However, our work only evaluates responses to loading and inflows
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\vspace{6pt}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
-A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\funding{This project was funded in part by a Texas Coastal Management
-Program grant approved by the Texas Land Commissioner, providing
-financial assistance under the Coastal Zone Management Act of 1972, as
-amended, awarded by the National Oceanic and Atmospheric Administration
-(NOAA), Office for Coastal Management, pursuant to NOAA Award
+\backmatter
+
+\bmhead{Acknowledgments}
+
+The author extends thanks to Dr.~Mike Wetz (Harte Research Institute,
+Texas A\&M Corpus Christi), Chad Kinsfather and Partick Brzozowski
+(Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water
+Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason
+Pinchbeck (Texas General Land Office) and the Lavaca Bay Foundation for
+supporting development of this project and providing valuable feedback.
+
+\hypertarget{declarations}{%
+\section*{Declarations}\label{declarations}}
+\addcontentsline{toc}{section}{Declarations}
+
+\bmhead{Funding}
+
+This project was funded in part by a Texas Coastal Management Program
+grant approved by the Texas Land Commissioner, providing financial
+assistance under the Coastal Zone Management Act of 1972, as amended,
+awarded by the National Oceanic and Atmospheric Administration (NOAA),
+Office for Coastal Management, pursuant to NOAA Award
 No.~NA21NOS4190136. The views expressed herein are those of the
 author(s) and do not necessarily reflect the views of NOAA, the U.S.
-Department of Commerce, or any of their subagencies}
-
-\dataavailability{Data and code are openly available in Zenodo at
-\url{https://doi.org/10.5281/zenodo.7330754}.}
-
-\acknowledgments{The author extends thanks to Dr.~Mike Wetz (Harte
-Research Institute, Texas A\&M Corpus Christi), Chad Kinsfather and
-Partick Brzozowski (Lavaca-Navidad River Authority), Brian Koch (Texas
-State Soil and Water Conservation Board), Bill Balboa (Matagorda Bay
-Foundation), Jason Pinchbeck (Texas General Land Office) and the Lavaca
-Bay Foundation for supporting development of this project and providing
-valuable feedback.}
-
-\conflictsofinterest{The author declares no conflict of interest. The
-project sponsors had no role in the design of the study; in the
-collection, analyses, or interpretation of data; in the writing of the
-manuscript, or in the decision to publish the results.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-
-%% Only for journal Encyclopedia
-%\entrylink{The Link to this entry published on the encyclopedia platform.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{adjustwidth}{-\extralength}{0cm}
-
-%\printendnotes[custom] % Un-comment to print a list of endnotes
-
-\reftitle{References}
-
-% Please provide either the correct journal abbreviation (e.g.
-according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
-% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
-
-%=====================================
-% References, variant A: external bibliography
-%=====================================
-\externalbibliography{yes}
-\bibliography{mybibfile.bib}
-
-% If authors have biography, please use the format below
-%\section*{Short Biography of Authors}
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
-%{\textbf{Firstname Lastname} Biography of first author}
-%
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
-%{\textbf{Firstname Lastname} Biography of second author}
-
-% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
-This produces: Wong (1999, p.
-328; 2000, p.
-475)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\PublishersNote{}
-\end{adjustwidth}
+Department of Commerce, or any of their subagencies.
+
+\bmhead{Data Availability}
+
+Reproducible code and datasets generated during this study are available
+in the Zenodo repository, \url{https://doi.org/10.5281/zenodo.733075}.
+
+\bibliography{bibliography.bib}
 

### 2023-05-20 map.R

@@ -117,7 +117,7 @@ waterbody_labs <- tibble(lab = c("Lake\nTexana", "Lavaca Bay", "Matagorda Bay"),
 
 p1 <- ggplot() +
   ## watersheds
-  geom_sf(data = ws, aes(fill = label, color = label), alpha = 0.3, size = 0.2, show.legend = FALSE) +
+  geom_sf(data = ws, aes(fill = label, color = label), alpha = 0.2, linewidth = 0.2, show.legend = FALSE) +
   ## colorado, wharton, dewit, calhoun county label
   geom_sf_text(data = counties_sub |> 
                  filter(CNTY_NM == "Colorado"|CNTY_NM == "Lavaca"| 
@@ -140,12 +140,12 @@ p1 <- ggplot() +
                size = 2, family = "Atkinson Hyperlegible", 
                alpha = 0.5, fontface = "bold",
                nudge_x = -10000, nudge_y = -10000) +
-  geom_sf(data = urban, fill = "azure4", size = 0.2, alpha = 0.5) +
-  geom_sf(data = Area, color = alpha("steelblue",0.25), alpha = 0.25, fill = "steelblue", size = 0.15) +
-  geom_sf(data = counties_sub, fill = "transparent", linetype = 3, size = 0.15, show.legend = FALSE) +
-  geom_sf(data = Flowlines, alpha = 0.25, size = 0.15, color = "steelblue") +
-  geom_sf(data = rivers, color = "steelblue", size = 0.15) +
-  geom_sf(data = waterbody, alpha = 1, fill = "slategray3",color = alpha("slategray3",1), size = 0.15) +
+  geom_sf(data = urban, fill = "azure4", linewidth = 0.2, alpha = 0.5) +
+  geom_sf(data = Area, color = alpha("steelblue",0.25), alpha = 0.25, fill = "steelblue", linewidth = 0.15) +
+  geom_sf(data = counties_sub, fill = "transparent", linetype = 3, linewidth = 0.15, show.legend = FALSE) +
+  geom_sf(data = Flowlines, alpha = 0.25, linewidth = 0.15, color = "steelblue") +
+  geom_sf(data = rivers, color = "steelblue", linewidth = 0.15) +
+  geom_sf(data = waterbody, alpha = 1, fill = "slategray3",color = alpha("slategray3",1), linewidth = 0.15) +
   geom_sf(data = df_sites, aes(shape = "Freshwater Sites")) +
[…]

### 2023-05-20 map_final_report.R

@@ -120,7 +120,7 @@ waterbody_labs <- tibble(lab = c("Lake\nTexana", "Lavaca Bay", "Matagorda Bay"),
 
 p1 <- ggplot() +
   ## watersheds
-  geom_sf(data = ws, aes(fill = label, color = label), alpha = 0.3, size = 0.2, show.legend = FALSE) +
+  geom_sf(data = ws, aes(fill = label, color = label), alpha = 0.3, linewidth = 0.2, show.legend = FALSE) +
   ## colorado, wharton, dewit, calhoun county label
   geom_sf_text(data = counties_sub |> 
                  filter(CNTY_NM == "Colorado"|CNTY_NM == "Lavaca"| 
@@ -143,12 +143,12 @@ p1 <- ggplot() +
                size = 2, family = "Atkinson Hyperlegible", 
                alpha = 0.5, fontface = "bold",
                nudge_x = -10000, nudge_y = -10000) +
-  geom_sf(data = urban, fill = "azure4", size = 0.2, alpha = 0.5) +
-  geom_sf(data = Area, color = alpha("steelblue",0.25), alpha = 0.25, fill = "steelblue", size = 0.15) +
-  geom_sf(data = counties_sub, fill = "transparent", linetype = 3, size = 0.15, show.legend = FALSE) +
-  geom_sf(data = Flowlines, alpha = 0.25, size = 0.15, color = "steelblue") +
-  geom_sf(data = rivers, color = "steelblue", size = 0.15) +
-  geom_sf(data = waterbody, alpha = 1, fill = "slategray3",color = alpha("slategray3",1), size = 0.15) +
+  geom_sf(data = urban, fill = "azure4", linewidth = 0.2, alpha = 0.5) +
+  geom_sf(data = Area, color = alpha("steelblue",0.25), alpha = 0.25, fill = "steelblue", linewidth = 0.15) +
+  geom_sf(data = counties_sub, fill = "transparent", linetype = 3, linewidth = 0.15, show.legend = FALSE) +
+  geom_sf(data = Flowlines, alpha = 0.25, linewidth = 0.15, color = "steelblue") +
+  geom_sf(data = rivers, color = "steelblue", linewidth = 0.15) +
+  geom_sf(data = waterbody, alpha = 1, fill = "slategray3",color = alpha("slategray3",1), linewidth = 0.15) +
   geom_sf(data = df_sites, aes(shape = "Freshwater Sites")) +
   geom_sf(data = tceq_sites, aes(shape = "Lavaca Bay Sites")) +
[…]

### 2023-05-20 Schramm-2023-05-AS/Schramm-2023-05-AS.Rmd
fix cross refs

@@
 effect on changes in estuary water quality a series of GAM models were fit at
-each site relating parameter concentration to temporal trends ([@eq-eq5]), 
-temporal trends and inflow (@eq-eq6), and temporal trends, inflow, and
-nutrient loads (@eq-eq7):
+each site relating parameter concentration to temporal trends (Eq.\ \ref{eq5}), 
+temporal trends and inflow (Eq.\ \ref{eq6}), and temporal trends, inflow, and
+nutrient loads (Eq.\ \ref{eq7}):
 \begin{align}
@@
 `r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 
-(Figure\ \ref{tab:fig4}).
+(Figure\ \ref{fig:fig4}).
 
@@
     geom_hline(yintercept = 0, linetype = 2, alpha = 0.25) +
-    # geom_text_repel(aes(label = label,
-    #                     color = model), 
-    #                 nudge_x = 5,
-    #                 hjust = "left",
-    #                 size = 2) +
     scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
@@
     theme_TWRI_print(base_family = "sans") +
-    theme(axis.title.y = element_text(size = 8),
+    theme(axis.title.y = element_markdown(size = 8),
           axis.title.x = element_text(size = 8),
@@
                              tar_read(no3_lavaca_13384_temporal, store = store),
-                             ylab = "NOx [mg/L]", 
+                             ylab = "NO<sub>*x*</sub> [mg/L]", 
                              subtitle = "B")
@@
                              tar_read(chla_lavaca_13384_temporal, store = store),
-                             ylab = expression("Chlorophyll-"~italic(a)~"[μg/L]"), 
+                             ylab = "Chlorophyll-*a* [μg/L]", 
                              subtitle = "D")
@@
   mutate(Parameter = case_when(
-    Parameter == "Nitrite+Nitrate" ~ "NOx",
-    Parameter == "Chlorophyll-a" ~ "Chlorophyll-*a*",
+    Parameter == "Nitrite+Nitrate" ~ "NO\\textsubscript{x}",
+    Parameter == "Chlorophyll-a" ~ "Chlorophyll-\\textit{a}",
     TRUE ~ as.character(Parameter)
   )) |>
-  mutate(Parameter = fct_relevel(Parameter, "TP", "NOx", "Chlorophyll-*a*","TKN", "DO"),
+  mutate(Parameter = fct_relevel(Parameter, "TP", "NO\\textsubscript{x}", "Chlorophyll-\\textit{a}","TKN", "DO"),
          Model = fct_relevel(Model, "Temporal", "Flow", "Flow + Load")) |>
@@
   mutate(model_prob = case_when(
-    model_prob == "NA (NA)" ~ "NA",
+    model_prob == "NA (NA)" ~ "-",
     model_prob != "NA (NA)" ~ model_prob
@@
   kbl(format = "latex",
-      caption = "Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
+      caption = "Estuary GAM AIC\\textsubscript{c} values and associated model probabilities (in parenthesis).
 Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.",
-      booktabs = TRUE)
-  # flextable() |>   
-  # set_caption(caption = "Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
-Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.",
-  #             autonum = run_autonum(seq_id = "tab", bkm = "table4")) |> 
-  # style(pr_t = fp_text_default(font.size = 9), part = "body") |>
-  # style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
-  # set_table_properties(layout = "autofit") |>
-  # padding(padding.top = 0, padding.bottom = 0, part = "body")
+      booktabs = TRUE, escape = FALSE) |> 
+  collapse_rows(columns = 1, latex_hline = "major")
 
@@
     geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
-    # geom_text_repel(aes(label = label,
-    #                     color = model),
-    #                 nudge_x = 5,
-    #                 hjust = "left",
-    #                 size = 2) +
     scale_x_continuous(expand = expansion(mult = c(0.05,0.05))) +
@@
                       xlab = "Residuals log(Inflow) [cfs]",
-                      ylab = "NO*<sub>x</sub>* Smooth<br>Estimate [mg/L]",
-                      subtitle = "B: NO*<sub>x</sub>*")
+                      ylab = "NO<sub>*x*</sub> Smooth<br>Estimate [mg/L]",
+                      subtitle = "B: NO<sub>*x*</sub>")
 
@@
 lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
-provided only small improvements to the model (@tbl-table4). As noted
+provided only small improvements to the model (Table\ \ref{tab:table4}). As noted
 above, nutrient loadings did not provide any explanation in changes in the
@@
                       xlab = "Residuals log(NO<sub>3</sub>) Load [kg]",
-                      ylab = "NO*<sub>x</sub>* [mg/L]",
-                      subtitle = "B: NO*<sub>x</sub>*")
+                      ylab = "NO<sub>*x*</sub> [mg/L]",
+                      subtitle = "B: NO<sub>*x*</sub>")
 
@@
 Lavaca River are within the ranges in previous published studies
-(@tbl-table5; [@dunnTrendsNutrientInflows1996; @rebichSourcesDeliveryNutrients2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]). It isn't obvious
+[Table\ \ref{tab:table5}, @dunnTrendsNutrientInflows1996; @rebichSourcesDeliveryNutrients2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]. It isn't obvious
 why TP estimates in @dunnTrendsNutrientInflows1996 were notably lower. Given that none

### 2023-05-20 Schramm-2023-05-AS/Schramm-2023-05-AS.tex
fix cross refs

@@
 models were fit at each site relating parameter concentration to
-temporal trends (\citep{eq-eq5}), temporal trends and inflow
-(\citet{eq-eq6}), and temporal trends, inflow, and nutrient loads
-(\citet{eq-eq7}): \begin{align}
+temporal trends (Eq.~\ref{eq5}), temporal trends and inflow
+(Eq.~\ref{eq6}), and temporal trends, inflow, and nutrient loads
+(Eq.~\ref{eq7}): \begin{align}
 g(\mu) &= \alpha + f_1(ddate) + f_2(yday), \label{eq5} \\
@@
 watershed with the Navidad River only accounting for 15\% and 25\% of
-NO\textsubscript{3} and TP loads in 2011 (Figure~\ref{tab:fig4}).
+NO\textsubscript{3} and TP loads in 2011 (Figure~\ref{fig:fig4}).
 
@@
 
-\caption{\label{tab:table4}Estuary GAM AIC~c~ values and associated model probabilities (in parenthesis).
+\caption{\label{tab:table4}Estuary GAM AIC\textsubscript{c} values and associated model probabilities (in parenthesis).
 Models with the highest probability for each site and water quality parameter combination are bolded and italicized for emphasis.}
@@
 \midrule
-TP & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & -158.2 (0.72)\\
-TP & TCEQ-13384 & -194.4 (0.03) & -200.2 (0.49) & -200.2 (0.49)\\
-TP & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & -157.3 (0.59)\\
-NOx & TCEQ-13383 & -218.9 (0) & -244.8 (0.5) & -244.8 (0.5)\\
-NOx & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & -311.9 (0.52)\\
-\addlinespace
-NOx & TCEQ-13563 & -175.1 (0) & -190.2 (0.5) & -190.2 (0.5)\\
-Chlorophyll-*a* & TCEQ-13383 & 279.7 (0.18) & 278.1 (0.41) & 278.1 (0.41)\\
-Chlorophyll-*a* & TCEQ-13384 & 268.2 (0.33) & 268.2 (0.33) & 268.2 (0.33)\\
-Chlorophyll-*a* & TCEQ-13563 & 289.5 (0.08) & 286.1 (0.46) & 286.1 (0.46)\\
-TKN & TCEQ-13383 & 42.2 (0.66) & 43.5 (0.34) & NA\\
-\addlinespace
-TKN & TCEQ-13384 & 34.3 (0.57) & 34.8 (0.43) & NA\\
-TKN & TCEQ-13563 & 31.1 (0.22) & 28.7 (0.78) & NA\\
-DO & TCEQ-13383 & 146.4 (0.34) & 146.4 (0.34) & 146.5 (0.32)\\
-DO & TCEQ-13384 & 135.9 (0.47) & 137 (0.27) & 137 (0.27)\\
-DO & TCEQ-13563 & 138.3 (0.25) & 137.2 (0.43) & 137.8 (0.32)\\
+ & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & -158.2 (0.72)\\
+
+ & TCEQ-13384 & -194.4 (0.03) & -200.2 (0.49) & -200.2 (0.49)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash TP} & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & -157.3 (0.59)\\
+\cmidrule{1-5}
+ & TCEQ-13383 & -218.9 (0) & -244.8 (0.5) & -244.8 (0.5)\\
+
+ & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & -311.9 (0.52)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & TCEQ-13563 & -175.1 (0) & -190.2 (0.5) & -190.2 (0.5)\\
+\cmidrule{1-5}
+ & TCEQ-13383 & 279.7 (0.18) & 278.1 (0.41) & 278.1 (0.41)\\
+
+ & TCEQ-13384 & 268.2 (0.33) & 268.2 (0.33) & 268.2 (0.33)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & TCEQ-13563 & 289.5 (0.08) & 286.1 (0.46) & 286.1 (0.46)\\
+\cmidrule{1-5}
+ & TCEQ-13383 & 42.2 (0.66) & 43.5 (0.34) & -\\
+
+ & TCEQ-13384 & 34.3 (0.57) & 34.8 (0.43) & -\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash TKN} & TCEQ-13563 & 31.1 (0.22) & 28.7 (0.78) & -\\
+\cmidrule{1-5}
+ & TCEQ-13383 & 146.4 (0.34) & 146.4 (0.34) & 146.5 (0.32)\\
+
+ & TCEQ-13384 & 135.9 (0.47) & 137 (0.27) & 137 (0.27)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash DO} & TCEQ-13563 & 138.3 (0.25) & 137.2 (0.43) & 137.8 (0.32)\\
 \bottomrule
@@
 TCEQ-13384 site. The effect was quite small compared to streamflow and
-provided only small improvements to the model (\citet{tbl-table4}). As
-noted above, nutrient loadings did not provide any explanation in
+provided only small improvements to the model (Table~\ref{tab:table4}).
+As noted above, nutrient loadings did not provide any explanation in
 changes in the remaining assessed water quality parameters.
@@
 the Lavaca River are within the ranges in previous published studies
-(\citet{tbl-table5};
-\citep{dunnTrendsNutrientInflows1996, rebichSourcesDeliveryNutrients2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019}).
+\citep[Table~\ref{tab:table5},][]{dunnTrendsNutrientInflows1996, rebichSourcesDeliveryNutrients2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019}.
 It isn't obvious why TP estimates in

### 2023-05-20 Schramm-2023-05-AS/Schramm-2023-05-AS.Rmd
fix captions

@@
 
-```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525."}
+```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525."}
 
@@
 
-```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."}
+```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO\\textsubscript{\\textit{x}}, chlorophyll-\\textit{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."}
 
@@
 
-```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.45, fig.cap="Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs."}
+```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.45, fig.cap="Estimated effects of nutrient load residuals on TP and NO\\textsubscript{\\textit{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs."}
 

### 2023-05-20 Schramm-2023-05-AS/Schramm-2023-05-AS.tex
fix captions

@@
 
-\caption{Aggregated estimated annual and flow-normalized annual NO~3~ and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
+\caption{Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
 \end{figure}
@@
 
-\caption{Estimated effects of mean daily inflow residuals on mean TP, NO~*x*~, chlorophyll-*a*, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.}\label{fig:fig6}
+\caption{Estimated effects of mean daily inflow residuals on mean TP, NO\textsubscript{\textit{x}}, chlorophyll-\textit{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.}\label{fig:fig6}
 \end{figure}
@@
 
-\caption{Estimated effects of nutrient load residuals on TP and NO~*x*~ concentrations in Lavaca Bay obtained from flow+load estuary GAMs.}\label{fig:fig7}
+\caption{Estimated effects of nutrient load residuals on TP and NO\textsubscript{\textit{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.}\label{fig:fig7}
 \end{figure}

### 2023-05-23 Schramm-2023-05-AS/Schramm-2023-05-AS.Rmd
lots of edits

@@
 abstract: |
-  Lavaca Bay is a small secondary embayment on the Texas coast that is displaying early signals of water quality degradation.
-This study applied a semiparametric approach to assess both watershed nutrient loads and the responses in estuary water quality to nutrient loading and streamflow.
+  Lavaca Bay is a secondary embayment on the Texas coast displaying early signals of water quality degradation.
+This study applied a semiparametric approach to assess both watershed nutrient loads and responses in estuary water quality to nutrient loading and streamflow.
 Cross-validation indicated that, despite data constraints, semiparametric models performed well at nutrient load prediction.
 Based on these models, delivered annual nutrient loads varied substantially from year to year.
-In contrast, minimal changes in calculated flow-normalized loads indicate that changes in nutrient loading were driven by natural variation in precipitation and runoff as opposed to changes in nonpoint sources.
-Estuary water quality models did not identify significant long-term changes within Lavaca Bay for dissolved oxygen or chlorophyll-*a*.
+In contrast, minimal changes in flow-normalized loads indicate that nutrient loadings were driven by natural variation in precipitation and runoff as opposed to changes in management of nonpoint sources.
+Models indicated there was no evidence of long-term changes in dissolved oxygen or chlorophyll-*a* within Lavaca Bay.
 However, site specific long-term increases in both organic and inorganic nitrogen are concerning.
-Further analysis found freshwater inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that changes in nutrient loading explained variation in dissolved oxygen or chlorophyll-*a* concentrations.
+Further analysis found freshwater inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that changes in watershed nutrient loading explained variation in dissolved oxygen or chlorophyll-*a* concentrations.
 In addition to providing a baseline assessment of watershed nutrient loading and water quality responses in the Lavaca Bay watershed, this study provides methodological support for the use of semiparametric methods in load regression models and estuary assessments.
@@
 
+*Remaining Items*
+
+- keywords
+- discussion
+- formalize language (reduce first person active?)
+- final language review
+- springer doesn't want figures in subdirectory, need to knit to same directory as tex file.
+
 # Introduction {#sec1}
 
-Like many coastal areas globally, the coastal watersheds along the Texas Gulf coast 
+Similar to many coastal areas globally, the coastal watersheds along the Texas Gulf coast 
 are facing pressures from increasing population, increases in point source 
@@
 estuary water quality dynamics in both agriculturally dominated and urban 
-watersheds within Texas are in fact expressing conditions that are increasingly 
+watersheds within Texas are expressing conditions that are increasingly 
 conducive to algal blooms and eutrophication [@wetzWaterQualityDynamics2016; 
 @wetz_exceptionally_2017; @bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022]. 
-With identification of localized areas of estuary water quality concern along 
+With identification of several localized areas of estuary water quality concern along 
 the Texas coast [@bugica_water_2020], localized studies are being prioritized 
@@
 
-This project aims to provide an assessment of nutrient loading and water quality
+This project aims to provide an assessment of watershed nutrient loading and resulting water quality
 responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
@@
 (DO), nutrients, and biological parameters have been well within state
-water quality standards. More recently identified long-term declines 
-in abundance, biomass, and diversity of benthic fauna in Lavaca Bay have been 
-linked to reductions in freshwater inflows and changes in estuary salinity 
+water quality standards. Despite largely meeting statewide water quality standards,
+there have been concerning recently identified declines in abundance, biomass, and diversity of 
+benthic fauna in Lavaca Bay [@beserespollackLongtermTrendsResponse2011]. These declines
+are partially attributed to reductions in freshwater inflow and changes in estuary salinity
+and are indicative of an already stressed system 
 [@beserespollackLongtermTrendsResponse2011; @palmerImpactsDroughtsLow2015;
-@montagnaAssessmentRelationshipFreshwater2020] and are a concern to local 
-stakeholders. Water quality assessments identified monotonic increases 
-in total phosphorus (TP), orthophosphate, total Kjeldahl nitrogen (TKN), and
-chlorophyll-*a* at sites within Lavaca Bay [@bugica_water_2020]. Although 
-long-term changes in DO concentrations were not identified, the trends in 
+@montagnaAssessmentRelationshipFreshwater2020]. More recently, significant
+linear increases in total phosphorus (TP), orthophosphate, 
+total Kjeldahl nitrogen (TKN), and chlorophyll-*a* concentrations were identified
+at monitoring sites within Lavaca Bay [@bugica_water_2020]. Although 
+long-term changes in DO concentrations have not been identified, the trends in 
 nutrient concentrations are concerning due to the role of nitrogen as a 
@@
 [@wazniakLinkingWaterQuality2007; @lloydMethodsDetectingChange2014]. 
-The development and application of statistical methods such as Weighted
+The development and application of flexible statistical methods such as Weighted
 Regressions on Time, Discharge and Season [WRTDS, @hirschWeightedRegressionsTime2010] 
-and Generalized Additive Models [GAMs, @woodFastStableRestricted2011] has 
+and Generalized Additive Models [GAMs, @woodFastStableRestricted2011] have 
 provided effective tools for researchers to quantify and
@@
 
-WRTDS calculates a time series of in stream concentrations or loads 
+WRTDS calculates a time series of in-stream concentrations or loads 
 (daily, monthly, or annually) and flow-normalized estimates of concentrations and loads
@@
 procedure of GAMs is substantially different than WRTDS, both the functional 
-form and results have been demonstrated to be similar 
-[@beckNumericalQualitativeContrasts2017]. Water quality applications of GAMs
-include river and catchment nutrient concentration and load models 
+form and results have been demonstrated to be similar when assessing nutrient
+concentration trends [@beckNumericalQualitativeContrasts2017]. 
+Water quality applications of GAMs have included include river and catchment 
+nutrient concentration and load estimation
 [@wangLoadEstimationUncertainties2011;@kroonRiverLoadsSuspended2012;
@@
 @hagemannEstimatingNutrientOrganic2016;@mcdowell_implications_2021;
-@biagi_novel_2022]. GAMs can also be used to identify non-linear temporal trends
-(including flow-normalized trends) in pollutant concentrations and loads 
-[@beckNumericalQualitativeContrasts2017; @murphyGeneralizedAdditiveModel2019]. 
+@biagi_novel_2022], assessment of temporal trends of nutrients
+[@beckNumericalQualitativeContrasts2017; @murphyGeneralizedAdditiveModel2019],
+phytoplankton [@bergbuschUnexpectedShiftPhytoplankton2021], 
+and cyanobacteria [@hayesEffectsLakeWarming2020]. 
 Recently GAMs have also been used to link water quality responses in 
-receiving water bodies to changes in nutrient inputs 
+receiving water bodies to changes in nonpoint source nutrient inputs 
 [@murphyNutrientImprovementsChesapeake2022]. @beckNumericalQualitativeContrasts2017
@@
 the response to linear predictors thus avoiding unneeded data transformations
-and bias corrections. The study also assesses the response of water quality parameters in
+and bias corrections. The exploratory study also assesses the response of water quality parameters in
 Lavaca Bay over time and in response to freshwater inflow controlled
@@
 are required to be collected under Quality Assurance Project Plans and lab 
-method procedures outlined by the TCEQ's procedures manual. 
-The QAPP and procedures manuals ensure the consistent collection and laboratory 
-methods are applied between samples collected by different entities and under 
-different projects. 
+method procedures outlined by the TCEQ's procedures manual to ensure
+consistent collection and laboratory methods are applied between samples 
+collected by different entities and under different projects. 
 All sites had varying lengths of and availability of data.
@@
            site_no == "usgs08164525") |> 
-  rename(`NO3 (mg/L)` = original_NO3,
+  rename(`NO\\textsubscript{3} (mg/L)` = original_NO3,
          `TP (mg/L)` = original_TP,
@@
 
-datasummary(`Station ID` * (`TP (mg/L)` + `NO3 (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
+datasummary(`Station ID` * (`TP (mg/L)` + `NO\\textsubscript{3} (mg/L)` + `Mean Daily Streamflow (cfs)`) ~
                Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-            data = df, output = "kableExtra", format = "latex",
-            caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.")
+            data = df, output = "data.frame") |> 
+  mutate(`Station ID` = c("USGS-08164000", "USGS-08164000", "USGS-08164000",
+                          "USGS-08164525", "USGS-08164525", "USGS-08164525")) |> 
+  rename("Parameter" = " ") |> 
+  kbl(format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.") |>
+  collapse_rows(columns = 1, latex_hline = "major")
 
@@
            case_when(
-             parameter_code == "70953" ~ "Chlorophyll-a (ug/L)",
+             parameter_code == "70953" ~ "Chlorophyll-\\textit{a} (ug/L)",
              parameter_code == "00625" ~ "TKN (mg/L)",
-             parameter_code == "00630" ~ "NOx (mg/L)",
+             parameter_code == "00630" ~ "NO\\textsubscript{x} (mg/L)",
              parameter_code == "00665" ~ "TP (mg/L)",
@@
 modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
-                           `NOx (mg/L)` +
+                           `NO\\textsubscript{x} (mg/L)` +
                            `TKN (mg/L)` +
-                           `Chlorophyll-a (ug/L)` +
+                           `Chlorophyll-\\textit{a} (ug/L)` +
                            `DO (mg/L)`) ~
                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-                         data = df, output = "kableExtra", format = "latex",
-                         caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.")
+                         data = df, output = "data.frame") |> 
+  mutate(`Station ID` = c(rep("TCEQ-13383", 5), rep("TCEQ-13384", 5),
+                          rep("TCEQ-13563", 5))) |> 
+  rename("Parameter" = " ") |> 
+  kbl(format = "latex", booktabs = TRUE, escape = FALSE,
+      caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.") |> 
+  collapse_rows(columns = 1, latex_hline = "major")
+
+# modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
+#                            `NO\\textsubscript{x} (mg/L)` +
+#                            `TKN (mg/L)` +
+#                            `Chlorophyll-\\textit{a} (ug/L)` +
+#                            `DO (mg/L)`) ~
+#                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
+#                          data = df, output = "kableExtra", format = "latex",
+#                          escape = FALSE,
+#                          caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.")
 ```
@@
 *ltfa* and moderate improvements in TP regression models that incorporated 
-*stfa* [@zhang_improving_2017].
+*stfa* [@zhang_improving_2017]. Moving averages and flow anomalies were calculated
+with the *adc* R package [@schrammAdcCalculateAntecedant2023].
 
@@
 each day considering all possible flow values.
-Standard deviations and credible intervals were obtained by 
+Standard deviations and 90% credible intervals were obtained by 
 drawing samples from the multivariate normal posterior distribution of the 
@@
 @marraCoveragePropertiesConfidence2012; @mcdowell_implications_2021]. 
-Uncertainty in loads were calculated as 90% credible intervals estimated by 
-drawing 1000 realizations of parameter estimates from the multivariate normal 
-posterior distribution of the model parameters. GAM performance was evaluated 
+GAM performance was evaluated 
 using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
@@
 methodology implemented by @murphyNutrientImprovementsChesapeake2022,
-both streamflow and nutrient loads were prepossessed to account for season
-and flow. Freshwater inflow and nutrient loads were replaced by 
-seasonally adjusted log transformed inflow and flow-adjusted 
-log transformed nutrient loads obtained by fitting a GAM relating season 
-(day of year) to log transformed daily freshwater inflow values
-and a GAM relating log transformed NO~3~ or TP loads to log transformed daily
-inflow.Response residuals from the respective GAM models were 
+both freshwater inflow and nutrient loads were prepossessed to account for season
+and streamflow respectively. Raw freshwater inflow values were replaced by 
+seasonally adjusted log transformed inflow obtained from the residuals of a GAM model 
+fit between season(day of year) and log transformed daily freshwater inflow. Raw nutrient
+loads were replaced with flow-adjusted values obtained from the residuals of a
+GAM model relating log transformed NO~3~ or TP loads to log transformed daily
+inflow. Response residuals from the respective GAM models were 
 used as *Q* and *Load* in Eq.\ \ref{eq6} and Eq.\ \ref{eq7}.
@@
 This study used an information theoretic approach to evaluate 
-if nutrient loads and/or freshwater inflows provided evidence of effects on 
-water quality concentrations in Lavaca Bay. Model probabilities were calculated 
-and compared using the AIC~c~ scores between each group of temporal, flow, and 
-flow+load models [@burnhamAICModelSelection2011]. Improvements in model 
+evidence of model covariate effects on Lavaca Bay water quality.
+Model probabilities were calculated 
+and compared using the AIC~c~ scores between each group of temporal, inflow, and 
+inflow+load models [@burnhamAICModelSelection2011]. Improvements in model 
 probabilities provide evidence that the terms explain additional variation in 
-the response variable. If model probabilities were tied, there wasn't evidence 
-the more complicated model explains additional variation in water quality.
+the response variable.
 
@@
 
-Using evaluation criteria recommended by @moriasiHydrologicWaterQuality2015, 
-predictive performance of nutrient loading GAMs ranged from "satisfactory" to
-"very good" based on median NSE, *r*, and PBIAS metrics calculated using 5-fold 
-cross validation. Median goodness-of-fit metrics for NO~3~ models in the Lavaca 
+Predictive performance of nutrient loads ranged from "satisfactory" to "very good"
+based on standardized evaluation metrics of NSE, *r*, and PBIAS 
+[@moriasiHydrologicWaterQuality2015] calculated using 5-fold cross validation.
+Median goodness-of-fit metrics for NO~3~ models in the Lavaca 
 River were 0.34 NSE, 0.70 *r*, and 2.00 PBIAS. Navidad River NO~3~ models 
@@
 
-Predicted annual NO~3~ and TP loads show considerable variation, generally following
+Annual NO~3~ and TP loads show considerable variation, generally following
 patterns in discharge (Figures\ \ref{fig:fig3}, \ref{fig:fig4}). 
@@
 
-GAMs did not identify significant changes in TP or DO concentrations at any
-of the Lavaca Bay sites from 2005 through 2020 (Figure\ \ref{fig:fig5}). The
-upper-bay site, TCEQ-13563, had a linear increase in NO~x~ concentration and
-and decrease in chlorophyll-*a* from 2005 through 2014. The mid-bay site,
-TCEQ-13383, showed a periodic pattern in NO~x~ concentration that appeared
-similar to precipitation/inflow patterns, as well as a post 2011 increase
-in TKN concentrations. No significant long-term trends in concentrations 
-were identified by GAMs for the lower-bay TCEQ-13384 site.
-
-```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs."}
+There is no evidence of long-term changes in TP or DO concentrations at any
+Lavaca Bay site (Figure\ \ref{fig:fig5}). The upper-bay site, TCEQ-13563, shows
+evidence of a long-term linear increase in NO~x~ while chlorophyll-*a*
+decreased from 2005 through 2014 (Figure\ \ref{fig:fig5}). NO~x~ concentration
+at the mid-bay site, TCEQ-13383, displayed an unusual periodic pattern that is
+indicative of a strong influence from inflow or precipitation. The temporal GAMs
+did not provide evidence of long-term trends in any of the water quality constituents
+at the lower-bay TCEQ-13384 site.
+
+```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{x}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
 
@@
     smooths = "s(ddate)",
-    data = newdata)
+    data = newdata,
+    overall_uncertainty = FALSE)
 
@@
     ggplot(aes(x = ddate, y = est, group = model)) +
+    geom_line(aes(color = model, linetype = model), alpha = 0.75) +
     geom_ribbon(aes(ymin = lower_ci,
@@
                     fill = model),
-                alpha = 0.15) +
-    geom_line(aes(color = model)) +
-    geom_hline(yintercept = 0, linetype = 2, alpha = 0.25) +
+                alpha = 0.25) +
     scale_x_continuous(breaks = c(2005, 2010, 2015, 2020),
                        expand = expansion(mult = c(0.05,0.05))) +
-    colorspace::scale_fill_discrete_qualitative(name = "") +
-    colorspace::scale_color_discrete_qualitative(name = "") +
+    scale_fill_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_color_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_linetype_manual(name = "", values = c("solid","twodash","dashed")) +
     labs(x = "Year", y = ylab,
          subtitle = subtitle) +
-    theme_TWRI_print(base_family = "sans") +
+    theme_TWRI_print(base_family = "Arial") +
     theme(axis.title.y = element_markdown(size = 8),
@@
           axis.text.y = element_text(size = 8),
+          panel.grid = element_blank(),
           plot.subtitle = element_text(size = 8),
           legend.direction = "vertical",
-          legend.text = element_text(size = 8))
+          legend.text = element_text(size = 8),
+          legend.key.width = unit(25, "points"))
 
@@
                              ylab = "TKN [mg/L]", 
-                             subtitle = "C")
+                             subtitle = "D")
 
@@
                              ylab = "Chlorophyll-*a* [μg/L]", 
-                             subtitle = "D")
+                             subtitle = "C")
 
@@
     Parameter == "Chlorophyll-a" ~ "Chlorophyll-\\textit{a}",
-    TRUE ~ as.character(Parameter)
+    TRUE ~ as.character(Parameter)),
+    Model = case_when(
+      Model == "Temporal" ~ "Temporal",
+      Model == "Flow" ~ "Inflow",
+      Model == "Flow + Load" ~ "Inflow + Load",
+      TRUE ~ as.character(Model)
     )) |>
   mutate(Parameter = fct_relevel(Parameter, "TP", "NO\\textsubscript{x}", "Chlorophyll-\\textit{a}","TKN", "DO"),
-         Model = fct_relevel(Model, "Temporal", "Flow", "Flow + Load")) |>
+         Model = fct_relevel(Model, "Temporal", "Inflow", "Inflow + Load")) |>
   tidyr::complete(Site, Model, nesting(Parameter)) |>
@@
   pivot_wider(names_from = Model, values_from = model_prob) |>
-  select(Parameter, Site, Temporal, Flow, `Flow + Load`) |>
+  select(Parameter, Site, Temporal, Inflow, `Inflow + Load`) |>
   arrange(Parameter, Site) |>
@@
       booktabs = TRUE, escape = FALSE) |> 
+  column_spec(column = 3, 
+              bold = c(F,F,F,F,F,F,F,T,F,T,T,F,T,T,F),
+              italic = c(F,F,F,F,F,F,F,T,F,T,T,F,T,T,F)) |> 
+  column_spec(column = 4, 
+              bold = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T),
+              italic = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T)) |> 
+  column_spec(column = 5,
+              bold = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F),
+              italic = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F)) |> 
   collapse_rows(columns = 1, latex_hline = "major")
@@
 
-```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Estimated effects of mean daily inflow residuals on mean TP, NO\\textsubscript{\\textit{x}}, chlorophyll-\\textit{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs."}
+```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\\textit{\\textsubscript{x}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
 
@@
     ggplot(aes(x = {{ x }}, y = est, group = model)) +
+    geom_line(aes(color = model, linetype = model)) +
     geom_ribbon(aes(ymin = lower_ci,
@@
                     fill = model),
-                alpha = 0.15) +
-    geom_line(aes(color = model)) +
-    geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
+                alpha = 0.25) +
     scale_x_continuous(expand = expansion(mult = c(0.05,0.05))) +
-    colorspace::scale_fill_discrete_qualitative(name = "") +
-    colorspace::scale_color_discrete_qualitative(name = "") +
+    scale_fill_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_color_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_linetype_manual(name = "", values = c("solid","twodash","dashed")) +
     labs(x = xlab, y = ylab,
          subtitle = subtitle) +
-    theme_TWRI_print(base_family = "sans") +
+    theme_TWRI_print(base_family = "Arial") +
     theme(axis.title.y = element_markdown(size = 8),
           axis.title.x = element_markdown(size = 8),
-          plot.subtitle = element_markdown(size = 8),
+          axis.text.x = element_text(size = 8),
+          axis.text.y = element_text(size = 8),
+          panel.grid = element_blank(),
+          plot.subtitle = element_text(size = 8),
           legend.direction = "vertical",
-          legend.text = element_text(size = 8))
+          legend.text = element_text(size = 8),
+          legend.key.width = unit(25, "points"))
 
@@
                       `TCEQ-13563` = tar_read(tp_lavaca_13563_flow, store = store),
-                      smooths = c("s(flw_res)"))
+                      smooths = c("s(flw_res)"),
+                      overall_uncertainty = FALSE)
 p_tp <- draw_smoothed_surface(comp,
@@
                       ylab =  "TP Smooth<br>Estimate [mg/L]",
-                      subtitle = "A: TP")
+                      subtitle = "A")
 
@@
                       `TCEQ-13563` = tar_read(no3_lavaca_13563_flow, store = store),
-                      smooths = c("s(flw_res)"))
+                      smooths = c("s(flw_res)"),
+                      overall_uncertainty = FALSE)
 p_no3 <- draw_smoothed_surface(comp,
@@
                       ylab = "NO<sub>*x*</sub> Smooth<br>Estimate [mg/L]",
-                      subtitle = "B: NO<sub>*x*</sub>")
+                      subtitle = "B")
 
@@
                       `TCEQ-13563` = tar_read(chla_lavaca_13563_flow, store = store),
-                      smooths = c("s(flw_res)"))
+                      smooths = c("s(flw_res)"),
+                      overall_uncertainty = FALSE)
 p_chla <- draw_smoothed_surface(comp,
@@
                       ylab = "Chlorophyll-*a* Smooth<br>Estimate [µg/L]",
-                      subtitle = "C: Chlorophyll-*a*")
+                      subtitle = "C")
 
@@
                       `TCEQ-13563` = tar_read(tkn_lavaca_13563_flow, store = store),
-                      smooths = c("s(flw_res)"))
+                      smooths = c("s(flw_res)"),
+                      overall_uncertainty = FALSE)
 p_tkn <- draw_smoothed_surface(comp,
@@
                       ylab = "TKN Smooth<br>Estimate [mg/L]",
-                      subtitle = "D: TKN")
+                      subtitle = "D")
 
@@
                       `TCEQ-13563` = tar_read(do_lavaca_13563_flow, store = store),
-                      smooths = c("s(flw_res)"))
+                      smooths = c("s(flw_res)"),
+                      overall_uncertainty = FALSE)
 p_do <- draw_smoothed_surface(comp,
@@
                       ylab = "DO Smooth<br>Estimate [mg/L]",
-                      subtitle = "E: DO")
+                      subtitle = "E")
 
@@
 
-```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.45, fig.cap="Estimated effects of nutrient load residuals on TP and NO\\textsubscript{\\textit{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs."}
+```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.4, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\\textsubscript{3} loads on TP (A) and NO\\textsubscript{\\textit{x}} (B) concentrations at each site in Lavaca Bay."}
 
@@
                       `TCEQ-13563` = tar_read(tp_lavaca_13563_full, store = store),
-                      smooths = c("s(TP_resid)"))
+                      smooths = c("s(TP_resid)"),
+                      overall_uncertainty = FALSE)
 p_tp <- draw_smoothed_surface(comp,
@@
                       ylab = "TP [mg/L]",
-                      subtitle = "A: TP")
+                      subtitle = "A")
 
@@
                       `TCEQ-13563` = tar_read(no3_lavaca_13563_full, store = store),
-                      smooths = c("s(NO3_resid)"))
+                      smooths = c("s(NO3_resid)"),
+                      overall_uncertainty = FALSE)
 p_no3 <- draw_smoothed_surface(comp,
@@
                       ylab = "NO<sub>*x*</sub> [mg/L]",
-                      subtitle = "B: NO<sub>*x*</sub>")
+                      subtitle = "B")
 
@@
 
-# Discussion
+# Discussion {#sec4}
 
@@
 
+## Watershed Nutrient Loads
+
 TP and NO~3~ loadings from the Lavaca Bay watershed showed high inter-annual
-variability tied with changes in discharge. There is little evidence for changes
-in flow-normalized TP loads in either rivers. There is some evidence of recent
-decreases in flow-normalized NO~3~ loads in the Lavaca River. Although there is
-no work directly correlating water quality planning and implementation efforts
-in the watershed to water quality outcomes, efforts to increase agricultural 
-producer participation in the watershed have been ongoing since 2016
-[@schrammLavacaRiverWatershed2018; @bertholdDirectMailingEducation2021]. The 
-decrease in flow-normalized NO~3~ loads could be a reflection of those 
-collective efforts but further data collection and research is required to 
-support that statement.
+variability driven primarily by fluctuations in discharge. Notably, there were
+no indications of trends in flow-normalized NO~3~ and TP loads in the Navidad River.
+Freshwater discharges in the Navidad River are regulated by the Palmetto Dam forming
+Lake Texana at the lower extent of the river. While the dominant agricultural land uses
+differ between the Lavaca (primarily grazed pasture and rangeland) and Navidad 
+(mix of pasture and row crops) catchments, we did not have a reason to expect 
+different flow normalized trends between the two systems from land use alone.
+Lentic nitrogen uptake and cycling may have regulating effects that mask changes
+in upstream nitrogen loadings. Additional nutrient data collection in the contributing 
+tributaries of Lake Texana is needed to fully assess the role of Lake Texana
+in regulating nutrient delivery to the Lavaca Bay system. However, these results
+suggest that there have been no changes in the NO~3~ or TP loading from Lake Texana
+when accounting for variations in year to year discharge.
+
+There is some evidence of recent decreases in flow-normalized NO~3~ loads in 
+the Lavaca River. Although there is no work directly correlating water quality 
+planning and implementation efforts in the watershed to water quality outcomes, 
+efforts to increase agricultural producer participation in the watershed have 
+been ongoing since 2016 [@schrammLavacaRiverWatershed2018; 
+@bertholdDirectMailingEducation2021]. The decrease in flow-normalized NO~3~ 
+loads could be a reflection of those collective efforts but the lack of evidence
+for similar changes in flow-normalized TP loads provide contrary support. The
+inconsistent flow-normalized trends may also reflect some of the weakness of 
+the water quality dataset that is primarily composed of ambient water quality
+measurements.
+
+Cross-validation of the nutrient loading models highlights that predictions are
+prone to high bias, owing to the lack of targeted storm or flow biased measurements.
+The high biases are indicative that subsets of values were unable to capture the
+functional relationships with the flow based dependent variables. It was beyond the scope
+of the current study to evaluate the subsets of cross-validation data and scores.
+However, the cross-validation procedure is indicative that more robust sampling
+would be beneficial for reducing prediction variance. Supplementary flow-biased 
+monitoring targeting storm- or high-flow conditions is critical to
+improve model performance and strength of evidence produced by these models [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017].
+There is no absolute guidance for the number of samples needed...
+Would be useful to develop guidance for apply GAMs to estimate loads and estimate trends of different effect sizes.
 
@@
 
-Cross-validation of the GAM loading models indicated that GAMs performed well on
-average at predicting daily nutrient loading values. The variance in scores was
-very high indicating subsets of values were problematic at characterizing functional
-relationships between nutrients and predictors. Because all of the water quality
-data for these two locations in the TCEQ databases were ambient water quality
-data, collected to be representative of typical flow conditions, there were few
-data at the highest portions of the flow-duration curve. It was beyond the scope
-of the current study to evaluate the subsets of cross-validation data and scores.
-However, the cross-validation procedure is indicative that more robust sampling
-would be beneficial for reducing prediction variance. Supplementary flow-biased 
-monitoring targeting storm- or high-flow conditions is recommended here to
-improve the precision of GAM predictions [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017].
+## Estuary Water Quality Responses
 

### 2023-05-23 Schramm-2023-05-AS/Schramm-2023-05-AS.tex
lots of edits

@@
 \usepackage{xcolor}
-\usepackage{siunitx}
-
-  \newcolumntype{d}{S[
-    input-open-uncertainty=,
-    input-close-uncertainty=,
-    parse-numbers = false,
-    table-align-text-pre=false,
-    table-align-text-post=false
-  ]}
 
@@
 
-\abstract{Lavaca Bay is a small secondary embayment on the Texas coast
-that is displaying early signals of water quality degradation. This
-study applied a semiparametric approach to assess both watershed
-nutrient loads and the responses in estuary water quality to nutrient
-loading and streamflow. Cross-validation indicated that, despite data
-constraints, semiparametric models performed well at nutrient load
-prediction. Based on these models, delivered annual nutrient loads
-varied substantially from year to year. In contrast, minimal changes in
-calculated flow-normalized loads indicate that changes in nutrient
-loading were driven by natural variation in precipitation and runoff as
-opposed to changes in nonpoint sources. Estuary water quality models did
-not identify significant long-term changes within Lavaca Bay for
-dissolved oxygen or chlorophyll-\emph{a}. However, site specific
-long-term increases in both organic and inorganic nitrogen are
-concerning. Further analysis found freshwater inflow was a strong driver
-of nutrient and chlorophyll-\emph{a} concentrations within Lavaca Bay
-but there was no evidence that changes in nutrient loading explained
-variation in dissolved oxygen or chlorophyll-\emph{a} concentrations. In
-addition to providing a baseline assessment of watershed nutrient
-loading and water quality responses in the Lavaca Bay watershed, this
-study provides methodological support for the use of semiparametric
-methods in load regression models and estuary assessments.}
+\abstract{Lavaca Bay is a secondary embayment on the Texas coast
+displaying early signals of water quality degradation. This study
+applied a semiparametric approach to assess both watershed nutrient
+loads and responses in estuary water quality to nutrient loading and
+streamflow. Cross-validation indicated that, despite data constraints,
+semiparametric models performed well at nutrient load prediction. Based
+on these models, delivered annual nutrient loads varied substantially
+from year to year. In contrast, minimal changes in flow-normalized loads
+indicate that nutrient loadings were driven by natural variation in
+precipitation and runoff as opposed to changes in management of nonpoint
+sources. Models indicated there was no evidence of long-term changes in
+dissolved oxygen or chlorophyll-\emph{a} within Lavaca Bay. However,
+site specific long-term increases in both organic and inorganic nitrogen
+are concerning. Further analysis found freshwater inflow was a strong
+driver of nutrient and chlorophyll-\emph{a} concentrations within Lavaca
+Bay but there was no evidence that changes in watershed nutrient loading
+explained variation in dissolved oxygen or chlorophyll-\emph{a}
+concentrations. In addition to providing a baseline assessment of
+watershed nutrient loading and water quality responses in the Lavaca Bay
+watershed, this study provides methodological support for the use of
+semiparametric methods in load regression models and estuary
+assessments.}
 
@@
 
+\emph{Remaining Items}
+
+\begin{itemize}
+\tightlist
+\item
+  keywords
+\item
+  discussion
+\item
+  formalize language (reduce first person active?)
+\item
+  final language review
+\item
+  springer doesn't want figures in subdirectory, need to knit to same
+  directory as tex file.
+\end{itemize}
+
 \hypertarget{sec1}{%
@@
 
-Like many coastal areas globally, the coastal watersheds along the Texas
-Gulf coast are facing pressures from increasing population, increases in
-point source and non-point source pollution and alterations to
-freshwater flows that degrade water quality in downstream estuaries
+Similar to many coastal areas globally, the coastal watersheds along the
+Texas Gulf coast are facing pressures from increasing population,
+increases in point source and non-point source pollution and alterations
+to freshwater flows that degrade water quality in downstream estuaries
 \citep{bricker_effects_2008, kennicuttWaterQualityGulf2017, bugica_water_2020}.
@@
 recent studies indicates that estuary water quality dynamics in both
-agriculturally dominated and urban watersheds within Texas are in fact
+agriculturally dominated and urban watersheds within Texas are
 expressing conditions that are increasingly conducive to algal blooms
@@
 \citep{wetzWaterQualityDynamics2016, wetz_exceptionally_2017, bugica_water_2020, chinPhytoplanktonBiomassCommunity2022}.
-With identification of localized areas of estuary water quality concern
-along the Texas coast \citep{bugica_water_2020}, localized studies are
-being prioritized to better inform management actions.
+With identification of several localized areas of estuary water quality
+concern along the Texas coast \citep{bugica_water_2020}, localized
+studies are being prioritized to better inform management actions.
 
-This project aims to provide an assessment of nutrient loading and water
-quality responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in
-the larger Matagorda Bay system located roughly halfway between Houston,
-Texas and Corpus Christi, Texas. Lavaca Bay faces substantial challenges
-associated with legacy contamination but general water quality
-parameters such as dissolved oxygen (DO), nutrients, and biological
-parameters have been well within state water quality standards. More
-recently identified long-term declines in abundance, biomass, and
-diversity of benthic fauna in Lavaca Bay have been linked to reductions
-in freshwater inflows and changes in estuary salinity
-\citep{beserespollackLongtermTrendsResponse2011, palmerImpactsDroughtsLow2015, montagnaAssessmentRelationshipFreshwater2020}
-and are a concern to local stakeholders. Water quality assessments
-identified monotonic increases in total phosphorus (TP), orthophosphate,
-total Kjeldahl nitrogen (TKN), and chlorophyll-\emph{a} at sites within
-Lavaca Bay \citep{bugica_water_2020}. Although long-term changes in DO
-concentrations were not identified, the trends in nutrient
+This project aims to provide an assessment of watershed nutrient loading
+and resulting water quality responses in Lavaca Bay, Texas. Lavaca Bay
+is a secondary bay in the larger Matagorda Bay system located roughly
+halfway between Houston, Texas and Corpus Christi, Texas. Lavaca Bay
+faces substantial challenges associated with legacy contamination but
+general water quality parameters such as dissolved oxygen (DO),
+nutrients, and biological parameters have been well within state water
+quality standards. Despite largely meeting statewide water quality
+standards, there have been concerning recently identified declines in
+abundance, biomass, and diversity of benthic fauna in Lavaca Bay
+\citep{beserespollackLongtermTrendsResponse2011}. These declines are
+partially attributed to reductions in freshwater inflow and changes in
+estuary salinity and are indicative of an already stressed system
+\citep{beserespollackLongtermTrendsResponse2011, palmerImpactsDroughtsLow2015, montagnaAssessmentRelationshipFreshwater2020}.
+More recently, significant linear increases in total phosphorus (TP),
+orthophosphate, total Kjeldahl nitrogen (TKN), and chlorophyll-\emph{a}
+concentrations were identified at monitoring sites within Lavaca Bay
+\citep{bugica_water_2020}. Although long-term changes in DO
+concentrations have not been identified, the trends in nutrient
 concentrations are concerning due to the role of nitrogen as a limiting
@@
 \citep{wazniakLinkingWaterQuality2007, lloydMethodsDetectingChange2014}.
-The development and application of statistical methods such as Weighted
-Regressions on Time, Discharge and Season
+The development and application of flexible statistical methods such as
+Weighted Regressions on Time, Discharge and Season
 \citep[WRTDS,][]{hirschWeightedRegressionsTime2010} and Generalized
-Additive Models \citep[GAMs,][]{woodFastStableRestricted2011} has
+Additive Models \citep[GAMs,][]{woodFastStableRestricted2011} have
 provided effective tools for researchers to quantify and communicate
@@
 
-WRTDS calculates a time series of in stream concentrations or loads
+WRTDS calculates a time series of in-stream concentrations or loads
 (daily, monthly, or annually) and flow-normalized estimates of
@@
 the functional form and results have been demonstrated to be similar
+when assessing nutrient concentration trends
 \citep{beckNumericalQualitativeContrasts2017}. Water quality
-applications of GAMs include river and catchment nutrient concentration
-and load models
-\citep{wangLoadEstimationUncertainties2011, kroonRiverLoadsSuspended2012, kuhnertQuantifyingTotalSuspended2012, robson_prediction_2015-1, hagemannEstimatingNutrientOrganic2016, mcdowell_implications_2021, biagi_novel_2022}.
-GAMs can also be used to identify non-linear temporal trends (including
-flow-normalized trends) in pollutant concentrations and loads
-\citep{beckNumericalQualitativeContrasts2017, murphyGeneralizedAdditiveModel2019}.
-Recently GAMs have also been used to link water quality responses in
-receiving water bodies to changes in nutrient inputs
+applications of GAMs have included include river and catchment nutrient
+concentration and load estimation
+\citep{wangLoadEstimationUncertainties2011, kroonRiverLoadsSuspended2012, kuhnertQuantifyingTotalSuspended2012, robson_prediction_2015-1, hagemannEstimatingNutrientOrganic2016, mcdowell_implications_2021, biagi_novel_2022},
+assessment of temporal trends of nutrients
+\citep{beckNumericalQualitativeContrasts2017, murphyGeneralizedAdditiveModel2019},
+phytoplankton \citep{bergbuschUnexpectedShiftPhytoplankton2021}, and
+cyanobacteria \citep{hayesEffectsLakeWarming2020}. Recently GAMs have
+also been used to link water quality responses in receiving water bodies
+to changes in nonpoint source nutrient inputs
 \citep{murphyNutrientImprovementsChesapeake2022}.
@@
 thus avoiding unneeded data transformations and bias corrections. The
-study also assesses the response of water quality parameters in Lavaca
-Bay over time and in response to freshwater inflow controlled for
-seasonality and to watershed nutrient loads that are controlled for
+exploratory study also assesses the response of water quality parameters
+in Lavaca Bay over time and in response to freshwater inflow controlled
+for seasonality and to watershed nutrient loads that are controlled for
 environmentally driven variation.
@@
 Assurance Project Plans and lab method procedures outlined by the TCEQ's
-procedures manual. The QAPP and procedures manuals ensure the consistent
-collection and laboratory methods are applied between samples collected
-by different entities and under different projects. All sites had
-varying lengths of and availability of data. For freshwater locations,
-TP from January 2000 through December 2020 and nitrate-nitrogen
-(NO\textsubscript{3}) data from January 2005 through December 2020 were
-downloaded (Table~\ref{tab:table1}). Less than 5-years of total nitrogen
-and TKN concentration data were available at the freshwater sites and
-deemed insufficient to develop load estimation models
+procedures manual to ensure consistent collection and laboratory methods
+are applied between samples collected by different entities and under
+different projects. All sites had varying lengths of and availability of
+data. For freshwater locations, TP from January 2000 through December
+2020 and nitrate-nitrogen (NO\textsubscript{3}) data from January 2005
+through December 2020 were downloaded (Table~\ref{tab:table1}). Less
+than 5-years of total nitrogen and TKN concentration data were available
+at the freshwater sites and deemed insufficient to develop load
+estimation models
 \citep{horowitzEvaluationSedimentRating2003, snelderEstimationCatchmentNutrient2017}.
@@
 \centering
-\begin{tabular}[t]{llrrr}
+\begin{tabular}[t]{lllll}
 \toprule
-Station ID &   & Mean & SD & N\\
+Station ID & Parameter & Mean & SD & N\\
 \midrule
-USGS-08164000 & TP (mg/L) & \num{0.21} & \num{0.09} & 80\\
- & NO3 (mg/L) & \num{0.18} & \num{0.24} & 74\\
- & Mean Daily Streamflow (cfs) & \num{332.78} & \num{1667.47} & 7671\\
-USGS-08164525 & TP (mg/L) & \num{0.20} & \num{0.08} & 81\\
- & NO3 (mg/L) & \num{0.29} & \num{0.26} & 62\\
- & Mean Daily Streamflow (cfs) & \num{666.14} & \num{2957.79} & 7671\\
+ & TP (mg/L) & 0.21 & 0.09 & 80\\
+
+ & NO\textsubscript{3} (mg/L) & 0.18 & 0.24 & 74\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash USGS-08164000} & Mean Daily Streamflow (cfs) & 332.78 & 1667.47 & 7671\\
+\cmidrule{1-5}
+ & TP (mg/L) & 0.20 & 0.08 & 81\\
+
+ & NO\textsubscript{3} (mg/L) & 0.29 & 0.26 & 62\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash USGS-08164525} & Mean Daily Streamflow (cfs) & 666.14 & 2957.79 & 7671\\
 \bottomrule
@@
 \centering
-\begin{tabular}[t]{llrrr}
+\begin{tabular}[t]{lllll}
 \toprule
-Station ID &   & Mean & SD & N\\
+Station ID & Parameter & Mean & SD & N\\
 \midrule
-TCEQ-13383 & TP (mg/L) & \num{0.11} & \num{0.05} & 47\\
- & NOx (mg/L) & \num{0.07} & \num{0.15} & 51\\
- & TKN (mg/L) & \num{0.94} & \num{0.49} & 45\\
- & Chlorophyll-a (ug/L) & \num{9.43} & \num{5.31} & 47\\
- & DO (mg/L) & \num{7.22} & \num{1.35} & 55\\
-TCEQ-13384 & TP (mg/L) & \num{0.08} & \num{0.03} & 51\\
- & NOx (mg/L) & \num{0.06} & \num{0.08} & 52\\
- & TKN (mg/L) & \num{0.76} & \num{0.40} & 48\\
- & Chlorophyll-a (ug/L) & \num{8.22} & \num{6.44} & 46\\
- & DO (mg/L) & \num{7.51} & \num{1.32} & 54\\
-TCEQ-13563 & TP (mg/L) & \num{0.13} & \num{0.06} & 50\\
- & NOx (mg/L) & \num{0.09} & \num{0.13} & 53\\
- & TKN (mg/L) & \num{0.94} & \num{0.37} & 49\\
- & Chlorophyll-a (ug/L) & \num{9.67} & \num{5.33} & 49\\
- & DO (mg/L) & \num{7.91} & \num{1.34} & 56\\
+ & TP (mg/L) & 0.11 & 0.05 & 47\\
+
+ & NO\textsubscript{x} (mg/L) & 0.07 & 0.15 & 51\\
+
+ & TKN (mg/L) & 0.94 & 0.49 & 45\\
+
+ & Chlorophyll-\textit{a} (ug/L) & 9.43 & 5.31 & 47\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13383} & DO (mg/L) & 7.22 & 1.35 & 55\\
+\cmidrule{1-5}
+ & TP (mg/L) & 0.08 & 0.03 & 51\\
+
+ & NO\textsubscript{x} (mg/L) & 0.06 & 0.08 & 52\\
+
+ & TKN (mg/L) & 0.76 & 0.40 & 48\\
+
+ & Chlorophyll-\textit{a} (ug/L) & 8.22 & 6.44 & 46\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13384} & DO (mg/L) & 7.51 & 1.32 & 54\\
+\cmidrule{1-5}
+ & TP (mg/L) & 0.13 & 0.06 & 50\\
+
+ & NO\textsubscript{x} (mg/L) & 0.09 & 0.13 & 53\\
+
+ & TKN (mg/L) & 0.94 & 0.37 & 49\\
+
+ & Chlorophyll-\textit{a} (ug/L) & 9.67 & 5.33 & 49\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13563} & DO (mg/L) & 7.91 & 1.34 & 56\\
 \bottomrule
@@
 \emph{ltfa} and moderate improvements in TP regression models that
-incorporated \emph{stfa} \citep{zhang_improving_2017}.
+incorporated \emph{stfa} \citep{zhang_improving_2017}. Moving averages
+and flow anomalies were calculated with the \emph{adc} R package
+\citep{schrammAdcCalculateAntecedant2023}.
 
@@
 estimate was calculated as the mean of all the predictions for each day
-considering all possible flow values. Standard deviations and credible
-intervals were obtained by drawing samples from the multivariate normal
-posterior distribution of the fitted GAM
+considering all possible flow values. Standard deviations and 90\%
+credible intervals were obtained by drawing samples from the
+multivariate normal posterior distribution of the fitted GAM
 \citep{woodConfidenceIntervalsGeneralized2006, marraCoveragePropertiesConfidence2012, mcdowell_implications_2021}.
-Uncertainty in loads were calculated as 90\% credible intervals
-estimated by drawing 1000 realizations of parameter estimates from the
-multivariate normal posterior distribution of the model parameters. GAM
-performance was evaluated using repeated 5-fold cross validation
+GAM performance was evaluated using repeated 5-fold cross validation
 \citep{burmanComparativeStudyOrdinary1989} and average Nash-Sutcliffe
@@
 methodology implemented by
-\citet{murphyNutrientImprovementsChesapeake2022}, both streamflow and
-nutrient loads were prepossessed to account for season and flow.
-Freshwater inflow and nutrient loads were replaced by seasonally
-adjusted log transformed inflow and flow-adjusted log transformed
-nutrient loads obtained by fitting a GAM relating season (day of year)
-to log transformed daily freshwater inflow values and a GAM relating log
+\citet{murphyNutrientImprovementsChesapeake2022}, both freshwater inflow
+and nutrient loads were prepossessed to account for season and
+streamflow respectively. Raw freshwater inflow values were replaced by
+seasonally adjusted log transformed inflow obtained from the residuals
+of a GAM model fit between season(day of year) and log transformed daily
+freshwater inflow. Raw nutrient loads were replaced with flow-adjusted
+values obtained from the residuals of a GAM model relating log
 transformed NO\textsubscript{3} or TP loads to log transformed daily
-inflow.Response residuals from the respective GAM models were used as
+inflow. Response residuals from the respective GAM models were used as
 \emph{Q} and \emph{Load} in Eq.~\ref{eq6} and Eq.~\ref{eq7}.
 
-This study used an information theoretic approach to evaluate if
-nutrient loads and/or freshwater inflows provided evidence of effects on
-water quality concentrations in Lavaca Bay. Model probabilities were
-calculated and compared using the AIC\textsubscript{c} scores between
-each group of temporal, flow, and flow+load models
-\citep{burnhamAICModelSelection2011}. Improvements in model
-probabilities provide evidence that the terms explain additional
-variation in the response variable. If model probabilities were tied,
-there wasn't evidence the more complicated model explains additional
-variation in water quality.
+This study used an information theoretic approach to evaluate evidence
+of model covariate effects on Lavaca Bay water quality. Model
+probabilities were calculated and compared using the
+AIC\textsubscript{c} scores between each group of temporal, inflow, and
+inflow+load models \citep{burnhamAICModelSelection2011}. Improvements in
+model probabilities provide evidence that the terms explain additional
+variation in the response variable.
 
@@
 
-Using evaluation criteria recommended by
-\citet{moriasiHydrologicWaterQuality2015}, predictive performance of
-nutrient loading GAMs ranged from ``satisfactory'' to ``very good''
-based on median NSE, \emph{r}, and PBIAS metrics calculated using 5-fold
-cross validation. Median goodness-of-fit metrics for NO\textsubscript{3}
-models in the Lavaca River were 0.34 NSE, 0.70 \emph{r}, and 2.00 PBIAS.
-Navidad River NO\textsubscript{3} models appeared to perform slightly
-better with 0.48 NSE and 0.87 \emph{r} but with higher bias at 10.90
-PBIAS. Generally, TP models performed better than NO\textsubscript{3}
-models. Median goodness-of-fit metrics for TP in the Lavaca River were
-0.81 NSE, 0.93 \emph{r}, and -7.20 PBIAS. Navidad River TP models has
-similar performance with 0.91 NSE, 0.99 \emph{r}, and -3.30 PBIAS.
-Density plots of metrics show similar distribution of values between
-sites for the same parameter, with the exception \emph{r} values for
-NO\textsubscript{3} loads where Lavaca River had a much larger variance
-in values compared to the Navidad River (Figure~\ref{fig:fig2}). TP GAMS
-had higher average NSE and \emph{r} values and lower variance in metric
-values compared to NO\textsubscript{3}.
+Predictive performance of nutrient loads ranged from ``satisfactory'' to
+``very good'' based on standardized evaluation metrics of NSE, \emph{r},
+and PBIAS \citep{moriasiHydrologicWaterQuality2015} calculated using
+5-fold cross validation. Median goodness-of-fit metrics for
+NO\textsubscript{3} models in the Lavaca River were 0.34 NSE, 0.70
+\emph{r}, and 2.00 PBIAS. Navidad River NO\textsubscript{3} models
+appeared to perform slightly better with 0.48 NSE and 0.87 \emph{r} but
+with higher bias at 10.90 PBIAS. Generally, TP models performed better
+than NO\textsubscript{3} models. Median goodness-of-fit metrics for TP
+in the Lavaca River were 0.81 NSE, 0.93 \emph{r}, and -7.20 PBIAS.
+Navidad River TP models has similar performance with 0.91 NSE, 0.99
+\emph{r}, and -3.30 PBIAS. Density plots of metrics show similar
+distribution of values between sites for the same parameter, with the
+exception \emph{r} values for NO\textsubscript{3} loads where Lavaca
+River had a much larger variance in values compared to the Navidad River
+(Figure~\ref{fig:fig2}). TP GAMS had higher average NSE and \emph{r}
+values and lower variance in metric values compared to
+NO\textsubscript{3}.
 
@@
 
-Predicted annual NO\textsubscript{3} and TP loads show considerable
-variation, generally following patterns in discharge
-(Figures~\ref{fig:fig3}, \ref{fig:fig4}). Flow-normalized TP loads at
-both sites and flow-normalized NO\textsubscript{3} loads in the Lavaca
-River indicated watershed based loads did not change much over time when
-accounting for variation driven by streamflow (Figure~\ref{fig:fig3}).
-Flow-normalized loads in the Lavaca River showed small variation over
-time with some decreases in NO\textsubscript{3} loads since 2013.
+Annual NO\textsubscript{3} and TP loads show considerable variation,
+generally following patterns in discharge (Figures~\ref{fig:fig3},
+\ref{fig:fig4}). Flow-normalized TP loads at both sites and
+flow-normalized NO\textsubscript{3} loads in the Lavaca River indicated
+watershed based loads did not change much over time when accounting for
+variation driven by streamflow (Figure~\ref{fig:fig3}). Flow-normalized
+loads in the Lavaca River showed small variation over time with some
+decreases in NO\textsubscript{3} loads since 2013.
 
@@
 
-GAMs did not identify significant changes in TP or DO concentrations at
-any of the Lavaca Bay sites from 2005 through 2020
-(Figure~\ref{fig:fig5}). The upper-bay site, TCEQ-13563, had a linear
-increase in NO\textsubscript{x} concentration and and decrease in
-chlorophyll-\emph{a} from 2005 through 2014. The mid-bay site,
-TCEQ-13383, showed a periodic pattern in NO\textsubscript{x}
-concentration that appeared similar to precipitation/inflow patterns, as
-well as a post 2011 increase in TKN concentrations. No significant
-long-term trends in concentrations were identified by GAMs for the
-lower-bay TCEQ-13384 site.
+There is no evidence of long-term changes in TP or DO concentrations at
+any Lavaca Bay site (Figure~\ref{fig:fig5}). The upper-bay site,
+TCEQ-13563, shows evidence of a long-term linear increase in
+NO\textsubscript{x} while chlorophyll-\emph{a} decreased from 2005
+through 2014 (Figure~\ref{fig:fig5}). NO\textsubscript{x} concentration
+at the mid-bay site, TCEQ-13383, displayed an unusual periodic pattern
+that is indicative of a strong influence from inflow or precipitation.
+The temporal GAMs did not provide evidence of long-term trends in any of
+the water quality constituents at the lower-bay TCEQ-13384 site.
 
@@
 
-\caption{Smoothed temporal trend component for water quality paramaters obtained from temporal estuary GAMs.}\label{fig:fig5}
+\caption{Fitted splines (shaded regions indicate 90\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\textit{\textsubscript{x}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig5}
 \end{figure}
@@
 \centering
-\begin{tabular}[t]{lllll}
+\begin{tabular}[t]{ll>{}l>{}l>{}l}
 \toprule
-Parameter & Site & Temporal & Flow & Flow + Load\\
+Parameter & Site & Temporal & Inflow & Inflow + Load\\
 \midrule
- & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & -158.2 (0.72)\\
+ & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & \em{\textbf{-158.2 (0.72)}}\\
 
- & TCEQ-13384 & -194.4 (0.03) & -200.2 (0.49) & -200.2 (0.49)\\
+ & TCEQ-13384 & -194.4 (0.03) & \em{\textbf{-200.2 (0.49)}} & -200.2 (0.49)\\
 
-\multirow{-3}{*}{\raggedright\arraybackslash TP} & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & -157.3 (0.59)\\
+\multirow{-3}{*}{\raggedright\arraybackslash TP} & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & \em{\textbf{-157.3 (0.59)}}\\
 \cmidrule{1-5}
- & TCEQ-13383 & -218.9 (0) & -244.8 (0.5) & -244.8 (0.5)\\
+ & TCEQ-13383 & -218.9 (0) & \em{\textbf{-244.8 (0.5)}} & -244.8 (0.5)\\
 
- & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & -311.9 (0.52)\\
+ & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & \em{\textbf{-311.9 (0.52)}}\\
 
-\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & TCEQ-13563 & -175.1 (0) & -190.2 (0.5) & -190.2 (0.5)\\
+\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & TCEQ-13563 & -175.1 (0) & \em{\textbf{-190.2 (0.5)}} & -190.2 (0.5)\\
 \cmidrule{1-5}
- & TCEQ-13383 & 279.7 (0.18) & 278.1 (0.41) & 278.1 (0.41)\\
+ & TCEQ-13383 & 279.7 (0.18) & \em{\textbf{278.1 (0.41)}} & 278.1 (0.41)\\
 
- & TCEQ-13384 & 268.2 (0.33) & 268.2 (0.33) & 268.2 (0.33)\\
+ & TCEQ-13384 & \em{\textbf{268.2 (0.33)}} & 268.2 (0.33) & 268.2 (0.33)\\
 
-\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & TCEQ-13563 & 289.5 (0.08) & 286.1 (0.46) & 286.1 (0.46)\\
+\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & TCEQ-13563 & 289.5 (0.08) & \em{\textbf{286.1 (0.46)}} & 286.1 (0.46)\\
 \cmidrule{1-5}
- & TCEQ-13383 & 42.2 (0.66) & 43.5 (0.34) & -\\
+ & TCEQ-13383 & \em{\textbf{42.2 (0.66)}} & 43.5 (0.34) & -\\
 
- & TCEQ-13384 & 34.3 (0.57) & 34.8 (0.43) & -\\
+ & TCEQ-13384 & \em{\textbf{34.3 (0.57)}} & 34.8 (0.43) & -\\
 
-\multirow{-3}{*}{\raggedright\arraybackslash TKN} & TCEQ-13563 & 31.1 (0.22) & 28.7 (0.78) & -\\
+\multirow{-3}{*}{\raggedright\arraybackslash TKN} & TCEQ-13563 & 31.1 (0.22) & \em{\textbf{28.7 (0.78)}} & -\\
 \cmidrule{1-5}
- & TCEQ-13383 & 146.4 (0.34) & 146.4 (0.34) & 146.5 (0.32)\\
+ & TCEQ-13383 & \em{\textbf{146.4 (0.34)}} & 146.4 (0.34) & 146.5 (0.32)\\
 
- & TCEQ-13384 & 135.9 (0.47) & 137 (0.27) & 137 (0.27)\\
+ & TCEQ-13384 & \em{\textbf{135.9 (0.47)}} & 137 (0.27) & 137 (0.27)\\
 
-\multirow{-3}{*}{\raggedright\arraybackslash DO} & TCEQ-13563 & 138.3 (0.25) & 137.2 (0.43) & 137.8 (0.32)\\
+\multirow{-3}{*}{\raggedright\arraybackslash DO} & TCEQ-13563 & 138.3 (0.25) & \em{\textbf{137.2 (0.43)}} & 137.8 (0.32)\\
 \bottomrule
@@
 
-\caption{Estimated effects of mean daily inflow residuals on mean TP, NO\textsubscript{\textit{x}}, chlorophyll-\textit{a}, TKN, and DO concentrations in Lavaca Bay obtained from flow estuary GAMs.}\label{fig:fig6}
+\caption{Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\textit{\textsubscript{x}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig6}
 \end{figure}
@@
 
-\caption{Estimated effects of nutrient load residuals on TP and NO\textsubscript{\textit{x}} concentrations in Lavaca Bay obtained from flow+load estuary GAMs.}\label{fig:fig7}
+\caption{Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\textsubscript{3} loads on TP (A) and NO\textsubscript{\textit{x}} (B) concentrations at each site in Lavaca Bay.}\label{fig:fig7}
 \end{figure}
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\hypertarget{sec4}{%
+\section{Discussion}\label{sec4}}
 
@@
 
+\hypertarget{watershed-nutrient-loads-1}{%
+\subsection{Watershed Nutrient Loads}\label{watershed-nutrient-loads-1}}
+
 TP and NO\textsubscript{3} loadings from the Lavaca Bay watershed showed
-high inter-annual variability tied with changes in discharge. There is
-little evidence for changes in flow-normalized TP loads in either
-rivers. There is some evidence of recent decreases in flow-normalized
-NO\textsubscript{3} loads in the Lavaca River. Although there is no work
-directly correlating water quality planning and implementation efforts
-in the watershed to water quality outcomes, efforts to increase
-agricultural producer participation in the watershed have been ongoing
-since 2016
+high inter-annual variability tied closely with changes in discharge.
+There is little evidence for changes in flow-normalized TP loads in
+either river. There is some evidence of recent decreases in
+flow-normalized NO\textsubscript{3} loads in the Lavaca River. Although
+there is no work directly correlating water quality planning and
+implementation efforts in the watershed to water quality outcomes,
+efforts to increase agricultural producer participation in the watershed
+have been ongoing since 2016
 \citep{schrammLavacaRiverWatershed2018, bertholdDirectMailingEducation2021}.
@@
 \citep{horowitzEvaluationSedimentRating2003, snelderEstimationCatchmentNutrient2017}.
+
+\hypertarget{estuary-water-quality-responses}{%
+\subsection{Estuary Water Quality
+Responses}\label{estuary-water-quality-responses}}
 

### 2023-05-31 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.Rmd
trends plot plus lots of edits

--- 2023-05-23 Schramm-2023-05-AS/Schramm-2023-05-AS.Rmd
@@
 ---
-# Supported options: 
-#   sn-nature:       Style for submissions to Nature Portfolio journals
-#   sn-basic:        Basic Springer Nature Reference Style/Chemistry Reference Style
-#   sn-mathphys:     Math and Physical Sciences Reference Style
-#   sn-aps:          American Physical Society (APS) Reference Style
-#   sn-vancouver:    Vancouver Reference Style
-#   sn-apa:          APA Reference Style 
-#   sn-chicago:      Chicago-based Humanities Reference Style
-#   default:         Default
-
-classoptions: 
-  - sn-basic      
-  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
-  - referee       # Optional: Use double line spacing 
-  - lineno        # Optional: Add line numbers
-  # - iicol         # Optional: Double column layour
-
-title: Assessing linkages between watershed nutrient loading and water quality in a subtropical estuary with semiparametric models.
-titlerunning: Watershed nutrient loading and water quality in a subtropical estuary
-
-authors: 
-  - firstname: Michael
-    lastname: Schramm
+title: Linking watershed nutrient loading to estuary water quality with semiparametric models
+preprint: false
+author: 
+  - name: Michael Schramm
+    affiliation: 1
+    corresponding: true
     email: michael.schramm@ag.tamu.edu
-    affiliation: [1]
-    corresponding: TRUE
-
-affiliations:
-  - number: 1
-    corresponding: TRUE
-    info:
-      orgdiv: Texas Water Resources Institute
-      orgname: Texas A&M AgriLife Research
-    address:
-        street: 1001 Holleman Dr. E.
-        city: College Station
-        postcode: 77840
-        state: TX
-        country: United States
-
-keywords:
-  - key
-  - dictionary
-  - word
-
-# Sample for unstructured abstract 
-abstract: |
-  Lavaca Bay is a secondary embayment on the Texas coast displaying early signals of water quality degradation.
-This study applied a semiparametric approach to assess both watershed nutrient loads and responses in estuary water quality to nutrient loading and streamflow.
+affiliation:
+  - code: 1
+    address: Texas Water Resources Institute, Texas A\&M AgriLife Research, College Station, Texas, United States
+abstract: >
+  Evaluating estuary water quality responses to reductions (or increases) in nutrient loading attributed to on the ground management actions can be challenging due to the strong influence of environmental drivers on nutrient loads and non-linear relationships.
+This study applied semiparametric approaches to calculate watershed nutrient loads and assess responses in estuary water quality to seasonally-adjusted freshwater inflow and flow-adjusted nutrient loads in Lavaca Bay, Texas.
+Lavaca Bay is a secondary embayment on the Texas coast displaying early potential for eutrophication and water quality degredation.
+Use of flow-adjusted nutrient loads allowed the study to evaluate the response in water quality to changes in nutrient loads driven by non-natural variation.
 Cross-validation indicated that, despite data constraints, semiparametric models performed well at nutrient load prediction.
@@
 In contrast, minimal changes in flow-normalized loads indicate that nutrient loadings were driven by natural variation in precipitation and runoff as opposed to changes in management of nonpoint sources.
-Models indicated there was no evidence of long-term changes in dissolved oxygen or chlorophyll-*a* within Lavaca Bay.
-However, site specific long-term increases in both organic and inorganic nitrogen are concerning.
-Further analysis found freshwater inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that changes in watershed nutrient loading explained variation in dissolved oxygen or chlorophyll-*a* concentrations.
+Models indicated no evidence of long-term changes in dissolved oxygen or chlorophyll-*a* within Lavaca Bay.
+However, site specific long-term increases in both organic and inorganic nitrogen are causes of concern as potential causes of future eutrophication.
+Further analysis found freshwater inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that changes in watershed nutrient loading explained additional variation in dissolved oxygen or chlorophyll-*a* concentrations.
 In addition to providing a baseline assessment of watershed nutrient loading and water quality responses in the Lavaca Bay watershed, this study provides methodological support for the use of semiparametric methods in load regression models and estuary assessments.
-
 bibliography: bibliography.bib
-output: rticles::springer_article
+output:
+  rticles::peerj_article:
+    citation_package: biblatex
+biblio-style: apa
+header-includes:
+    - \usepackage{csquotes}
+
 ---
@@
 
-*Remaining Items*
-
-- keywords
-- discussion
-- formalize language (reduce first person active?)
-- final language review
-- springer doesn't want figures in subdirectory, need to knit to same directory as tex file.
-
-# Introduction {#sec1}
+# Introduction {.unnumbered}
 
@@
 
-This project aims to provide an assessment of watershed nutrient loading and resulting water quality
-responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
+This goal of this project is to assess watershed nutrient loading and the 
+resulting water quality responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
 Matagorda Bay system located roughly halfway between Houston, Texas and Corpus
@@
 water quality standards. Despite largely meeting statewide water quality standards,
-there have been concerning recently identified declines in abundance, biomass, and diversity of 
+there have been concerning declines in abundance, biomass, and diversity of 
 benthic fauna in Lavaca Bay [@beserespollackLongtermTrendsResponse2011]. These declines
@@
 [@woodFastStableRestricted2011]. Although the underlying parameter estimation 
-procedure of GAMs is substantially different than WRTDS, both the functional 
+procedure of GAMs are substantially different than WRTDS, both the functional 
 form and results have been demonstrated to be similar when assessing nutrient
 concentration trends [@beckNumericalQualitativeContrasts2017]. 
-Water quality applications of GAMs have included include river and catchment 
+Water quality applications of GAMs have included river and catchment 
 nutrient concentration and load estimation
@@
 @hagemannEstimatingNutrientOrganic2016;@mcdowell_implications_2021;
-@biagi_novel_2022], assessment of temporal trends of nutrients
+@biagi_novel_2022], and assessment of temporal trends of nutrients
 [@beckNumericalQualitativeContrasts2017; @murphyGeneralizedAdditiveModel2019],
@@
 receiving water bodies to changes in nonpoint source nutrient inputs 
-[@murphyNutrientImprovementsChesapeake2022]. @beckNumericalQualitativeContrasts2017
-provides a substantial discussion on the differences (and similarities) between 
-GAMs and WRTDS for water quality applications.
+[@murphyNutrientImprovementsChesapeake2022]. For a substantial discussion on the 
+differences (and similarities) between GAMs and WRTDS for water quality 
+applications reders are referred to @beckNumericalQualitativeContrasts2017.
 
@@
 This study utilizes GAMs to develop estimates of delivered 
-and flow-normalized nutrient loads and assess changes in loads delivered to 
+and flow-normalized nutrient loads and assess estuary water quality responses
+to changes in loads delivered to 
 Lavaca Bay. GAMs were chosen over other regression-based
 approached for use in this study due to; (1) the ability to easily explore and 
-incorporate different model terms; (2) the incorporation of non-liner smooth
+incorporate different model terms; (2) the incorporation of non-linear smooth
 functions that do not require explicit a priori knowledge of the expected
@@
 the response to linear predictors thus avoiding unneeded data transformations
-and bias corrections. The exploratory study also assesses the response of water quality parameters in
-Lavaca Bay over time and in response to freshwater inflow controlled
-for seasonality and to watershed nutrient loads that are controlled for
-environmentally driven variation.
-
-# Methods {#sec2}
-
-## Location and Data
-Lavaca Bay is 190 km^2^ with the majority of freshwater inflow provided by the 
-Lavaca and Navidad River systems (Figure\ \ref{fig:fig1}). The Garcitas-Arenosa, 
-Placedo Creek, and Cox Bay watersheds provide additional freshwater inflows. 
-The entire watershed land area is 8,149 km^2^ and primarily rural. Watershed 
+and bias corrections. The exploratory study also assesses the response of water 
+quality parameters in Lavaca Bay over time and in response to freshwater inflow 
+controlled for seasonality and to watershed nutrient loads that are controlled 
+for environmentally driven variation.
+
+# Methods and Materials {.unnumbered}
+
+## Location and Data {.unnumbered}
+Lavaca Bay is a 190 km^2^ estuary with the majority of freshwater inflow provided by the 
+Lavaca and Navidad River systems (Fig.\ \ref{fig:fig1}). The Garcitas-Arenosa, 
+Placedo Creek, and Cox Bay subwatersheds provide additional freshwater inflows. 
+The entire watershed area is 8,149 km^2^ and primarily rural. Watershed 
 land cover and land use is 50% grazed pasture and rangeland, 20% cultivated 
@@
 
-Daily discharges for the Lavaca River (USGS-08164000, Figure\ \ref{fig:fig1}) 
+Daily discharges for the Lavaca River (USGS-08164000, Fig.\ \ref{fig:fig1}) 
 were obtained from the United States Geologic Survey (USGS) National Water 
@@
 
-Water quality sample data for the two freshwater and three estuary locations were 
+Water quality data for the two freshwater and three estuary locations were 
 obtained from the Texas Commission on Environmental Quality (TCEQ) Surface Water 
@@
 lower Lavaca Bay site near the mouth of the Bay (TCEQ-13384). 
-For estuary locations, we obtained data for TP, Nitrite+Nitrate (NO*~x~*), TKN, 
+For estuary locations, we obtained data for TP, Nitrite+Nitrate (NO~*X*~), TKN, 
 chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
@@
 
-## Estimating Watershed Based Nutrient Loads
+## Estimating Watershed Based Nutrient Loads {.unnumbered}
 
@@
 
-\begin{align}
-ma(\delta) &= d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}, \nonumber \\
-\kappa_{i} &=\sum_{m=1}^{i}\hat{Q}_m
+\begin{equation}
+ma(t) = \frac{\sum^{J}_{j=1}{d^{t-t_j}Q_{t_{j}}}}{\sum^{J}_{j=1}d^{t-t_j}}
 \label{eq2}
-\end{align}
-
-where $\delta$ is the discount factor (here, set equal to 0.95), $\kappa_i$
-is the cumulative flow (*Q*) up to the *i*th day.
+\end{equation}
+
+where *t* is the day of estimation, *d* is the discount factor (set to 0.95),
+*J* is the total number of daily discharge observations, $t-t_j$ is the lag time
+between the current day and some historical day of observation ($t_j$). Here,
+as *d* approaches zero, the time series becomes the daily observed values and 
+as *d* gets closer to one, the time series becomes the mean flow.
 
@@
 @zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
-the previous year relative to the entire period [@zhang_improving_2017] and 
-calculated using Eq. \ref{eq3}:
-
-\begin{equation}
-ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}
-\label{eq3}
-\end{equation}
-
+the previous year relative to the entire period [Eq.\ \ref{eq3}, @zhang_improving_2017] and 
 and the short-term flow anomaly (*stfa*) calculated as the current day flow 
-compared to the preceding 1-month streamflow using \ref{eq4}:
-
-\begin{equation}
-stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t)
-\label{eq4}
-\end{equation}
+compared to the preceding 1-month streamflow [Eq.\ \ref{eq4}, @zhang_improving_2017]:
+
+\begin{align}
+ltfa(t) &= \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}, \label{eq3}\\
+stfa(t) &= x_{current\,day}(t) - \bar{x}_{1\,month}(t) \label{eq4}
+\end{align}
 
@@
 and *stfa* in TP models based on previous work demonstrating major improvements 
-in NO~x~ regression models that incorporated 
+in NO~*X*~ regression models that incorporated 
 *ltfa* and moderate improvements in TP regression models that incorporated 
@@
 
-## Linking Estuary Water Quality to Hydrology and Nutrient Loads
+## Linking Estuary Water Quality to Hydrology and Nutrient Loads {.unnumbered}
 
@@
 Because streamflow and nutrient loads are tightly correlated, freshwater inflow
-can mask signals from nutrient loads alone. Following the 
+can mask signals due to chnages in nutrient loads alone. Following the 
 methodology implemented by @murphyNutrientImprovementsChesapeake2022,
@@
 
-# Results {#sec3}
-
-## Watershed Nutrient Loads
+# Results {.unnumbered}
+
+## Watershed Nutrient Loads {.unnumbered}
 
@@
 Lavaca River had a much larger variance in values compared to the Navidad River
-(Figure\ \ref{fig:fig2}). TP GAMS had higher average NSE and *r* values and 
+(Fig.\ \ref{fig:fig2}). TP GAMS had higher average NSE and *r* values and 
 lower variance in metric values compared to NO~3~.
@@
 Annual NO~3~ and TP loads show considerable variation, generally following
-patterns in discharge (Figures\ \ref{fig:fig3}, \ref{fig:fig4}). 
+patterns in discharge (Fig.\ \ref{fig:fig3}, Fig. \ref{fig:fig4}). 
 Flow-normalized TP loads at both sites and flow-normalized NO~3~ loads in the Lavaca River
 indicated watershed based loads did not change much over time
-when accounting for variation driven by streamflow (Figure\ \ref{fig:fig3}). 
+when accounting for variation driven by streamflow (Fig.\ \ref{fig:fig3}). 
 Flow-normalized loads in the Lavaca River showed small variation over time with 
@@
 `r min_no3_prop` and `r min_tp_prop` of NO~3~ and TP loads in 2011 
-(Figure\ \ref{fig:fig4}).
+(Fig.\ \ref{fig:fig4}).
 
@@
 
-## Linkages Between Water Quality and Watershed Flows and Loads
-
-There is no evidence of long-term changes in TP or DO concentrations at any
-Lavaca Bay site (Figure\ \ref{fig:fig5}). The upper-bay site, TCEQ-13563, shows
-evidence of a long-term linear increase in NO~x~ while chlorophyll-*a*
-decreased from 2005 through 2014 (Figure\ \ref{fig:fig5}). NO~x~ concentration
+## Linkages Between Water Quality and Watershed Flows and Loads {.unnumbered}
+
+There was no evidence of long-term changes in TP or DO concentrations at any
+Lavaca Bay site (Fig.\ \ref{fig:fig5}). The upper-bay site, TCEQ-13563, had
+evidence of a long-term linear increase in NO~*X*~ while chlorophyll-*a*
+decreased from 2005 through 2014 (Fig.\ \ref{fig:fig5}). NO~*X*~ concentration
 at the mid-bay site, TCEQ-13383, displayed an unusual periodic pattern that is
@@
 
-```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{x}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
+```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
 
@@
 Freshwater inflow provided additional explanation for changes in TP and 
-NO~*x*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
+NO~*X*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
 model probability values (Table\ \ref{tab:table4}).
@@
 
-GAMs showed that increases in freshwater inflow resulted in nearly linear increases 
-in TP and NO~*x*~ concentration at all three sites (Figure\ \ref{fig:fig6}). 
-At the upper-bay TCEQ-13563 site, GAMs showed that increases in freshwater inflow initially 
+Increases in freshwater inflow resulted in nearly linear increases 
+in TP and NO~*X*~ concentration at all three sites (Fig.\ \ref{fig:fig6}). 
+At the upper-bay TCEQ-13563 site, increases in freshwater inflow initially 
 increased chlorophyll-*a* and DO concentration but concentrations leveled and
-potentially decreased at higher flows. The mid-bay TCEQ-13383 site showed a
-nearly linear increased in chlorophyll-*a* concentration in response to increases
-freshwater inflow. Freshwater flow did not have significant
-effects on chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.
-
-```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\\textit{\\textsubscript{x}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
+potentially decreased at higher flows. The mid-bay TCEQ-13383 site had
+nearly linear increases in chlorophyll-*a* concentration in response to increased
+freshwater inflow. There was no evidence of freshwater inflow effects on 
+chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.
+
+```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
 
@@
 the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively 
-(Figure\ \ref{fig:fig7}).
-The relative effect size appeared to much smaller than the effect of freshwater
+(Fig.\ \ref{fig:fig7}).
+The relative effect size appeared much smaller than the effect of freshwater
 inflow alone. Increased NO~3~ loads only showed an effect at the 
@@
 
-```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.4, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\\textsubscript{3} loads on TP (A) and NO\\textsubscript{\\textit{x}} (B) concentrations at each site in Lavaca Bay."}
+```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.4, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\\textsubscript{3} loads on TP (A) and NO\\textsubscript{\\textit{X}} (B) concentrations at each site in Lavaca Bay."}
 
@@
 
-# Discussion {#sec4}
-
-**resume here**
-
-## Watershed Nutrient Loads
+# Discussion {.unnumbered}
+
+## Nutrient Loads {.unnumbered}
 
@@
 no indications of trends in flow-normalized NO~3~ and TP loads in the Navidad River.
-Freshwater discharges in the Navidad River are regulated by the Palmetto Dam forming
-Lake Texana at the lower extent of the river. While the dominant agricultural land uses
-differ between the Lavaca (primarily grazed pasture and rangeland) and Navidad 
-(mix of pasture and row crops) catchments, we did not have a reason to expect 
-different flow normalized trends between the two systems from land use alone.
-Lentic nitrogen uptake and cycling may have regulating effects that mask changes
+In comparison, there was weak evidence of more recent decreases in flow-normalized NO~3~ 
+(but not TP) loads in the Lavaca River watershed. While the dominant agricultural
+land uses differ between the Lavaca (primarily 
+grazed pasture and rangeland) and Navidad (mix of pasture and row crops) 
+catchments, we did not have a reason to expect different flow normalized trends 
+between the two systems from land use alone. Freshwater discharges in the 
+Navidad River are regulated by the Palmetto Bend Dam forming
+Lake Texana at the lower extent of the river. Lentic nitrogen uptake and 
+cycling may have regulating effects that mask changes
 in upstream nitrogen loadings. Additional nutrient data collection in the contributing 
@@
 in regulating nutrient delivery to the Lavaca Bay system. However, these results
-suggest that there have been no changes in the NO~3~ or TP loading from Lake Texana
-when accounting for variations in year to year discharge.
-
-There is some evidence of recent decreases in flow-normalized NO~3~ loads in 
-the Lavaca River. Although there is no work directly correlating water quality 
-planning and implementation efforts in the watershed to water quality outcomes, 
-efforts to increase agricultural producer participation in the watershed have 
-been ongoing since 2016 [@schrammLavacaRiverWatershed2018; 
-@bertholdDirectMailingEducation2021]. The decrease in flow-normalized NO~3~ 
+suggest that there have been no changes in the NO~3~ or TP loading from the
+Navidad River system at the Lake Texana discharge point when accounting for 
+variations in year to year discharge.
+
+The evidence of decreased Lavaca River NO~3~ loading, although weak, is a 
+potential positive sign for water quality managers working to implement 
+practices that improve water quality in the freshwater sections of the Lavaca 
+River watershed. Planning and implementation efforts to increase agricultural 
+producer participation in water quality protection practices have 
+been ongoing in the watershed since 2016 [@schrammLavacaRiverWatershed2018; 
+@bertholdDirectMailingEducation2021], however little work has been conducted
+to directly link these efforts with water quality outcomes. The decrease in 
+flow-normalized NO~3~ 
 loads could be a reflection of those collective efforts but the lack of evidence
@@
 the water quality dataset that is primarily composed of ambient water quality
-measurements.
-
-Cross-validation of the nutrient loading models highlights that predictions are
-prone to high bias, owing to the lack of targeted storm or flow biased measurements.
-The high biases are indicative that subsets of values were unable to capture the
-functional relationships with the flow based dependent variables. It was beyond the scope
-of the current study to evaluate the subsets of cross-validation data and scores.
-However, the cross-validation procedure is indicative that more robust sampling
-would be beneficial for reducing prediction variance. Supplementary flow-biased 
-monitoring targeting storm- or high-flow conditions is critical to
-improve model performance and strength of evidence produced by these models [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017].
-There is no absolute guidance for the number of samples needed...
-Would be useful to develop guidance for apply GAMs to estimate loads and estimate trends of different effect sizes.
-
-Converted to average annual yield, the estimates of annual TP loads for the
-Lavaca River are within the ranges in previous published studies
-[Table\ \ref{tab:table5}, @dunnTrendsNutrientInflows1996; @rebichSourcesDeliveryNutrients2011; @omaniEstimationSedimentNutrient2014; @wise_spatially_2019]. It isn't obvious
-why TP estimates in @dunnTrendsNutrientInflows1996 were notably lower. Given that none
-of the studies identify substantially sized trends in TP, it is is possible that 
-the period used in @dunnTrendsNutrientInflows1996 was drier on average than the
-other studies. The SPARROW models used in @rebichSourcesDeliveryNutrients2011 and 
-@wise_spatially_2019 utilize a version of LOADEST in the underlying load
-estimation procedure, so a difference due to methodology alone is unlikely.
+measurements. The issues associated with the lack of flow-biased measurements 
+is further discussed later in this section.
+
+Some prior studies have generated estimates of mean annual TP yields in the 
+Lavaca River watershed [Table\ \ref{tab:table5}, @dunnTrendsNutrientInflows1996; 
+@rebichSourcesDeliveryNutrients2011; @omaniEstimationSedimentNutrient2014;
+@wise_spatially_2019]. Although these studies differ in time periods and 
+methodologies, they provide a sanity check for the reasonableness of the 
+annual estimates generated in the current study. In a regional assessment of
+nutrient loading in river's along the Gulf of Mexico, @dunnTrendsNutrientInflows1996
+used the LOADEST model to develop an estimated mean annual yield of 28.9 kg/km^2^.
+LOADEST is a multiple linear regression model that fits log transformed pollutant concentrations
+to long term, seasonal, and flow based predictors and includes methods for bias correction
+when exponentiating the response variable. While the annual yield estimated in 
+@dunnTrendsNutrientInflows1996 is lower than the estimate of 35.2 kg/km^2^
+in this study, it is important to note that both hydrology and land use was
+different for a proportion of the study period. For almost the first ten years of the study period
+in @dunnTrendsNutrientInflows1996, the Navidad River was still undammed. More recently
+the eastern portion of the watershed has seen large conversions of rice fields to
+more traditional row crop production [@kulatHolisticWaterEnergyFoodNexus2019]. 
+The apparent increase in mean TP yield might be suggestive of wetter recent conditions.
+To evaluate long-term trends in discharge, we fit a GAM relating log-transformed daily
+discharges on the Lavaca and Navidad Rivers to season and time (Fig.\ \ref{fig:fig8}).
+The long-term trends in discharge indicate watershed discharges were at or above average
+from 1972 though the early and mid-1980s. In comparison watershed discharges
+since the mid-2000s are at or below average. The trend in mean streamflow alone
+is counter to expectations based on changes in nutrient yeild estimates. Studies of
+long-term precipitation patterns indicate both the total amount [@dixonTrendDetectionTexas2011] 
+and variability of rainfall [@mishraChangesExtremePrecipitation2010;
+@fagnantCharacterizingSpatiotemporalTrends2020] has increased in the region 
+which provide plausible reasons for increased runoff driven nutrient loading.
+Further assessments of the frequency and occurrence of periodic high flow events 
+that mobilize sediments and are responsible for the bulk of loading volume would
+provide further insight to this discrepancy.
+
+```{r fig8, fig.width=6.85, fig.asp=0.75, fig.cap="Measured daily discharges (log-transformed) and smoothed long-term trends for the Lavaca and Navidad Rivers form 1972 though 2001."}
+
+knitr::include_graphics("../fig8.png")
+
+```
+
+@rebichSourcesDeliveryNutrients2011 and @wise_spatially_2019 used SPARROW to provide
+a more recent assessment of regional catchment based loadings to the Gulf of Mexico
+(Table\ \ref{tab:table5}).
+SPARROW is a hybrid statistical-process model with the underlying nutrient
+load estimation methods based on the previously described LOADEST 
+[@schwarzSPARROWSurfaceWaterQuality2006]. The functional form of the LOADEST
+regression model is similar to the terms applied in the GAMs used in the current
+study. The only study to apply a mechanistic watershed model (SWAT) to estimate
+nutrient loadings in the Lavaca River watershed 
+@omaniEstimationSedimentNutrient2014 developed estimated yields (42 kg/km^2^) 
+similar to the two SPARROW models. The base periods used by 
+@rebichSourcesDeliveryNutrients2011, @omaniEstimationSedimentNutrient2014 and
+@wise_spatially_2019 differ from the current study, making direct comparisons 
+difficult. It is probable that the lower than average discharges observed from 2010 
+through 2021 (Fig.\ \ref{fig:fig8}) bias our estimates downward compared to other studies which included
+higher than average streamflow periods. Overall, the ranges of estimated yields
+among different studies along with the apparent large variability in 
+streamflow driven loadings (Fig.\ \ref{fig:fig3}, Fig. \ref{fig:fig4}) suggest that the current estimates
+of TP loading are reasonable.
 
@@
     caption = "Comparisons of previously published estimates of mean annual TP yield at the Lavaca River site.") |>
-  footnote(alphabet = c("Mean of the annual point estimates and the lower and upper 95% credible intervals.",
+  footnote(alphabet = c("Mean of the annual point estimates and the lower and upper 90% credible intervals.",
                         "Represents a binned value range from a choropleth map."))
@@
 
-## Estuary Water Quality Responses
-
-The non-linear temporal water quality trends identified using GAMs differed
-slightly from previously identified trends [@bugica_water_2020]. This is
-not unexpected due to the different time periods, different methodology, and
+## Estuary Water Quality {.unnumbered}
+
+The non-linear estuary water quality trends identified in the current
+study differed slightly from previously identified trends [@bugica_water_2020]. 
+This is not unexpected due to the different time periods, different methodology, and
 generally small slopes identified for most of the significant water quality
-parameters in prior work. The trend in DO and 
-cholorophyll-*a* concentrations are stable in comparison to other Texas
-estuaries that are facing larger demands for freshwater diversions, higher
-population growth, and more intense agricultural production 
-[@wetzWaterQualityDynamics2016; @bugica_water_2020]. The trend of 
-increasing NO~*x*~ concentration at the upper-bay TCEQ-13563 site and recent 
-increases in TKN concentration at the mid-bay TCEQ-13383
-site are concerning due to the nitrogen limitation identified in many Texas 
+parameters in prior work. Both DO and cholorophyll-*a* concentrations at all three
+Lavaca Bay sites were stable from 2005 through 2020. This is a positive outcome 
+in comparison to other Texas estuaries that are facing larger demands for 
+freshwater diversions, higher population growth, and more intense agricultural 
+production which have resulted in more direct signs of eutrophication
+[@wetzWaterQualityDynamics2016; @bugica_water_2020]. Despite the stability of 
+DO and cholorophyll-*a*, there are concerning site specific increases in 
+NO~*X*~ and TKN concentration over the same time period. These trends are espcially
+concerning due to the nitrogen limitation identified in many Texas 
 estuaries [@gardnerNitrogenFixationDissimilatory2006; 
@@
 
-The strong positive effect of freshwater inflow on NO~*x*~, TKN, and TP are
-suggestive of nonpoint watershed sources, consistent with watershed uses and
+The strong positive effect of freshwater inflow on NO~*X*~, TKN, and TP concentration
+are suggestive of nonpoint watershed sources, consistent with watershed uses and
 with other studies relating freshwater inflow with nutrient 
@@
 freshwater inflow were observed at TCEQ-13384 located in the lower reach of 
-Lavaca Bay. Tidal flushing from Matagorda Bay could be responsible for 
-diluting TKN and acting as a control on the effects of freshwater inflow in 
-lower reaches of Lavaca Bay. Previous work suggests the processing of 
-organic loads in the upper portions of Lavaca Bay might reduce the transport of 
-nutrients into the lower reaches of the Bay [@russell_effect_2006].
-
-Freshwater inflow had a strong positive effect on chlorophyll-*a* at 
+Lavaca Bay. The results suggest that tidal flushing maybe diluting TKN and act as 
+a control on the effects of freshwater inflow in the lower reaches of Lavaca Bay.
+The results are also consistent with previous work that suggest processing of
+organic loads in the upper Lavaca Bay or tidal portions of the Lavaca River reduce
+transport of nutrients to the lower reaches of Lavaca Bay [@russell_effect_2006].
+
+Freshwater inflow also displayed a strong positive effect on chlorophyll-*a* at 
 the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed decreases
@@
 
-GAMs did not identify responses in DO concentration to inflows or nutrient loads.
-The seasonality term in the temporal GAM models explained a substantial amount of DO
-variation at all of the sites. Responses of estuary metabolic processes and 
-resulting DO concentrations can be quite complicated and often locally specific [@caffreyFactorsControllingNet2004]. While the lack of total nitrogen or TKN 
-loading data hinders interpretation, the large seasonal effect on DO suggests 
-physical factors play an important role and should be included in future models.
+There was no evidence that adjusted freshwater inflow and nutrient loads had effects on
+DO concentration in Lavaca Bay. The seasonality term in the temporal GAM models 
+explained a substantial amount of DO variation at all of the sites. Responses of 
+estuary metabolic processes and resulting DO concentrations can be quite 
+complicated and often locally specific [@caffreyFactorsControllingNet2004]. 
+While the lack of total nitrogen or TKN loading data hinders interpretation, 
+the large seasonal effect on DO concentration indicates physical factors 
+(such as temperature, wind, and turbidity) play 
+an important role and should be included in future models.
 Prior work suggests that Lavaca Bay may not be limited by nutrients alone, 
@@
 is reasonable to assume that fluctuations in DO may not occur immediately in 
-response to nutrient pulses or freshwater inflow. Work has has shown that various
+response to nutrient pulses or freshwater inflow. Work has has shown that many
 water quality parameters may have lagged effects lasting days or even months following
-storms and large discharge events [@mooneyWatershedExportEvents2012a; @wetzExtremeFutureEstuaries2013; @bukaveckasInfluenceStormEvents2020; @walkerTimescalesMagnitudeWater2021]. 
-However, our work only evaluates responses to
-loading and inflows occuring the day of water quality observations.
-
-# Conclusion
-
-GAM models appear to provide reliable estimates of nutrient loads in the Lavaca
-Bay watershed. However, additional flow-biased data collection efforts would
+storms and large discharge events [@mooneyWatershedExportEvents2012a; 
+@wetzExtremeFutureEstuaries2013; @bukaveckasInfluenceStormEvents2020; 
+@walkerTimescalesMagnitudeWater2021]. 
+However, this study only evaluated responses to
+loading and inflows occurring the day of water quality observations.
+
+Overall, this study suggests that DO and chlorophyll-*a* concentrations
+have been relatively stable in Lavaca Bay. 
+Site-specific increases in TKN and NO~*X*~ concentrations are
+a cause of concern for increasing risks of eutrophication within Lavaca Bay
+which might be currently attenuated by changes in freshwater flow, turbidity,
+and other physical processes. While loading models indicate that there are
+large annual fluctuations in NO~3~ loads, these changes have been largely driven
+environmental conditions (changes in runoff and river discharge). These models
+also provide evidence that estuary NO~*X*~ and TP
+concentrations are strongly driven by freshwater inflow and to a lesser extent
+fluctuation in flow-adjusted riverine loadings. Site-specific changes in the
+relationships between freshwater inflow and responses in both chlorophyll-*a* 
+and TKN concentrations are indicative nutrient processing and or tidal flushing
+effects moving from the river discharge point to the mouth of Lavaca Bay.
+This study does not completely
+explain site specific increases in NO~*X*~ and TKN concentrations in Lavaca Bay.
+The freshwater study sites did not quantify nutrient loadings from tidal
+contribution areas or ungauaged watersheds. Nutrient contributions from wastewater
+facility discharges, septic systems, and stormwater could be considerable
+contributors to nutrient loadings in Lavaca Bay since they are not processed by
+a tidal river reach prior to entering the Bay.
+The Garcitas-Arenosa Creek, Placedo Creek, and Cox Bay subwatersheds are currently undersampled
+but compose approximately 27% of the watershed area. The contribution of nutrient
+loadings from these undersampled areas is unknown.
+
+## Limitations {.unnumbered}
+
+The GAM approach proved useful for both estimating loads and assessing downstream
+responses in water quality. Although we did not compare other models, it is likely similar
+estimates of loadings would be obtained by methods suggest as LOADEST, WRTDS, or SPARROW
+given the functionally similar dependent variable structures. The underlying weakness 
+in the estimates of loading in the current study is the
+reliance on ambient water quality data used for statewide water quality assessments.
+Cross-validation of the nutrient loading models highlights that predictions are
+prone to high bias, owing to the lack of targeted storm or flow biased measurements.
+The high biases are indicative that subsets of values were unable to capture the
+functional relationships with the flow based dependent variables. It was beyond the scope
+of the current study to evaluate the subsets of cross-validation data and scores.
+However, the cross-validation procedure is indicative that more robust sampling
+is needed. Supplementary flow-biased 
+monitoring targeting storm- or high-flow conditions is critical to
+improve model performance and strength of evidence produced by these models [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017].
+Although there is existing work on the samples sizes and sample design required
+for reliable performance of both LOADEST [@parkUsePollutantLoad2014] and WRTDS 
+[@kumarValueIntensiveSampling2019] models, similar work does not appear to
+have been extended to water quality applications of GAMs.
+
+Due to the 
+concerning increases in eutrophication associated parameters in Lavaca Bay and 
+other Texas estuaries [@bugica_water_2020], 
+and the desire to quantify linkages between environmental outcomes
+and on the ground management actions [@schrammTotalMaximumDaily2022]
+there is a strong need for reliable estimates of pollutant loadings and 
+responses along the Texas coast. Within Texas, state wide water quality monitoring programs have focused on 
+collection of ambient condition data. A framework for establishing pollutant 
+load monitoring programs across catchments that explicitly incorporate
+flow biased data is needed for assessing nutrient loading and estuary health along the Texas coast. 
+Additional efforts focused on identifying relevant effect sizes, sampling 
+designs, and funding mechanisms that can support long term efforts are also 
+needed to adequately design such a framework. Large long-term monitoring programs 
+in and around the Chesapeake Bay, San Francisco Bay, and along the Mississippi 
+River have proven extremely effective at informing management actions and 
+tracking progress towards long-term pollutant reduction 
+goals. Similar coordinated efforts across Texas coastal watersheds would provide
+useful for resource management efforts intended to protect biological and water
+quality integrity of Texas's estuaries.
+
+# Conclusions {.unnumbered}
+
+The primary purpose of this study was to provide estimates of watershed
+nutrient loadings and assess water quality responses to changes in nutrient loads.
+GAMs provided reliable estimates of watershed NO~3~ and TP loads.
+However, additional flow-biased data collection efforts is needed to
 decrease the prediction variance and improve accuracy at critical 
-high flow events. Ongoing projects will fill data gaps for total nitrogen and 
-TKN loading. This study, consistent with others along the Texas coast, found
+high-flow loading events. While some ongoing projects will fill data gaps for 
+total nitrogen and TKN loading, additional efforts are needed to coordinate data 
+collection efforts specifically for load estimation across Texas estuaries. Despite
+these data gaps, this study identified high annual fluctuations in nutrient loads
+driven primarily by discharge. No evidence was identified to indicate that on the
+ground management had changed nutrient loading in the Navidad River subwatershed.
+There was weak evidence for recent reductions in flow-normalized NO~3~ loading in the Lavaca
+River subwatershed although the results are at odds with flow-normalized trends in TP loads.
+
+This study, consistent with others along the Texas coast, found
 strong effects of freshwater flow on nutrient and chlorophyll-*a* 
 concentrations. DO concentrations, dominated by seasonal effects, did not show
-strong direct responses to freshwater flow. Small variance in flow-adjusted
-nutrient loads indicates that (1) there have been limited changes in non-point
+strong direct responses to freshwater flow. Small variances in flow-adjusted
+nutrient loads indicate that (1) there have been limited changes in non-point
 sources of nutrients and (2) that there isn't strong evidence that those small 
 changes have had effects on chlorophyll-*a* or dissolved oxygen in Lavaca Bay.
-Although the study did not identify strong responses to changes in nutrient 
-loading, this does provide a baseline assessment for future water quality 
-management activities in the watershed.
-
-\backmatter
-
-\bmhead{Acknowledgments}
-The author extends thanks to Dr. Mike Wetz (Harte Research Institute, 
-Texas A&M Corpus Christi), Chad Kinsfather and Partick Brzozowski 
-(Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water 
-Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason Pinchbeck 
-(Texas General Land Office) and the Lavaca Bay Foundation for supporting 
-development of this project and providing valuable feedback.
-
-# Declarations {.unnumbered}
-
-\bmhead{Funding}
+Althogh this study did not identify changes in DO or chlorophyll-*a* concentrations
+in Lavaca Bay, site specific increases in NO~*X*~ and TKN are a cause for water 
+quality concern. The study provides a baseline assessment for future water quality 
+management activities in the watershed. In order to effectively track and link 
+improvements or degradation of water quality conditions in Lavaca Bay and other coastal
+Texas watersheds with on the ground effort more robust sampling networks are needed
+to improve spatial coverage of undersampled areas and explicitly incorporate
+flow-biased sampling.
+
+# Funding {.unnumbered}
 This project was funded in part by a Texas Coastal Management Program grant 
@@
 
-\bmhead{Data Availability}
+# Data Availability {.unnumbered}
 Reproducible code and datasets generated during this study are available in

### 2023-05-31 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.tex
trends plot plus lots of edits

--- 2023-05-23 Schramm-2023-05-AS/Schramm-2023-05-AS.tex
@@
-%Version 2.1 April 2023
-% See section 11 of the User Manual for version history
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\documentclass[sn-basic,referee,lineno,pdflatex]{sn-jnl}
-
-%%%% Standard Packages
-%%<additional latex packages if required can be included here>
-
-\usepackage{graphicx}%
-\usepackage{multirow}%
-\usepackage{amsmath,amssymb,amsfonts}%
-\usepackage{amsthm}%
-\usepackage{mathrsfs}%
-\usepackage[title]{appendix}%
-\usepackage{xcolor}%
-\usepackage{textcomp}%
-\usepackage{manyfoot}%
-\usepackage{booktabs}%
-\usepackage{algorithm}%
-\usepackage{algorithmicx}%
-\usepackage{algpseudocode}%
-\usepackage{listings}%
-%%%%
-
-%%%%%=============================================================================%%%%
-%%%%  Remarks: This template is provided to aid authors with the preparation
-%%%%  of original research articles intended for submission to journals published
-%%%%  by Springer Nature. The guidance has been prepared in partnership with
-%%%%  production teams to conform to Springer Nature technical requirements.
-%%%%  Editorial and presentation requirements differ among journal portfolios and
-%%%%  research disciplines. You may find sections in this template are irrelevant
-%%%%  to your work and are empowered to omit any such section if allowed by the
-%%%%  journal you intend to submit to. The submission guidelines and policies
-%%%%  of the journal take precedence. A detailed User Manual is available in the
-%%%%  template package for technical guidance.
-%%%%%=============================================================================%%%%
-
+%% Submissions for peer-review must enable line-numbering
+%% using the lineno option in the \documentclass command.
+%%
+%% Preprints and camera-ready submissions do not need
+%% line numbers, and should have this option removed.
+%%
+%% Please note that the line numbering option requires
+%% version 1.1 or newer of the wlpeerj.cls file, and
+%% the corresponding author info requires v1.2
+
+\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
+
+% ZNK -- Adding headers for pandoc
+
+\setlength{\emergencystretch}{3em}
+\usepackage[unicode=true]{hyperref}
+
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\usepackage{csquotes}
 \usepackage{booktabs}
@@
 
-\raggedbottom
-
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\begin{document}
-
-\title[Watershed nutrient loading and water quality in a subtropical
-estuary]{Assessing linkages between watershed nutrient loading and water
-quality in a subtropical estuary with semiparametric models.}
-
-%%=============================================================%%
-%% Prefix	-> \pfx{Dr}
-%% GivenName	-> \fnm{Joergen W.}
-%% Particle	-> \spfx{van der} -> surname prefix
-%% FamilyName	-> \sur{Ploeg}
-%% Suffix	-> \sfx{IV}
-%% NatureName	-> \tanm{Poet Laureate} -> Title after name
-%% Degrees	-> \dgr{MSc, PhD}
-%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate}
-%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
-%%=============================================================%%
-
-\author*[1]{\fnm{Michael} \sur{Schramm} }\email{\href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}
-
-  \affil*[1]{\orgdiv{Texas Water Resources Institute}, \orgname{Texas
-A\&M AgriLife Research}, \orgaddress{\city{College
-Station}, \country{United
-States}, \postcode{77840}, \state{TX}, \street{1001 Holleman Dr.~E.}}}
-
-\abstract{Lavaca Bay is a secondary embayment on the Texas coast
-displaying early signals of water quality degradation. This study
-applied a semiparametric approach to assess both watershed nutrient
-loads and responses in estuary water quality to nutrient loading and
-streamflow. Cross-validation indicated that, despite data constraints,
+\title{Linking watershed nutrient loading to estuary water quality with
+semiparametric models}
+
+\author[1]{Michael Schramm}
+
+\corrauthor[1]{Michael
+Schramm}{\href{mailto:michael.schramm@ag.tamu.edu}{\nolinkurl{michael.schramm@ag.tamu.edu}}}
+
+\affil[1]{Texas Water Resources Institute, Texas A\&M AgriLife Research,
+College Station, Texas, United States}
+
+%
+% \author[1]{First Author}
+% \author[2]{Second Author}
+% \affil[1]{Address of first author}
+% \affil[2]{Address of second author}
+% \corrauthor[1]{First Author}{f.author@email.com}
+
+% 
+\usepackage[style=apa,]{biblatex}
+\addbibresource{bibliography.bib}
+
+\begin{abstract}
+Evaluating estuary water quality responses to reductions (or increases)
+in nutrient loading attributed to on the ground management actions can
+be challenging due to the strong influence of environmental drivers on
+nutrient loads and non-linear relationships. This study applied
+semiparametric approaches to calculate watershed nutrient loads and
+assess responses in estuary water quality to seasonally-adjusted
+freshwater inflow and flow-adjusted nutrient loads in Lavaca Bay, Texas.
+Lavaca Bay is a secondary embayment on the Texas coast displaying early
+potential for eutrophication and water quality degredation. Use of
+flow-adjusted nutrient loads allowed the study to evaluate the response
+in water quality to changes in nutrient loads driven by non-natural
+variation. Cross-validation indicated that, despite data constraints,
 semiparametric models performed well at nutrient load prediction. Based
@@
 precipitation and runoff as opposed to changes in management of nonpoint
-sources. Models indicated there was no evidence of long-term changes in
-dissolved oxygen or chlorophyll-\emph{a} within Lavaca Bay. However,
-site specific long-term increases in both organic and inorganic nitrogen
-are concerning. Further analysis found freshwater inflow was a strong
-driver of nutrient and chlorophyll-\emph{a} concentrations within Lavaca
-Bay but there was no evidence that changes in watershed nutrient loading
-explained variation in dissolved oxygen or chlorophyll-\emph{a}
-concentrations. In addition to providing a baseline assessment of
-watershed nutrient loading and water quality responses in the Lavaca Bay
-watershed, this study provides methodological support for the use of
-semiparametric methods in load regression models and estuary
-assessments.}
-
-\keywords{key, dictionary, word}
-
+sources. Models indicated no evidence of long-term changes in dissolved
+oxygen or chlorophyll-\emph{a} within Lavaca Bay. However, site specific
+long-term increases in both organic and inorganic nitrogen are causes of
+concern as potential causes of future eutrophication. Further analysis
+found freshwater inflow was a strong driver of nutrient and
+chlorophyll-\emph{a} concentrations within Lavaca Bay but there was no
+evidence that changes in watershed nutrient loading explained additional
+variation in dissolved oxygen or chlorophyll-\emph{a} concentrations. In
+addition to providing a baseline assessment of watershed nutrient
+loading and water quality responses in the Lavaca Bay watershed, this
+study provides methodological support for the use of semiparametric
+methods in load regression models and estuary assessments.
+% Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+Dummy abstract text.
+\end{abstract}
+
+\begin{document}
+
+\flushbottom
 \maketitle
-
-\emph{Remaining Items}
-
-\begin{itemize}
-\tightlist
-\item
-  keywords
-\item
-  discussion
-\item
-  formalize language (reduce first person active?)
-\item
-  final language review
-\item
-  springer doesn't want figures in subdirectory, need to knit to same
-  directory as tex file.
-\end{itemize}
-
-\hypertarget{sec1}{%
-\section{Introduction}\label{sec1}}
+\thispagestyle{empty}
+
+\hypertarget{introduction}{%
+\section*{Introduction}\label{introduction}}
+\addcontentsline{toc}{section}{Introduction}
 
@@
 to freshwater flows that degrade water quality in downstream estuaries
-\citep{bricker_effects_2008, kennicuttWaterQualityGulf2017, bugica_water_2020}.
+\autocite{bricker_effects_2008,kennicuttWaterQualityGulf2017,bugica_water_2020}.
 Despite these escalating pressures, national scale assessments have
 classified coastal estuaries in Texas as moderate or low risk for
-eutrophic conditions \citep{bricker_effects_2008}. However, a suite of
-recent studies indicates that estuary water quality dynamics in both
+eutrophic conditions \autocite{bricker_effects_2008}. However, a suite
+of recent studies indicates that estuary water quality dynamics in both
 agriculturally dominated and urban watersheds within Texas are
@@
 and eutrophication
-\citep{wetzWaterQualityDynamics2016, wetz_exceptionally_2017, bugica_water_2020, chinPhytoplanktonBiomassCommunity2022}.
+\autocite{wetzWaterQualityDynamics2016,wetz_exceptionally_2017,bugica_water_2020,chinPhytoplanktonBiomassCommunity2022}.
 With identification of several localized areas of estuary water quality
-concern along the Texas coast \citep{bugica_water_2020}, localized
+concern along the Texas coast \autocite{bugica_water_2020}, localized
 studies are being prioritized to better inform management actions.
 
-This project aims to provide an assessment of watershed nutrient loading
-and resulting water quality responses in Lavaca Bay, Texas. Lavaca Bay
+This goal of this project is to assess watershed nutrient loading and
+the resulting water quality responses in Lavaca Bay, Texas. Lavaca Bay
 is a secondary bay in the larger Matagorda Bay system located roughly
@@
 quality standards. Despite largely meeting statewide water quality
-standards, there have been concerning recently identified declines in
-abundance, biomass, and diversity of benthic fauna in Lavaca Bay
-\citep{beserespollackLongtermTrendsResponse2011}. These declines are
+standards, there have been concerning declines in abundance, biomass,
+and diversity of benthic fauna in Lavaca Bay
+\autocite{beserespollackLongtermTrendsResponse2011}. These declines are
 partially attributed to reductions in freshwater inflow and changes in
 estuary salinity and are indicative of an already stressed system
-\citep{beserespollackLongtermTrendsResponse2011, palmerImpactsDroughtsLow2015, montagnaAssessmentRelationshipFreshwater2020}.
+\autocite{beserespollackLongtermTrendsResponse2011,palmerImpactsDroughtsLow2015,montagnaAssessmentRelationshipFreshwater2020}.
 More recently, significant linear increases in total phosphorus (TP),
@@
 concentrations were identified at monitoring sites within Lavaca Bay
-\citep{bugica_water_2020}. Although long-term changes in DO
+\autocite{bugica_water_2020}. Although long-term changes in DO
 concentrations have not been identified, the trends in nutrient
@@
 factor for primary production in many Texas estuaries
-\citep{gardnerNitrogenFixationDissimilatory2006, houTransformationFateNitrate2012, doradoUnderstandingInteractionsFreshwater2015, paudelRelationshipSuspendedSolids2019, wetz_exceptionally_2017}
+\autocite{gardnerNitrogenFixationDissimilatory2006,houTransformationFateNitrate2012,doradoUnderstandingInteractionsFreshwater2015,paudelRelationshipSuspendedSolids2019,wetz_exceptionally_2017}
 and the ramifications that changes in nitrogen loadings could have for
@@
 Lavaca Bay watershed
-\citep{jainTechnicalSupportDocument2021, schrammLavacaRiverWatershed2018, bertholdDirectMailingEducation2021}.
+\autocite{jainTechnicalSupportDocument2021,schrammLavacaRiverWatershed2018,bertholdDirectMailingEducation2021}.
 However, at a statewide scale, these approaches have shown limited
@@
 replicate effective management actions across the state
-\citep{schrammTotalMaximumDaily2022}. The identification and
+\autocite{schrammTotalMaximumDaily2022}. The identification and
 communication of changes and trends in water quality is complicated by
@@
 precipitation and runoff that hinder traditional analysis
-\citep{wazniakLinkingWaterQuality2007, lloydMethodsDetectingChange2014}.
+\autocite{wazniakLinkingWaterQuality2007,lloydMethodsDetectingChange2014}.
 The development and application of flexible statistical methods such as
 Weighted Regressions on Time, Discharge and Season
-\citep[WRTDS,][]{hirschWeightedRegressionsTime2010} and Generalized
-Additive Models \citep[GAMs,][]{woodFastStableRestricted2011} have
+\autocite[WRTDS,][]{hirschWeightedRegressionsTime2010} and Generalized
+Additive Models \autocite[GAMs,][]{woodFastStableRestricted2011} have
 provided effective tools for researchers to quantify and communicate
@@
 to assess and identify trends in riverine nutrients
-\citep{oelsner_recent_2019, stackpooleLongTermMississippi2021},
-chlorides \citep{stetsIncreasingChlorideRivers2018}, and other
-pollutants of concern \citep{shodaWaterqualityTrendsRivers2019}. WRTDS
-has also been succesfully adapted to assess trends in estuarine water
-quality concentrations \citep{beckFourDecadesWater2018}.
+\autocite{oelsner_recent_2019,stackpooleLongTermMississippi2021},
+chlorides \autocite{stetsIncreasingChlorideRivers2018}, and other
+pollutants of concern \autocite{shodaWaterqualityTrendsRivers2019}.
+WRTDS has also been succesfully adapted to assess trends in estuarine
+water quality concentrations \autocite{beckFourDecadesWater2018}.
 
@@
 smooth functions and parametric linear predictors
-\citep{woodFastStableRestricted2011}. Although the underlying parameter
-estimation procedure of GAMs is substantially different than WRTDS, both
-the functional form and results have been demonstrated to be similar
-when assessing nutrient concentration trends
-\citep{beckNumericalQualitativeContrasts2017}. Water quality
-applications of GAMs have included include river and catchment nutrient
+\autocite{woodFastStableRestricted2011}. Although the underlying
+parameter estimation procedure of GAMs are substantially different than
+WRTDS, both the functional form and results have been demonstrated to be
+similar when assessing nutrient concentration trends
+\autocite{beckNumericalQualitativeContrasts2017}. Water quality
+applications of GAMs have included river and catchment nutrient
 concentration and load estimation
-\citep{wangLoadEstimationUncertainties2011, kroonRiverLoadsSuspended2012, kuhnertQuantifyingTotalSuspended2012, robson_prediction_2015-1, hagemannEstimatingNutrientOrganic2016, mcdowell_implications_2021, biagi_novel_2022},
-assessment of temporal trends of nutrients
-\citep{beckNumericalQualitativeContrasts2017, murphyGeneralizedAdditiveModel2019},
-phytoplankton \citep{bergbuschUnexpectedShiftPhytoplankton2021}, and
-cyanobacteria \citep{hayesEffectsLakeWarming2020}. Recently GAMs have
+\autocite{wangLoadEstimationUncertainties2011,kroonRiverLoadsSuspended2012,kuhnertQuantifyingTotalSuspended2012,robson_prediction_2015-1,hagemannEstimatingNutrientOrganic2016,mcdowell_implications_2021,biagi_novel_2022},
+and assessment of temporal trends of nutrients
+\autocite{beckNumericalQualitativeContrasts2017,murphyGeneralizedAdditiveModel2019},
+phytoplankton \autocite{bergbuschUnexpectedShiftPhytoplankton2021}, and
+cyanobacteria \autocite{hayesEffectsLakeWarming2020}. Recently GAMs have
 also been used to link water quality responses in receiving water bodies
 to changes in nonpoint source nutrient inputs
-\citep{murphyNutrientImprovementsChesapeake2022}.
-\citet{beckNumericalQualitativeContrasts2017} provides a substantial
+\autocite{murphyNutrientImprovementsChesapeake2022}. For a substantial
 discussion on the differences (and similarities) between GAMs and WRTDS
-for water quality applications.
+for water quality applications reders are referred to
+\textcite{beckNumericalQualitativeContrasts2017}.
 
@@
 human impacts. This study utilizes GAMs to develop estimates of
-delivered and flow-normalized nutrient loads and assess changes in loads
-delivered to Lavaca Bay. GAMs were chosen over other regression-based
-approached for use in this study due to; (1) the ability to easily
-explore and incorporate different model terms; (2) the incorporation of
-non-liner smooth functions that do not require explicit a priori
-knowledge of the expected shape; and (3) inclusion of a link function
-that related the expected value of the response to linear predictors
-thus avoiding unneeded data transformations and bias corrections. The
-exploratory study also assesses the response of water quality parameters
-in Lavaca Bay over time and in response to freshwater inflow controlled
-for seasonality and to watershed nutrient loads that are controlled for
-environmentally driven variation.
-
-\hypertarget{sec2}{%
-\section{Methods}\label{sec2}}
+delivered and flow-normalized nutrient loads and assess estuary water
+quality responses to changes in loads delivered to Lavaca Bay. GAMs were
+chosen over other regression-based approached for use in this study due
+to; (1) the ability to easily explore and incorporate different model
+terms; (2) the incorporation of non-linear smooth functions that do not
+require explicit a priori knowledge of the expected shape; and (3)
+inclusion of a link function that related the expected value of the
+response to linear predictors thus avoiding unneeded data
+transformations and bias corrections. The exploratory study also
+assesses the response of water quality parameters in Lavaca Bay over
+time and in response to freshwater inflow controlled for seasonality and
+to watershed nutrient loads that are controlled for environmentally
+driven variation.
+
+\hypertarget{methods-and-materials}{%
+\section*{Methods and Materials}\label{methods-and-materials}}
+\addcontentsline{toc}{section}{Methods and Materials}
 
 \hypertarget{location-and-data}{%
-\subsection{Location and Data}\label{location-and-data}}
-
-Lavaca Bay is 190 km\textsuperscript{2} with the majority of freshwater
-inflow provided by the Lavaca and Navidad River systems
-(Figure~\ref{fig:fig1}). The Garcitas-Arenosa, Placedo Creek, and Cox
-Bay watersheds provide additional freshwater inflows. The entire
-watershed land area is 8,149 km\textsuperscript{2} and primarily rural.
+\subsection*{Location and Data}\label{location-and-data}}
+\addcontentsline{toc}{subsection}{Location and Data}
+
+Lavaca Bay is a 190 km\textsuperscript{2} estuary with the majority of
+freshwater inflow provided by the Lavaca and Navidad River systems
+(Fig.~\ref{fig:fig1}). The Garcitas-Arenosa, Placedo Creek, and Cox Bay
+subwatersheds provide additional freshwater inflows. The entire
+watershed area is 8,149 km\textsuperscript{2} and primarily rural.
 Watershed land cover and land use is 50\% grazed pasture and rangeland,
@@
 Daily discharges for the Lavaca River (USGS-08164000,
-Figure~\ref{fig:fig1}) were obtained from the United States Geologic
+Fig.~\ref{fig:fig1}) were obtained from the United States Geologic
 Survey (USGS) National Water Information System using the
 \emph{dataRetrieval} R package
-\citep{deciccoDataRetrievalPackagesDiscovering2022}. Gaged daily
+\autocite{deciccoDataRetrievalPackagesDiscovering2022}. Gaged daily
 discharges from the outlet of Lake Texana on the Navidad River
@@
 
-Water quality sample data for the two freshwater and three estuary
-locations were obtained from the Texas Commission on Environmental
-Quality (TCEQ) Surface Water Quality Monitoring Information System. Data
-submitted through the system are required to be collected under Quality
-Assurance Project Plans and lab method procedures outlined by the TCEQ's
+Water quality data for the two freshwater and three estuary locations
+were obtained from the Texas Commission on Environmental Quality (TCEQ)
+Surface Water Quality Monitoring Information System. Data submitted
+through the system are required to be collected under Quality Assurance
+Project Plans and lab method procedures outlined by the TCEQ's
 procedures manual to ensure consistent collection and laboratory methods
@@
 estimation models
-\citep{horowitzEvaluationSedimentRating2003, snelderEstimationCatchmentNutrient2017}.
+\autocite{horowitzEvaluationSedimentRating2003,snelderEstimationCatchmentNutrient2017}.
 The three estuary sites included an upper Lavaca Bay site near the
@@
 (TCEQ-13384). For estuary locations, we obtained data for TP,
-Nitrite+Nitrate (NO\emph{\textsubscript{x}}), TKN, chlorophyll-\emph{a},
+Nitrite+Nitrate (NO\textsubscript{\emph{X}}), TKN, chlorophyll-\emph{a},
 and DO concentrations from January 2005 through December 2020
@@
 \hypertarget{estimating-watershed-based-nutrient-loads}{%
-\subsection{Estimating Watershed Based Nutrient
+\subsection*{Estimating Watershed Based Nutrient
 Loads}\label{estimating-watershed-based-nutrient-loads}}
+\addcontentsline{toc}{subsection}{Estimating Watershed Based Nutrient
+Loads}
 
@@
 functions with automatic smoothness selection
-\citep{woodFastStableRestricted2011}. The general form of the model
+\autocite{woodFastStableRestricted2011}. The general form of the model
 related NO\textsubscript{3} or TP concentration to a long term tend,
@@
 concentration at the current time period
-\citep{wangLoadEstimationUncertainties2011, kuhnertQuantifyingTotalSuspended2012, zhang_improving_2017},
+\autocite{wangLoadEstimationUncertainties2011,kuhnertQuantifyingTotalSuspended2012,zhang_improving_2017},
 using Eq. \ref{eq2}:
 
-\begin{align}
-ma(\delta) &= d{\kappa_{i-1}}+(1-\delta)\hat{q}_{i-1}, \nonumber \\
-\kappa_{i} &=\sum_{m=1}^{i}\hat{Q}_m
+\begin{equation}
+ma(t) = \frac{\sum^{J}_{j=1}{d^{t-t_j}Q_{t_{j}}}}{\sum^{J}_{j=1}d^{t-t_j}}
 \label{eq2}
-\end{align}
-
-where \(\delta\) is the discount factor (here, set equal to 0.95),
-\(\kappa_i\) is the cumulative flow (\emph{Q}) up to the \emph{i}th day.
+\end{equation}
+
+where \emph{t} is the day of estimation, \emph{d} is the discount factor
+(set to 0.95), \emph{J} is the total number of daily discharge
+observations, \(t-t_j\) is the lag time between the current day and some
+historical day of observation (\(t_j\)). Here, as \emph{d} approaches
+zero, the time series becomes the daily observed values and as \emph{d}
+gets closer to one, the time series becomes the mean flow.
 
@@
 dry the current time period is from a previous time period
-\citep{vecchiaTrendsConcentrationsUse2009, zhang_improving_2017}.
+\autocite{vecchiaTrendsConcentrationsUse2009,zhang_improving_2017}.
 Long-term flow anomaly (\emph{ltfa}) is the streamflow over the previous
-year relative to the entire period \citep{zhang_improving_2017} and
-calculated using Eq. \ref{eq3}:
-
-\begin{equation}
-ltfa(t) = \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}
-\label{eq3}
-\end{equation}
-
-and the short-term flow anomaly (\emph{stfa}) calculated as the current
-day flow compared to the preceding 1-month streamflow using \ref{eq4}:
-
-\begin{equation}
-stfa(t) = x_{current\,day}(t) - \bar{x}_{1\,month}(t)
-\label{eq4}
-\end{equation}
+year relative to the entire period
+\autocite[Eq.~\ref{eq3},][]{zhang_improving_2017} and and the short-term
+flow anomaly (\emph{stfa}) calculated as the current day flow compared
+to the preceding 1-month streamflow
+\autocite[Eq.~\ref{eq4},][]{zhang_improving_2017}:
+
+\begin{align}
+ltfa(t) &= \bar{x}_{1\,year}(t) - \bar{x}_{entire\,period}, \label{eq3}\\
+stfa(t) &= x_{current\,day}(t) - \bar{x}_{1\,month}(t) \label{eq4}
+\end{align}
 
@@
 \emph{stfa} in TP models based on previous work demonstrating major
-improvements in NO\textsubscript{x} regression models that incorporated
-\emph{ltfa} and moderate improvements in TP regression models that
-incorporated \emph{stfa} \citep{zhang_improving_2017}. Moving averages
-and flow anomalies were calculated with the \emph{adc} R package
-\citep{schrammAdcCalculateAntecedant2023}.
+improvements in NO\textsubscript{\emph{X}} regression models that
+incorporated \emph{ltfa} and moderate improvements in TP regression
+models that incorporated \emph{stfa} \autocite{zhang_improving_2017}.
+Moving averages and flow anomalies were calculated with the \emph{adc} R
+package \autocite{schrammAdcCalculateAntecedant2023}.
 
@@
 simple approach can introduce bias
-\citep{hornungEstimationAverageConcentration1990}, we considered it
+\autocite{hornungEstimationAverageConcentration1990}, we considered it
 acceptable because high concentrations and loadings are associated with
 high-flow events and low-flow/low-concentration events will account for
-a small proportion of total loadings \citep{mcdowell_implications_2021}.
+a small proportion of total loadings
+\autocite{mcdowell_implications_2021}.
 
@@
 each of the historical values for that day of the year over the study
-period \citep{hirschWeightedRegressionsTime2010}. The flow-normalized
+period \autocite{hirschWeightedRegressionsTime2010}. The flow-normalized
 estimate was calculated as the mean of all the predictions for each day
@@
 multivariate normal posterior distribution of the fitted GAM
-\citep{woodConfidenceIntervalsGeneralized2006, marraCoveragePropertiesConfidence2012, mcdowell_implications_2021}.
+\autocite{woodConfidenceIntervalsGeneralized2006,marraCoveragePropertiesConfidence2012,mcdowell_implications_2021}.
 GAM performance was evaluated using repeated 5-fold cross validation
-\citep{burmanComparativeStudyOrdinary1989} and average Nash-Sutcliffe
+\autocite{burmanComparativeStudyOrdinary1989} and average Nash-Sutcliffe
 Efficiency (NSE), Pearson sample correlation (\emph{r}) and percent bias
@@
 \hypertarget{linking-estuary-water-quality-to-hydrology-and-nutrient-loads}{%
-\subsection{Linking Estuary Water Quality to Hydrology and Nutrient
+\subsection*{Linking Estuary Water Quality to Hydrology and Nutrient
 Loads}\label{linking-estuary-water-quality-to-hydrology-and-nutrient-loads}}
+\addcontentsline{toc}{subsection}{Linking Estuary Water Quality to
+Hydrology and Nutrient Loads}
 
@@
 Because streamflow and nutrient loads are tightly correlated, freshwater
-inflow can mask signals from nutrient loads alone. Following the
-methodology implemented by
-\citet{murphyNutrientImprovementsChesapeake2022}, both freshwater inflow
-and nutrient loads were prepossessed to account for season and
+inflow can mask signals due to chnages in nutrient loads alone.
+Following the methodology implemented by
+\textcite{murphyNutrientImprovementsChesapeake2022}, both freshwater
+inflow and nutrient loads were prepossessed to account for season and
 streamflow respectively. Raw freshwater inflow values were replaced by
@@
 AIC\textsubscript{c} scores between each group of temporal, inflow, and
-inflow+load models \citep{burnhamAICModelSelection2011}. Improvements in
-model probabilities provide evidence that the terms explain additional
-variation in the response variable.
-
-\hypertarget{sec3}{%
-\section{Results}\label{sec3}}
+inflow+load models \autocite{burnhamAICModelSelection2011}. Improvements
+in model probabilities provide evidence that the terms explain
+additional variation in the response variable.
+
+\hypertarget{results}{%
+\section*{Results}\label{results}}
+\addcontentsline{toc}{section}{Results}
 
 \hypertarget{watershed-nutrient-loads}{%
-\subsection{Watershed Nutrient Loads}\label{watershed-nutrient-loads}}
+\subsection*{Watershed Nutrient Loads}\label{watershed-nutrient-loads}}
+\addcontentsline{toc}{subsection}{Watershed Nutrient Loads}
 
@@
 ``very good'' based on standardized evaluation metrics of NSE, \emph{r},
-and PBIAS \citep{moriasiHydrologicWaterQuality2015} calculated using
+and PBIAS \autocite{moriasiHydrologicWaterQuality2015} calculated using
 5-fold cross validation. Median goodness-of-fit metrics for
@@
 River had a much larger variance in values compared to the Navidad River
-(Figure~\ref{fig:fig2}). TP GAMS had higher average NSE and \emph{r}
+(Fig.~\ref{fig:fig2}). TP GAMS had higher average NSE and \emph{r}
 values and lower variance in metric values compared to
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig2-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig2-1}
 
@@
 Annual NO\textsubscript{3} and TP loads show considerable variation,
-generally following patterns in discharge (Figures~\ref{fig:fig3},
+generally following patterns in discharge (Fig.~\ref{fig:fig3}, Fig.
 \ref{fig:fig4}). Flow-normalized TP loads at both sites and
@@
 watershed based loads did not change much over time when accounting for
-variation driven by streamflow (Figure~\ref{fig:fig3}). Flow-normalized
+variation driven by streamflow (Fig.~\ref{fig:fig3}). Flow-normalized
 loads in the Lavaca River showed small variation over time with some
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig3-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig3-1}
 
@@
 watershed with the Navidad River only accounting for 15\% and 25\% of
-NO\textsubscript{3} and TP loads in 2011 (Figure~\ref{fig:fig4}).
+NO\textsubscript{3} and TP loads in 2011 (Fig.~\ref{fig:fig4}).
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig4-1}
+{\centering \includegraphics[width=0.5\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig4-1}
 
@@
 \hypertarget{linkages-between-water-quality-and-watershed-flows-and-loads}{%
-\subsection{Linkages Between Water Quality and Watershed Flows and
+\subsection*{Linkages Between Water Quality and Watershed Flows and
 Loads}\label{linkages-between-water-quality-and-watershed-flows-and-loads}}
-
-There is no evidence of long-term changes in TP or DO concentrations at
-any Lavaca Bay site (Figure~\ref{fig:fig5}). The upper-bay site,
-TCEQ-13563, shows evidence of a long-term linear increase in
-NO\textsubscript{x} while chlorophyll-\emph{a} decreased from 2005
-through 2014 (Figure~\ref{fig:fig5}). NO\textsubscript{x} concentration
-at the mid-bay site, TCEQ-13383, displayed an unusual periodic pattern
-that is indicative of a strong influence from inflow or precipitation.
-The temporal GAMs did not provide evidence of long-term trends in any of
-the water quality constituents at the lower-bay TCEQ-13384 site.
+\addcontentsline{toc}{subsection}{Linkages Between Water Quality and
+Watershed Flows and Loads}
+
+There was no evidence of long-term changes in TP or DO concentrations at
+any Lavaca Bay site (Fig.~\ref{fig:fig5}). The upper-bay site,
+TCEQ-13563, had evidence of a long-term linear increase in
+NO\textsubscript{\emph{X}} while chlorophyll-\emph{a} decreased from
+2005 through 2014 (Fig.~\ref{fig:fig5}). NO\textsubscript{\emph{X}}
+concentration at the mid-bay site, TCEQ-13383, displayed an unusual
+periodic pattern that is indicative of a strong influence from inflow or
+precipitation. The temporal GAMs did not provide evidence of long-term
+trends in any of the water quality constituents at the lower-bay
+TCEQ-13384 site.
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig5-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig5-1}
 
@@
 
-\caption{Fitted splines (shaded regions indicate 90\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\textit{\textsubscript{x}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig5}
+\caption{Fitted splines (shaded regions indicate 90\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\textit{\textsubscript{X}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig5}
 \end{figure}
@@
 Freshwater inflow provided additional explanation for changes in TP and
-NO\textsubscript{\emph{x}} concentration at all of the Lavaca Bay sites
+NO\textsubscript{\emph{X}} concentration at all of the Lavaca Bay sites
 according to AIC\textsubscript{c} and model probability values
@@
 
-GAMs showed that increases in freshwater inflow resulted in nearly
-linear increases in TP and NO\textsubscript{\emph{x}} concentration at
-all three sites (Figure~\ref{fig:fig6}). At the upper-bay TCEQ-13563
-site, GAMs showed that increases in freshwater inflow initially
-increased chlorophyll-\emph{a} and DO concentration but concentrations
-leveled and potentially decreased at higher flows. The mid-bay
-TCEQ-13383 site showed a nearly linear increased in chlorophyll-\emph{a}
-concentration in response to increases freshwater inflow. Freshwater
-flow did not have significant effects on chlorophyll-\emph{a}, TKN, or
-DO at the lower-bay TCEQ-13384 site.
+Increases in freshwater inflow resulted in nearly linear increases in TP
+and NO\textsubscript{\emph{X}} concentration at all three sites
+(Fig.~\ref{fig:fig6}). At the upper-bay TCEQ-13563 site, increases in
+freshwater inflow initially increased chlorophyll-\emph{a} and DO
+concentration but concentrations leveled and potentially decreased at
+higher flows. The mid-bay TCEQ-13383 site had nearly linear increases in
+chlorophyll-\emph{a} concentration in response to increased freshwater
+inflow. There was no evidence of freshwater inflow effects on
+chlorophyll-\emph{a}, TKN, or DO at the lower-bay TCEQ-13384 site.
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig6-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig6-1}
 
@@
 
-\caption{Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\textit{\textsubscript{x}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig6}
+\caption{Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\textit{\textsubscript{X}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig6}
 \end{figure}
@@
 concentration at the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383
-respectively (Figure~\ref{fig:fig7}). The relative effect size appeared
-to much smaller than the effect of freshwater inflow alone. Increased
+respectively (Fig.~\ref{fig:fig7}). The relative effect size appeared
+much smaller than the effect of freshwater inflow alone. Increased
 NO\textsubscript{3} loads only showed an effect at the lower-bay
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-05-AS_files/figure-latex/fig7-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig7-1}
 
@@
 
-\caption{Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\textsubscript{3} loads on TP (A) and NO\textsubscript{\textit{x}} (B) concentrations at each site in Lavaca Bay.}\label{fig:fig7}
+\caption{Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\textsubscript{3} loads on TP (A) and NO\textsubscript{\textit{X}} (B) concentrations at each site in Lavaca Bay.}\label{fig:fig7}
 \end{figure}
 
-\hypertarget{sec4}{%
-\section{Discussion}\label{sec4}}
-
-\textbf{resume here}
-
-\hypertarget{watershed-nutrient-loads-1}{%
-\subsection{Watershed Nutrient Loads}\label{watershed-nutrient-loads-1}}
+\hypertarget{discussion}{%
+\section*{Discussion}\label{discussion}}
+\addcontentsline{toc}{section}{Discussion}
+
+\hypertarget{nutrient-loads}{%
+\subsection*{Nutrient Loads}\label{nutrient-loads}}
+\addcontentsline{toc}{subsection}{Nutrient Loads}
 
 TP and NO\textsubscript{3} loadings from the Lavaca Bay watershed showed
-high inter-annual variability tied closely with changes in discharge.
-There is little evidence for changes in flow-normalized TP loads in
-either river. There is some evidence of recent decreases in
-flow-normalized NO\textsubscript{3} loads in the Lavaca River. Although
-there is no work directly correlating water quality planning and
-implementation efforts in the watershed to water quality outcomes,
-efforts to increase agricultural producer participation in the watershed
-have been ongoing since 2016
-\citep{schrammLavacaRiverWatershed2018, bertholdDirectMailingEducation2021}.
-The decrease in flow-normalized NO\textsubscript{3} loads could be a
-reflection of those collective efforts but further data collection and
-research is required to support that statement.
-
-Converted to average annual yield, the estimates of annual TP loads for
-the Lavaca River are within the ranges in previous published studies
-\citep[Table~\ref{tab:table5},][]{dunnTrendsNutrientInflows1996, rebichSourcesDeliveryNutrients2011, omaniEstimationSedimentNutrient2014, wise_spatially_2019}.
-It isn't obvious why TP estimates in
-\citet{dunnTrendsNutrientInflows1996} were notably lower. Given that
-none of the studies identify substantially sized trends in TP, it is is
-possible that the period used in \citet{dunnTrendsNutrientInflows1996}
-was drier on average than the other studies. The SPARROW models used in
-\citet{rebichSourcesDeliveryNutrients2011} and
-\citet{wise_spatially_2019} utilize a version of LOADEST in the
-underlying load estimation procedure, so a difference due to methodology
-alone is unlikely.
+high inter-annual variability driven primarily by fluctuations in
+discharge. Notably, there were no indications of trends in
+flow-normalized NO\textsubscript{3} and TP loads in the Navidad River.
+In comparison, there was weak evidence of more recent decreases in
+flow-normalized NO\textsubscript{3} (but not TP) loads in the Lavaca
+River watershed. While the dominant agricultural land uses differ
+between the Lavaca (primarily grazed pasture and rangeland) and Navidad
+(mix of pasture and row crops) catchments, we did not have a reason to
+expect different flow normalized trends between the two systems from
+land use alone. Freshwater discharges in the Navidad River are regulated
+by the Palmetto Bend Dam forming Lake Texana at the lower extent of the
+river. Lentic nitrogen uptake and cycling may have regulating effects
+that mask changes in upstream nitrogen loadings. Additional nutrient
+data collection in the contributing tributaries of Lake Texana is needed
+to fully assess the role of Lake Texana in regulating nutrient delivery
+to the Lavaca Bay system. However, these results suggest that there have
+been no changes in the NO\textsubscript{3} or TP loading from the
+Navidad River system at the Lake Texana discharge point when accounting
+for variations in year to year discharge.
+
+The evidence of decreased Lavaca River NO\textsubscript{3} loading,
+although weak, is a potential positive sign for water quality managers
+working to implement practices that improve water quality in the
+freshwater sections of the Lavaca River watershed. Planning and
+implementation efforts to increase agricultural producer participation
+in water quality protection practices have been ongoing in the watershed
+since 2016
+\autocite{schrammLavacaRiverWatershed2018,bertholdDirectMailingEducation2021},
+however little work has been conducted to directly link these efforts
+with water quality outcomes. The decrease in flow-normalized
+NO\textsubscript{3} loads could be a reflection of those collective
+efforts but the lack of evidence for similar changes in flow-normalized
+TP loads provide contrary support. The inconsistent flow-normalized
+trends may also reflect some of the weakness of the water quality
+dataset that is primarily composed of ambient water quality
+measurements. The issues associated with the lack of flow-biased
+measurements is further discussed later in this section.
+
+Some prior studies have generated estimates of mean annual TP yields in
+the Lavaca River watershed
+\autocites[Table~\ref{tab:table5},][]{dunnTrendsNutrientInflows1996,rebichSourcesDeliveryNutrients2011,omaniEstimationSedimentNutrient2014,wise_spatially_2019}.
+Although these studies differ in time periods and methodologies, they
+provide a sanity check for the reasonableness of the annual estimates
+generated in the current study. In a regional assessment of nutrient
+loading in river's along the Gulf of Mexico,
+\textcite{dunnTrendsNutrientInflows1996} used the LOADEST model to
+develop an estimated mean annual yield of 28.9 kg/km\textsuperscript{2}.
+LOADEST is a multiple linear regression model that fits log transformed
+pollutant concentrations to long term, seasonal, and flow based
+predictors and includes methods for bias correction when exponentiating
+the response variable. While the annual yield estimated in
+\textcite{dunnTrendsNutrientInflows1996} is lower than the estimate of
+35.2 kg/km\textsuperscript{2} in this study, it is important to note
+that both hydrology and land use was different for a proportion of the
+study period. For almost the first ten years of the study period in
+\textcite{dunnTrendsNutrientInflows1996}, the Navidad River was still
+undammed. More recently the eastern portion of the watershed has seen
+large conversions of rice fields to more traditional row crop production
+\autocite{kulatHolisticWaterEnergyFoodNexus2019}. The apparent increase
+in mean TP yield might be suggestive of wetter recent conditions. To
+evaluate long-term trends in discharge, we fit a GAM relating
+log-transformed daily discharges on the Lavaca and Navidad Rivers to
+season and time (Fig.~\ref{fig:fig8}). The long-term trends in discharge
+indicate watershed discharges were at or above average from 1972 though
+the early and mid-1980s. In comparison watershed discharges since the
+mid-2000s are at or below average. The trend in mean streamflow alone is
+counter to expectations based on changes in nutrient yeild estimates.
+Studies of long-term precipitation patterns indicate both the total
+amount \autocite{dixonTrendDetectionTexas2011} and variability of
+rainfall
+\autocite{mishraChangesExtremePrecipitation2010,fagnantCharacterizingSpatiotemporalTrends2020}
+has increased in the region which provide plausible reasons for
+increased runoff driven nutrient loading. Further assessments of the
+frequency and occurrence of periodic high flow events that mobilize
+sediments and are responsible for the bulk of loading volume would
+provide further insight to this discrepancy.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1\linewidth,]{../fig8}
+
+}
+
+\caption{Measured daily discharges (log-transformed) and smoothed long-term trends for the Lavaca and Navidad Rivers form 1972 though 2001.}\label{fig:fig8}
+\end{figure}
+
+\textcite{rebichSourcesDeliveryNutrients2011} and
+\textcite{wise_spatially_2019} used SPARROW to provide a more recent
+assessment of regional catchment based loadings to the Gulf of Mexico
+(Table~\ref{tab:table5}). SPARROW is a hybrid statistical-process model
+with the underlying nutrient load estimation methods based on the
+previously described LOADEST
+\autocite{schwarzSPARROWSurfaceWaterQuality2006}. The functional form of
+the LOADEST regression model is similar to the terms applied in the GAMs
+used in the current study. The only study to apply a mechanistic
+watershed model (SWAT) to estimate nutrient loadings in the Lavaca River
+watershed \textcite{omaniEstimationSedimentNutrient2014} developed
+estimated yields (42 kg/km\textsuperscript{2}) similar to the two
+SPARROW models. The base periods used by
+\textcite{rebichSourcesDeliveryNutrients2011},
+\textcite{omaniEstimationSedimentNutrient2014} and
+\textcite{wise_spatially_2019} differ from the current study, making
+direct comparisons difficult. It is probable that the lower than average
+discharges observed from 2010 through 2021 (Fig.~\ref{fig:fig8}) bias
+our estimates downward compared to other studies which included higher
+than average streamflow periods. Overall, the ranges of estimated yields
+among different studies along with the apparent large variability in
+streamflow driven loadings (Fig.~\ref{fig:fig3}, Fig. \ref{fig:fig4})
+suggest that the current estimates of TP loading are reasonable.
 
@@
 \bottomrule
-\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{a} Mean of the annual point estimates and the lower and upper 95\% credible intervals.}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{a} Mean of the annual point estimates and the lower and upper 90\% credible intervals.}\\
 \multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{b} Represents a binned value range from a choropleth map.}\\
@@
 
-Cross-validation of the GAM loading models indicated that GAMs performed
-well on average at predicting daily nutrient loading values. The
-variance in scores was very high indicating subsets of values were
-problematic at characterizing functional relationships between nutrients
-and predictors. Because all of the water quality data for these two
-locations in the TCEQ databases were ambient water quality data,
-collected to be representative of typical flow conditions, there were
-few data at the highest portions of the flow-duration curve. It was
-beyond the scope of the current study to evaluate the subsets of
-cross-validation data and scores. However, the cross-validation
-procedure is indicative that more robust sampling would be beneficial
-for reducing prediction variance. Supplementary flow-biased monitoring
-targeting storm- or high-flow conditions is recommended here to improve
-the precision of GAM predictions
-\citep{horowitzEvaluationSedimentRating2003, snelderEstimationCatchmentNutrient2017}.
-
-\hypertarget{estuary-water-quality-responses}{%
-\subsection{Estuary Water Quality
-Responses}\label{estuary-water-quality-responses}}
-
-The non-linear temporal water quality trends identified using GAMs
-differed slightly from previously identified trends
-\citep{bugica_water_2020}. This is not unexpected due to the different
-time periods, different methodology, and generally small slopes
-identified for most of the significant water quality parameters in prior
-work. The trend in DO and cholorophyll-\emph{a} concentrations are
-stable in comparison to other Texas estuaries that are facing larger
-demands for freshwater diversions, higher population growth, and more
-intense agricultural production
-\citep{wetzWaterQualityDynamics2016, bugica_water_2020}. The trend of
-increasing NO\textsubscript{\emph{x}} concentration at the upper-bay
-TCEQ-13563 site and recent increases in TKN concentration at the mid-bay
-TCEQ-13383 site are concerning due to the nitrogen limitation identified
-in many Texas estuaries
-\citep{gardnerNitrogenFixationDissimilatory2006, houTransformationFateNitrate2012, doradoUnderstandingInteractionsFreshwater2015, paudelRelationshipSuspendedSolids2019, wetz_exceptionally_2017}
+\hypertarget{estuary-water-quality}{%
+\subsection*{Estuary Water Quality}\label{estuary-water-quality}}
+\addcontentsline{toc}{subsection}{Estuary Water Quality}
+
+The non-linear estuary water quality trends identified in the current
+study differed slightly from previously identified trends
+\autocite{bugica_water_2020}. This is not unexpected due to the
+different time periods, different methodology, and generally small
+slopes identified for most of the significant water quality parameters
+in prior work. Both DO and cholorophyll-\emph{a} concentrations at all
+three Lavaca Bay sites were stable from 2005 through 2020. This is a
+positive outcome in comparison to other Texas estuaries that are facing
+larger demands for freshwater diversions, higher population growth, and
+more intense agricultural production which have resulted in more direct
+signs of eutrophication
+\autocite{wetzWaterQualityDynamics2016,bugica_water_2020}. Despite the
+stability of DO and cholorophyll-\emph{a}, there are concerning site
+specific increases in NO\textsubscript{\emph{X}} and TKN concentration
+over the same time period. These trends are espcially concerning due to
+the nitrogen limitation identified in many Texas estuaries
+\autocite{gardnerNitrogenFixationDissimilatory2006,houTransformationFateNitrate2012,doradoUnderstandingInteractionsFreshwater2015,paudelRelationshipSuspendedSolids2019,wetz_exceptionally_2017}
 and the relatively low ambient concentrations observed in Lavaca Bay.
@@
 The strong positive effect of freshwater inflow on
-NO\textsubscript{\emph{x}}, TKN, and TP are suggestive of nonpoint
-watershed sources, consistent with watershed uses and with other studies
-relating freshwater inflow with nutrient concentrations in Lavaca Bay
-and other estuaries
-\citep{russell_effect_2006, caffreyHighNutrientPulses2007, peierlsNonmonotonicResponsesPhytoplankton2012, palmerImpactsDroughtsLow2015, ciraPhytoplanktonDynamicsLowinflow2021}.
+NO\textsubscript{\emph{X}}, TKN, and TP concentration are suggestive of
+nonpoint watershed sources, consistent with watershed uses and with
+other studies relating freshwater inflow with nutrient concentrations in
+Lavaca Bay and other estuaries
+\autocite{russell_effect_2006,caffreyHighNutrientPulses2007,peierlsNonmonotonicResponsesPhytoplankton2012,palmerImpactsDroughtsLow2015,ciraPhytoplanktonDynamicsLowinflow2021}.
 Inflow had a non-linear relationship with TKN at the two upstream sites,
@@
 inflow. No relationship between TKN and freshwater inflow were observed
-at TCEQ-13384 located in the lower reach of Lavaca Bay. Tidal flushing
-from Matagorda Bay could be responsible for diluting TKN and acting as a
-control on the effects of freshwater inflow in lower reaches of Lavaca
-Bay. Previous work suggests the processing of organic loads in the upper
-portions of Lavaca Bay might reduce the transport of nutrients into the
-lower reaches of the Bay \citep{russell_effect_2006}.
-
-Freshwater inflow had a strong positive effect on chlorophyll-\emph{a}
-at the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed
-decreases in chlorophyll-\emph{a} at the highest freshwater inflow
-volumes. Freshwater flushing or increases in turbidity are associated
-with decreases in chlorophyll-\emph{a} in other estuaries
-\citep{peierlsNonmonotonicResponsesPhytoplankton2012, cloernPhytoplanktonPrimaryProduction2014}.
+at TCEQ-13384 located in the lower reach of Lavaca Bay. The results
+suggest that tidal flushing maybe diluting TKN and act as a control on
+the effects of freshwater inflow in the lower reaches of Lavaca Bay. The
+results are also consistent with previous work that suggest processing
+of organic loads in the upper Lavaca Bay or tidal portions of the Lavaca
+River reduce transport of nutrients to the lower reaches of Lavaca Bay
+\autocite{russell_effect_2006}.
+
+Freshwater inflow also displayed a strong positive effect on
+chlorophyll-\emph{a} at the upper- and mid-bay sites. The upper-bay
+site, TCEQ-13563, showed decreases in chlorophyll-\emph{a} at the
+highest freshwater inflow volumes. Freshwater flushing or increases in
+turbidity are associated with decreases in chlorophyll-\emph{a} in other
+estuaries
+\autocite{peierlsNonmonotonicResponsesPhytoplankton2012,cloernPhytoplanktonPrimaryProduction2014}.
 No relationships between inorganic nitrogen or TP loadings with
@@
 chlorophyll-\emph{a} concentrations in Texas estuaries
-\citep{ornolfsdottirNutrientPulsingRegulator2004, doradoUnderstandingInteractionsFreshwater2015, ciraPhytoplanktonDynamicsLowinflow2021, tominackVariabilityPhytoplanktonBiomass2022},
+\autocite{ornolfsdottirNutrientPulsingRegulator2004,doradoUnderstandingInteractionsFreshwater2015,ciraPhytoplanktonDynamicsLowinflow2021,tominackVariabilityPhytoplanktonBiomass2022},
 this study specifically used flow-adjusted freshwater derived nutrient
@@
 
-GAMs did not identify responses in DO concentration to inflows or
-nutrient loads. The seasonality term in the temporal GAM models
-explained a substantial amount of DO variation at all of the sites.
-Responses of estuary metabolic processes and resulting DO concentrations
-can be quite complicated and often locally specific
-\citep{caffreyFactorsControllingNet2004}. While the lack of total
-nitrogen or TKN loading data hinders interpretation, the large seasonal
-effect on DO suggests physical factors play an important role and should
-be included in future models. Prior work suggests that Lavaca Bay may
-not be limited by nutrients alone, with high turbidity or nutrient
+There was no evidence that adjusted freshwater inflow and nutrient loads
+had effects on DO concentration in Lavaca Bay. The seasonality term in
+the temporal GAM models explained a substantial amount of DO variation
+at all of the sites. Responses of estuary metabolic processes and
+resulting DO concentrations can be quite complicated and often locally
+specific \autocite{caffreyFactorsControllingNet2004}. While the lack of
+total nitrogen or TKN loading data hinders interpretation, the large
+seasonal effect on DO concentration indicates physical factors (such as
+temperature, wind, and turbidity) play an important role and should be
+included in future models. Prior work suggests that Lavaca Bay may not
+be limited by nutrients alone, with high turbidity or nutrient
 processing in upper portions of the Bay or intertidal river limiting
-production \citep{russell_effect_2006}. Finally, it is reasonable to
+production \autocite{russell_effect_2006}. Finally, it is reasonable to
 assume that fluctuations in DO may not occur immediately in response to
-nutrient pulses or freshwater inflow. Work has has shown that various
-water quality parameters may have lagged effects lasting days or even
-months following storms and large discharge events
-\citep{mooneyWatershedExportEvents2012a, wetzExtremeFutureEstuaries2013, bukaveckasInfluenceStormEvents2020, walkerTimescalesMagnitudeWater2021}.
-However, our work only evaluates responses to loading and inflows
-occuring the day of water quality observations.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-GAM models appear to provide reliable estimates of nutrient loads in the
-Lavaca Bay watershed. However, additional flow-biased data collection
-efforts would decrease the prediction variance and improve accuracy at
-critical high flow events. Ongoing projects will fill data gaps for
-total nitrogen and TKN loading. This study, consistent with others along
-the Texas coast, found strong effects of freshwater flow on nutrient and
-chlorophyll-\emph{a} concentrations. DO concentrations, dominated by
-seasonal effects, did not show strong direct responses to freshwater
-flow. Small variance in flow-adjusted nutrient loads indicates that (1)
-there have been limited changes in non-point sources of nutrients and
-(2) that there isn't strong evidence that those small changes have had
-effects on chlorophyll-\emph{a} or dissolved oxygen in Lavaca Bay.
-Although the study did not identify strong responses to changes in
-nutrient loading, this does provide a baseline assessment for future
-water quality management activities in the watershed.
-
-\backmatter
-
-\bmhead{Acknowledgments}
-
-The author extends thanks to Dr.~Mike Wetz (Harte Research Institute,
-Texas A\&M Corpus Christi), Chad Kinsfather and Partick Brzozowski
-(Lavaca-Navidad River Authority), Brian Koch (Texas State Soil and Water
-Conservation Board), Bill Balboa (Matagorda Bay Foundation), Jason
-Pinchbeck (Texas General Land Office) and the Lavaca Bay Foundation for
-supporting development of this project and providing valuable feedback.
-
-\hypertarget{declarations}{%
-\section*{Declarations}\label{declarations}}
-\addcontentsline{toc}{section}{Declarations}
-
-\bmhead{Funding}
+nutrient pulses or freshwater inflow. Work has has shown that many water
+quality parameters may have lagged effects lasting days or even months
+following storms and large discharge events
+\autocite{mooneyWatershedExportEvents2012a,wetzExtremeFutureEstuaries2013,bukaveckasInfluenceStormEvents2020,walkerTimescalesMagnitudeWater2021}.
+However, this study only evaluated responses to loading and inflows
+occurring the day of water quality observations.
+
+Overall, this study suggests that DO and chlorophyll-\emph{a}
+concentrations have been relatively stable in Lavaca Bay. Site-specific
+increases in TKN and NO\textsubscript{\emph{X}} concentrations are a
+cause of concern for increasing risks of eutrophication within Lavaca
+Bay which might be currently attenuated by changes in freshwater flow,
+turbidity, and other physical processes. While loading models indicate
+that there are large annual fluctuations in NO\textsubscript{3} loads,
+these changes have been largely driven environmental conditions (changes
+in runoff and river discharge). These models also provide evidence that
+estuary NO\textsubscript{\emph{X}} and TP concentrations are strongly
+driven by freshwater inflow and to a lesser extent fluctuation in
+flow-adjusted riverine loadings. Site-specific changes in the
+relationships between freshwater inflow and responses in both
+chlorophyll-\emph{a} and TKN concentrations are indicative nutrient
+processing and or tidal flushing effects moving from the river discharge
+point to the mouth of Lavaca Bay. This study does not completely explain
+site specific increases in NO\textsubscript{\emph{X}} and TKN
+concentrations in Lavaca Bay. The freshwater study sites did not
+quantify nutrient loadings from tidal contribution areas or ungauaged
+watersheds. Nutrient contributions from wastewater facility discharges,
+septic systems, and stormwater could be considerable contributors to
+nutrient loadings in Lavaca Bay since they are not processed by a tidal
+river reach prior to entering the Bay. The Garcitas-Arenosa Creek,
+Placedo Creek, and Cox Bay subwatersheds are currently undersampled but
+compose approximately 27\% of the watershed area. The contribution of
+nutrient loadings from these undersampled areas is unknown.
+
+\hypertarget{limitations}{%
+\subsection*{Limitations}\label{limitations}}
+\addcontentsline{toc}{subsection}{Limitations}
+
+The GAM approach proved useful for both estimating loads and assessing
+downstream responses in water quality. Although we did not compare other
+models, it is likely similar estimates of loadings would be obtained by
+methods suggest as LOADEST, WRTDS, or SPARROW given the functionally
+similar dependent variable structures. The underlying weakness in the
+estimates of loading in the current study is the reliance on ambient
+water quality data used for statewide water quality assessments.
+Cross-validation of the nutrient loading models highlights that
+predictions are prone to high bias, owing to the lack of targeted storm
+or flow biased measurements. The high biases are indicative that subsets
+of values were unable to capture the functional relationships with the
+flow based dependent variables. It was beyond the scope of the current
+study to evaluate the subsets of cross-validation data and scores.
+However, the cross-validation procedure is indicative that more robust
+sampling is needed. Supplementary flow-biased monitoring targeting
+storm- or high-flow conditions is critical to improve model performance
+and strength of evidence produced by these models
+\autocite{horowitzEvaluationSedimentRating2003,snelderEstimationCatchmentNutrient2017}.
+Although there is existing work on the samples sizes and sample design
+required for reliable performance of both LOADEST
+\autocite{parkUsePollutantLoad2014} and WRTDS
+\autocite{kumarValueIntensiveSampling2019} models, similar work does not
+appear to have been extended to water quality applications of GAMs.
+
+Due to the concerning increases in eutrophication associated parameters
+in Lavaca Bay and other Texas estuaries \autocite{bugica_water_2020},
+and the desire to quantify linkages between environmental outcomes and
+on the ground management actions \autocite{schrammTotalMaximumDaily2022}
+there is a strong need for reliable estimates of pollutant loadings and
+responses along the Texas coast. Within Texas, state wide water quality
+monitoring programs have focused on collection of ambient condition
+data. A framework for establishing pollutant load monitoring programs
+across catchments that explicitly incorporate flow biased data is needed
+for assessing nutrient loading and estuary health along the Texas coast.
+Additional efforts focused on identifying relevant effect sizes,
+sampling designs, and funding mechanisms that can support long term
+efforts are also needed to adequately design such a framework. Large
+long-term monitoring programs in and around the Chesapeake Bay, San
+Francisco Bay, and along the Mississippi River have proven extremely
+effective at informing management actions and tracking progress towards
+long-term pollutant reduction goals. Similar coordinated efforts across
+Texas coastal watersheds would provide useful for resource management
+efforts intended to protect biological and water quality integrity of
+Texas's estuaries.
+
+\hypertarget{conclusions}{%
+\section*{Conclusions}\label{conclusions}}
+\addcontentsline{toc}{section}{Conclusions}
+
+The primary purpose of this study was to provide estimates of watershed
+nutrient loadings and assess water quality responses to changes in
+nutrient loads. GAMs provided reliable estimates of watershed
+NO\textsubscript{3} and TP loads. However, additional flow-biased data
+collection efforts is needed to decrease the prediction variance and
+improve accuracy at critical high-flow loading events. While some
+ongoing projects will fill data gaps for total nitrogen and TKN loading,
+additional efforts are needed to coordinate data collection efforts
+specifically for load estimation across Texas estuaries. Despite these
+data gaps, this study identified high annual fluctuations in nutrient
+loads driven primarily by discharge. No evidence was identified to
+indicate that on the ground management had changed nutrient loading in
+the Navidad River subwatershed. There was weak evidence for recent
+reductions in flow-normalized NO\textsubscript{3} loading in the Lavaca
+River subwatershed although the results are at odds with flow-normalized
+trends in TP loads.
+
+This study, consistent with others along the Texas coast, found strong
+effects of freshwater flow on nutrient and chlorophyll-\emph{a}
+concentrations. DO concentrations, dominated by seasonal effects, did
+not show strong direct responses to freshwater flow. Small variances in
+flow-adjusted nutrient loads indicate that (1) there have been limited
+changes in non-point sources of nutrients and (2) that there isn't
+strong evidence that those small changes have had effects on
+chlorophyll-\emph{a} or dissolved oxygen in Lavaca Bay. Althogh this
+study did not identify changes in DO or chlorophyll-\emph{a}
+concentrations in Lavaca Bay, site specific increases in
+NO\textsubscript{\emph{X}} and TKN are a cause for water quality
+concern. The study provides a baseline assessment for future water
+quality management activities in the watershed. In order to effectively
+track and link improvements or degradation of water quality conditions
+in Lavaca Bay and other coastal Texas watersheds with on the ground
+effort more robust sampling networks are needed to improve spatial
+coverage of undersampled areas and explicitly incorporate flow-biased
+sampling.
+
+\hypertarget{funding}{%
+\section*{Funding}\label{funding}}
+\addcontentsline{toc}{section}{Funding}
 
@@
 
-\bmhead{Data Availability}
+\hypertarget{data-availability}{%
+\section*{Data Availability}\label{data-availability}}
+\addcontentsline{toc}{section}{Data Availability}
 
@@
 
-\bibliography{bibliography.bib}
+\printbibliography
 

### 2023-05-31 flow-trends.R

@@ -1,0 +1,309 @@
+library(dataRetrieval)
+library(arrow)
+library(tidyverse)
+library(mgcv)
+library(gratia)
+library(twriTemplates)
+library(patchwork)
+library(ragg)
+library(ggrepel)
+## plot discharge trends over time
+
+download_Q_data <- function(usgs_dir = "data/USGS",
+                            twdb_dir = "data/twdb") {
+  ## download or read USGS data
+  if(!file.exists(paste0(usgs_dir, "/usgs_flows.csv"))) {
+    q <- dataRetrieval::readNWISdv(
+      siteNumbers = c("08164000"), 
+      parameterCd = "00060", #discharge
+      startDate = "1972-01-01",
+      endDate = "2020-12-31"
+    )
+    q <- dataRetrieval::renameNWISColumns(q)
+    
+    arrow::write_csv_arrow(q, paste0(usgs_dir,"/usgs_flows.csv"))
+    
+  } else {
+    q <- arrow::read_csv_arrow(paste0(usgs_dir,"/usgs_flows.csv"))
+  }
+  q <- q |> 
+    mutate(site_no = as.character(site_no))
+  ## agency_cd, site_no, Date, Flow, Flow_cd
+  
+  ## read TWDB data
+  
+  # texana is gaged flow and can be used as is.
+  # need to convert from acre-feet per day to cubic feet per second!
+  texana <- read_table("data/TWDB/lktexanag", col_types = "nnnn_") |> 
+    pivot_longer(cols = lktexana_g, names_to = "site_no", values_to = "Flow") |> 
+    mutate(Date = lubridate::ymd(paste0(year,"-", month,"-", day)),
+           agency_cd = "TWDB") |> 
+    select(-c(year, month, day)) |> 
+    #convert to cubic feet per second
+    mutate(Flow = (Flow * (43560/86400)))
+  
+  q |> 
+    bind_rows(texana)
+  
+}
+
+
+
+df <- download_Q_data()
+
+
+df |> 
+  filter(site_no == "8164000") |> 
+  mutate(year = lubridate::year(Date),
+         month = lubridate::month(Date),
+         doy = lubridate::yday(Date),
+         ddate = lubridate::decimal_date(Date),
+         Flow = case_when(
+           Flow == 0 ~ 0.001,
+           .default = as.numeric(Flow)),
+         norm_flow = Flow/mean(Flow)) -> lv_df
+
+hist(lv_df$norm_flow)
+hist(log(lv_df$norm_flow))
+
[…]

### 2023-06-23 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.Rmd
peerj submission files

@@
 This study applied semiparametric approaches to calculate watershed nutrient loads and assess responses in estuary water quality to seasonally-adjusted freshwater inflow and flow-adjusted nutrient loads in Lavaca Bay, Texas.
-Lavaca Bay is a secondary embayment on the Texas coast displaying early potential for eutrophication and water quality degredation.
+Lavaca Bay is a secondary embayment on the Texas coast displaying early potential for eutrophication and water quality degradation.
 Use of flow-adjusted nutrient loads allowed the study to evaluate the response in water quality to changes in nutrient loads driven by non-natural variation.
@@
 Models indicated no evidence of long-term changes in dissolved oxygen or chlorophyll-*a* within Lavaca Bay.
-However, site specific long-term increases in both organic and inorganic nitrogen are causes of concern as potential causes of future eutrophication.
-Further analysis found freshwater inflow was a strong driver of nutrient and chlorophyll-*a* concentrations within Lavaca Bay but there was no evidence that changes in watershed nutrient loading explained additional variation in dissolved oxygen or chlorophyll-*a* concentrations.
-In addition to providing a baseline assessment of watershed nutrient loading and water quality responses in the Lavaca Bay watershed, this study provides methodological support for the use of semiparametric methods in load regression models and estuary assessments.
+However, site specific long-term increases in both organic and inorganic nitrogen are concerning for their potential to fuel eutrophication.
+Further analysis found freshwater inflow had strong influences on nutrient and chlorophyll-*a* concentrations but there was no evidence that changes in watershed nutrient loading explained additional variation in dissolved oxygen or chlorophyll-*a* concentrations.
+In addition to providing a baseline assessment of watershed nutrient loading and water quality responses in the Lavaca Bay watershed, this study provides methodological support for the use of semiparametric models in load regression models and estuary assessments.
 bibliography: bibliography.bib
@@
                       fig.align = "center",
-                      out.width = "100%")
+                      out.width = "100%",
+                      fig.process = 
+                        function(filename) {
+                          new_filename <- stringr::str_remove(string = filename,
+                                                              pattern = "-1")
+                          fs::file_move(path = filename, new_path = new_filename)
+                          ifelse(fs::file_exists(new_filename), new_filename, filename)
+                        }
+                      )
 
@@
 
-Similar to many coastal areas globally, the coastal watersheds along the Texas Gulf coast 
-are facing pressures from increasing population, increases in point source 
+Similar to many coastal areas globally, the coastal watersheds along the Texas 
+Gulf coast are facing pressures from growing populations, increases in point source 
 and non-point source pollution and alterations to freshwater flows that degrade
-water quality in downstream estuaries [@bricker_effects_2008; @kennicuttWaterQualityGulf2017; @bugica_water_2020]. 
+water quality in downstream estuaries [@bricker_effects_2008; 
+@kennicuttWaterQualityGulf2017; @bugica_water_2020]. 
 Despite these escalating pressures, national scale assessments have classified 
@@
 [@bricker_effects_2008]. However, a suite of recent studies indicates that 
-estuary water quality dynamics in both agriculturally dominated and urban 
+estuary water quality dynamics in both agricultural and urban dominated
 watersheds within Texas are expressing conditions that are increasingly 
@@
 @wetz_exceptionally_2017; @bugica_water_2020; @chinPhytoplanktonBiomassCommunity2022]. 
-With identification of several localized areas of estuary water quality concern along 
-the Texas coast [@bugica_water_2020], localized studies are being prioritized 
+With identification of several localized areas of estuary water quality concern 
+along the Texas coast [@bugica_water_2020], localized studies are being prioritized 
 to better inform management actions.
 
-This goal of this project is to assess watershed nutrient loading and the 
+The goal of this project is to assess watershed nutrient loading and the 
 resulting water quality responses in Lavaca Bay, Texas. Lavaca Bay is a secondary bay in the larger
@@
 (DO), nutrients, and biological parameters have been well within state
-water quality standards. Despite largely meeting statewide water quality standards,
-there have been concerning declines in abundance, biomass, and diversity of 
-benthic fauna in Lavaca Bay [@beserespollackLongtermTrendsResponse2011]. These declines
-are partially attributed to reductions in freshwater inflow and changes in estuary salinity
-and are indicative of an already stressed system 
+water quality standards. Despite largely meeting state designated water quality 
+thresholds, there have been concerning declines in abundance, biomass, and 
+diversity of benthic fauna in Lavaca Bay [@beserespollackLongtermTrendsResponse2011]. 
+These declines are partially attributed to reductions in freshwater inflow and 
+changes in estuary salinity and are indicative of an already stressed system 
 [@beserespollackLongtermTrendsResponse2011; @palmerImpactsDroughtsLow2015;
@@
 total Kjeldahl nitrogen (TKN), and chlorophyll-*a* concentrations were identified
-at monitoring sites within Lavaca Bay [@bugica_water_2020]. Although 
-long-term changes in DO concentrations have not been identified, the trends in 
+at monitoring sites within Lavaca Bay [@bugica_water_2020]. These trends in 
 nutrient concentrations are concerning due to the role of nitrogen as a 
@@
 @schrammLavacaRiverWatershed2018; @bertholdDirectMailingEducation2021]. 
-However, at a statewide scale, these approaches have shown limited success and 
-emphasize a need for improved efforts at assessing and linking management 
-actions with downstream water quality to identify and replicate effective 
-management actions across the state [@schrammTotalMaximumDaily2022]. The 
-identification and communication of changes and trends in water quality is 
+However, on a statewide scale, these approaches have shown limited success and 
+emphasize a need for improved methods to assess and link management 
+actions with downstream water quality [@schrammTotalMaximumDaily2022].  
+Such methods could help identify and replicate effective water quality 
+management actions across the state . The identification and communication of 
+changes and trends in water quality is 
 complicated by the fact that trends are often non-linear and confounded by 
@@
 (daily, monthly, or annually) and flow-normalized estimates of concentrations and loads
-using locally weighted regresssion for unique combinations of time, discharge, and season. 
+using locally weighted regression for unique combinations of time, discharge, and season. 
 WRTDS has been widely used to assess and identify trends in riverine nutrients 
@@
 chlorides [@stetsIncreasingChlorideRivers2018], and other pollutants of concern 
-[@shodaWaterqualityTrendsRivers2019]. WRTDS has also been succesfully adapted to 
-assess trends in estuarine water quality concentrations [@beckFourDecadesWater2018].
+[@shodaWaterqualityTrendsRivers2019]. WRTDS has also been successfully adapted to 
+assess trends in estuary water quality concentrations [@beckFourDecadesWater2018].
 
@@
 differences (and similarities) between GAMs and WRTDS for water quality 
-applications reders are referred to @beckNumericalQualitativeContrasts2017.
+applications readers are referred to @beckNumericalQualitativeContrasts2017.
 
@@
 
-knitr::include_graphics("../fig1.png")
+knitr::include_graphics("Schramm-2023-06-PeerJ_files/figure-latex/fig1.png")
 
@@
 are required to be collected under Quality Assurance Project Plans and lab 
-method procedures outlined by the TCEQ's procedures manual to ensure
-consistent collection and laboratory methods are applied between samples 
-collected by different entities and under different projects. 
-All sites had varying lengths of and availability of data.
+method procedures outlined by the TCEQ's procedures manual. These operating 
+procedures ensure consistent collection and laboratory methods are applied 
+between samples collected by different entities and under different projects. 
+All sites had varying time periods and availability of water quality data.
 For freshwater locations, TP from January 2000 through December 2020 and 
@@
 \begin{align}
-g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log1p(Q)) + f_4(ma) + f_5(fa), \nonumber \\
+g(\mu) &= \alpha + f_1(ddate) + f_2(yday) + f_3(log(1+Q)) + f_4(ma) + f_5(fa), \nonumber \\
 y &\sim \mathcal{N}(\mu,\,\sigma^{2})
@@
 with mean $\mu$ and standard deviation $\sigma$. *ddate* is the date converted to 
-decimal notation, *yday* is numeric day of year (1-366), and *log1p(Q)* is the 
-natural log of mean daily streamflow plus 1.
+decimal notation, *yday* is numeric day of year (1-366), and *Q* is the 
+mean daily streamflow.
 
@@
 GAM performance was evaluated 
-using repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
-and average Nash-Sutcliffe Efficiency (NSE), Pearson sample correlation (*r*) 
-and percent bias (PBIAS) metrics across folds were calculated for each model.
+with repeated 5-fold cross validation  [@burmanComparativeStudyOrdinary1989] 
+using average Nash-Sutcliffe Efficiency (NSE), Pearson sample correlation (*r*) 
+and percent bias (PBIAS) metrics.
 
@@
 Because streamflow and nutrient loads are tightly correlated, freshwater inflow
-can mask signals due to chnages in nutrient loads alone. Following the 
+can mask signals due to changes in nutrient loads alone. Following the 
 methodology implemented by @murphyNutrientImprovementsChesapeake2022,
@@
 Median goodness-of-fit metrics for TP in the Lavaca River were 0.81 NSE, 
-0.93 *r*, and -7.20 PBIAS. Navidad River TP models has similar performance 
+0.93 *r*, and -7.20 PBIAS. Navidad River TP models had similar performance 
 with 0.91 NSE, 0.99 *r*, and -3.30 PBIAS. Density plots of
@@
 ```{r fig2, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.65, fig.cap="Density plots of goodness-of-fit metrics (NSE, \\textit{r}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
-Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics."}
+Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics.", dev='cairo_pdf'}
 
@@
 
-```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525."}
+```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.", dev='cairo_pdf'}
 
@@
 Aggregated across both sites, the mean 
-annual NO~3~ load 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
+annual NO~3~ load from 2005 through 2020 was `r mean_no3` kg (`r mean_no3_low` kg - 
 `r mean_no3_upper` kg, 90% CI). Annual NO~3~ loads ranged from `r min_no3` kg 
@@
 
-```{r fig4, error=FALSE, message=FALSE, warning=FALSE, fig.width=3.3, fig.asp=2, fig.cap="Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers.", out.width="0.5\\linewidth"}
+```{r fig4, error=FALSE, message=FALSE, warning=FALSE, fig.width=3.3, fig.asp=2, fig.cap="Comparison of delivered annual loads and annual discharge at the Lavaca (USGS-08164000) and Navidad (USGS-08164525) Rivers.", out.width="0.5\\linewidth", dev='cairo_pdf'}
 
@@
 
-```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
+```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.", dev='cairo_pdf'}
 
@@
 
-```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay."}
+```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.", dev='cairo_pdf'}
 
@@
 
-```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.4, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\\textsubscript{3} loads on TP (A) and NO\\textsubscript{\\textit{X}} (B) concentrations at each site in Lavaca Bay."}
+```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.44, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\\textsubscript{3} loads on TP (A) and NO\\textsubscript{\\textit{X}} (B) concentrations at each site in Lavaca Bay.",  dev='cairo_pdf'}
 
@@
 
-knitr::include_graphics("../fig8.png")
+knitr::include_graphics("Schramm-2023-06-PeerJ_files/figure-latex/fig8.png")
 
@@
 DO and cholorophyll-*a*, there are concerning site specific increases in 
-NO~*X*~ and TKN concentration over the same time period. These trends are espcially
+NO~*X*~ and TKN concentration over the same time period. These trends are especially
 concerning due to the nitrogen limitation identified in many Texas 
@@
 freshwater inflow were observed at TCEQ-13384 located in the lower reach of 
-Lavaca Bay. The results suggest that tidal flushing maybe diluting TKN and act as 
+Lavaca Bay. The results suggest that tidal flushing might dilute TKN and act as 
 a control on the effects of freshwater inflow in the lower reaches of Lavaca Bay.
@@
 there is a strong need for reliable estimates of pollutant loadings and 
-responses along the Texas coast. Within Texas, state wide water quality monitoring programs have focused on 
+responses along the Texas coast. Within Texas, statewide water quality 
+monitoring programs have focused on 
 collection of ambient condition data. A framework for establishing pollutant 
 load monitoring programs across catchments that explicitly incorporate
-flow biased data is needed for assessing nutrient loading and estuary health along the Texas coast. 
+flow biased data is needed for assessing nutrient loading and estuary health 
+along the Texas coast.
+
 Additional efforts focused on identifying relevant effect sizes, sampling 
@@
 goals. Similar coordinated efforts across Texas coastal watersheds would provide
-useful for resource management efforts intended to protect biological and water
-quality integrity of Texas's estuaries.
+useful for resource management efforts intended to protect the biological and 
+water quality integrity of Texas's estuaries.
 
@@
 GAMs provided reliable estimates of watershed NO~3~ and TP loads.
-However, additional flow-biased data collection efforts is needed to
+However, additional flow-biased data collection efforts are needed to
 decrease the prediction variance and improve accuracy at critical 
@@
 nutrient loads indicate that (1) there have been limited changes in non-point
-sources of nutrients and (2) that there isn't strong evidence that those small 
+sources of nutrients and (2) that there is not strong evidence that those small 
 changes have had effects on chlorophyll-*a* or dissolved oxygen in Lavaca Bay.
-Althogh this study did not identify changes in DO or chlorophyll-*a* concentrations
+Although this study did not identify changes in DO or chlorophyll-*a* concentrations
 in Lavaca Bay, site specific increases in NO~*X*~ and TKN are a cause for water 
@@
 improvements or degradation of water quality conditions in Lavaca Bay and other coastal
-Texas watersheds with on the ground effort more robust sampling networks are needed
+Texas watersheds with on the ground efforts, more robust sampling networks are needed
 to improve spatial coverage of undersampled areas and explicitly incorporate
 flow-biased sampling.
+
+# Acknowledgements
+The author thanks Stephanie deVilleneuve for comments on an early draft and
+members of the Coastal Nutrient Input Repository project committee for
+supporting project development and their valuable input.
 

### 2023-06-23 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.tex
peerj submission files

@@
 Lavaca Bay is a secondary embayment on the Texas coast displaying early
-potential for eutrophication and water quality degredation. Use of
+potential for eutrophication and water quality degradation. Use of
 flow-adjusted nutrient loads allowed the study to evaluate the response
@@
 oxygen or chlorophyll-\emph{a} within Lavaca Bay. However, site specific
-long-term increases in both organic and inorganic nitrogen are causes of
-concern as potential causes of future eutrophication. Further analysis
-found freshwater inflow was a strong driver of nutrient and
-chlorophyll-\emph{a} concentrations within Lavaca Bay but there was no
-evidence that changes in watershed nutrient loading explained additional
-variation in dissolved oxygen or chlorophyll-\emph{a} concentrations. In
-addition to providing a baseline assessment of watershed nutrient
-loading and water quality responses in the Lavaca Bay watershed, this
-study provides methodological support for the use of semiparametric
-methods in load regression models and estuary assessments.
+long-term increases in both organic and inorganic nitrogen are
+concerning for their potential to fuel eutrophication. Further analysis
+found freshwater inflow had strong influences on nutrient and
+chlorophyll-\emph{a} concentrations but there was no evidence that
+changes in watershed nutrient loading explained additional variation in
+dissolved oxygen or chlorophyll-\emph{a} concentrations. In addition to
+providing a baseline assessment of watershed nutrient loading and water
+quality responses in the Lavaca Bay watershed, this study provides
+methodological support for the use of semiparametric models in load
+regression models and estuary assessments.
 % Dummy abstract text.
@@
 Similar to many coastal areas globally, the coastal watersheds along the
-Texas Gulf coast are facing pressures from increasing population,
+Texas Gulf coast are facing pressures from growing populations,
 increases in point source and non-point source pollution and alterations
@@
 of recent studies indicates that estuary water quality dynamics in both
-agriculturally dominated and urban watersheds within Texas are
-expressing conditions that are increasingly conducive to algal blooms
-and eutrophication
+agricultural and urban dominated watersheds within Texas are expressing
+conditions that are increasingly conducive to algal blooms and
+eutrophication
 \autocite{wetzWaterQualityDynamics2016,wetz_exceptionally_2017,bugica_water_2020,chinPhytoplanktonBiomassCommunity2022}.
@@
 
-This goal of this project is to assess watershed nutrient loading and
-the resulting water quality responses in Lavaca Bay, Texas. Lavaca Bay
-is a secondary bay in the larger Matagorda Bay system located roughly
-halfway between Houston, Texas and Corpus Christi, Texas. Lavaca Bay
-faces substantial challenges associated with legacy contamination but
-general water quality parameters such as dissolved oxygen (DO),
-nutrients, and biological parameters have been well within state water
-quality standards. Despite largely meeting statewide water quality
-standards, there have been concerning declines in abundance, biomass,
+The goal of this project is to assess watershed nutrient loading and the
+resulting water quality responses in Lavaca Bay, Texas. Lavaca Bay is a
+secondary bay in the larger Matagorda Bay system located roughly halfway
+between Houston, Texas and Corpus Christi, Texas. Lavaca Bay faces
+substantial challenges associated with legacy contamination but general
+water quality parameters such as dissolved oxygen (DO), nutrients, and
+biological parameters have been well within state water quality
+standards. Despite largely meeting state designated water quality
+thresholds, there have been concerning declines in abundance, biomass,
 and diversity of benthic fauna in Lavaca Bay
@@
 concentrations were identified at monitoring sites within Lavaca Bay
-\autocite{bugica_water_2020}. Although long-term changes in DO
-concentrations have not been identified, the trends in nutrient
-concentrations are concerning due to the role of nitrogen as a limiting
-factor for primary production in many Texas estuaries
+\autocite{bugica_water_2020}. These trends in nutrient concentrations
+are concerning due to the role of nitrogen as a limiting factor for
+primary production in many Texas estuaries
 \autocite{gardnerNitrogenFixationDissimilatory2006,houTransformationFateNitrate2012,doradoUnderstandingInteractionsFreshwater2015,paudelRelationshipSuspendedSolids2019,wetz_exceptionally_2017}
@@
 \autocite{jainTechnicalSupportDocument2021,schrammLavacaRiverWatershed2018,bertholdDirectMailingEducation2021}.
-However, at a statewide scale, these approaches have shown limited
-success and emphasize a need for improved efforts at assessing and
-linking management actions with downstream water quality to identify and
-replicate effective management actions across the state
-\autocite{schrammTotalMaximumDaily2022}. The identification and
+However, on a statewide scale, these approaches have shown limited
+success and emphasize a need for improved methods to assess and link
+management actions with downstream water quality
+\autocite{schrammTotalMaximumDaily2022}.\\
+Such methods could help identify and replicate effective water quality
+management actions across the state . The identification and
 communication of changes and trends in water quality is complicated by
@@
 (daily, monthly, or annually) and flow-normalized estimates of
-concentrations and loads using locally weighted regresssion for unique
+concentrations and loads using locally weighted regression for unique
 combinations of time, discharge, and season. WRTDS has been widely used
@@
 pollutants of concern \autocite{shodaWaterqualityTrendsRivers2019}.
-WRTDS has also been succesfully adapted to assess trends in estuarine
+WRTDS has also been successfully adapted to assess trends in estuary
 water quality concentrations \autocite{beckFourDecadesWater2018}.
@@
 discussion on the differences (and similarities) between GAMs and WRTDS
-for water quality applications reders are referred to
+for water quality applications readers are referred to
 \textcite{beckNumericalQualitativeContrasts2017}.
@@
 
-{\centering \includegraphics[width=5.2in,]{../fig1}
+{\centering \includegraphics[width=5.2in,]{Schramm-2023-06-PeerJ_files/figure-latex/fig1}
 
@@
 Project Plans and lab method procedures outlined by the TCEQ's
-procedures manual to ensure consistent collection and laboratory methods
-are applied between samples collected by different entities and under
-different projects. All sites had varying lengths of and availability of
-data. For freshwater locations, TP from January 2000 through December
-2020 and nitrate-nitrogen (NO\textsubscript{3}) data from January 2005
-through December 2020 were downloaded (Table~\ref{tab:table1}). Less
-than 5-years of total nitrogen and TKN concentration data were available
-at the freshwater sites and deemed insufficient to develop load
-estimation models
+procedures manual. These operating procedures ensure consistent
+collection and laboratory methods are applied between samples collected
+by different entities and under different projects. All sites had
+varying time periods and availability of water quality data. For
+freshwater locations, TP from January 2000 through December 2020 and
+nitrate-nitrogen (NO\textsubscript{3}) data from January 2005 through
+December 2020 were downloaded (Table~\ref{tab:table1}). Less than
+5-years of total nitrogen and TKN concentration data were available at
+the freshwater sites and deemed insufficient to develop load estimation
+models
 \autocite{horowitzEvaluationSedimentRating2003,snelderEstimationCatchmentNutrient2017}.
@@
 \emph{ddate} is the date converted to decimal notation, \emph{yday} is
-numeric day of year (1-366), and \emph{log1p(Q)} is the natural log of
-mean daily streamflow plus 1.
+numeric day of year (1-366), and \emph{Q} is the mean daily streamflow.
 
@@
 \autocite{woodConfidenceIntervalsGeneralized2006,marraCoveragePropertiesConfidence2012,mcdowell_implications_2021}.
-GAM performance was evaluated using repeated 5-fold cross validation
-\autocite{burmanComparativeStudyOrdinary1989} and average Nash-Sutcliffe
-Efficiency (NSE), Pearson sample correlation (\emph{r}) and percent bias
-(PBIAS) metrics across folds were calculated for each model.
+GAM performance was evaluated with repeated 5-fold cross validation
+\autocite{burmanComparativeStudyOrdinary1989} using average
+Nash-Sutcliffe Efficiency (NSE), Pearson sample correlation (\emph{r})
+and percent bias (PBIAS) metrics.
 
@@
 Because streamflow and nutrient loads are tightly correlated, freshwater
-inflow can mask signals due to chnages in nutrient loads alone.
+inflow can mask signals due to changes in nutrient loads alone.
 Following the methodology implemented by
@@
 in the Lavaca River were 0.81 NSE, 0.93 \emph{r}, and -7.20 PBIAS.
-Navidad River TP models has similar performance with 0.91 NSE, 0.99
+Navidad River TP models had similar performance with 0.91 NSE, 0.99
 \emph{r}, and -3.30 PBIAS. Density plots of metrics show similar
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig2-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig2}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig3-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig3}
 
@@
 Aggregated across both sites, the mean annual NO\textsubscript{3} load
-2005 through 2020 was 205,405 kg (126,867 kg - 341,569 kg, 90\% CI).
-Annual NO\textsubscript{3} loads ranged from 12,574 kg in 2011 to
+from 2005 through 2020 was 205,405 kg (126,867 kg - 341,569 kg, 90\%
+CI). Annual NO\textsubscript{3} loads ranged from 12,574 kg in 2011 to
 794,510 kg in 2007. Total annual TP loads ranged from 7,839 kg in 2011
@@
 
-{\centering \includegraphics[width=0.5\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig4-1}
+{\centering \includegraphics[width=0.5\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig4}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig5-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig5}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig6-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig6}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig7-1}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig7}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{../fig8}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig8}
 
@@
 specific increases in NO\textsubscript{\emph{X}} and TKN concentration
-over the same time period. These trends are espcially concerning due to
+over the same time period. These trends are especially concerning due to
 the nitrogen limitation identified in many Texas estuaries
@@
 at TCEQ-13384 located in the lower reach of Lavaca Bay. The results
-suggest that tidal flushing maybe diluting TKN and act as a control on
-the effects of freshwater inflow in the lower reaches of Lavaca Bay. The
+suggest that tidal flushing might dilute TKN and act as a control on the
+effects of freshwater inflow in the lower reaches of Lavaca Bay. The
 results are also consistent with previous work that suggest processing
@@
 there is a strong need for reliable estimates of pollutant loadings and
-responses along the Texas coast. Within Texas, state wide water quality
+responses along the Texas coast. Within Texas, statewide water quality
 monitoring programs have focused on collection of ambient condition
@@
 for assessing nutrient loading and estuary health along the Texas coast.
+
 Additional efforts focused on identifying relevant effect sizes,
@@
 Texas coastal watersheds would provide useful for resource management
-efforts intended to protect biological and water quality integrity of
-Texas's estuaries.
+efforts intended to protect the biological and water quality integrity
+of Texas's estuaries.
 
@@
 NO\textsubscript{3} and TP loads. However, additional flow-biased data
-collection efforts is needed to decrease the prediction variance and
+collection efforts are needed to decrease the prediction variance and
 improve accuracy at critical high-flow loading events. While some
@@
 flow-adjusted nutrient loads indicate that (1) there have been limited
-changes in non-point sources of nutrients and (2) that there isn't
+changes in non-point sources of nutrients and (2) that there is not
 strong evidence that those small changes have had effects on
-chlorophyll-\emph{a} or dissolved oxygen in Lavaca Bay. Althogh this
+chlorophyll-\emph{a} or dissolved oxygen in Lavaca Bay. Although this
 study did not identify changes in DO or chlorophyll-\emph{a}
@@
 in Lavaca Bay and other coastal Texas watersheds with on the ground
-effort more robust sampling networks are needed to improve spatial
+efforts, more robust sampling networks are needed to improve spatial
 coverage of undersampled areas and explicitly incorporate flow-biased
 sampling.
+
+\hypertarget{acknowledgements}{%
+\section{Acknowledgements}\label{acknowledgements}}
+
+The author thanks Stephanie deVilleneuve for comments on an early draft
+and members of the Coastal Nutrient Input Repository project committee
+for supporting project development and their valuable input.
 

### 2023-06-23 flow-trends.R

@@ -265,7 +265,8 @@ p1 <- bind_rows(lv_preds, lt_preds) |>
                 y = log(Flow),
                 color = "Measured Discharge",
                 alpha = "Measured Discharge",
-                linetype = "Measured Discharge")) +
+                linetype = "Measured Discharge"),
+            linewidth = 0.25) +
   geom_ribbon(aes(x = ddate,
                   ymin = ci_low,
                   ymax = ci_hi),
@@ -273,12 +274,14 @@ p1 <- bind_rows(lv_preds, lt_preds) |>
   geom_line(aes(x = ddate, y = preds,
                 color = "Smoothed Long-term Trend",
                 alpha = "Smoothed Long-term Trend",
-                linetype = "Smoothed Long-term Trend")) +
+                linetype = "Smoothed Long-term Trend"),
+            linewidth = 0.25) +
   geom_hline(data = intercepts_df,
              aes(yintercept = intercept,
                  color = "Intercept",
                  alpha = "Intercept",
-                 linetype = "Intercept")) +
+                 linetype = "Intercept"),
+             linewidth = 0.25) +
   scale_color_manual(name = "", values = c("black", "steelblue", "black")) +
   scale_alpha_manual(name = "", values = c(0.5, 0.2, 1)) +
   scale_linetype_manual(name = "", values = c(2, 1, 1)) +
@@ -293,17 +296,25 @@ p1 <- bind_rows(lv_preds, lt_preds) |>
         plot.subtitle = element_text(size = 8),
         legend.text = element_text(size = 8),
         legend.key.width = unit(25, "points"),
-        strip.background = element_rect(fill = "white",
-                                        color = "black",
-                                        linewidth = .5),
+        strip.background = element_rect(fill = "white", color = NULL),
         strip.text = element_text(size = 8))
 
+p1
+# ragg::agg_png(filename = "fig8.png",
+#               width = 6.85,
+#               height = 6.85*0.65,
+#               units = "in",
+#               res = 600)
+# Cairo::CairoPDF(file = "fig8.pdf",
+#                 width = 6.85*0.65,
+#                 height = 3,
+#                 pointsize = 8,
+#                 family = "Arial")
+cairo_pdf(file = "fig8.pdf",
+    width = 6.85,
+    height = 6.85*0.65,
+    family = "Arial")
 
-ragg::agg_png(filename = "fig8.png",
-              width = 6.85,
-              height = 6.85*0.65,
-              units = "in",
-              res = 600)
 p1
 
 dev.off()

### 2023-06-23 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.Rmd
final formatting

@@
 
-# Acknowledgements
+# Acknowledgements {.unnumbered}
 The author thanks Stephanie deVilleneuve for comments on an early draft and
@@
 supporting project development and their valuable input.
-
-# Funding {.unnumbered}
-This project was funded in part by a Texas Coastal Management Program grant 
-approved by the Texas Land Commissioner, providing financial assistance under 
-the Coastal Zone Management Act of 1972, as amended, awarded by the National 
-Oceanic and Atmospheric Administration (NOAA), Office for Coastal Management,
-pursuant to NOAA Award No. NA21NOS4190136. The views expressed herein are those 
-of the author(s) and do not necessarily reflect the views of NOAA, the U.S. 
-Department of Commerce, or any of their subagencies.
-
-# Data Availability {.unnumbered}
-Reproducible code and datasets generated during this study are available in
-the Zenodo repository, https://doi.org/10.5281/zenodo.733075.

### 2023-06-23 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.tex
final formatting

@@
 \hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section*{Acknowledgements}\label{acknowledgements}}
+\addcontentsline{toc}{section}{Acknowledgements}
 
@@
 
-\hypertarget{funding}{%
-\section*{Funding}\label{funding}}
-\addcontentsline{toc}{section}{Funding}
-
-This project was funded in part by a Texas Coastal Management Program
-grant approved by the Texas Land Commissioner, providing financial
-assistance under the Coastal Zone Management Act of 1972, as amended,
-awarded by the National Oceanic and Atmospheric Administration (NOAA),
-Office for Coastal Management, pursuant to NOAA Award
-No.~NA21NOS4190136. The views expressed herein are those of the
-author(s) and do not necessarily reflect the views of NOAA, the U.S.
-Department of Commerce, or any of their subagencies.
-
-\hypertarget{data-availability}{%
-\section*{Data Availability}\label{data-availability}}
-\addcontentsline{toc}{section}{Data Availability}
-
-Reproducible code and datasets generated during this study are available
-in the Zenodo repository, \url{https://doi.org/10.5281/zenodo.733075}.
-
 \printbibliography

### 2023-08-07 Rebuttal_Letter/Rebuttal_Letter.Rmd
start working on revisions

---
author: Michael Schramm
return-address: 
  - Texas A&M AgriLife
  - Texas Water Resources Institute
  - 1001 Holleman Drive E.
  - College Station, Texas 77840
return-short: Michael Schramm, 2118 TAMU, College Station, TX 77840
address:
  - Dr. James Latimer
  - Academic Editor
  - PeerJ Life & Environment

subject: Revised Manuscript
opening: Dear Dr. Latimer,
closing: Best regards,
signature: Michael Schramm

output: komaletter::komaletter
---

Thank you for the prompt and thorough review of my submitted manuscript, "Linking watershed nutrient loading to estuary water quality with semiparametric models." I really appreciate the substantive and helpful comments from all the reviewers. Line by line comments and responses are included below (reviewer comments are italicized).

**Reviewer 1**

*Very well-structured abstract – introduced the problem, the objectives of the research, the results obtained, and the conclusion. No changes are recommended.*
*The introduction was well-structured and easy to follow. No changes are recommended.*

**Response:** Noted, and thank you.

*Line 114: A link to the USGS NWIS system could be added.*

**Response:** The url has been added.

*A brief description of the sample collection and laboratory analyses described in the QAPP and TCEQ procedures manual, detection and recovery limits, and instrumentation would seem appropriate.*

**Response:** I added the EPA methods and TCEQ's required AWRL's to Tables 1 and 2. I believe this is sufficient documentation.

*Also, what are the state-recommended guideline values? Include them in Table 2*

**Response:** I've added the associated state water quality criteria or screening levels in both Tables 1 and 2.

*Line 371 “has” repeated twice*

**Response:** Thank you, this is fixed.

*The author adequately answered the research questions and fulfilled all the stated objectives. The results from the GAM model provided estimates of watershed nutrient loadings and an outlook of nutrient loading's influence on water quality.*
*The whole of the paper was well structured, easy to read, and contributes sufficiently to the existing body of knowledge on estuary anthropogenic pollution and water quality monitoring.*
*Through this paper, the author has created an addition to the body of work on using models for the prediction and assessment of pollution in estuaries and aquatic bodies, as a means of recommending and establishing policy actions for pollution mitigation.*

**Response:** Thank you for the review.

**Reviewer 3**

*Table 1: It would be helpful if you would include the watershed or station name under the USGS station ID in this table.*

**Response:** I've added the associated watersheds to each USGS station in Table 1.

*Line 164: Not sure if this is a typo, or if I’m missing what log1p(Q) is.*

**Response:** This was shorthand for *log(1+Q)*. I've changed this to *log(1+Q)* to be consistent with equation 1.

*Line 392, "Limitation" section: I expected this to be a discussion of limitations of the approach like matching daily loads to concentrations, the complex environmental factors that could be involved in chlorophyll and DO trends, changing climate, etc. However, since you focus on the (very valid) problem of limited sampling, you may want to rename this sub-section of the Discussion. Maybe call it "Data Needs" or "Suggested Management Focus" or something like that?*

*Line 53-54: Since only one of the trends you mention from Bugica is for a nitrogen species, I suggest rephrasing or expanding this sentence to explain why we’d be concerned about all of the parameters that are increasing (or clarify if we are only concerned about the TKN).*

*Lines 129-131: It would be helpful if you described how deep the water column is at the estuary sample locations and from what depth the samples that you analyzed were taken. A little more explanation of the estuary dynamics would be helpful too -- is this a stratified or well-mixed system, what is the range in salinity, is there a key season of concern for eutrophication?*

**Response:** I have added additional description of the estuary dynamics *(lines 101-108)*. Specifically mixing, turbidity, and salinity. Given the preliminary nature of this study, there is not a season of concern identified yet, so it is not included. *still need to add sample depth here*

*Lines 194-201: Some discussion is needed of how you dealt with temporal matching of flow and loads to concentrations in the estuary. I see in the discussion (line 373-374) that you matched flow and load estimated for the day of each sample. This approach should be explained in the methods, and I think you need some explanation is needed as to why that would be appropriate and/or how it may impact the results. As you discuss below, you might get better relationships (for Table 4, Lines 244-248) if you either lagged the match or matched an average load and flow from some reasonable period before the estuary sample date. If you had the ability to try some models like that, I think it may be a useful addition.*

*Line 296: Would damming of the Navidad River impact the loads through the Lavaca River watershed? I don’t see why if the data was from the USGS-08164000 site.*

**Response:** You are absolutely correct. I mistakenly wrote this section considering the
combined watershed loads, when these estimates were for the Lavaca River watershed only.
I've removed this and incorporated your statement below. Thank you very much for catching this!

*Lines 286-321: I think this discussion is too long comparing the TP yield first to the Dunn study, and then to 3 other studies. The TP yield estimate from your study is in the middle of the others, so you can condense this 2 paragraph discussion and simply say that you fall in the middle of other estimates and state briefly that there are reasons such as land use and flow variability that the estimates vary.*

**Response:** I agree, this section has been considerably condensed.

*Lines 360-361: It is probably worthwhile to note that the effect of flow you see in chlorophyll is very likely due to more nutrients coming in with high flows. I wouldn’t want a reader to think that since your flow-adjusted nutrient loads are not linked to chlorophyll that this means there is definitely no impact of nutrients on chlorophyll concentrations. Along those lines, I’d suggest in Line 350-251 you specific that “No relationships between flow-adjusted inorganic nitrogen or TP loadings with chlorophyll-a were observed”*

**Reviewer: Marcus Beck**

*I have some concerns about the methods starting on line 184 that describe how the load estimates are related to the estuarine data. First, I strongly encourage the authors to develop these models using cumulative or lagged estimates for the flow and load variables. It appears that these are currently modeled as the estimates for the same day the water quality parameter was sampled. I would not expect much of a relationship between water quality and the independent variables at that time scale. I suspect that the relationships the author shows regarding flow/load with the response reflects some level of temporal autocorrelation between the daily estimates and preceding conditions/loads, where the latter may be a truly causative relationship. I see two potential approaches to remedy the issue. One, the author could test different lagged values for flow and load, although this would require a bit of trial and error. A potentially better solution is to use the aggregate flow volume and load between the sample dates as a sum of the volume and mass, respectively, between each observation of the time series for the response variable. I would expect these aggregate measurements to be much more strongly correlated to the response variables than the daily values that are currently used. This limitation is noted on lines 371-374, so I’m curious why the authors did not use this approach.*

*My second concern about the models in eqns. 5-7 also relates to the flow and load variables. It’s stated on line 199-201 that flow-adjusted load values were used, which makes sense, but I don’t understand how this additional analysis differs from the flow-normalization in the previous section (lines 176 – 179). Why not just use the flow-adjusted loads from the previous models? Once this is clarified (or corrected), it might make more sense to change eqn 6 to include smoothers for ddate, yday, and flow-normalized load, and eqn 7 to included smoothers for ddate, yday, and load. These wouldn’t be “nested” models for AIC comparison, but would perhaps provide a better approach to evaluate loads related to natural or non-natural variation. I don’t see the value in including flow as a separate variable since it’s basically a surrogate for total load.*

*I also strongly encourage the author to consider replacing “semiparametric models” with “Generalized Additive Models” in the title. It’s a more accurate description of what’s included in the paper.*

*Line 18: Maybe replace “non-natural variation” with “anthropogenic sources”.*

**Response:** Suggested changed made.

*Line 61: Accidental line break. *

**Response:** Fixed.

*Line 77: Suggest rephrasing as “…generalized linear models where the response variable is modeled as the sum of…”*

**Response:** Suggested change made.

*Tables 1 and 2: It would be useful to indicate the approximate sampling interval for the water quality parameters. It looks like quarterly. *

*Line 102: Some more information about the Bay would be helpful. What is the average depth? Any information on residence time? Is the water column mixed, stratified, or seasonally-stratified?*

**Response:** I have not identified any information on residence time. I have added more information on average depth, mixing and stratification of the water column. **add line numbers**

*Lines 128 – 129: The interpretation of the figures, tables, and text would be much easier if the estuarine sites were simply referred to as “upper”, “mid”, and “lower”. I suggest replacing all station names with these terms throughout the manuscript. *

*Line 129 – 131: Some additional information about the sampling methods for the estuarine sites could be useful. Where in the water column were these samples taken? Is the water column mixed? I’m thinking specifically about dissolved oxygen. The results show that it was not related to load, but you may have different outcomes depending on surface vs bottom DO. *

*Line 138: Typo “tend”.*

**Response** Fixed typo.

*Figure 1: Remove “contribution” from the legend. *

**Response:** Suggested change made.

*Eqn 2: I think some additional details are needed in the text below this equation. Why was d set at 0.95? Based on the text, this seems awfully close to 1 or the “mean flow”. Also, the equation includes a lag term, t – t_j, but the size of this lag is not stated, i.e., “some historical day of observation”. Was this set explicitly by the author, and if so, is there a justification? *

*Line 151: It’s not clear to me how fa is a unitless term. The equations show these as simple differences of average streamflow, so wouldn’t the units be in log-cfs? *

*Line 169: What percentage of the observations for each parameter were non-detects? I realize that setting the values as ½ the limit can introduce bias, but this may not be a big deal if there aren’t many non-detects. There does appear to be a GAM package that handles non-detects with Tobit models, but it is quite old (by software standards). It may be worth exploring. https://cran.r-project.org/web/packages/cenGAM/index.html*

*Line 236: This might be a personal preference, but “unusual periodic pattern” seems like regular periodicity and not a more random periodicity as one would expect with rain events. Consider rephrasing this text. *

*Figure 2 caption: Consider adding a sentence like “Values closer to 1 for NSE and r and values closer to 0 for PBIAS represent more ideal goodness-of-fit assessments.”*

*Figure 3 caption: Consider adding Lavaca and Navidad with the station name, as in the caption for figure 4. *

*Figure 5 caption: Consider including a reference to Table 3, as it wasn’t immediately clear to me that these models didn’t include flow or load. *

*Table 4: As for my previous comment, I’d replace the site names with “upper”, “mid”, and “lower” and arrange them accordingly. The current arrangement is “mid”, “lower”, and “upper”. Also, it looks like several of the statistical summaries are identical, e.g. the inflow and inflow+load model for TP at TCEQ-13384. Are these simply rounding issues? And if so, are these models really all that different in terms of explained variance? *

*Line 272: The text “contributing tributaries” seems redundant. Maybe remove “contributing”. *

*Line 341: Replace “were” with “was”. *

*Line 395: Replace “suggest” with “such as”.*

*Line 420: Replace “provide” with “prove” or “be”.*

### 2023-08-07 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.Rmd
start working on revisions

--- 2023-06-23 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.Rmd
@@
 Lavaca Bay is a secondary embayment on the Texas coast displaying early potential for eutrophication and water quality degradation.
-Use of flow-adjusted nutrient loads allowed the study to evaluate the response in water quality to changes in nutrient loads driven by non-natural variation.
+Use of flow-adjusted nutrient loads allowed the study to evaluate the response in water quality to changes in nutrient loads driven by anthropogenic sources.
 Cross-validation indicated that, despite data constraints, semiparametric models performed well at nutrient load prediction.
@@
     - \usepackage{csquotes}
+    - \usepackage{booktabs}
+    - \usepackage{longtable}
+    - \usepackage{array}
+    - \usepackage{multirow}
+    - \usepackage{wrapfig}
+    - \usepackage{float}
+    - \usepackage{colortbl}
+    - \usepackage{pdflscape}
+    - \usepackage{tabu}
+    - \usepackage{threeparttable}
+    - \usepackage{threeparttablex}
+    - \usepackage[normalem]{ulem}
+    - \usepackage{makecell}
 
@@
 
+options(knitr.kable.NA = '')
 store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
@@
 emphasize a need for improved methods to assess and link management 
-actions with downstream water quality [@schrammTotalMaximumDaily2022].  
-Such methods could help identify and replicate effective water quality 
+actions with downstream water quality [@schrammTotalMaximumDaily2022]. Such 
+methods could help identify and replicate effective water quality 
 management actions across the state . The identification and communication of 
@@
 GAMs are a broadly applicable statistical method. GAMs are a semiparametric 
-extension of generalized linear models where the linear predictor is represented
+extension of generalized linear models where the response variable is modeled
 as the sum of multiple unknown smooth functions and parametric linear predictors
@@
 Lavaca Bay is a 190 km^2^ estuary with the majority of freshwater inflow provided by the 
-Lavaca and Navidad River systems (Fig.\ \ref{fig:fig1}). The Garcitas-Arenosa, 
-Placedo Creek, and Cox Bay subwatersheds provide additional freshwater inflows. 
+Lavaca and Navidad River systems (Fig.\ \ref{fig:fig1}). Lavaca Bay is relatively
+shallow with an average depth of 1.2 m and a generally well mixed and turbid water column 
+[@beserespollackLongtermTrendsResponse2011; @montagnaAssessmentRelationshipFreshwater2020].
+The Garcitas-Arenosa, Placedo Creek, and Cox Bay subwatersheds provide additional 
+freshwater inflows. The proximity to freshwater inputs in the upper Lavaca Bay
+results in a strong mean salinity and variance gradient 
+[@montagnaAssessmentRelationshipFreshwater2020]. Average salinity ranges from 
+12 psu near the Lavaca River mouth to 23 psu at the mouth of Lavaca Bay. The
+variance is inversely correlated with mean salinity.
 The entire watershed area is 8,149 km^2^ and primarily rural. Watershed 
@@
 
-```{r fig1, fig.width=6.75, fig.width=5.2, out.width="5.2in", fig.cap="Map of Lavaca Bay and the contribution watershed.
+```{r fig1, fig.width=6.75, fig.width=5.2, out.width="5.2in", fig.cap="Map of Lavaca Bay watershed.
 The freshwater sites are the most downstream freshwater stream locations with water quality and streamflow data used for nutrient load models.
@@
 
-knitr::include_graphics("Schramm-2023-06-PeerJ_files/figure-latex/fig1.png")
+knitr::include_graphics("Schramm-2023-08-PeerJ_files/figure-latex/fig1.png")
 
@@
 were obtained from the United States Geologic Survey (USGS) National Water 
-Information System using the *dataRetrieval* R package 
+Information System (https://waterdata.usgs.gov/nwis) using the *dataRetrieval* R package 
 [@deciccoDataRetrievalPackagesDiscovering2022]. Gaged daily discharges from 
@@
             data = df, output = "data.frame") |> 
-  mutate(`Station ID` = c("USGS-08164000", "USGS-08164000", "USGS-08164000",
-                          "USGS-08164525", "USGS-08164525", "USGS-08164525")) |> 
-  rename("Parameter" = " ") |> 
+  mutate(`Station ID` = c("USGS-08164000, Lavaca River", "USGS-08164000, Lavaca River", "USGS-08164000, Lavaca River",
+                          "USGS-08164525, Navidad River", "USGS-08164525, Navidad River", "USGS-08164525, Navidad River")) |> 
+  rename("Parameter" = " ",
+         "Station ID and Watershed" = "Station ID") |> 
+  mutate(Method = c("EPA 365.1", "EPA 353.2", "", "EPA 365.1", "EPA 353.2", ""),
+         `AWRL\\textsuperscript{a}`= c(0.06, 0.05, NA, 0.06, 0.05, NA),
+         `Criteria\\textsuperscript{b}` = c(0.69,1.95,NA,0.2,0.37,NA)) |> 
   kbl(format = "latex", booktabs = TRUE, escape = FALSE,
       caption = "Summary of gauged streamflow and freshwater water quality samples between January 1, 2000 and December 31, 2020.") |>
-  collapse_rows(columns = 1, latex_hline = "major")
+  column_spec(1:2, width = "2.5cm") |>
+  collapse_rows(columns = 1, latex_hline = "major") |> 
+  footnote(alphabet = c("Ambient water reporting limit (AWRL) represents the highest concentration that can be used as a reporting limit by a laboratory for inclusion in the state's monitoring data.",
+                        "The state of Texas has not adopted numeric criteria for nutrients, these criteria represent screening levels used by the state for assessment purposes."),
+           threeparttable = TRUE)
 
@@
   rename("Parameter" = " ") |> 
+  mutate(Method = rep(c("EPA 365.1", "EPA 353.2", "EPA 351.2", "EPA 445.0", "ASTM D888-09(C) and TCEQ SOP V1"),3),
+         `AWRL\\textsuperscript{a}`= rep(c(0.06, 0.05, 0.2, 3, NA),3),
+         `Criteria\\textsuperscript{b}` = rep(c(0.21,NA,NA,11.6,5),3)) |> 
   kbl(format = "latex", booktabs = TRUE, escape = FALSE,
       caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.") |> 
-  collapse_rows(columns = 1, latex_hline = "major")
+  column_spec(6, width = "3cm") |> 
+  collapse_rows(columns = 1, latex_hline = "major") |> 
+  footnote(alphabet = c("Ambient water reporting limit (AWRL) represents the highest concentration that can be used as a reporting limit by a laboratory for inclusion in the state's monitoring data.",
+                        "The state of Texas has not adopted numeric criteria for nutrients, these criteria represent screening levels used by the state for assessment purposes.
+NO\\\\textsubscript{x} and TKN parameters are not currently screened by the state and do not have an associated screening level."),
+           threeparttable = T,
+           escape = F)
 
@@
 The general form of the model related NO~3~ or TP concentration to a long term
-tend, season, streamflow, and two different antecedent discharge terms, 
+trend, season, streamflow, and two different antecedent discharge terms, 
 shown in Eq. \ref{eq1}:
@@
 the 4 major tributaries to the lake. Thin-plate regression splines were used 
-for *ddate*, *log1p(Q)*, *fa*, and *ma*. A cyclic cubic regression spline was 
+for *ddate*, *log(1+Q)*, *fa*, and *ma*. A cyclic cubic regression spline was 
 used for *yday* to ensure the ends of the spline match 
@@
 to long term, seasonal, and flow based predictors and includes methods for bias correction
-when exponentiating the response variable. While the annual yield estimated in 
-@dunnTrendsNutrientInflows1996 is lower than the estimate of 35.2 kg/km^2^
-in this study, it is important to note that both hydrology and land use was
-different for a proportion of the study period. For almost the first ten years of the study period
-in @dunnTrendsNutrientInflows1996, the Navidad River was still undammed. More recently
-the eastern portion of the watershed has seen large conversions of rice fields to
-more traditional row crop production [@kulatHolisticWaterEnergyFoodNexus2019]. 
-The apparent increase in mean TP yield might be suggestive of wetter recent conditions.
-To evaluate long-term trends in discharge, we fit a GAM relating log-transformed daily
-discharges on the Lavaca and Navidad Rivers to season and time (Fig.\ \ref{fig:fig8}).
-The long-term trends in discharge indicate watershed discharges were at or above average
-from 1972 though the early and mid-1980s. In comparison watershed discharges
-since the mid-2000s are at or below average. The trend in mean streamflow alone
-is counter to expectations based on changes in nutrient yeild estimates. Studies of
-long-term precipitation patterns indicate both the total amount [@dixonTrendDetectionTexas2011] 
-and variability of rainfall [@mishraChangesExtremePrecipitation2010;
-@fagnantCharacterizingSpatiotemporalTrends2020] has increased in the region 
-which provide plausible reasons for increased runoff driven nutrient loading.
-Further assessments of the frequency and occurrence of periodic high flow events 
-that mobilize sediments and are responsible for the bulk of loading volume would
-provide further insight to this discrepancy.
-
-```{r fig8, fig.width=6.85, fig.asp=0.75, fig.cap="Measured daily discharges (log-transformed) and smoothed long-term trends for the Lavaca and Navidad Rivers form 1972 though 2001."}
-
-knitr::include_graphics("Schramm-2023-06-PeerJ_files/figure-latex/fig8.png")
-
-```
-
-@rebichSourcesDeliveryNutrients2011 and @wise_spatially_2019 used SPARROW to provide
-a more recent assessment of regional catchment based loadings to the Gulf of Mexico
-(Table\ \ref{tab:table5}).
+when exponentiating the response variable. @rebichSourcesDeliveryNutrients2011 
+and @wise_spatially_2019 used SPARROW to provide a more recent assessment of 
+regional catchment based loadings to the Gulf of Mexico (Table\ \ref{tab:table5}).
 SPARROW is a hybrid statistical-process model with the underlying nutrient
@@
 @omaniEstimationSedimentNutrient2014 developed estimated yields (42 kg/km^2^) 
-similar to the two SPARROW models. The base periods used by 
-@rebichSourcesDeliveryNutrients2011, @omaniEstimationSedimentNutrient2014 and
-@wise_spatially_2019 differ from the current study, making direct comparisons 
-difficult. It is probable that the lower than average discharges observed from 2010 
-through 2021 (Fig.\ \ref{fig:fig8}) bias our estimates downward compared to other studies which included
-higher than average streamflow periods. Overall, the ranges of estimated yields
-among different studies along with the apparent large variability in 
-streamflow driven loadings (Fig.\ \ref{fig:fig3}, Fig. \ref{fig:fig4}) suggest that the current estimates
+similar to the two SPARROW models. Although direct comparisons are complicated
+by varying time periods, the estimates in this study do fall within the range of
+of previously developed estimates. 
+To evaluate changes in long-term trends in discharge might be associated with
+the nutrient yield estimates covering different time periods, 
+we fit a GAM relating log-transformed daily discharges on the Lavaca River to 
+season and time (Fig.\ \ref{fig:fig8}).
+The long-term trends in discharge indicate watershed discharges were at or above average
+from 1972 though the early and mid-1980s. In comparison watershed discharges
+since the mid-2000s are at or below average. 
+It is probable that the lower than average discharges observed from 2010 
+through 2021 (Fig.\ \ref{fig:fig8}) bias our estimates downward compared to studies
+that included higher than average streamflow periods (1995-2005). Overall, the 
+ranges of estimated yields among different studies along with the apparent 
+large variability in streamflow driven loadings (Fig.\ \ref{fig:fig3}, 
+Fig. \ref{fig:fig4}) suggest that the current estimates
 of TP loading are reasonable.
+
+```{r fig8, out.width="70%", fig.cap="Measured daily discharges (log-transformed) and smoothed long-term trends for the Lavaca River form 1972 though 2001."}
+
+knitr::include_graphics("Schramm-2023-08-PeerJ_files/figure-latex/fig8.pdf")
+
+```
 
@@
 is reasonable to assume that fluctuations in DO may not occur immediately in 
-response to nutrient pulses or freshwater inflow. Work has has shown that many
+response to nutrient pulses or freshwater inflow. Work has shown that many
 water quality parameters may have lagged effects lasting days or even months following

### 2023-08-07 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.tex
start working on revisions

--- 2023-06-23 Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.tex
@@
 \usepackage{makecell}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
 \usepackage{xcolor}
@@
 flow-adjusted nutrient loads allowed the study to evaluate the response
-in water quality to changes in nutrient loads driven by non-natural
-variation. Cross-validation indicated that, despite data constraints,
+in water quality to changes in nutrient loads driven by anthropogenic
+sources. Cross-validation indicated that, despite data constraints,
 semiparametric models performed well at nutrient load prediction. Based
@@
 management actions with downstream water quality
-\autocite{schrammTotalMaximumDaily2022}.\\
-Such methods could help identify and replicate effective water quality
-management actions across the state . The identification and
-communication of changes and trends in water quality is complicated by
-the fact that trends are often non-linear and confounded by
-precipitation and runoff that hinder traditional analysis
+\autocite{schrammTotalMaximumDaily2022}. Such methods could help
+identify and replicate effective water quality management actions across
+the state . The identification and communication of changes and trends
+in water quality is complicated by the fact that trends are often
+non-linear and confounded by precipitation and runoff that hinder
+traditional analysis
 \autocite{wazniakLinkingWaterQuality2007,lloydMethodsDetectingChange2014}.
@@
 GAMs are a semiparametric extension of generalized linear models where
-the linear predictor is represented as the sum of multiple unknown
-smooth functions and parametric linear predictors
+the response variable is modeled as the sum of multiple unknown smooth
+functions and parametric linear predictors
 \autocite{woodFastStableRestricted2011}. Although the underlying
@@
 freshwater inflow provided by the Lavaca and Navidad River systems
-(Fig.~\ref{fig:fig1}). The Garcitas-Arenosa, Placedo Creek, and Cox Bay
-subwatersheds provide additional freshwater inflows. The entire
-watershed area is 8,149 km\textsuperscript{2} and primarily rural.
-Watershed land cover and land use is 50\% grazed pasture and rangeland,
-20\% cultivated cropland (primarily rows crops such as corn, cotton, and
-sorghum), and 5\% suburban/urban. Pasture and rangeland is concentrated
-in the Lavaca River watershed, while cultivated crops are generally
-located along the eastern tributaries of the Navidad river. The Lavaca
-and Navidad River watersheds are a combined 5,966 km\textsuperscript{2},
-or approximately 73\% of the entire Lavaca Bay watershed area. Discharge
-from the Navidad River is regulated by Lake Texana which has been in
-operation since 1980. Lake Texana provides 0.210 km\textsuperscript{3}
-of water storage and discharges into the tidal section of the Navidad
-River which ultimately joins the tidal section of the Lavaca River 15 km
-upstream of the confluence with the Lavaca Bay.
+(Fig.~\ref{fig:fig1}). Lavaca Bay is relatively shallow with an average
+depth of 1.2 m and a generally well mixed and turbid water column
+\autocite{beserespollackLongtermTrendsResponse2011,montagnaAssessmentRelationshipFreshwater2020}.
+The Garcitas-Arenosa, Placedo Creek, and Cox Bay subwatersheds provide
+additional freshwater inflows. The proximity to freshwater inputs in the
+upper Lavaca Bay results in a strong mean salinity and variance gradient
+\autocite{montagnaAssessmentRelationshipFreshwater2020}. Average
+salinity ranges from 12 psu near the Lavaca River mouth to 23 psu at the
+mouth of Lavaca Bay. The variance is inversely correlated with mean
+salinity. The entire watershed area is 8,149 km\textsuperscript{2} and
+primarily rural. Watershed land cover and land use is 50\% grazed
+pasture and rangeland, 20\% cultivated cropland (primarily rows crops
+such as corn, cotton, and sorghum), and 5\% suburban/urban. Pasture and
+rangeland is concentrated in the Lavaca River watershed, while
+cultivated crops are generally located along the eastern tributaries of
+the Navidad river. The Lavaca and Navidad River watersheds are a
+combined 5,966 km\textsuperscript{2}, or approximately 73\% of the
+entire Lavaca Bay watershed area. Discharge from the Navidad River is
+regulated by Lake Texana which has been in operation since 1980. Lake
+Texana provides 0.210 km\textsuperscript{3} of water storage and
+discharges into the tidal section of the Navidad River which ultimately
+joins the tidal section of the Lavaca River 15 km upstream of the
+confluence with the Lavaca Bay.
 
@@
 
-{\centering \includegraphics[width=5.2in,]{Schramm-2023-06-PeerJ_files/figure-latex/fig1}
+{\centering \includegraphics[width=5.2in,]{Schramm-2023-08-PeerJ_files/figure-latex/fig1}
 
@@
 
-\caption{Map of Lavaca Bay and the contribution watershed.
+\caption{Map of Lavaca Bay watershed.
 The freshwater sites are the most downstream freshwater stream locations with water quality and streamflow data used for nutrient load models.
@@
 Fig.~\ref{fig:fig1}) were obtained from the United States Geologic
-Survey (USGS) National Water Information System using the
-\emph{dataRetrieval} R package
-\autocite{deciccoDataRetrievalPackagesDiscovering2022}. Gaged daily
-discharges from the outlet of Lake Texana on the Navidad River
+Survey (USGS) National Water Information System
+(\url{https://waterdata.usgs.gov/nwis}) using the \emph{dataRetrieval} R
+package \autocite{deciccoDataRetrievalPackagesDiscovering2022}. Gaged
+daily discharges from the outlet of Lake Texana on the Navidad River
 (USGS-0816425) were provided by the Texas Water Development Board (TWDB)
@@
 \centering
-\begin{tabular}[t]{lllll}
+\begin{threeparttable}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2.5cm}llllrr}
 \toprule
-Station ID & Parameter & Mean & SD & N\\
+Station ID and Watershed & Parameter & Mean & SD & N & Method & AWRL\textsuperscript{a} & Criteria\textsuperscript{b}\\
 \midrule
- & TP (mg/L) & 0.21 & 0.09 & 80\\
-
- & NO\textsubscript{3} (mg/L) & 0.18 & 0.24 & 74\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash USGS-08164000} & Mean Daily Streamflow (cfs) & 332.78 & 1667.47 & 7671\\
-\cmidrule{1-5}
- & TP (mg/L) & 0.20 & 0.08 & 81\\
-
- & NO\textsubscript{3} (mg/L) & 0.29 & 0.26 & 62\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash USGS-08164525} & Mean Daily Streamflow (cfs) & 666.14 & 2957.79 & 7671\\
+ & TP (mg/L) & 0.21 & 0.09 & 80 & EPA 365.1 & 0.06 & 0.69\\
+
+ & NO\textsubscript{3} (mg/L) & 0.18 & 0.24 & 74 & EPA 353.2 & 0.05 & 1.95\\
+
+\multirow{-3}{2.5cm}{\raggedright\arraybackslash USGS-08164000, Lavaca River} & Mean Daily Streamflow (cfs) & 332.78 & 1667.47 & 7671 &  &  & \\
+\cmidrule{1-8}
+ & TP (mg/L) & 0.20 & 0.08 & 81 & EPA 365.1 & 0.06 & 0.20\\
+
+ & NO\textsubscript{3} (mg/L) & 0.29 & 0.26 & 62 & EPA 353.2 & 0.05 & 0.37\\
+
+\multirow{-3}{2.5cm}{\raggedright\arraybackslash USGS-08164525, Navidad River} & Mean Daily Streamflow (cfs) & 666.14 & 2957.79 & 7671 &  &  & \\
 \bottomrule
 \end{tabular}
+\begin{tablenotes}
+\item[a] Ambient water reporting limit (AWRL) represents the highest concentration that can be used as a reporting limit by a laboratory for inclusion in the state's monitoring data.
+\item[b] The state of Texas has not adopted numeric criteria for nutrients, these criteria represent screening levels used by the state for assessment purposes.
+\end{tablenotes}
+\end{threeparttable}
 \end{table}
@@
 \centering
-\begin{tabular}[t]{lllll}
+\begin{threeparttable}
+\begin{tabular}[t]{lllll>{\raggedright\arraybackslash}p{3cm}rr}
 \toprule
-Station ID & Parameter & Mean & SD & N\\
+Station ID & Parameter & Mean & SD & N & Method & AWRL\textsuperscript{a} & Criteria\textsuperscript{b}\\
 \midrule
- & TP (mg/L) & 0.11 & 0.05 & 47\\
-
- & NO\textsubscript{x} (mg/L) & 0.07 & 0.15 & 51\\
-
- & TKN (mg/L) & 0.94 & 0.49 & 45\\
-
- & Chlorophyll-\textit{a} (ug/L) & 9.43 & 5.31 & 47\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13383} & DO (mg/L) & 7.22 & 1.35 & 55\\
-\cmidrule{1-5}
- & TP (mg/L) & 0.08 & 0.03 & 51\\
-
- & NO\textsubscript{x} (mg/L) & 0.06 & 0.08 & 52\\
-
- & TKN (mg/L) & 0.76 & 0.40 & 48\\
-
- & Chlorophyll-\textit{a} (ug/L) & 8.22 & 6.44 & 46\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13384} & DO (mg/L) & 7.51 & 1.32 & 54\\
-\cmidrule{1-5}
- & TP (mg/L) & 0.13 & 0.06 & 50\\
-
- & NO\textsubscript{x} (mg/L) & 0.09 & 0.13 & 53\\
-
- & TKN (mg/L) & 0.94 & 0.37 & 49\\
-
- & Chlorophyll-\textit{a} (ug/L) & 9.67 & 5.33 & 49\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13563} & DO (mg/L) & 7.91 & 1.34 & 56\\
+ & TP (mg/L) & 0.11 & 0.05 & 47 & EPA 365.1 & 0.06 & 0.21\\
+
+ & NO\textsubscript{x} (mg/L) & 0.07 & 0.15 & 51 & EPA 353.2 & 0.05 & \\
+
+ & TKN (mg/L) & 0.94 & 0.49 & 45 & EPA 351.2 & 0.20 & \\
+
+ & Chlorophyll-\textit{a} (ug/L) & 9.43 & 5.31 & 47 & EPA 445.0 & 3.00 & 11.60\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13383} & DO (mg/L) & 7.22 & 1.35 & 55 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
+\cmidrule{1-8}
+ & TP (mg/L) & 0.08 & 0.03 & 51 & EPA 365.1 & 0.06 & 0.21\\
+
+ & NO\textsubscript{x} (mg/L) & 0.06 & 0.08 & 52 & EPA 353.2 & 0.05 & \\
+
+ & TKN (mg/L) & 0.76 & 0.40 & 48 & EPA 351.2 & 0.20 & \\
+
+ & Chlorophyll-\textit{a} (ug/L) & 8.22 & 6.44 & 46 & EPA 445.0 & 3.00 & 11.60\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13384} & DO (mg/L) & 7.51 & 1.32 & 54 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
+\cmidrule{1-8}
+ & TP (mg/L) & 0.13 & 0.06 & 50 & EPA 365.1 & 0.06 & 0.21\\
+
+ & NO\textsubscript{x} (mg/L) & 0.09 & 0.13 & 53 & EPA 353.2 & 0.05 & \\
+
+ & TKN (mg/L) & 0.94 & 0.37 & 49 & EPA 351.2 & 0.20 & \\
+
+ & Chlorophyll-\textit{a} (ug/L) & 9.67 & 5.33 & 49 & EPA 445.0 & 3.00 & 11.60\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13563} & DO (mg/L) & 7.91 & 1.34 & 56 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
 \bottomrule
 \end{tabular}
+\begin{tablenotes}
+\item[a] Ambient water reporting limit (AWRL) represents the highest concentration that can be used as a reporting limit by a laboratory for inclusion in the state's monitoring data.
+\item[b] The state of Texas has not adopted numeric criteria for nutrients, these criteria represent screening levels used by the state for assessment purposes.
+NO\textsubscript{x} and TKN parameters are not currently screened by the state and do not have an associated screening level.
+\end{tablenotes}
+\end{threeparttable}
 \end{table}
@@
 \autocite{woodFastStableRestricted2011}. The general form of the model
-related NO\textsubscript{3} or TP concentration to a long term tend,
+related NO\textsubscript{3} or TP concentration to a long term trend,
 season, streamflow, and two different antecedent discharge terms, shown
@@
 the lake. Thin-plate regression splines were used for \emph{ddate},
-\emph{log1p(Q)}, \emph{fa}, and \emph{ma}. A cyclic cubic regression
+\emph{log(1+Q)}, \emph{fa}, and \emph{ma}. A cyclic cubic regression
 spline was used for \emph{yday} to ensure the ends of the spline match
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig2}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-08-PeerJ_files/figure-latex/fig2}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig3}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-08-PeerJ_files/figure-latex/fig3}
 
@@
 
-{\centering \includegraphics[width=0.5\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig4}
+{\centering \includegraphics[width=0.5\linewidth,]{Schramm-2023-08-PeerJ_files/figure-latex/fig4}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig5}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-08-PeerJ_files/figure-latex/fig5}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig6}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-08-PeerJ_files/figure-latex/fig6}
 
@@
 
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig7}
+{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-08-PeerJ_files/figure-latex/fig7}
 
@@
 predictors and includes methods for bias correction when exponentiating
-the response variable. While the annual yield estimated in
-\textcite{dunnTrendsNutrientInflows1996} is lower than the estimate of
-35.2 kg/km\textsuperscript{2} in this study, it is important to note
-that both hydrology and land use was different for a proportion of the
-study period. For almost the first ten years of the study period in
-\textcite{dunnTrendsNutrientInflows1996}, the Navidad River was still
-undammed. More recently the eastern portion of the watershed has seen
-large conversions of rice fields to more traditional row crop production
-\autocite{kulatHolisticWaterEnergyFoodNexus2019}. The apparent increase
-in mean TP yield might be suggestive of wetter recent conditions. To
-evaluate long-term trends in discharge, we fit a GAM relating
-log-transformed daily discharges on the Lavaca and Navidad Rivers to
-season and time (Fig.~\ref{fig:fig8}). The long-term trends in discharge
-indicate watershed discharges were at or above average from 1972 though
-the early and mid-1980s. In comparison watershed discharges since the
-mid-2000s are at or below average. The trend in mean streamflow alone is
-counter to expectations based on changes in nutrient yeild estimates.
-Studies of long-term precipitation patterns indicate both the total
-amount \autocite{dixonTrendDetectionTexas2011} and variability of
-rainfall
-\autocite{mishraChangesExtremePrecipitation2010,fagnantCharacterizingSpatiotemporalTrends2020}
-has increased in the region which provide plausible reasons for
-increased runoff driven nutrient loading. Further assessments of the
-frequency and occurrence of periodic high flow events that mobilize
-sediments and are responsible for the bulk of loading volume would
-provide further insight to this discrepancy.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth,]{Schramm-2023-06-PeerJ_files/figure-latex/fig8}
-
-}
-
-\caption{Measured daily discharges (log-transformed) and smoothed long-term trends for the Lavaca and Navidad Rivers form 1972 though 2001.}\label{fig:fig8}
-\end{figure}
-
-\textcite{rebichSourcesDeliveryNutrients2011} and
+the response variable. \textcite{rebichSourcesDeliveryNutrients2011} and
 \textcite{wise_spatially_2019} used SPARROW to provide a more recent
@@
 estimated yields (42 kg/km\textsuperscript{2}) similar to the two
-SPARROW models. The base periods used by
-\textcite{rebichSourcesDeliveryNutrients2011},
-\textcite{omaniEstimationSedimentNutrient2014} and
-\textcite{wise_spatially_2019} differ from the current study, making
-direct comparisons difficult. It is probable that the lower than average
+SPARROW models. Although direct comparisons are complicated by varying
+time periods, the estimates in this study do fall within the range of of
+previously developed estimates. To evaluate changes in long-term trends
+in discharge might be associated with the nutrient yield estimates
+covering different time periods, we fit a GAM relating log-transformed
+daily discharges on the Lavaca River to season and time
+(Fig.~\ref{fig:fig8}). The long-term trends in discharge indicate
+watershed discharges were at or above average from 1972 though the early
+and mid-1980s. In comparison watershed discharges since the mid-2000s
+are at or below average. It is probable that the lower than average
 discharges observed from 2010 through 2021 (Fig.~\ref{fig:fig8}) bias
-our estimates downward compared to other studies which included higher
-than average streamflow periods. Overall, the ranges of estimated yields
-among different studies along with the apparent large variability in
-streamflow driven loadings (Fig.~\ref{fig:fig3}, Fig. \ref{fig:fig4})
+our estimates downward compared to studies that included higher than
+average streamflow periods (1995-2005). Overall, the ranges of estimated
+yields among different studies along with the apparent large variability
+in streamflow driven loadings (Fig.~\ref{fig:fig3}, Fig. \ref{fig:fig4})
 suggest that the current estimates of TP loading are reasonable.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth,]{Schramm-2023-08-PeerJ_files/figure-latex/fig8}
+
+}
+
+\caption{Measured daily discharges (log-transformed) and smoothed long-term trends for the Lavaca River form 1972 though 2001.}\label{fig:fig8}
+\end{figure}
 
@@
 assume that fluctuations in DO may not occur immediately in response to
-nutrient pulses or freshwater inflow. Work has has shown that many water
+nutrient pulses or freshwater inflow. Work has shown that many water
 quality parameters may have lagged effects lasting days or even months

### 2023-08-07 Schramm-2023-06-PeerJ/diff.R

@@ -1,0 +1,17 @@
+library(latexdiffr)
+library(fs)
+
+
+file_1 <- fs::path_wd("Schramm-2023-06-PeerJ/Schramm-2023-06-PeerJ.tex")
+file_2 <- fs::path_wd("Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.tex")
+
+latexdiff(file_1, file_2, output = "Schramm-2023-06-PeerJ/diff", 
+          compile = FALSE
+          )
+# tinytex::latexmk(fs::path_wd("Schramm-2023-06-PeerJ/diff.tex"), 
+#                  engine = "pdflatex",
+#                  bib_engine = "biber",
+#                  pdf_file = fs::path_wd("Schramm-2023-06-PeerJ/diff.pdf"))
+
+system2(command = "latexmk",
+        args = c("-pdf", "-cd Schramm-2023-06-PeerJ/diff.tex"))

### 2023-08-07 flow-trends.R

@@ -153,6 +153,54 @@ p1 <- ggplot() +
 
 
 
+p1 <- ggplot(lv_preds) +
+  geom_line(aes(x = ddate,
+                y = log(Flow),
+                color = "Measured Discharge",
+                alpha = "Measured Discharge",
+                linetype = "Measured Discharge"),
+            linewidth = 0.5) +
+  geom_ribbon(aes(x = ddate,
+                  ymin = ci_low,
+                  ymax = ci_hi),
+              alpha = 0.25) +
+  geom_line(aes(x = ddate, y = preds,
+                color = "Smoothed Long-term Trend",
+                alpha = "Smoothed Long-term Trend",
+                linetype = "Smoothed Long-term Trend"),
+            linewidth = 0.5) +
+  geom_hline(data = as.data.frame(lv_int),
+             aes(yintercept = lv_int,
+                 color = "Intercept",
+                 alpha = "Intercept",
+                 linetype = "Intercept"),
+             linewidth = 0.5) +
+  scale_color_manual(name = "", values = c("black", "steelblue", "black")) +
+  scale_alpha_manual(name = "", values = c(0.5, 0.2, 1)) +
+  scale_linetype_manual(name = "", values = c(2, 1, 1)) +
+  labs(x = "", y = "log(Discharge) [cfs]") +
+  theme_TWRI_print(base_family = "Arial") +
+  theme(axis.title.y = element_text(size = 11),
+        axis.title.x = element_text(size = 11),
+        axis.text.x = element_text(size = 9),
+        axis.text.y = element_text(size = 9),
+        panel.grid = element_blank(),
+        plot.subtitle = element_text(size = 14),
+        legend.text = element_text(size = 7),
+        legend.key.width = unit(25, "points"),
+        strip.background = element_rect(fill = "white", color = NULL),
+        strip.text = element_text(size = 9))
+
+cairo_pdf(file = "fig8.pdf",
+          width = 11*0.4,
+          height = 8.5*0.4,
+          family = "Arial",
+          antialias = "subpixel")
+
+p1
+
+dev.off()
+
 df |> 
   filter(site_no == "lktexana_g") |> 
   mutate(year = lubridate::year(Date),
@@ -285,7 +333,7 @@ p1 <- bind_rows(lv_preds, lt_preds) |>
   scale_color_manual(name = "", values = c("black", "steelblue", "black")) +
[…]

### 2023-08-08 Rebuttal_Letter/Rebuttal_Letter.Rmd
incorporate new estuary models

@@
 
-Thank you for the prompt and thorough review of my submitted manuscript, "Linking watershed nutrient loading to estuary water quality with semiparametric models." I really appreciate the substantive and helpful comments from all the reviewers.
+Thank you for the prompt and thorough review of my manuscript, "Linking watershed nutrient loading to estuary water quality with semiparametric models." I really appreciate the substantive and helpful comments from each of the reviewers.
+The final draft and
+draft with differences marked are attached. Please note that the latexdiff tool
+did not highlight the very minor changes in the abstract or references. I also would like to note the updated title based on comments from one reviewer.
+
 Line by line comments and responses are included below (reviewer comments are italicized).
+I have incorporated nearly all of the requested changes from the reviewers.
 
@@
 
+**Response:** Dr. Marcus Beck made a similar comment in his review. This is now addressed
+by incorporating cumulative flows and loads into the estuary models. **need to address this and justify it in the manuscript now**.
+
 *Line 296: Would damming of the Navidad River impact the loads through the Lavaca River watershed? I don’t see why if the data was from the USGS-08164000 site.*
@@
 **Reviewer: Marcus Beck**
+*This article describes the use of Generalized Additive Models (GAMs) to estimate nutrient loads entering Lavica Bay from the main inflows.
+The author uses these models to evaluate the anthropogenic contribution of nutrients to the bay through flow-normalization and to evaluate if these loads are related to long-term variation in estuarine water quality.
+Overall, this is a valuable paper that further validates the use of GAMs for long-term trend analysis and will have particular value for the management of this system.
+My comments below relate to either clarification or adjustment of the methods that authors should consider.
+If pursued, I consider these changes appropriate for a major revision.*
+
+**Response:** I'm highly appreciative of these comments, they have substantially improved the manuscript.
 
@@
 I don’t see the value in including flow as a separate variable since it’s basically a surrogate for total load.*
+
+**Response:** I am combining my response to both these comments since they are related.
+These are valid concerns and related to a comment made by Reviewer 3 (probably indicative that I better make a change here).
+
+(1) I agree that it makes more sense to utilize the flow normalized loading estimates generated in the previous section. The methodology and results have been adjusted to correct this.
+
+(2) I struggled with the concept of incorporating lag effects while developing these models, it didn't occur to me to utilize aggregated loads/flows.
+I believe this is an appropriate approach and will incorporate it into the models.
+However, since the estuary water quality data is not measured at equal intervals, the suggested aggregation method is potentially inappropriate.
+For example, if a quarterly measurement was missed you can end up aggregating 6 months of loads for a given data point instead of 3 months.
+Walker et al.
+(2021) found that after Hurricane Harvey (a massive magnitude storm) water quality impacts in Lavaca Bay and adjacent estuaries were restricted to a relatively short period, somewhere between 20 days and a few months.
+Based on this info, I am incorporating 1-day lagged, 20-day cumulative loads and flows into this set of models instead of the same day measures.
+
+(3) In regard to the suggestion for changing equations 6 and 7 to utilize loads and flow-normalized loads instead of flow: I suspect there are a number of different approaches available here.
+While it is correct that the flow variable is essentially a surrogate for total load, inclusion does allow us to assess the relative contribution of "natural" or streamflow/load driven variation compared to the variation driven by flow-normalized loads or human activities.
+The lack of co-variation between flow normalized loads (in particular the aggregated loads) and the flow variable (shown below)
+
+```{r out.width="75%"}
+knitr::include_graphics("TP_plots.png")
+```
 
@@
 *Line 420: Replace “provide” with “prove” or “be”.*
+
+## References
+
+Walker, Lily M., Paul A.
+Montagna, Xinping Hu, and Michael S.
+Wetz.
+2021.
+“Timescales and Magnitude of Water Quality Change in Three Texas Estuaries Induced by Passage of Hurricane Harvey.” Estuaries and Coasts 44 (4): 960–71.
+https://doi.org/10.1007/s12237-020-00846-6.

### 2023-08-08 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.Rmd
incorporate new estuary models

@@
 ---
-title: Linking watershed nutrient loading to estuary water quality with semiparametric models
+title: Linking watershed nutrient loading to estuary water quality with Generalized Additive Models
 preprint: false
@@
   rename("Parameter" = " ") |> 
-  mutate(Method = rep(c("EPA 365.1", "EPA 353.2", "EPA 351.2", "EPA 445.0", "ASTM D888-09(C) and TCEQ SOP V1"),3),
+  mutate(`Station ID` = case_when(
+    `Station ID` == "TCEQ-13383" ~ "TCEQ-13383, Upper-Bay",
+    `Station ID` == "TCEQ-13384" ~ "TCEQ-13384, Lower-Bay",
+    `Station ID` == "TCEQ-13563" ~ "TCEQ-13563, Mid-Bay"),
+         Method = rep(c("EPA 365.1", "EPA 353.2", "EPA 351.2", "EPA 445.0", "ASTM D888-09(C) and TCEQ SOP V1"),3),
          `AWRL\\textsuperscript{a}`= rep(c(0.06, 0.05, 0.2, 3, NA),3),
          `Criteria\\textsuperscript{b}` = rep(c(0.21,NA,NA,11.6,5),3)) |>
+  mutate(`Station ID` = fct_relevel(`Station ID`,
+                                    "TCEQ-13383, Upper-Bay",
+                                    "TCEQ-13563, Mid-Bay",
+                                    "TCEQ-13384, Lower-Bay")) |> 
   kbl(format = "latex", booktabs = TRUE, escape = FALSE,
       caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.") |> 
-  column_spec(6, width = "3cm") |> 
+  column_spec(1, width = "2cm") |>
+  column_spec(6, width = "2.5cm") |> 
   collapse_rows(columns = 1, latex_hline = "major") |> 
@@
 thin-plate regression spline, $f_2(yday)$ is the numeric day of year smoothed 
-with a cyclic cubic regression spline, $f_3(Q)$ is mean daily inflow (the 
-combined measurements from Lavaca River and Navidad River) and $f_4(Load)$ is 
-the total NO~3~ or TP watershed load. The set of models specified for each water
+with a cyclic cubic regression spline.
+Prior work has shown that many water quality parameters may have lagged effects 
+lasting days or even months following storms and large discharge events [@mooneyWatershedExportEvents2012a; 
+@wetzExtremeFutureEstuaries2013; @bukaveckasInfluenceStormEvents2020; 
+@walkerTimescalesMagnitudeWater2021]. @walkerTimescalesMagnitudeWater2021 showed
+the impact of large flow events like Hurricane Harvey had somewhat short
+water quality impacts in Lavaca Bay and adjacent estuary systems, ranging 
+somewhere around 20 days to a few months. We incorporate this lag effect by using
+cumulative totals for $f_3(Q)$ and $f_4(Load)$. $f_3(Q)$ is the cumulative of the 
+previous 20 days of inflow (combined measurements from Lavaca River and 
+Navidad River) and $f_4(Load)$ is the cumulative of the previous 20 days 
+total NO~3~ or TP watershed load, both smoothed with thin plate regression 
+splines. The set of models specified for each water
 quality response are in Table\ \ref{tab:table3}.
@@
 both freshwater inflow and nutrient loads were prepossessed to account for season
-and streamflow respectively. Raw freshwater inflow values were replaced by 
-seasonally adjusted log transformed inflow obtained from the residuals of a GAM model 
-fit between season(day of year) and log transformed daily freshwater inflow. Raw nutrient
-loads were replaced with flow-adjusted values obtained from the residuals of a
-GAM model relating log transformed NO~3~ or TP loads to log transformed daily
-inflow. Response residuals from the respective GAM models were 
-used as *Q* and *Load* in Eq.\ \ref{eq6} and Eq.\ \ref{eq7}.
+and streamflow respectively. Cumulative inflow values were replaced by 
+seasonally adjusted cumulative inflow obtained from the residuals of a GAM model 
+fit between season(day of year) and log transformed cumulative inflow. Nutrient
+loads utilized the flow-normalized loads estimated in the previous section.
 
@@
       Model == "Flow + Load" ~ "Inflow + Load",
-      TRUE ~ as.character(Model)
+      TRUE ~ as.character(Model)),
+    Site = case_when(
+      Site == "TCEQ-13563" ~ "Mid-Bay",
+      Site == "TCEQ-13383" ~ "Upper-Bay",
+      Site == "TCEQ-13384" ~ "Lower-Bay"
     )) |>
   mutate(Parameter = fct_relevel(Parameter, "TP", "NO\\textsubscript{x}", "Chlorophyll-\\textit{a}","TKN", "DO"),
-         Model = fct_relevel(Model, "Temporal", "Inflow", "Inflow + Load")) |>
+         Model = fct_relevel(Model, "Temporal", "Inflow", "Inflow + Load"),
+         Site = fct_relevel(Site, "Upper-Bay", "Mid-Bay", "Lower-Bay")) |>
   tidyr::complete(Site, Model, nesting(Parameter)) |>
@@
   column_spec(column = 3,
-              bold = c(F,F,F,F,F,F,F,T,F,T,T,F,T,T,F),
-              italic = c(F,F,F,F,F,F,F,T,F,T,T,F,T,T,F)) |> 
+              bold = c(F,F,F,F,F,F,F,F,T,T,F,T,T,F,T),
+              italic = c(F,F,F,F,F,F,F,F,T,T,F,T,T,F,T)) |>
   column_spec(column = 4,
-              bold = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T),
-              italic = c(F,T,F,T,F,T,T,F,T,F,F,T,F,F,T)) |> 
+              bold = c(F,F,T,T,T,F,T,T,F,F,T,F,F,T,F),
+              italic = c(F,F,T,T,T,F,T,T,F,F,T,F,F,T,F)) |>
   column_spec(column = 5,
-              bold = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F),
-              italic = c(T,F,T,F,T,F,F,F,F,F,F,F,F,F,F)) |> 
+              bold = c(T,T,F,F,F,T,F,F,F,F,F,F,F,F,F),
+              italic = c(T,T,F,F,F,T,F,F,F,F,F,F,F,F,F)) |>
   collapse_rows(columns = 1, latex_hline = "major")
@@
                       `TCEQ-13563` = tar_read(tp_lavaca_13563_full, store = store),
-                      smooths = c("s(TP_resid)"),
+                      smooths = c("s(sum_TP_FN)"),
                       overall_uncertainty = FALSE)
 p_tp <- draw_smoothed_surface(comp,
-                      x = TP_resid,
+                      x = sum_TP_FN,
                       xlab = "Residuals log(TP) Load [kg]",
@@
                       `TCEQ-13563` = tar_read(no3_lavaca_13563_full, store = store),
-                      smooths = c("s(NO3_resid)"),
+                      smooths = c("s(sum_NO3_FN)"),
                       overall_uncertainty = FALSE)
 p_no3 <- draw_smoothed_surface(comp,
-                      x = NO3_resid,
+                      x = sum_NO3_FN,
                       xlab = "Residuals log(NO<sub>3</sub>) Load [kg]",

### 2023-08-08 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.tex
incorporate new estuary models

@@
 \title{Linking watershed nutrient loading to estuary water quality with
-semiparametric models}
+Generalized Additive Models}
 
@@
 \begin{threeparttable}
-\begin{tabular}[t]{lllll>{\raggedright\arraybackslash}p{3cm}rr}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}llll>{\raggedright\arraybackslash}p{2.5cm}rr}
 \toprule
@@
 
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13383} & DO (mg/L) & 7.22 & 1.35 & 55 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
+\multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13383, Upper-Bay} & DO (mg/L) & 7.22 & 1.35 & 55 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
 \cmidrule{1-8}
@@
 
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13384} & DO (mg/L) & 7.51 & 1.32 & 54 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
+\multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13384, Lower-Bay} & DO (mg/L) & 7.51 & 1.32 & 54 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
 \cmidrule{1-8}
@@
 
-\multirow{-5}{*}{\raggedright\arraybackslash TCEQ-13563} & DO (mg/L) & 7.91 & 1.34 & 56 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
+\multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13563, Mid-Bay} & DO (mg/L) & 7.91 & 1.34 & 56 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
 \bottomrule
@@
 spline, \(f_2(yday)\) is the numeric day of year smoothed with a cyclic
-cubic regression spline, \(f_3(Q)\) is mean daily inflow (the combined
-measurements from Lavaca River and Navidad River) and \(f_4(Load)\) is
-the total NO\textsubscript{3} or TP watershed load. The set of models
-specified for each water quality response are in Table~\ref{tab:table3}.
+cubic regression spline. Prior work has shown that many water quality
+parameters may have lagged effects lasting days or even months following
+storms and large discharge events
+\autocite{mooneyWatershedExportEvents2012a,wetzExtremeFutureEstuaries2013,bukaveckasInfluenceStormEvents2020,walkerTimescalesMagnitudeWater2021}.
+\textcite{walkerTimescalesMagnitudeWater2021} showed the impact of large
+flow events like Hurricane Harvey had somewhat short water quality
+impacts in Lavaca Bay and adjacent estuary systems, ranging somewhere
+around 20 days to a few months. We incorporate this lag effect by using
+cumulative totals for \(f_3(Q)\) and \(f_4(Load)\). \(f_3(Q)\) is the
+cumulative of the previous 20 days of inflow (combined measurements from
+Lavaca River and Navidad River) and \(f_4(Load)\) is the cumulative of
+the previous 20 days total NO\textsubscript{3} or TP watershed load. The
+set of models specified for each water quality response are in
+Table~\ref{tab:table3}.
 
@@
 inflow and nutrient loads were prepossessed to account for season and
-streamflow respectively. Raw freshwater inflow values were replaced by
-seasonally adjusted log transformed inflow obtained from the residuals
-of a GAM model fit between season(day of year) and log transformed daily
-freshwater inflow. Raw nutrient loads were replaced with flow-adjusted
-values obtained from the residuals of a GAM model relating log
-transformed NO\textsubscript{3} or TP loads to log transformed daily
-inflow. Response residuals from the respective GAM models were used as
-\emph{Q} and \emph{Load} in Eq.~\ref{eq6} and Eq.~\ref{eq7}.
+streamflow respectively. Cumulative inflow values were replaced by
+seasonally adjusted cumulative inflow obtained from the residuals of a
+GAM model fit between season(day of year) and log transformed cumulative
+inflow. Nutrient loads utilized the flow-normalized loads estimated in
+the previous section.
 
@@
 \midrule
- & TCEQ-13383 & -152.1 (0.03) & -156.1 (0.24) & \em{\textbf{-158.2 (0.72)}}\\
-
- & TCEQ-13384 & -194.4 (0.03) & \em{\textbf{-200.2 (0.49)}} & -200.2 (0.49)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash TP} & TCEQ-13563 & -145.3 (0) & -156.6 (0.41) & \em{\textbf{-157.3 (0.59)}}\\
+ & Upper-Bay & -152.1 (0.02) & -155.1 (0.07) & \em{\textbf{-160.2 (0.91)}}\\
+
+ & Mid-Bay & -145.3 (0.01) & -150 (0.12) & \em{\textbf{-154 (0.87)}}\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash TP} & Lower-Bay & -194.4 (0.39) & \em{\textbf{-194 (0.31)}} & -194 (0.31)\\
 \cmidrule{1-5}
- & TCEQ-13383 & -218.9 (0) & \em{\textbf{-244.8 (0.5)}} & -244.8 (0.5)\\
-
- & TCEQ-13384 & -263.4 (0) & -311.7 (0.48) & \em{\textbf{-311.9 (0.52)}}\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & TCEQ-13563 & -175.1 (0) & \em{\textbf{-190.2 (0.5)}} & -190.2 (0.5)\\
+ & Upper-Bay & -218.9 (0) & \em{\textbf{-241.7 (0.5)}} & -241.7 (0.5)\\
+
+ & Mid-Bay & -175.1 (0.01) & \em{\textbf{-183.8 (0.5)}} & -183.8 (0.5)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & Lower-Bay & -263.4 (0) & -298.1 (0.5) & \em{\textbf{-298.1 (0.5)}}\\
 \cmidrule{1-5}
- & TCEQ-13383 & 279.7 (0.18) & \em{\textbf{278.1 (0.41)}} & 278.1 (0.41)\\
-
- & TCEQ-13384 & \em{\textbf{268.2 (0.33)}} & 268.2 (0.33) & 268.2 (0.33)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & TCEQ-13563 & 289.5 (0.08) & \em{\textbf{286.1 (0.46)}} & 286.1 (0.46)\\
+ & Upper-Bay & 279.7 (0.24) & \em{\textbf{279.6 (0.26)}} & 278.2 (0.5)\\
+
+ & Mid-Bay & 289.5 (0.37) & \em{\textbf{289.5 (0.38)}} & 290.3 (0.25)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & Lower-Bay & \em{\textbf{268.2 (0.03)}} & 262.7 (0.48) & 262.7 (0.48)\\
 \cmidrule{1-5}
- & TCEQ-13383 & \em{\textbf{42.2 (0.66)}} & 43.5 (0.34) & -\\
-
- & TCEQ-13384 & \em{\textbf{34.3 (0.57)}} & 34.8 (0.43) & -\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash TKN} & TCEQ-13563 & 31.1 (0.22) & \em{\textbf{28.7 (0.78)}} & -\\
+ & Upper-Bay & \em{\textbf{42.2 (0.5)}} & 42.2 (0.5) & -\\
+
+ & Mid-Bay & 31.1 (0.5) & \em{\textbf{31.1 (0.5)}} & -\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash TKN} & Lower-Bay & \em{\textbf{34.3 (0.5)}} & 34.3 (0.5) & -\\
 \cmidrule{1-5}
- & TCEQ-13383 & \em{\textbf{146.4 (0.34)}} & 146.4 (0.34) & 146.5 (0.32)\\
-
- & TCEQ-13384 & \em{\textbf{135.9 (0.47)}} & 137 (0.27) & 137 (0.27)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash DO} & TCEQ-13563 & 138.3 (0.25) & \em{\textbf{137.2 (0.43)}} & 137.8 (0.32)\\
+ & Upper-Bay & \em{\textbf{146.4 (0.37)}} & 146.8 (0.29) & 146.5 (0.34)\\
+
+ & Mid-Bay & 138.3 (0.17) & \em{\textbf{136.4 (0.42)}} & 136.4 (0.42)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash DO} & Lower-Bay & \em{\textbf{135.9 (0.04)}} & 130.6 (0.48) & 130.6 (0.48)\\
 \bottomrule

### 2023-08-08 Rebuttal_Letter/Rebuttal_Letter.Rmd
reorder by estuary location

@@
 
-**Response:** Thank you, this is fixed.
+**Response:** This is fixed.
 
@@
 **Response:** I am combining my response to both these comments since they are related.
-These are valid concerns and related to a comment made by Reviewer 3 (probably indicative that I better make a change here).
+These are valid concerns and related to a comment made by Reviewer 3 (indicative that I better make a change here).
 
@@
 
+**Response:** Title has been updated.
+
 *Line 18: Maybe replace “non-natural variation” with “anthropogenic sources”.*
@@
 *Tables 1 and 2: It would be useful to indicate the approximate sampling interval for the water quality parameters. It looks like quarterly. *
+
+**Response:** I'm out of space in the tables, but include this in the text **(addline number)**.
 
@@
 
+**Response:** (1) I have updated the equation to (hopefully) be more clear about
+the calculation.
+For each day in the record, the full record of flows is utilized (an admittedly not efficient calculation, but not an issue for a period of record the size of this data set).
+(2) I set *d* to 0.95 based on results from Robson and Dourdet (2015) and Zhang and Ball (2017), which is now discussed in **lines ...**
+
 *Line 151: It’s not clear to me how fa is a unitless term. The equations show these as simple differences of average streamflow, so wouldn’t the units be in log-cfs? *
+
+**Response:** I worded this incorrectly. I should have phrased this "dimensionless." This is fixed in **lines ...**
 
@@
 
+**Response:** 4% of TP samples were below detection limits and 20% of NO~3~ 
+samples fell below detection limits.
+Early in this project 
+I explored the use of the Tobit I family using the linked cenGAM package and 
+unfortunately found the models result in large overestimation of concentrations.
+I similarly explored the use of censored Gamma models in brms with similar 
+results. Bergbusch et al. (2021) report similar findings. Conversely, Hayes 
+et al. (2020) apply a censored GAM using brms without issue on microcystin 
+concentration in lakes. brms appears promising for fitting censored water 
+quality data, but I was frankly unable to fit reasonable models for this project
+which might be due to sample size or my misunderstanding of model application.
+
 *Line 236: This might be a personal preference, but “unusual periodic pattern” seems like regular periodicity and not a more random periodicity as one would expect with rain events.
@@
 
+**Response:** Rephrased to simply "periodic pattern."
+
 *Figure 2 caption: Consider adding a sentence like “Values closer to 1 for NSE and r and values closer to 0 for PBIAS represent more ideal goodness-of-fit assessments.”*
+
+**Response:** Incorporated the suggested sentence.
 
@@
 *Figure 5 caption: Consider including a reference to Table 3, as it wasn’t immediately clear to me that these models didn’t include flow or load. *
+
+**Response:** Added a reference to Table 3.
 
@@
 
+**Respones:** I've replaced the site names and reordered the table based on location.
+You are correct, the statistical summaries for several models are identical indicating that the more complicated models are not explaining additional variance.
+
 *Line 272: The text “contributing tributaries” seems redundant. Maybe remove “contributing”. *
+
+**Response:** Removed "contributing."
 
@@
 
+**Response:** Fixed.
+
 *Line 395: Replace “suggest” with “such as”.*
+
+**Response:** Fixed.
 
@@
 
+**Response:** Fixed.
+
 ## References
+
+Hayes, Nicole M., Heather A.
+Haig, Gavin L.
+Simpson, and Peter R.
+Leavitt.
+2020.
+“Effects of Lake Warming on the Seasonal Risk of Toxic Cyanobacteria Exposure.” Limnology and Oceanography Letters 5 (6): 393–402.
+https://doi.org/10.1002/lol2.10164.
+
+Robson, Barbara J., and Vincent Dourdet.
+2015.
+“Prediction of Sediment, Particulate Nutrient and Dissolved Nutrient Concentrations in a Dry Tropical River to Provide Input to a Mechanistic Coastal Water Quality Model.” Environmental Modelling & Software 63 (January): 97–108.
+https://doi.org/10.1016/j.envsoft.2014.08.009.
 
@@
 https://doi.org/10.1007/s12237-020-00846-6.
+
+Zhang, Qian, and William P.
+Ball.
+2017.
+“Improving Riverine Constituent Concentration and Flux Estimation by Accounting for Antecedent Discharge Conditions.” Journal of Hydrology 547 (April): 387–402.
+https://doi.org/10.1016/j.jhydrol.2016.12.052.

### 2023-08-08 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.Rmd
reorder by estuary location

@@
 [@horowitzEvaluationSedimentRating2003; @snelderEstimationCatchmentNutrient2017]. 
-The three estuary sites included an upper Lavaca Bay site near the outlet of the
-Lavaca River system (TCEQ-13563), a mid-Lavaca Bay site (TCEQ-13383), and the 
-lower Lavaca Bay site near the mouth of the Bay (TCEQ-13384). 
+The three estuary sites included an upper-bay site near the outlet of the
+Lavaca River system (TCEQ-13563), a mid-bay site (TCEQ-13383), and the 
+lower-bay site near the mouth of the Bay (TCEQ-13384). 
 For estuary locations, we obtained data for TP, Nitrite+Nitrate (NO~*X*~), TKN, 
 chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
-(Table\ \ref{tab:table2}).
+(Table\ \ref{tab:table2}). All water quality data was collected approximately 
+quarterly.
 
@@
   mutate(`Station ID` = case_when(
-    `Station ID` == "TCEQ-13383" ~ "TCEQ-13383, Upper-Bay",
+    `Station ID` == "TCEQ-13383" ~ "TCEQ-13383, Mid-Bay",
     `Station ID` == "TCEQ-13384" ~ "TCEQ-13384, Lower-Bay",
-    `Station ID` == "TCEQ-13563" ~ "TCEQ-13563, Mid-Bay"),
+    `Station ID` == "TCEQ-13563" ~ "TCEQ-13563, Upper-Bay"),
          Method = rep(c("EPA 365.1", "EPA 353.2", "EPA 351.2", "EPA 445.0", "ASTM D888-09(C) and TCEQ SOP V1"),3),
@@
   mutate(`Station ID` = fct_relevel(`Station ID`,
-                                    "TCEQ-13383, Upper-Bay",
-                                    "TCEQ-13563, Mid-Bay",
+                                    "TCEQ-13563, Upper-Bay",
+                                    "TCEQ-13383, Mid-Bay",
                                     "TCEQ-13384, Lower-Bay")) |> 
+  arrange(`Station ID`) |> 
   kbl(format = "latex", booktabs = TRUE, escape = FALSE,
@@
            escape = F)
-
-# modelsummary::datasummary(`Station ID` * (`TP (mg/L)`+
-#                            `NO\\textsubscript{x} (mg/L)` +
-#                            `TKN (mg/L)` +
-#                            `Chlorophyll-\\textit{a} (ug/L)` +
-#                            `DO (mg/L)`) ~
-#                            Mean*Arguments(na.rm=TRUE) + SD*Arguments(na.rm=TRUE) + N,
-#                          data = df, output = "kableExtra", format = "latex",
-#                          escape = FALSE,
-#                          caption = "Summary of estuary water quality samples collected between January 1, 2005 and December 31, 2020.")
 ```
@@
 current time period [@wangLoadEstimationUncertainties2011;  
-@kuhnertQuantifyingTotalSuspended2012; @zhang_improving_2017], using Eq. \ref{eq2}:
+@kuhnertQuantifyingTotalSuspended2012; @wangSedimentConcentrationPrediction2013a; @zhang_improving_2017], using Eq. \ref{eq2}:
 
 \begin{equation}
-ma(t) = \frac{\sum^{J}_{j=1}{d^{t-t_j}Q_{t_{j}}}}{\sum^{J}_{j=1}d^{t-t_j}}
+ma(d) = \frac{Q_j + dQ_{j-1} + \mathellipsis + d^{j-1}Q_1}{1+d+...+d^{j-1}}
 \label{eq2}
@@
 
-where *t* is the day of estimation, *d* is the discount factor (set to 0.95),
-*J* is the total number of daily discharge observations, $t-t_j$ is the lag time
-between the current day and some historical day of observation ($t_j$). Here,
-as *d* approaches zero, the time series becomes the daily observed values and 
-as *d* gets closer to one, the time series becomes the mean flow.
-
-Flow anomaly (*fa*) is a unitless term that represents how wet or dry the 
-current time period is from a previous time period [@vecchiaTrendsConcentrationsUse2009; 
-@zhang_improving_2017]. Long-term flow anomaly (*ltfa*) is the streamflow over 
-the previous year relative to the entire period [Eq.\ \ref{eq3}, @zhang_improving_2017] and 
+where *ma* at discount factor *d* is calculated using flows $(Q_1,Q_2,...,Q_j)$
+for days 1 through *j*. Here, as *d* approaches zero, the time series 
+becomes the daily observed values and as *d* gets closer to one, the time series 
+becomes the mean flow. Although @kuhnertQuantifyingTotalSuspended2012 and 
+@wangSedimentConcentrationPrediction2013a suggest using multiple covariates with
+*d* set at different values from 0.75 to 0.99, @robson_prediction_2015-1 and 
+@zhang_improving_2017 found substantial improvements in statistical nutrient 
+models with *d*  equal to 0.95 which was adopted in this study.
+
+Flow anomaly (*fa*) is a dimensionless term that represents how wet or dry 
+(or how anomalous) the current time period is from previous time periods
+[@vecchiaTrendsConcentrationsUse2009; @zhang_improving_2017]. 
+Long-term flow anomaly (*ltfa*) is the streamflow over 
+the previous year relative to the entire period 
+[Eq.\ \ref{eq3}, @zhang_improving_2017] and 
 and the short-term flow anomaly (*stfa*) calculated as the current day flow 
@@
 
-# flextable::flextable(df) |> 
-#   set_caption(caption = "Set of GAM models specified for each water quality parameter response.",
-#               autonum = run_autonum(seq_id = "tab", bkm = "table3")) |> 
-#   style(pr_t = fp_text_default(font.size = 9), part = "body") |>
-#   style(pr_t = fp_text_default(font.size = 9, bold = TRUE), part = "header") |>
-#   set_table_properties(layout = "autofit") |>
-#   padding(padding.top = 0, padding.bottom = 0, part = "body")
-
 ```
@@
 ```{r fig2, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.65, fig.cap="Density plots of goodness-of-fit metrics (NSE, \\textit{r}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
-Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics.", dev='cairo_pdf'}
+Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics.
+Values closer to 1 for NSE and r and values closer to 0 for PBIAS represent more ideal goodness-of-fit assessments.", dev='cairo_pdf'}
 
@@
 decreased from 2005 through 2014 (Fig.\ \ref{fig:fig5}). NO~*X*~ concentration
-at the mid-bay site, TCEQ-13383, displayed an unusual periodic pattern that is
+at the mid-bay site, TCEQ-13383, displayed a periodic pattern that is
 indicative of a strong influence from inflow or precipitation. The temporal GAMs
@@
 
-```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.", dev='cairo_pdf'}
+```{r fig5, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines (shaded regions indicate 90\\% confidence intervals) from the temporal estuary GAM (Table \\ref{tab:table3}) display the marginal smoothed effect of date on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.", dev='cairo_pdf'}
 
@@
 
-  `TCEQ-13563` <- model1
-  `TCEQ-13383` <- model2
-  `TCEQ-13384` <- model3
+  `Upper-Bay` <- model1
+  `Mid-Bay` <- model2
+  `Lower-Bay` <- model3
 
@@
   comp <- gratia::compare_smooths(
-    `TCEQ-13563`,
-    `TCEQ-13383`,
-    `TCEQ-13384`,
+    `Upper-Bay`,
+    `Mid-Bay`,
+    `Lower-Bay`,
     smooths = "s(ddate)",
@@
                            as.character(model),
-                           NA_character_)) |> 
+                           NA_character_),
+           model = forcats::fct_relevel(model, "Upper-Bay", "Mid-Bay", "Lower-Bay")) |> 
     ggplot(aes(x = ddate, y = est, group = model)) +
@@
                        expand = expansion(mult = c(0.05,0.05))) +
-    scale_fill_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
-    scale_color_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
-    scale_linetype_manual(name = "", values = c("solid","twodash","dashed")) +
+    scale_fill_manual(name = "", 
+                      values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_color_manual(name = "", 
+                       values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_linetype_manual(name = "", 
+                          values = c("solid","twodash","dashed")) +
     labs(x = "Year", y = ylab,
@@
     Site = case_when(
-      Site == "TCEQ-13563" ~ "Mid-Bay",
-      Site == "TCEQ-13383" ~ "Upper-Bay",
+      Site == "TCEQ-13563" ~ "Upper-Bay",
+      Site == "TCEQ-13383" ~ "Mid-Bay",
       Site == "TCEQ-13384" ~ "Lower-Bay"
@@
   column_spec(column = 3,
-              bold = c(F,F,F,F,F,F,F,F,T,T,F,T,T,F,T),
-              italic = c(F,F,F,F,F,F,F,F,T,T,F,T,T,F,T)) |>
+              bold = c(F,F,T,F,F,F,F,F,F,T,T,T,F,T,F),
+              italic = c(F,F,T,F,F,F,F,F,F,T,T,T,F,T,F)) |>
   column_spec(column = 4,
-              bold = c(F,F,T,T,T,F,T,T,F,F,T,F,F,T,F),
-              italic = c(F,F,T,T,T,F,T,T,F,F,T,F,F,T,F)) |>
+              bold = c(F,F,F,T,T,T,T,F,T,F,F,F,T,F,T),
+              italic = c(F,F,F,T,T,T,T,F,T,F,F,F,T,F,T)) |>
   column_spec(column = 5,
-              bold = c(T,T,F,F,F,T,F,F,F,F,F,F,F,F,F),
-              italic = c(T,T,F,F,F,T,F,F,F,F,F,F,F,F,F)) |>
+              bold = c(T,T,F,F,F,F,F,T,F,F,F,F,F,F,F),
+              italic = c(T,T,F,F,F,F,F,T,F,F,F,F,F,F,F)) |>
   collapse_rows(columns = 1, latex_hline = "major")
@@
     scale_x_continuous(expand = expansion(mult = c(0.05,0.05))) +
-    scale_fill_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
-    scale_color_manual(name = "", values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
-    scale_linetype_manual(name = "", values = c("solid","twodash","dashed")) +
+    scale_fill_manual(name = "", 
+                      values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_color_manual(name = "", 
+                       values = unname(palette.colors(palette = "Okabe-Ito")[2:4])) +
+    scale_linetype_manual(name = "", 
+                          values = c("solid","twodash","dashed")) +
     labs(x = xlab, y = ylab,
@@
 cycling may have regulating effects that mask changes
-in upstream nitrogen loadings. Additional nutrient data collection in the contributing 
+in upstream nitrogen loadings. Additional nutrient data collection in the 
 tributaries of Lake Texana is needed to fully assess the role of Lake Texana
@@
 flushing effect at higher freshwater inflow. No relationship between TKN and 
-freshwater inflow were observed at TCEQ-13384 located in the lower reach of 
+freshwater inflow was observed at TCEQ-13384 located in the lower reach of 
 Lavaca Bay. The results suggest that tidal flushing might dilute TKN and act as 
@@
 responses in water quality. Although we did not compare other models, it is likely similar
-estimates of loadings would be obtained by methods suggest as LOADEST, WRTDS, or SPARROW
+estimates of loadings would be obtained by methods such as LOADEST, WRTDS, or SPARROW
 given the functionally similar dependent variable structures. The underlying weakness 
@@
 tracking progress towards long-term pollutant reduction 
-goals. Similar coordinated efforts across Texas coastal watersheds would provide
+goals. Similar coordinated efforts across Texas coastal watersheds would prove
 useful for resource management efforts intended to protect the biological and 

### 2023-08-08 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.tex
reorder by estuary location

@@
 \autocite{horowitzEvaluationSedimentRating2003,snelderEstimationCatchmentNutrient2017}.
-The three estuary sites included an upper Lavaca Bay site near the
-outlet of the Lavaca River system (TCEQ-13563), a mid-Lavaca Bay site
-(TCEQ-13383), and the lower Lavaca Bay site near the mouth of the Bay
-(TCEQ-13384). For estuary locations, we obtained data for TP,
-Nitrite+Nitrate (NO\textsubscript{\emph{X}}), TKN, chlorophyll-\emph{a},
-and DO concentrations from January 2005 through December 2020
-(Table~\ref{tab:table2}).
+The three estuary sites included an upper-bay site near the outlet of
+the Lavaca River system (TCEQ-13563), a mid-bay site (TCEQ-13383), and
+the lower-bay site near the mouth of the Bay (TCEQ-13384). For estuary
+locations, we obtained data for TP, Nitrite+Nitrate
+(NO\textsubscript{\emph{X}}), TKN, chlorophyll-\emph{a}, and DO
+concentrations from January 2005 through December 2020
+(Table~\ref{tab:table2}). All water quality data was collected
+approximately quarterly.
 
@@
 \midrule
+ & TP (mg/L) & 0.13 & 0.06 & 50 & EPA 365.1 & 0.06 & 0.21\\
+
+ & NO\textsubscript{x} (mg/L) & 0.09 & 0.13 & 53 & EPA 353.2 & 0.05 & \\
+
+ & TKN (mg/L) & 0.94 & 0.37 & 49 & EPA 351.2 & 0.20 & \\
+
+ & Chlorophyll-\textit{a} (ug/L) & 9.67 & 5.33 & 49 & EPA 445.0 & 3.00 & 11.60\\
+
+\multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13563, Upper-Bay} & DO (mg/L) & 7.91 & 1.34 & 56 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
+\cmidrule{1-8}
  & TP (mg/L) & 0.11 & 0.05 & 47 & EPA 365.1 & 0.06 & 0.21\\
@@
 
-\multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13383, Upper-Bay} & DO (mg/L) & 7.22 & 1.35 & 55 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
+\multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13383, Mid-Bay} & DO (mg/L) & 7.22 & 1.35 & 55 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
 \cmidrule{1-8}
@@
 \multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13384, Lower-Bay} & DO (mg/L) & 7.51 & 1.32 & 54 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
-\cmidrule{1-8}
- & TP (mg/L) & 0.13 & 0.06 & 50 & EPA 365.1 & 0.06 & 0.21\\
-
- & NO\textsubscript{x} (mg/L) & 0.09 & 0.13 & 53 & EPA 353.2 & 0.05 & \\
-
- & TKN (mg/L) & 0.94 & 0.37 & 49 & EPA 351.2 & 0.20 & \\
-
- & Chlorophyll-\textit{a} (ug/L) & 9.67 & 5.33 & 49 & EPA 445.0 & 3.00 & 11.60\\
-
-\multirow{-5}{2cm}{\raggedright\arraybackslash TCEQ-13563, Mid-Bay} & DO (mg/L) & 7.91 & 1.34 & 56 & ASTM D888-09(C) and TCEQ SOP V1 &  & 5.00\\
 \bottomrule
@@
 concentration at the current time period
-\autocite{wangLoadEstimationUncertainties2011,kuhnertQuantifyingTotalSuspended2012,zhang_improving_2017},
+\autocite{wangLoadEstimationUncertainties2011,kuhnertQuantifyingTotalSuspended2012,wangSedimentConcentrationPrediction2013a,zhang_improving_2017},
 using Eq. \ref{eq2}:
@@
 \begin{equation}
-ma(t) = \frac{\sum^{J}_{j=1}{d^{t-t_j}Q_{t_{j}}}}{\sum^{J}_{j=1}d^{t-t_j}}
+ma(d) = \frac{Q_j + dQ_{j-1} + \mathellipsis + d^{j-1}Q_1}{1+d+...+d^{j-1}}
 \label{eq2}
@@
 
-where \emph{t} is the day of estimation, \emph{d} is the discount factor
-(set to 0.95), \emph{J} is the total number of daily discharge
-observations, \(t-t_j\) is the lag time between the current day and some
-historical day of observation (\(t_j\)). Here, as \emph{d} approaches
-zero, the time series becomes the daily observed values and as \emph{d}
-gets closer to one, the time series becomes the mean flow.
-
-Flow anomaly (\emph{fa}) is a unitless term that represents how wet or
-dry the current time period is from a previous time period
+where \emph{ma} at discount factor \emph{d} is calculated using flows
+\((Q_1,Q_2,...,Q_j)\) for days 1 through \emph{j}. Here, as \emph{d}
+approaches zero, the time series becomes the daily observed values and
+as \emph{d} gets closer to one, the time series becomes the mean flow.
+Although \textcite{kuhnertQuantifyingTotalSuspended2012} and
+\textcite{wangSedimentConcentrationPrediction2013a} suggest using
+multiple covariates with \emph{d} set at different values from 0.75 to
+0.99, \textcite{robson_prediction_2015-1} and
+\textcite{zhang_improving_2017} found substantial improvements in
+statistical nutrient models with \emph{d} equal to 0.95 which was
+adopted in this study.
+
+Flow anomaly (\emph{fa}) is a dimensionless term that represents how wet
+or dry (or how anomalous) the current time period is from previous time
+periods
 \autocite{vecchiaTrendsConcentrationsUse2009,zhang_improving_2017}.
@@
 Lavaca River and Navidad River) and \(f_4(Load)\) is the cumulative of
-the previous 20 days total NO\textsubscript{3} or TP watershed load. The
-set of models specified for each water quality response are in
-Table~\ref{tab:table3}.
+the previous 20 days total NO\textsubscript{3} or TP watershed load,
+both smoothed with thin plate regression splines. The set of models
+specified for each water quality response are in Table~\ref{tab:table3}.
 
@@
 \caption{Density plots of goodness-of-fit metrics (NSE, \textit{r}, and PBIAS) from repeated 5-fold cross validation between predicted nutrient loads from GAM models and measured nutrient loads.
-Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics.}\label{fig:fig2}
+Color indicates the tail probability calculated from the empirical cumulative distribution of the goodness-of-fit metrics.
+Values closer to 1 for NSE and r and values closer to 0 for PBIAS represent more ideal goodness-of-fit assessments.}\label{fig:fig2}
 \end{figure}
@@
 2005 through 2014 (Fig.~\ref{fig:fig5}). NO\textsubscript{\emph{X}}
-concentration at the mid-bay site, TCEQ-13383, displayed an unusual
-periodic pattern that is indicative of a strong influence from inflow or
+concentration at the mid-bay site, TCEQ-13383, displayed a periodic
+pattern that is indicative of a strong influence from inflow or
 precipitation. The temporal GAMs did not provide evidence of long-term
@@
 
-\caption{Fitted splines (shaded regions indicate 90\% confidence intervals) from the temporal estuary GAM display the marginal smoothed effect of date on TP (A), NO\textit{\textsubscript{X}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig5}
+\caption{Fitted splines (shaded regions indicate 90\% confidence intervals) from the temporal estuary GAM (Table \ref{tab:table3}) display the marginal smoothed effect of date on TP (A), NO\textit{\textsubscript{X}} (B), chlorophyll-\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.}\label{fig:fig5}
 \end{figure}
@@
 \midrule
- & Upper-Bay & -152.1 (0.02) & -155.1 (0.07) & \em{\textbf{-160.2 (0.91)}}\\
-
- & Mid-Bay & -145.3 (0.01) & -150 (0.12) & \em{\textbf{-154 (0.87)}}\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash TP} & Lower-Bay & -194.4 (0.39) & \em{\textbf{-194 (0.31)}} & -194 (0.31)\\
+ & Upper-Bay & -145.3 (0.01) & -150 (0.12) & \em{\textbf{-154 (0.87)}}\\
+
+ & Mid-Bay & -152.1 (0.02) & -155.1 (0.07) & \em{\textbf{-160.2 (0.91)}}\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash TP} & Lower-Bay & \em{\textbf{-194.4 (0.39)}} & -194 (0.31) & -194 (0.31)\\
 \cmidrule{1-5}
- & Upper-Bay & -218.9 (0) & \em{\textbf{-241.7 (0.5)}} & -241.7 (0.5)\\
-
- & Mid-Bay & -175.1 (0.01) & \em{\textbf{-183.8 (0.5)}} & -183.8 (0.5)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & Lower-Bay & -263.4 (0) & -298.1 (0.5) & \em{\textbf{-298.1 (0.5)}}\\
+ & Upper-Bay & -175.1 (0.01) & \em{\textbf{-183.8 (0.5)}} & -183.8 (0.5)\\
+
+ & Mid-Bay & -218.9 (0) & \em{\textbf{-241.7 (0.5)}} & -241.7 (0.5)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash NO\textsubscript{x}} & Lower-Bay & -263.4 (0) & \em{\textbf{-298.1 (0.5)}} & -298.1 (0.5)\\
 \cmidrule{1-5}
- & Upper-Bay & 279.7 (0.24) & \em{\textbf{279.6 (0.26)}} & 278.2 (0.5)\\
-
- & Mid-Bay & 289.5 (0.37) & \em{\textbf{289.5 (0.38)}} & 290.3 (0.25)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & Lower-Bay & \em{\textbf{268.2 (0.03)}} & 262.7 (0.48) & 262.7 (0.48)\\
+ & Upper-Bay & 289.5 (0.37) & \em{\textbf{289.5 (0.38)}} & 290.3 (0.25)\\
+
+ & Mid-Bay & 279.7 (0.24) & 279.6 (0.26) & \em{\textbf{278.2 (0.5)}}\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash Chlorophyll-\textit{a}} & Lower-Bay & 268.2 (0.03) & \em{\textbf{262.7 (0.48)}} & 262.7 (0.48)\\
 \cmidrule{1-5}
- & Upper-Bay & \em{\textbf{42.2 (0.5)}} & 42.2 (0.5) & -\\
-
- & Mid-Bay & 31.1 (0.5) & \em{\textbf{31.1 (0.5)}} & -\\
+ & Upper-Bay & \em{\textbf{31.1 (0.5)}} & 31.1 (0.5) & -\\
+
+ & Mid-Bay & \em{\textbf{42.2 (0.5)}} & 42.2 (0.5) & -\\
 
@@
 \cmidrule{1-5}
- & Upper-Bay & \em{\textbf{146.4 (0.37)}} & 146.8 (0.29) & 146.5 (0.34)\\
-
- & Mid-Bay & 138.3 (0.17) & \em{\textbf{136.4 (0.42)}} & 136.4 (0.42)\\
-
-\multirow{-3}{*}{\raggedright\arraybackslash DO} & Lower-Bay & \em{\textbf{135.9 (0.04)}} & 130.6 (0.48) & 130.6 (0.48)\\
+ & Upper-Bay & 138.3 (0.17) & \em{\textbf{136.4 (0.42)}} & 136.4 (0.42)\\
+
+ & Mid-Bay & \em{\textbf{146.4 (0.37)}} & 146.8 (0.29) & 146.5 (0.34)\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash DO} & Lower-Bay & 135.9 (0.04) & \em{\textbf{130.6 (0.48)}} & 130.6 (0.48)\\
 \bottomrule
@@
 that mask changes in upstream nitrogen loadings. Additional nutrient
-data collection in the contributing tributaries of Lake Texana is needed
-to fully assess the role of Lake Texana in regulating nutrient delivery
-to the Lavaca Bay system. However, these results suggest that there have
-been no changes in the NO\textsubscript{3} or TP loading from the
-Navidad River system at the Lake Texana discharge point when accounting
-for variations in year to year discharge.
+data collection in the tributaries of Lake Texana is needed to fully
+assess the role of Lake Texana in regulating nutrient delivery to the
+Lavaca Bay system. However, these results suggest that there have been
+no changes in the NO\textsubscript{3} or TP loading from the Navidad
+River system at the Lake Texana discharge point when accounting for
+variations in year to year discharge.
 
@@
 attenuated, possibly indicating a flushing effect at higher freshwater
-inflow. No relationship between TKN and freshwater inflow were observed
+inflow. No relationship between TKN and freshwater inflow was observed
 at TCEQ-13384 located in the lower reach of Lavaca Bay. The results
@@
 models, it is likely similar estimates of loadings would be obtained by
-methods suggest as LOADEST, WRTDS, or SPARROW given the functionally
+methods such as LOADEST, WRTDS, or SPARROW given the functionally
 similar dependent variable structures. The underlying weakness in the
@@
 long-term pollutant reduction goals. Similar coordinated efforts across
-Texas coastal watersheds would provide useful for resource management
+Texas coastal watersheds would prove useful for resource management
 efforts intended to protect the biological and water quality integrity

### 2023-08-11 Rebuttal_Letter/Rebuttal_Letter.Rmd
resubmission ready

@@
 return-address: 
-  - Texas A&M AgriLife
   - Texas Water Resources Institute
@@
   - College Station, Texas 77840
-return-short: Michael Schramm, 2118 TAMU, College Station, TX 77840
+return-email: michael.schramm@ag.tamu.edu
 address:
@@
 signature: Michael Schramm
+
+lco: UScommercial9
+papersize: letter
+
+inline-links: true
+
+komaoption:
+  - parskip=full
 
@@
 
+**Comment:**
 *Very well-structured abstract – introduced the problem, the objectives of the research, the results obtained, and the conclusion. No changes are recommended.*
@@
 
+**Comment:**
 *Line 114: A link to the USGS NWIS system could be added.*
 
-**Response:** The url has been added.
+**Response:** The url has been added (Line 122).
 
+**Comment:**
 *A brief description of the sample collection and laboratory analyses described in the QAPP and TCEQ procedures manual, detection and recovery limits, and instrumentation would seem appropriate.*
@@
 
+**Comment:**
 *Also, what are the state-recommended guideline values? Include them in Table 2*
@@
 
+**Comment:**
 *Line 371 “has” repeated twice*
@@
 
+**Comment:**
 *The author adequately answered the research questions and fulfilled all the stated objectives.
@@
 
+**Comment:**
 *Table 1: It would be helpful if you would include the watershed or station name under the USGS station ID in this table.*
@@
 
+**Comment:**
 *Line 164: Not sure if this is a typo, or if I’m missing what log1p(Q) is.*
 
-**Response:** This was shorthand for *log(1+Q)*. I've changed this to *log(1+Q)* to be consistent with equation 1.
+**Response:** This was shorthand for *log(1+Q)*. I've changed this to *log(1+Q)* to be consistent with equation 1 (line 178).
 
+**Comment:**
 *Line 392, "Limitation" section: I expected this to be a discussion of limitations of the approach like matching daily loads to concentrations, the complex environmental factors that could be involved in chlorophyll and DO trends, changing climate, etc.
@@
 
+**Response:**
+I agree, this subsection title is updated to "Future Management and Research Needs" (line 398).
+
+**Comment:**
 *Line 53-54: Since only one of the trends you mention from Bugica is for a nitrogen species, I suggest rephrasing or expanding this sentence to explain why we’d be concerned about all of the parameters that are increasing (or clarify if we are only concerned about the TKN).*
 
+**Response:**
+Added a few lines clarifying the need to assess both inorganic and organic species and phosphorus due to the complex roles they have in contributing to estuary health (lines 53-59).
+
+**Comment:**
 *Lines 129-131: It would be helpful if you described how deep the water column is at the estuary sample locations and from what depth the samples that you analyzed were taken.
@@
 
-**Response:** I have added additional description of the estuary dynamics *(lines 101-108)*.
+**Response:** I have added additional description of the estuary dynamics (lines 108-111).
 Specifically mixing, turbidity, and salinity.
 Given the preliminary nature of this study, there is not a season of concern identified yet, so it is not included.
-*still need to add sample depth here*
+Depth measurements are also detailed in lines 139-143.
 
+**Comment:**
 *Lines 194-201: Some discussion is needed of how you dealt with temporal matching of flow and loads to concentrations in the estuary.
@@
 **Response:** Dr. Marcus Beck made a similar comment in his review. This is now addressed
-by incorporating cumulative flows and loads into the estuary models. **need to address this and justify it in the manuscript now**.
+by incorporating cumulative flows and loads into the estuary models and further discussed below.
+See revisions in lines 208-214.
+Some minor changes to the results were made, but the conclusions are largely consistent with the initial submission.
 
+**Comment:**
 *Line 296: Would damming of the Navidad River impact the loads through the Lavaca River watershed? I don’t see why if the data was from the USGS-08164000 site.*
@@
 combined watershed loads, when these estimates were for the Lavaca River watershed only.
-I've removed this and incorporated your statement below. Thank you very much for catching this!
+I've removed this and incorporated your statement below (lines 301-325). Thank you very much for catching this!
 
+**Comment:**
 *Lines 286-321: I think this discussion is too long comparing the TP yield first to the Dunn study, and then to 3 other studies.
@@
 
-**Response:** I agree, this section has been considerably condensed.
+**Response:** I agree, this section has been considerably condensed. See response above.
 
+**Comment:**
 *Lines 360-361: It is probably worthwhile to note that the effect of flow you see in chlorophyll is very likely due to more nutrients coming in with high flows.
@@
 
+**Response:** This is one place where we had a slight change in the results due to incorporating cumulative flows and loads. This is reflected in revisions in lines 349-358.
+
 **Reviewer: Marcus Beck**
+
+**Comment:**
 *This article describes the use of Generalized Additive Models (GAMs) to estimate nutrient loads entering Lavica Bay from the main inflows.
@@
 
+**Comment:**
 *I have some concerns about the methods starting on line 184 that describe how the load estimates are related to the estuarine data.
@@
 
-(1) I agree that it makes more sense to utilize the flow normalized loading estimates generated in the previous section. The methodology and results have been adjusted to correct this.
+(1) I agree that it makes more sense to utilize the flow normalized loading estimates generated in the previous section. The methodology and results have been adjusted to correct this (lines 250-220).
 
@@
 Based on this info, I am incorporating 1-day lagged, 20-day cumulative loads and flows into this set of models instead of the same day measures.
+The aggregated flows are still seasonally adjusted (lines 206-214).
 
-(3) In regard to the suggestion for changing equations 6 and 7 to utilize loads and flow-normalized loads instead of flow: I suspect there are a number of different approaches available here.
-While it is correct that the flow variable is essentially a surrogate for total load, inclusion does allow us to assess the relative contribution of "natural" or streamflow/load driven variation compared to the variation driven by flow-normalized loads or human activities.
-The lack of co-variation between flow normalized loads (in particular the aggregated loads) and the flow variable (shown below)
+(3) In regard to the suggestion for changing equations 6 and 7 to utilize loads and flow-normalized loads instead of flow: I recognize there are a number of valid approaches available here.
+Since aggregated loads and estuary inflow are surrogates, I don't fully understand the value of replacing the seasonally adjust inflow values for aggregated loads.
+If I understand the models correctly, both the reviewer suggested approach and the approach used in the draft evaluate the relative variation driven by changes in inflow and anthropogenic changes in nonpoint source loads.
 
-```{r out.width="75%"}
-knitr::include_graphics("TP_plots.png")
-```
+In sum, the results have been slightly revised, but overall conclusions from the study remain the same.
+In particular, responses of chlorophyll-a at one site, and TKN at two of the sites are changed in response to flow and/or nutrient loads.
+This are updated in the relevant sections in the results and discussion.
 
+**Comment:**
 *I also strongly encourage the author to consider replacing “semiparametric models” with “Generalized Additive Models” in the title. It’s a more accurate description of what’s included in the paper.*
@@
 
+**Comment:**
 *Line 18: Maybe replace “non-natural variation” with “anthropogenic sources”.*
 
-**Response:** Suggested changed made.
+**Response:** Suggested changed made (line 18).
 
+**Comment:**
 *Line 61: Accidental line break. *
@@
 
+**Comment:**
 *Line 77: Suggest rephrasing as “…generalized linear models where the response variable is modeled as the sum of…”*
 
-**Response:** Suggested change made.
+**Response:** Suggested change made (line 80).
 
+**Comment:**
 *Tables 1 and 2: It would be useful to indicate the approximate sampling interval for the water quality parameters. It looks like quarterly. *
 
-**Response:** I'm out of space in the tables, but include this in the text **(addline number)**.
+**Response:** I'm out of space in the tables, but include this in the text (line 142-143).
 
+**Comment:**
 *Line 102: Some more information about the Bay would be helpful. What is the average depth? Any information on residence time? Is the water column mixed, stratified, or seasonally-stratified?*
 
-**Response:** I have not identified any information on residence time. I have added more information on average depth, mixing and stratification of the water column. **add line numbers**
+**Response:** I have not identified any information on residence time. I have added more information on average depth, mixing and stratification of the water column (lines 105-120)
 
+**Comment:**
 *Lines 128 – 129: The interpretation of the figures, tables, and text would be much easier if the estuarine sites were simply referred to as “upper”, “mid”, and “lower”.
@@
 
+**Response:**
+Fixed throughout.
+
+**Comment:**
 *Line 129 – 131: Some additional information about the sampling methods for the estuarine sites could be useful.
@@
 
+**Response:**
+Added in lines 139-143.
+
+**Comment:**
 *Line 138: Typo “tend”.*
@@
 
+**Comment:**
 *Figure 1: Remove “contribution” from the legend. *
@@
 
+**Comment:**
 *Eqn 2: I think some additional details are needed in the text below this equation.
@@
 For each day in the record, the full record of flows is utilized (an admittedly not efficient calculation, but not an issue for a period of record the size of this data set).
-(2) I set *d* to 0.95 based on results from Robson and Dourdet (2015) and Zhang and Ball (2017), which is now discussed in **lines ...**
+(2) I set *d* to 0.95 based on results from Robson and Dourdet (2015) and Zhang and Ball (2017), which is now discussed (lines 159-164).
 
+**Comment:**
 *Line 151: It’s not clear to me how fa is a unitless term. The equations show these as simple differences of average streamflow, so wouldn’t the units be in log-cfs? *
 
-**Response:** I worded this incorrectly. I should have phrased this "dimensionless." This is fixed in **lines ...**
+**Response:** I worded this incorrectly. I should have phrased this "dimensionless." This is fixed (line 165).
 
+**Comment:**
 *Line 169: What percentage of the observations for each parameter were non-detects?
@@
 **Response:** 4% of TP samples were below detection limits and 20% of NO~3~ 
-samples fell below detection limits.
+samples fell below detection limits (line 182).
 Early in this project 
@@
 
+**Comment:**
 *Line 236: This might be a personal preference, but “unusual periodic pattern” seems like regular periodicity and not a more random periodicity as one would expect with rain events.
@@
 
-**Response:** Rephrased to simply "periodic pattern."
+**Response:** Rephrased to simply "periodic pattern" (line 255).
 
+**Comment:**
 *Figure 2 caption: Consider adding a sentence like “Values closer to 1 for NSE and r and values closer to 0 for PBIAS represent more ideal goodness-of-fit assessments.”*
@@
 
+**Comment:**
 *Figure 3 caption: Consider adding Lavaca and Navidad with the station name, as in the caption for figure 4. *
 
+**Response:** Added Lavaca and Navidad to the captions.
+
+**Comment:**
 *Figure 5 caption: Consider including a reference to Table 3, as it wasn’t immediately clear to me that these models didn’t include flow or load. *
@@
 
+**Comment:**
 *Table 4: As for my previous comment, I’d replace the site names with “upper”, “mid”, and “lower” and arrange them accordingly.
@@
 
+**Comment:**
 *Line 272: The text “contributing tributaries” seems redundant. Maybe remove “contributing”. *
@@
 
+**Comment:**
 *Line 341: Replace “were” with “was”. *
@@
 
+**Comment:**
 *Line 395: Replace “suggest” with “such as”.*
@@
 
+**Comment:**
 *Line 420: Replace “provide” with “prove” or “be”.*
@@
 
-## References
+**References**
 

### 2023-08-11 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.Rmd
resubmission ready

@@
   Evaluating estuary water quality responses to reductions (or increases) in nutrient loading attributed to on the ground management actions can be challenging due to the strong influence of environmental drivers on nutrient loads and non-linear relationships.
-This study applied semiparametric approaches to calculate watershed nutrient loads and assess responses in estuary water quality to seasonally-adjusted freshwater inflow and flow-adjusted nutrient loads in Lavaca Bay, Texas.
+This study applied generalized additive models to calculate watershed nutrient loads and assess responses in estuary water quality to seasonally-adjusted freshwater inflow and flow-adjusted nutrient loads in Lavaca Bay, Texas.
 Lavaca Bay is a secondary embayment on the Texas coast displaying early potential for eutrophication and water quality degradation.
@@
 However, site specific long-term increases in both organic and inorganic nitrogen are concerning for their potential to fuel eutrophication.
-Further analysis found freshwater inflow had strong influences on nutrient and chlorophyll-*a* concentrations but there was no evidence that changes in watershed nutrient loading explained additional variation in dissolved oxygen or chlorophyll-*a* concentrations.
+Further analysis found freshwater inflow had strong influences on nutrient and chlorophyll-*a* concentrations but there was no evidence that changes in watershed nutrient loading explained additional variation in dissolved oxygen and limited evidence that watershed nutrient loadings explained chlorophyll-*a* concentrations.
 In addition to providing a baseline assessment of watershed nutrient loading and water quality responses in the Lavaca Bay watershed, this study provides methodological support for the use of semiparametric models in load regression models and estuary assessments.
@@
 total Kjeldahl nitrogen (TKN), and chlorophyll-*a* concentrations were identified
-at monitoring sites within Lavaca Bay [@bugica_water_2020]. These trends in 
-nutrient concentrations are concerning due to the role of nitrogen as a 
+at monitoring sites within Lavaca Bay [@bugica_water_2020]. While significant
+nitrogen increases were only identified in TKN, both organic and
+inorganic forms of nitrogen can serve as potential indicators of eutrophication
+status and estuary health [@jessenMarineEutrophication2015]. Nitrogen is 
+generally considered the 
 limiting factor for primary production in many Texas estuaries 
@@
 @doradoUnderstandingInteractionsFreshwater2015; 
-@paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017] and the 
-ramifications that changes in nitrogen loadings could have for 
-productivity and eutrophication in Lavaca Bay.
+@paudelRelationshipSuspendedSolids2019; @wetz_exceptionally_2017]. However, 
+estuaries can display nitrogen and phosphorus co-limitation that may vary
+temporally and spatially [@elserGlobalAnalysisNitrogen2007;
+@conleyControllingEutrophicationNitrogen2009], 
+emphasizing a need to assess a variety of nutrient species.
 
@@
 chlorophyll-*a*, and DO concentrations from January 2005 through December 2020 
-(Table\ \ref{tab:table2}). All water quality data was collected approximately 
-quarterly.
+(Table\ \ref{tab:table2}). Freshwater nutrient samples were collect 0.3 meters 
+below the surface or halfway between the surface and bottom if depth was 
+insufficient. Estuary nutrient samples were all collected 0.3 meters below the 
+surface. Estuary DO measurements were made at 3 locations along the water column 
+(0.3 meters, mid-depth, and bottom) and averaged. All water quality data was 
+collected approximately quarterly.
 
@@
 
-Left-censored data were not uncommon in this dataset. Several methods are
+Left-censored data were not uncommon in this dataset (4% of TP samples and 20%
+of NO~3~ samples). Several methods are
 available to account for censored data. We transformed left-censored nutrient 
@@
 
-```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.", dev='cairo_pdf'}
+```{r fig3, error=FALSE, warning=FALSE, message=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Aggregated estimated annual and flow-normalized annual NO\\textsubscript{3} and TP loads for the Lavaca (USGS-08164000) and and Navidad (USGS-08164525) Rivers.", dev='cairo_pdf'}
 
@@
 There was no evidence of long-term changes in TP or DO concentrations at any
-Lavaca Bay site (Fig.\ \ref{fig:fig5}). The upper-bay site, TCEQ-13563, had
+Lavaca Bay site (Fig.\ \ref{fig:fig5}). The upper-bay site (TCEQ-13563) had
 evidence of a long-term linear increase in NO~*X*~ while chlorophyll-*a*
 decreased from 2005 through 2014 (Fig.\ \ref{fig:fig5}). NO~*X*~ concentration
-at the mid-bay site, TCEQ-13383, displayed a periodic pattern that is
+at the mid-bay site (TCEQ-13383) displayed a periodic pattern that is
 indicative of a strong influence from inflow or precipitation. The temporal GAMs
 did not provide evidence of long-term trends in any of the water quality constituents
-at the lower-bay TCEQ-13384 site.
+at the lower-bay site (TCEQ-13384).
 
@@
 
-Freshwater inflow provided additional explanation for changes in TP and 
-NO~*X*~ concentration at all of the Lavaca Bay sites according to AIC~c~ and 
-model probability values (Table\ \ref{tab:table4}).
-TCEQ-13563, the site closest to the river outlet, was the only site that
-had improvements in the explanations of DO and TKN concentration with the inclusion 
-of inflow. Both TCEQ-13563 and TCEQ-13383, the mid-bay site, saw improvements 
-in explanations for variations in chlorophyll-*a* with the inclusion of freshwater
-inflow. The addition of nutrient loads (both TP and NO~3~) terms did not provide
-additional explanation for changes in chlorophyll-*a* or DO concentrations.
+Freshwater inflow provided additional explanation for changes in NO~x~ and 
+chlorophyll-*a* at all of the Lavaca Bay sites according to AIC~c~ and model 
+probability values (Table\ \ref{tab:table4}). Freshwater inflow also explained 
+additional variation in TP at the upper- and mid-bay sites but not the lower-bay 
+site. Freshwater inflow did not explain additional variation in TKN concentrations
+at any of the sites. The upper- and lower-bay sites saw improvements in explanation
+of DO concentration with the inclusion of inflow.
+
 Inclusion of TP loads provided additional explanation of TP concentrations at
-the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383. Inclusion of NO~3~
-loads only provided marginal improvements in the explanation of NO~*X*~ 
-concentration at the lower-bay TCEQ-13384 site.
+the upper- and mid-bay sites. Inclusion of NO~3~ loads did not provide model
+explanatory improvements at any of the sites. The addition of nutrient loads 
+(both TP and NO~3~) terms did not provide additional explanation for changes in 
+DO concentrations but did provide model improvement for the chlorophyll-*a* model
+at the mid-bay site.
+
+Increases in aggregated freshwater inflow resulted in increases 
+in TP at the upper- and mid-bay sites and increases in NO~*X*~ and 
+chlorophyll-*a*concentration at all three sites (Fig.\ \ref{fig:fig6}). Linear
+increases in DO concentration were observed with increasing flow at the upper-
+and lower-bay sites. TKN showed no response to changes in inflow at any of the sites.
+
+Increased TP loads resulted in nearly linear increases of TP concentration at
+the upper- and mid-bay sites (Fig.\ \ref{fig:fig7}).
+The relative effect size appeared much smaller than the effect of freshwater
+inflow alone. Increased NO~3~ loads did effect NO~x~ concentrations at any site.
+Linear increases in chlorophyll-*a* were observed in response to increased TP
+loads at the mid-bay site, but not the other sites or in response to NO~3~ loads.
 
@@
 
-Increases in freshwater inflow resulted in nearly linear increases 
-in TP and NO~*X*~ concentration at all three sites (Fig.\ \ref{fig:fig6}). 
-At the upper-bay TCEQ-13563 site, increases in freshwater inflow initially 
-increased chlorophyll-*a* and DO concentration but concentrations leveled and
-potentially decreased at higher flows. The mid-bay TCEQ-13383 site had
-nearly linear increases in chlorophyll-*a* concentration in response to increased
-freshwater inflow. There was no evidence of freshwater inflow effects on 
-chlorophyll-*a*, TKN, or DO at the lower-bay TCEQ-13384 site.
-
 ```{r fig6, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from estuary GAMs display the marginal smoothed effect of freshwater inflow (controlled for season) on TP (A), NO\\textit{\\textsubscript{X}} (B), chlorophyll-\\textit{a} (C), TKN (D), and DO (E) concentrations at each site in Lavaca Bay.", dev='cairo_pdf'}
@@
     mutate(lower_ci = est + (crit * se),
-           upper_ci = est - (crit * se)) |>
+           upper_ci = est - (crit * se),
+           model = forcats::fct_relevel(model, "Upper-Bay", "Mid-Bay", "Lower-Bay")) |>
     group_by(model) |> 
@@
 
-`TCEQ-13383` <- tar_read(tp_lavaca_13383_flow, store = store)
-
-comp <- compare_smooths(model = `TCEQ-13383`,
-                      `TCEQ-13384` = tar_read(tp_lavaca_13384_flow, store = store),
-                      `TCEQ-13563` = tar_read(tp_lavaca_13563_flow, store = store),
+`Mid-Bay` <- tar_read(tp_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(tp_lavaca_13384_flow, store = store),
+                      `Upper-Bay` = tar_read(tp_lavaca_13563_flow, store = store),
                       smooths = c("s(flw_res)"),
@@
 
-`TCEQ-13383` <- tar_read(no3_lavaca_13383_flow, store = store)
-
-comp <- compare_smooths(model = `TCEQ-13383`,
-                      `TCEQ-13384` = tar_read(no3_lavaca_13384_flow, store = store),
-                      `TCEQ-13563` = tar_read(no3_lavaca_13563_flow, store = store),
+`Mid-Bay` <- tar_read(no3_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(no3_lavaca_13384_flow, store = store),
+                      `Upper-Bay` = tar_read(no3_lavaca_13563_flow, store = store),
                       smooths = c("s(flw_res)"),
@@
 
-`TCEQ-13383` <- tar_read(chla_lavaca_13383_flow, store = store)
-
-comp <- compare_smooths(model = `TCEQ-13383`,
-                      `TCEQ-13384` = tar_read(chla_lavaca_13384_flow, store = store),
-                      `TCEQ-13563` = tar_read(chla_lavaca_13563_flow, store = store),
+`Mid-Bay` <- tar_read(chla_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(chla_lavaca_13384_flow, store = store),
+                      `Upper-Bay` = tar_read(chla_lavaca_13563_flow, store = store),
                       smooths = c("s(flw_res)"),
@@
 
-`TCEQ-13383` <- tar_read(tkn_lavaca_13383_flow, store = store)
-
-comp <- compare_smooths(model = `TCEQ-13383`,
-                      `TCEQ-13384` = tar_read(tkn_lavaca_13384_flow, store = store),
-                      `TCEQ-13563` = tar_read(tkn_lavaca_13563_flow, store = store),
+`Mid-Bay` <- tar_read(tkn_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(tkn_lavaca_13384_flow, store = store),
+                      `Upper-Bay` = tar_read(tkn_lavaca_13563_flow, store = store),
                       smooths = c("s(flw_res)"),
@@
 
-`TCEQ-13383` <- tar_read(do_lavaca_13383_flow, store = store)
-
-comp <- compare_smooths(model = `TCEQ-13383`,
-                      `TCEQ-13384` = tar_read(do_lavaca_13384_flow, store = store),
-                      `TCEQ-13563` = tar_read(do_lavaca_13563_flow, store = store),
+`Mid-Bay` <- tar_read(do_lavaca_13383_flow, store = store)
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(do_lavaca_13384_flow, store = store),
+                      `Upper-Bay` = tar_read(do_lavaca_13563_flow, store = store),
                       smooths = c("s(flw_res)"),
@@
 
-Increased TP loads resulted in nearly linear increases of TP concentration at
-the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383 respectively 
-(Fig.\ \ref{fig:fig7}).
-The relative effect size appeared much smaller than the effect of freshwater
-inflow alone. Increased NO~3~ loads only showed an effect at the 
-lower-bay TCEQ-13384 site. The effect was quite small compared to streamflow and
-provided only small improvements to the model (Table\ \ref{tab:table4}). As noted
-above, nutrient loadings did not provide any explanation in changes in the
-remaining assessed water quality parameters.
-
-```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.44, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\\textsubscript{3} loads on TP (A) and NO\\textsubscript{\\textit{X}} (B) concentrations at each site in Lavaca Bay.",  dev='cairo_pdf'}
-
-`TCEQ-13383` <- tar_read(tp_lavaca_13383_full, store = store)
-
-comp <- compare_smooths(model = `TCEQ-13383`,
-                      `TCEQ-13384` = tar_read(tp_lavaca_13384_full, store = store),
-                      `TCEQ-13563` = tar_read(tp_lavaca_13563_full, store = store),
+```{r fig7, message=FALSE, warning=FALSE, fig.width=6.85, fig.asp=0.7, fig.cap="Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of 20-day aggregated TP and NO\\textsubscript{3} flow-normalized loads on (A) TP, (B) NO\\textsubscript{\\textit{X}}, and (C,D) chlorophyll-\\textit{a} concentrations at each site in Lavaca Bay.",  dev='cairo_pdf'}
+
+`Mid-Bay` <- tar_read(tp_lavaca_13383_full, store = store)
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(tp_lavaca_13384_full, store = store),
+                      `Upper-Bay` = tar_read(tp_lavaca_13563_full, store = store),
                       smooths = c("s(sum_TP_FN)"),
@@
                       x = sum_TP_FN,
-                      xlab = "Residuals log(TP) Load [kg]",
+                      xlab = "Flow normalized TP Load [kg]",
                       ylab = "TP [mg/L]",
@@
 
-`TCEQ-13383` <- tar_read(no3_lavaca_13383_full, store = store)
-
-comp <- compare_smooths(model = `TCEQ-13383`,
-                      `TCEQ-13384` = tar_read(no3_lavaca_13384_full, store = store),
-                      `TCEQ-13563` = tar_read(no3_lavaca_13563_full, store = store),
+`Mid-Bay` <- tar_read(no3_lavaca_13383_full, store = store)
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(no3_lavaca_13384_full, store = store),
+                      `Upper-Bay` = tar_read(no3_lavaca_13563_full, store = store),
                       smooths = c("s(sum_NO3_FN)"),
@@
                       x = sum_NO3_FN,
-                      xlab = "Residuals log(NO<sub>3</sub>) Load [kg]",
+                      xlab = "Flow normalized NO<sub>3</sub> Load [kg]",
                       ylab = "NO<sub>*x*</sub> [mg/L]",
@@
 
+`Mid-Bay` <- tar_read(chla_lavaca_13383_full, store = store)
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(chla_lavaca_13384_full, store = store),
+                      `Upper-Bay` = tar_read(chla_lavaca_13563_full, store = store),
+                      smooths = c("s(sum_TP_FN)"),
+                      overall_uncertainty = FALSE)
+p_chla_tp <- draw_smoothed_surface(comp,
+                      x = sum_TP_FN,
+                      xlab = "Flow-normalized TP Load [kg]",
+                      ylab = "Chlorophyll-*a* [µg/L]",
+                      subtitle = "C")
+
+comp <- compare_smooths(model = `Mid-Bay`,
+                      `Lower-Bay` = tar_read(chla_lavaca_13384_full, store = store),
+                      `Upper-Bay` = tar_read(chla_lavaca_13563_full, store = store),
+                      smooths = c("s(sum_NO3_FN)"),
+                      overall_uncertainty = FALSE)
+p_chla_no3 <- draw_smoothed_surface(comp,
+                      x = sum_NO3_FN,
+                      xlab = "Flow-normalized NO<sub>3</sub> Load [kg]",
+                      ylab = "Chlorophyll-*a* [µg/L]",
+                      subtitle = "D")
+
 design <- "
-11223
+11225
+3344#
 "
 
-p_tp + p_no3 + guide_area() + plot_layout(guides = "collect",
+p_tp + p_no3 + p_chla_tp + p_chla_no3 + guide_area() + plot_layout(guides = "collect",
                                                    design = design)
@@
 
-The strong positive effect of freshwater inflow on NO~*X*~, TKN, and TP concentration
+The strong positive effect of freshwater inflow on NO~*X*~ and TP concentration
 are suggestive of nonpoint watershed sources, consistent with watershed uses and
@@
 @palmerImpactsDroughtsLow2015; @ciraPhytoplanktonDynamicsLowinflow2021]. Inflow 
-had a non-linear relationship with TKN at the two upstream sites, with TKN 
+had a non-linear relationship with NO~x~, with NO~x~ 
 increasing as freshwater inflow transitioned from low to moderate levels. 
-At higher freshwater inflows, the effect was attenuated, possibly indicating a 
-flushing effect at higher freshwater inflow. No relationship between TKN and 
-freshwater inflow was observed at TCEQ-13384 located in the lower reach of 
-Lavaca Bay. The results suggest that tidal flushing might dilute TKN and act as 
-a control on the effects of freshwater inflow in the lower reaches of Lavaca Bay.
-The results are also consistent with previous work that suggest processing of
-organic loads in the upper Lavaca Bay or tidal portions of the Lavaca River reduce
+At higher freshwater inflows, the NO~x~ decreased at the mid- and lower-bay sites, 
+possibly indicating a flushing effect at higher freshwater inflow. 
+No relationship between TKN and freshwater inflow was observed 
+at any of the sites. 
+The results coincide with previous work that suggest processing of
+organic loads in the tidal portions of the Lavaca River reduce
 transport of nutrients to the lower reaches of Lavaca Bay [@russell_effect_2006].
 
-Freshwater inflow also displayed a strong positive effect on chlorophyll-*a* at 
-the upper- and mid-bay sites. The upper-bay site, TCEQ-13563, showed decreases
+Freshwater inflow also displayed positive effects on chlorophyll-*a* at 
+each site, with the largest effect at the lower-bay site. 
+The lower-bay site also showed an attenuation in
 in chlorophyll-*a* at the highest freshwater inflow volumes. Freshwater flushing
-or increases in turbidity are associated with decreases in chlorophyll-*a* in
-other estuaries [@peierlsNonmonotonicResponsesPhytoplankton2012; 
+or increases in turbidity are associated with limitations or decreases in 
+chlorophyll-*a* in other estuaries [@peierlsNonmonotonicResponsesPhytoplankton2012; 
 @cloernPhytoplanktonPrimaryProduction2014]. No relationships between inorganic
-nitrogen or TP loadings with chlorophyll-*a* were observed. Due to the lack
+nitrogen and chlorophyll-*a* were observed. However, a small positive effect between
+flow-normalized TP loads and chlorophyll-*a* concentrations were detected at the
+mid-bay site. Although, this is not suggestive of a phosphorus limitation in the
+Bay, it is supportive of prior work emphasizing the importance of controlling both
+nitrogen and phosphorus runoff to protect water quality 
+[@conleyControllingEutrophicationNitrogen2009]. Due to the lack
 of TKN loading information, no assessment between organic nitrogen loads and
@@
 concentrations are strongly driven by freshwater inflow and to a lesser extent
-fluctuation in flow-adjusted riverine loadings. Site-specific changes in the
+fluctuation in flow-normalized riverine loadings. Site-specific changes in the
 relationships between freshwater inflow and responses in both chlorophyll-*a* 
-and TKN concentrations are indicative nutrient processing and or tidal flushing
+and NO~x~ concentrations are indicative of nutrient processing and or tidal flushing
 effects moving from the river discharge point to the mouth of Lavaca Bay.
@@
 
-## Limitations {.unnumbered}
+## Future Management and Research Needs {.unnumbered}
 
@@
 sources of nutrients and (2) that there is not strong evidence that those small 
-changes have had effects on chlorophyll-*a* or dissolved oxygen in Lavaca Bay.
+changes have had extensive effects on chlorophyll-*a* or dissolved oxygen in Lavaca Bay.
 Although this study did not identify changes in DO or chlorophyll-*a* concentrations

### 2023-08-11 Schramm-2023-06-PeerJ/Schramm-2023-08-PeerJ.tex
resubmission ready

@@
 nutrient loads and non-linear relationships. This study applied
-semiparametric approaches to calculate watershed nutrient loads and
+generalized additive models to calculate watershed nutrient loads and
 assess responses in estuary water quality to seasonally-adjusted
@@
 changes in watershed nutrient loading explained additional variation in
-dissolved oxygen or chlorophyll-\emph{a} concentrations. In addition to
-providing a baseline assessment of watershed nutrient loading and water
-quality responses in the Lavaca Bay watershed, this study provides
+dissolved oxygen and limited evidence that watershed nutrient loadings
+explained chlorophyll-\emph{a} concentrations. In addition to providing
+a baseline assessment of watershed nutrient loading and water quality
+responses in the Lavaca Bay watershed, this study provides
 methodological support for the use of semiparametric models in load
@@
 concentrations were identified at monitoring sites within Lavaca Bay
-\autocite{bugica_water_2020}. These trends in nutrient concentrations
-are concerning due to the role of nitrogen as a limiting factor for
-primary production in many Texas estuaries
-\autocite{gardnerNitrogenFixationDissimilatory2006,houTransformationFateNitrate2012,doradoUnderstandingInteractionsFreshwater2015,paudelRelationshipSuspendedSolids2019,wetz_exceptionally_2017}
-and the ramifications that changes in nitrogen loadings could have for
-productivity and eutrophication in Lavaca Bay.
+\autocite{bugica_water_2020}. While significant nitrogen increases were
+only identified in TKN, both organic and inorganic forms of nitrogen can
+serve as potential indicators of eutrophication status and estuary
+health \autocite{jessenMarineEutrophication2015}. Nitrogen is generally
+considered the limiting factor for primary production in many Texas
+estuaries
+\autocite{gardnerNitrogenFixationDissimilatory2006,houTransformationFateNitrate2012,doradoUnderstandingInteractionsFreshwater2015,paudelRelationshipSuspendedSolids2019,wetz_exceptionally_2017}.
+However, estuaries can display nitrogen and phosphorus co-limitation
+that may vary temporally and spatially
+\autocite{elserGlobalAnalysisNitrogen2007,conleyControllingEutrophicationNitrogen2009},
+emphasizing a need to assess a variety of nutrient species.
 
@@
 concentrations from January 2005 through December 2020
-(Table~\ref{tab:table2}). All water quality data was collected
-approximately quarterly.
+(Table~\ref{tab:table2}). Freshwater nutrient samples were collect 0.3
+meters below the surface or halfway between the surface and bottom if
+depth was insufficient. Estuary nutrient samples were all collected 0.3
+meters below the surface. Estuary DO measurements were made at 3
+locations along the water column (0.3 meters, mid-depth, and bottom) and
+averaged. All water quality data was collected approximately quarterly.
 
@@
 
-Left-censored data were not uncommon in this dataset. Several methods
-are available to account for censored data. We transformed left-censored
-nutrient concentrations to one-half the detection limit. Although this
-simple approach can introduce bias
+Left-censored data were not uncommon in this dataset (4\% of TP samples
+and 20\% of NO\textsubscript{3} samples). Several methods are available
+to account for censored data. We transformed left-censored nutrient
+concentrations to one-half the detection limit. Although this simple
+approach can introduce bias
 \autocite{hornungEstimationAverageConcentration1990}, we considered it
@@
 
-\caption{Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for USGS-08164000 and USGS-08164525.}\label{fig:fig3}
+\caption{Aggregated estimated annual and flow-normalized annual NO\textsubscript{3} and TP loads for the Lavaca (USGS-08164000) and and Navidad (USGS-08164525) Rivers.}\label{fig:fig3}
 \end{figure}
@@
 There was no evidence of long-term changes in TP or DO concentrations at
-any Lavaca Bay site (Fig.~\ref{fig:fig5}). The upper-bay site,
-TCEQ-13563, had evidence of a long-term linear increase in
+any Lavaca Bay site (Fig.~\ref{fig:fig5}). The upper-bay site
+(TCEQ-13563) had evidence of a long-term linear increase in
 NO\textsubscript{\emph{X}} while chlorophyll-\emph{a} decreased from
 2005 through 2014 (Fig.~\ref{fig:fig5}). NO\textsubscript{\emph{X}}
-concentration at the mid-bay site, TCEQ-13383, displayed a periodic
+concentration at the mid-bay site (TCEQ-13383) displayed a periodic
 pattern that is indicative of a strong influence from inflow or
 precipitation. The temporal GAMs did not provide evidence of long-term
-trends in any of the water quality constituents at the lower-bay
-TCEQ-13384 site.
+trends in any of the water quality constituents at the lower-bay site
+(TCEQ-13384).
 
@@
 
-Freshwater inflow provided additional explanation for changes in TP and
-NO\textsubscript{\emph{X}} concentration at all of the Lavaca Bay sites
-according to AIC\textsubscript{c} and model probability values
-(Table~\ref{tab:table4}). TCEQ-13563, the site closest to the river
-outlet, was the only site that had improvements in the explanations of
-DO and TKN concentration with the inclusion of inflow. Both TCEQ-13563
-and TCEQ-13383, the mid-bay site, saw improvements in explanations for
-variations in chlorophyll-\emph{a} with the inclusion of freshwater
-inflow. The addition of nutrient loads (both TP and NO\textsubscript{3})
-terms did not provide additional explanation for changes in
-chlorophyll-\emph{a} or DO concentrations. Inclusion of TP loads
-provided additional explanation of TP concentrations at the upper- and
-mid-bay sites, TCEQ-13563 and TCEQ-13383. Inclusion of
-NO\textsubscript{3} loads only provided marginal improvements in the
-explanation of NO\textsubscript{\emph{X}} concentration at the lower-bay
-TCEQ-13384 site.
+Freshwater inflow provided additional explanation for changes in
+NO\textsubscript{x} and chlorophyll-\emph{a} at all of the Lavaca Bay
+sites according to AIC\textsubscript{c} and model probability values
+(Table~\ref{tab:table4}). Freshwater inflow also explained additional
+variation in TP at the upper- and mid-bay sites but not the lower-bay
+site. Freshwater inflow did not explain additional variation in TKN
+concentrations at any of the sites. The upper- and lower-bay sites saw
+improvements in explanation of DO concentration with the inclusion of
+inflow.
+
+Inclusion of TP loads provided additional explanation of TP
+concentrations at the upper- and mid-bay sites. Inclusion of
+NO\textsubscript{3} loads did not provide model explanatory improvements
+at any of the sites. The addition of nutrient loads (both TP and
+NO\textsubscript{3}) terms did not provide additional explanation for
+changes in DO concentrations but did provide model improvement for the
+chlorophyll-\emph{a} model at the mid-bay site.
+
+Increases in aggregated freshwater inflow resulted in increases in TP at
+the upper- and mid-bay sites and increases in NO\textsubscript{\emph{X}}
+and chlorophyll-\emph{a}concentration at all three sites
+(Fig.~\ref{fig:fig6}). Linear increases in DO concentration were
+observed with increasing flow at the upper- and lower-bay sites. TKN
+showed no response to changes in inflow at any of the sites.
+
+Increased TP loads resulted in nearly linear increases of TP
+concentration at the upper- and mid-bay sites (Fig.~\ref{fig:fig7}). The
+relative effect size appeared much smaller than the effect of freshwater
+inflow alone. Increased NO\textsubscript{3} loads did effect
+NO\textsubscript{x} concentrations at any site. Linear increases in
+chlorophyll-\emph{a} were observed in response to increased TP loads at
+the mid-bay site, but not the other sites or in response to
+NO\textsubscript{3} loads.
 
@@
 
-Increases in freshwater inflow resulted in nearly linear increases in TP
-and NO\textsubscript{\emph{X}} concentration at all three sites
-(Fig.~\ref{fig:fig6}). At the upper-bay TCEQ-13563 site, increases in
-freshwater inflow initially increased chlorophyll-\emph{a} and DO
-concentration but concentrations leveled and potentially decreased at
-higher flows. The mid-bay TCEQ-13383 site had nearly linear increases in
-chlorophyll-\emph{a} concentration in response to increased freshwater
-inflow. There was no evidence of freshwater inflow effects on
-chlorophyll-\emph{a}, TKN, or DO at the lower-bay TCEQ-13384 site.
-
 \begin{figure}
@@
 
-Increased TP loads resulted in nearly linear increases of TP
-concentration at the upper- and mid-bay sites, TCEQ-13563 and TCEQ-13383
-respectively (Fig.~\ref{fig:fig7}). The relative effect size appeared
-much smaller than the effect of freshwater inflow alone. Increased
-NO\textsubscript{3} loads only showed an effect at the lower-bay
-TCEQ-13384 site. The effect was quite small compared to streamflow and
-provided only small improvements to the model (Table~\ref{tab:table4}).
-As noted above, nutrient loadings did not provide any explanation in
-changes in the remaining assessed water quality parameters.
-
 \begin{figure}
@@
 
-\caption{Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of TP and NO\textsubscript{3} loads on TP (A) and NO\textsubscript{\textit{X}} (B) concentrations at each site in Lavaca Bay.}\label{fig:fig7}
+\caption{Fitted splines from the nutrient loading GAMs display the marginal smoothed effect of 20-day aggregated TP and NO\textsubscript{3} flow-normalized loads on (A) TP, (B) NO\textsubscript{\textit{X}}, and (C,D) chlorophyll-\textit{a} concentrations at each site in Lavaca Bay.}\label{fig:fig7}
 \end{figure}
@@
 The strong positive effect of freshwater inflow on
-NO\textsubscript{\emph{X}}, TKN, and TP concentration are suggestive of
+NO\textsubscript{\emph{X}} and TP concentration are suggestive of
 nonpoint watershed sources, consistent with watershed uses and with
@@
 \autocite{russell_effect_2006,caffreyHighNutrientPulses2007,peierlsNonmonotonicResponsesPhytoplankton2012,palmerImpactsDroughtsLow2015,ciraPhytoplanktonDynamicsLowinflow2021}.
-Inflow had a non-linear relationship with TKN at the two upstream sites,
-with TKN increasing as freshwater inflow transitioned from low to
-moderate levels. At higher freshwater inflows, the effect was
-attenuated, possibly indicating a flushing effect at higher freshwater
-inflow. No relationship between TKN and freshwater inflow was observed
-at TCEQ-13384 located in the lower reach of Lavaca Bay. The results
-suggest that tidal flushing might dilute TKN and act as a control on the
-effects of freshwater inflow in the lower reaches of Lavaca Bay. The
-results are also consistent with previous work that suggest processing
-of organic loads in the upper Lavaca Bay or tidal portions of the Lavaca
-River reduce transport of nutrients to the lower reaches of Lavaca Bay
+Inflow had a non-linear relationship with NO\textsubscript{x}, with
+NO\textsubscript{x} increasing as freshwater inflow transitioned from
+low to moderate levels. At higher freshwater inflows, the
+NO\textsubscript{x} decreased at the mid- and lower-bay sites, possibly
+indicating a flushing effect at higher freshwater inflow. No
+relationship between TKN and freshwater inflow was observed at any of
+the sites. The results coincide with previous work that suggest
+processing of organic loads in the tidal portions of the Lavaca River
+reduce transport of nutrients to the lower reaches of Lavaca Bay
 \autocite{russell_effect_2006}.
 
-Freshwater inflow also displayed a strong positive effect on
-chlorophyll-\emph{a} at the upper- and mid-bay sites. The upper-bay
-site, TCEQ-13563, showed decreases in chlorophyll-\emph{a} at the
-highest freshwater inflow volumes. Freshwater flushing or increases in
-turbidity are associated with decreases in chlorophyll-\emph{a} in other
-estuaries
+Freshwater inflow also displayed positive effects on
+chlorophyll-\emph{a} at each site, with the largest effect at the
+lower-bay site. The lower-bay site also showed an attenuation in in
+chlorophyll-\emph{a} at the highest freshwater inflow volumes.
+Freshwater flushing or increases in turbidity are associated with
+limitations or decreases in chlorophyll-\emph{a} in other estuaries
 \autocite{peierlsNonmonotonicResponsesPhytoplankton2012,cloernPhytoplanktonPrimaryProduction2014}.
-No relationships between inorganic nitrogen or TP loadings with
-chlorophyll-\emph{a} were observed. Due to the lack of TKN loading
-information, no assessment between organic nitrogen loads and
-chlorophyll-\emph{a} were possible.
+No relationships between inorganic nitrogen and chlorophyll-\emph{a}
+were observed. However, a small positive effect between flow-normalized
+TP loads and chlorophyll-\emph{a} concentrations were detected at the
+mid-bay site. Although, this is not suggestive of a phosphorus
+limitation in the Bay, it is supportive of prior work emphasizing the
+importance of controlling both nitrogen and phosphorus runoff to protect
+water quality \autocite{conleyControllingEutrophicationNitrogen2009}.
+Due to the lack of TKN loading information, no assessment between
+organic nitrogen loads and chlorophyll-\emph{a} were possible.
 
@@
 driven by freshwater inflow and to a lesser extent fluctuation in
-flow-adjusted riverine loadings. Site-specific changes in the
+flow-normalized riverine loadings. Site-specific changes in the
 relationships between freshwater inflow and responses in both
-chlorophyll-\emph{a} and TKN concentrations are indicative nutrient
-processing and or tidal flushing effects moving from the river discharge
-point to the mouth of Lavaca Bay. This study does not completely explain
-site specific increases in NO\textsubscript{\emph{X}} and TKN
-concentrations in Lavaca Bay. The freshwater study sites did not
-quantify nutrient loadings from tidal contribution areas or ungauaged
-watersheds. Nutrient contributions from wastewater facility discharges,
-septic systems, and stormwater could be considerable contributors to
-nutrient loadings in Lavaca Bay since they are not processed by a tidal
-river reach prior to entering the Bay. The Garcitas-Arenosa Creek,
-Placedo Creek, and Cox Bay subwatersheds are currently undersampled but
-compose approximately 27\% of the watershed area. The contribution of
-nutrient loadings from these undersampled areas is unknown.
-
-\hypertarget{limitations}{%
-\subsection*{Limitations}\label{limitations}}
-\addcontentsline{toc}{subsection}{Limitations}
+chlorophyll-\emph{a} and NO\textsubscript{x} concentrations are
+indicative of nutrient processing and or tidal flushing effects moving
+from the river discharge point to the mouth of Lavaca Bay. This study
+does not completely explain site specific increases in
+NO\textsubscript{\emph{X}} and TKN concentrations in Lavaca Bay. The
+freshwater study sites did not quantify nutrient loadings from tidal
+contribution areas or ungauaged watersheds. Nutrient contributions from
+wastewater facility discharges, septic systems, and stormwater could be
+considerable contributors to nutrient loadings in Lavaca Bay since they
+are not processed by a tidal river reach prior to entering the Bay. The
+Garcitas-Arenosa Creek, Placedo Creek, and Cox Bay subwatersheds are
+currently undersampled but compose approximately 27\% of the watershed
+area. The contribution of nutrient loadings from these undersampled
+areas is unknown.
+
+\hypertarget{future-management-and-research-needs}{%
+\subsection*{Future Management and Research
+Needs}\label{future-management-and-research-needs}}
+\addcontentsline{toc}{subsection}{Future Management and Research Needs}
 
@@
 changes in non-point sources of nutrients and (2) that there is not
-strong evidence that those small changes have had effects on
+strong evidence that those small changes have had extensive effects on
 chlorophyll-\emph{a} or dissolved oxygen in Lavaca Bay. Although this

### 2023-08-11 map.R

@@ -41,6 +41,13 @@ tceq_sites <- st_as_sf(tceq_sites, coords = c("LongitudeMeasure", "LatitudeMeasu
 
 tceq_sites$MonitoringLocationIdentifier <- stringr::str_replace_all(tceq_sites$MonitoringLocationIdentifier, "MAIN", "")
 
+tceq_sites <- tceq_sites |> 
+  mutate(MonitoringLocationIdentifier = case_when(
+    MonitoringLocationIdentifier == "TCEQ-13563" ~ "TCEQ-13563\nUpper-Bay",
+    MonitoringLocationIdentifier == "TCEQ-13384" ~ "TCEQ-13384\nLower-Bay",
+    MonitoringLocationIdentifier == "TCEQ-13383" ~ "TCEQ-13383\nMid-Bay"
+  ))
+
 ## read rivers spatial data
 rivers <- read_sf("data/Spatial/lavaca.gpkg",
                   layer = "rivers")
@@ -211,7 +218,10 @@ p1 <- ggplot() +
                   hjust = 1,
                   min.segment.length = 10000,
                   color = "grey30",
-                  bg.color = "white") +
+                  bg.color = "white",
+                  nudge_x = 10000,
+                  nudge_y = -5000,
+                  ylim = c(-Inf, Inf)) +
   ## Lavaca Bay Label
   geom_text_repel(data = waterbody_labs |> filter(lab == "Lavaca Bay"), 
                   aes(label = lab, geometry = geometry), 

### 2023-08-11 map_final_report.R

@@ -44,6 +44,14 @@ tceq_sites <- st_as_sf(tceq_sites, coords = c("LongitudeMeasure", "LatitudeMeasu
 
 tceq_sites$MonitoringLocationIdentifier <- stringr::str_replace_all(tceq_sites$MonitoringLocationIdentifier, "MAIN", "")
 
+
+tceq_sites <- tceq_sites |> 
+  mutate(MonitoringLocationIdentifier = case_when(
+    MonitoringLocationIdentifier == "TCEQ-13563" ~ "TCEQ-13563\nUpper-Bay",
+    MonitoringLocationIdentifier == "TCEQ-13384" ~ "TCEQ-13384\nLower-Bay",
+    MonitoringLocationIdentifier == "TCEQ-13383" ~ "TCEQ-13383\nMid-Bay"
+  ))
+
 ## read rivers spatial data
 rivers <- read_sf("data/Spatial/lavaca.gpkg",
                   layer = "rivers")
@@ -207,6 +215,7 @@ p1 <- ggplot() +
   ## Matagorda Bay Label
   geom_text_repel(data = waterbody_labs |> filter(lab == "Matagorda Bay"), 
                   aes(label = lab, geometry = geometry), 
+                  point.size = NA,
                   size = 2, 
                   family = "Arial",
                   fontface = "bold",
@@ -214,7 +223,10 @@ p1 <- ggplot() +
                   hjust = 1,
                   min.segment.length = 10000,
                   color = "grey30",
-                  bg.color = "white") +
+                  bg.color = "white",
+                  nudge_x = 10000,
+                  nudge_y = -5000,
+                  ylim = c(-Inf, Inf)) +
   ## Lavaca Bay Label
   geom_text_repel(data = waterbody_labs |> filter(lab == "Lavaca Bay"), 
                   aes(label = lab, geometry = geometry), 

### 2023-08-14 figs-eps.R
clean up files

@@ -1,944 +1,0 @@
-## manuscript figs
-
-
-library(targets)
-library(knitr)
-library(tidyverse)
-library(kableExtra)
-library(patchwork)
-library(ggrepel)
-library(rcartocolor)
-library(twriTemplates)
-library(ggspatial)
-library(ragg)
-library(flextable)
-library(modelsummary)
-library(scico)
-library(ggridges)
-library(ggtext)
-library(lubridate)
-library(units)
-library(gratia)
-library(gt)
-library(extrafont)
-library(systemfonts)
-
-store <- "C:/Data-Analysis-Projects/lavaca-nutrients/_targets"
-# Fig 2 -----------------------------------------------------------------------
-
-## some functions to overide scales in facets via dewey dunnington
-#https://dewey.dunnington.ca/post/2018/modifying-facet-scales-in-ggplot2/
-scale_override <- function(which, scale) {
-  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
-    stop("which must be an integer of length 1")
-  }
-  
-  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
-    stop("scale must be an x or y position scale")
-  }
-  
-  structure(list(which = which, scale = scale), class = "scale_override")
-}
-
-CustomFacetWrap <- ggproto(
-  "CustomFacetWrap", FacetWrap,
-  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
-    # make the initial x, y scales list
-    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)
-    
-    if(is.null(params$scale_overrides)) return(scales)
-    
-    max_scale_x <- length(scales$x)
-    max_scale_y <- length(scales$y)
-    
-    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
-    for(scale_override in params$scale_overrides) {
-      which <- scale_override$which
-      scale <- scale_override$scale
-      
-      if("x" %in% scale$aesthetics) {
-        if(!is.null(scales$x)) {
-          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
-          scales$x[[which]] <- scale$clone()
-        }
-      } else if("y" %in% scale$aesthetics) {
-        if(!is.null(scales$y)) {
-          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
[…]

### 2023-08-14 README.Rmd
snapshot and readme

---
title: "Lavaca Bay nutrient loading manuscript files"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This repo contains the files used to generate:

Schramm, M. TBD. Linking watershed nutrient loading to estuary water quality with Generalized Additive Models.

The rmarkdown files require LaTeX and the [lavaca-nutrients](https://github.com/TxWRI/lavaca-nutrients) repo for targets to work properly.

```{r}
renv::diagnostics()
```

### 2023-08-24 map_final_report.R
accepted version

@@ -1,350 +1,0 @@
-library(targets)
-library(tidyverse)
-library(ragg)
-library(magick)
-library(sf)
-library(ggrepel)
-library(patchwork)
-library(dataRetrieval)
-#library(FedData)
-library(ggspatial)
-library(rcartocolor)
-library(twriTemplates)
-
-download_usgs_sites <- function() {
-  site_numbers <- c("08164000", #lavaca
-                    "08164390", #navidad
-                    "08164450", #sandy
-                    "08164503", #w mustang
-                    "08164504", #e mustang
-                    "08164525")
-  sites <- readNWISsite(site_numbers)
-  sites
-}
-
-download_tceq_sites <-function() {
-  site_numbers <- c("TCEQMAIN-13563",
-                    "TCEQMAIN-13383",
-                    "TCEQMAIN-13384")
-  sites <- whatWQPsites(siteid = site_numbers)
-  sites
-}
-
-
-## usgs sites
-df_sites <- download_usgs_sites()
-
-df_sites <- st_as_sf(df_sites, coords = c("dec_long_va", "dec_lat_va"),
-                     crs = 4326)  
-
-## tceq sites
-tceq_sites <- download_tceq_sites()
-tceq_sites <- st_as_sf(tceq_sites, coords = c("LongitudeMeasure", "LatitudeMeasure"),
-                       crs = 4326)
-
-tceq_sites$MonitoringLocationIdentifier <- stringr::str_replace_all(tceq_sites$MonitoringLocationIdentifier, "MAIN", "")
-
-
-tceq_sites <- tceq_sites |> 
-  mutate(MonitoringLocationIdentifier = case_when(
-    MonitoringLocationIdentifier == "TCEQ-13563" ~ "TCEQ-13563\nUpper-Bay",
-    MonitoringLocationIdentifier == "TCEQ-13384" ~ "TCEQ-13384\nLower-Bay",
-    MonitoringLocationIdentifier == "TCEQ-13383" ~ "TCEQ-13383\nMid-Bay"
-  ))
-
-## read rivers spatial data
-rivers <- read_sf("data/Spatial/lavaca.gpkg",
-                  layer = "rivers")
-
-## read lake texana polygon
-waterbody <- read_sf("data/Spatial/lavaca.gpkg",
-                     layer = "waterbody")
-
-## read watershed polygons
-lavaca_ws <- read_sf("data/Spatial/lavaca.gpkg",
[…]

### 2023-10-12 README.Rmd
accepted version
add manuscript and snapshot

@@
 
-This repo contains the files used to generate:
+This repo contains the rmarkdown manuscript files used to generate the following accepted manuscript:
 
-Schramm, M. TBD. Linking watershed nutrient loading to estuary water quality with Generalized Additive Models.
+Schramm MP. 2023. Linking watershed nutrient loading to estuary water quality with generalized additive models. PeerJ 11:e16073 https://doi.org/10.7717/peerj.16073
 
